### 2014-02-25 doc/proposal/proposal.Rmd
Proposal draft

intRo
========================================================

### Motivation 
To create a web-based tool to be used in introductory statistics class in lieu of JMP. Ultimately, our goal is to create an interest in coding while learning basic statistics, or at a minimum, help students to learn a bit more about working with data.

As a web-based application, this tool will be immediately more familiar to students than a desktop application. The need for dealing with software licenses, installation configuration, and supported platforms will be gone. This will allow students to spend more time working with the data and learning statistics than having to struggle to get the software running.

The application will combine several different modules which perform specific data analysis tasks.

### Modules
* Data Import (i.e., displaying data, subsets...)
* Descriptive Summaries (i.e., mean, median, variance, IQR...)
* Graphical Summaries (i.e., bar chart, pareto chart, scatterplot, histogram...)
* Statistical Tests (i.e., t test, z test...)
* Simulation (i.e., central limit theorem...)
* Linear Regression (i.e., fitting, plotting, least squares, prediction...)

Students will be able to view the R code that each module generates. When changing configuration options in each module, the students will be able to observe the specific changes occurring in the R code. Students will be able to import their own dataset, or work with datasets that are provided in the application.

### Class Deliverables
We aim to have a prototype of this application with Data Import, Descriptive Summaries, and Graphical Summaries completed for the class. We will include accompanying documentation for using intRo.

### 2014-11-10 modules/contingency/observe.R
Switch to selectize for inputs

@@ -1,4 +1,4 @@
     observe({
-        updateSelectInput(session, "xcont", choices = intro.categoricnames(), selected = ifelse(checkVariable(intro.data(), input$xcont), input$xcont, intro.categoricnames()[1]))
-        updateSelectInput(session, "ycont", choices = intro.categoricnames(), selected = ifelse(checkVariable(intro.data(), input$ycont), input$ycont, intro.categoricnames()[2]))
+        updateSelectizeInput(session, "xcont", choices = intro.categoricnames(), selected = ifelse(checkVariable(intro.data(), input$xcont), input$xcont, intro.categoricnames()[1]))
+        updateSelectizeInput(session, "ycont", choices = intro.categoricnames(), selected = ifelse(checkVariable(intro.data(), input$ycont), input$ycont, intro.categoricnames()[2]))
     })

### 2014-11-10 modules/contingency/ui.R

@@ -1,12 +1,12 @@
 contingency_ui <- tabPanel("Contingency",
                                                   column(4,
                                                          wellPanel(
-                                                             selectInput("xcont", "X Variable (x)", choices = categoricNames(mpg), selected = categoricNames(mpg)[5]),
-                                                             selectInput("ycont", "Y Variable (y)", choices = categoricNames(mpg), selected = categoricNames(mpg)[6]),
+                                                             selectizeInput("xcont", label = "X Variable (x)", choices = categoricNames(mpg), selected = categoricNames(mpg)[5]),
+                                                             selectizeInput("ycont", label = "Y Variable (y)", choices = categoricNames(mpg), selected = categoricNames(mpg)[6]),
                                                              
                                                              hr(),
                                                              
-                                                             radioButtons("conttype", "Type", choices = c("Counts" = "counts", "Row Percentages" = "rowpercs", "Column Percentages" = "columnpercs", "Total Percentages" = "totalpercs"))
+                                                             selectizeInput("conttype", label = "Type", choices = c("Counts" = "counts", "Row Percentages" = "rowpercs", "Column Percentages" = "columnpercs", "Total Percentages" = "totalpercs"))
                                                          )
                                                   ),
                                                   

### 2014-11-10 modules/graphical/observe.R

@@ -1,5 +1,5 @@
 observe({
     input$plottype
-    updateSelectInput(session, "x", choices = x_choices(), selected = x_selected())
-    updateSelectInput(session, "y", choices = y_choices(), selected = y_selected())
+    updateSelectizeInput(session, "x", choices = x_choices(), selected = x_selected())
+    updateSelectizeInput(session, "y", choices = y_choices(), selected = y_selected())
 })

### 2014-11-10 modules/graphical/ui.R

@@ -1,18 +1,18 @@
 graphical_ui <- tabPanel("Graphical",
                                                   column(4,
                                                          wellPanel(
-                                                             selectInput("plottype", "Plot Type", choices = c("Histogram" = "histogram", "Normal Quantile Plot" = "quantileplot", "Scatterplot" = "scatterplot", "Line Chart" = "linechart", "Boxplot" = "boxplot", "Bar Chart" = "barchart", "Pareto Chart" = "paretochart", "Mosaic Plot" = "mosaicplot")),
+                                                             selectizeInput("plottype", label = "Plot Type", choices = c("Histogram" = "histogram", "Normal Quantile Plot" = "quantileplot", "Scatterplot" = "scatterplot", "Line Chart" = "linechart", "Boxplot" = "boxplot", "Bar Chart" = "barchart", "Pareto Chart" = "paretochart", "Mosaic Plot" = "mosaicplot")),
                                                              
                                                              hr(),
                                                              
-                                                             selectInput("x", "X Variable (x)", choices = numericNames(mpg), selected = numericNames(mpg)[1]),
+                                                             selectizeInput("x", label = "X Variable (x)", choices = numericNames(mpg), selected = numericNames(mpg)[1]),
                                                              #conditionalPanel(
                                                              #    condition = "input.plottype == 'barchart' || input.plottype == 'paretochart'",
                                                              #    checkboxInput("addy", "Y Variable")
                                                              #),
                                                              conditionalPanel(
                                                                  condition = "(input.plottype != 'histogram' && input.plottype != 'quantileplot' && input.plottype != 'barchart' && input.plottype != 'paretochart') || input.addy == true",
-                                                                 selectInput("y", "Y Variable (y)", choices = numericNames(mpg), selected = numericNames(mpg)[2])
+                                                                 selectizeInput("y", label = "Y Variable (y)", choices = numericNames(mpg), selected = numericNames(mpg)[2])
                                                              ),
                                                              
                                                              hr(),
@@ -70,10 +70,10 @@ graphical_ui <- tabPanel("Graphical",
                                                                           plotOutput("mosaicplot")
                                                          ),
                                                          conditionalPanel(condition = "input.plottype == 'barchart'",
-                                                                          plotOutput("barchart")
+                                                                          ggvisOutput("barchart")
                                                          ),
                                                          conditionalPanel(condition = "input.plottype == 'paretochart'",
-                                                                          plotOutput("paretochart")
+                                                                          ggvisOutput("paretochart")
[…]

### 2014-11-10 modules/numerical/observe.R

@@ -1,4 +1,4 @@
     observe({
-        updateSelectInput(session, "grouping", choices = c("None" = "none", names(intro.data())))
-        updateCheckboxGroupInput(session, "tblvars", choices = names(intro.data()))
+        updateSelectizeInput(session, "grouping", choices = c("None" = "none", names(intro.data())))
+        updateSelectizeInput(session, "tblvars", choices = names(intro.data()))
     })

### 2014-11-10 modules/numerical/ui.R

@@ -1,8 +1,8 @@
 numerical_ui <- tabPanel("Numerical",
                                                   column(4,
                                                          wellPanel(
-                                                             checkboxGroupInput("tblvars", "Select Variables", choices = names(mpg)),
-                                                             selectInput("grouping", "Select Grouping Variable", choices = names(mpg))
+                                                             selectizeInput("tblvars", label = "Select Variables", choices = names(mpg), multiple = TRUE),
+                                                             selectizeInput("grouping", label = "Select Grouping Variable", choices = names(mpg))
                                                          )
                                                   ),
                                                   

### 2014-11-10 modules/regression/observe.R

@@ -1,5 +1,5 @@
 ## Regression Observers
 observe({
-  updateSelectInput(session, "xreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$xreg), input$xreg, intro.numericnames()[1]))
-  updateSelectInput(session, "yreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$yreg), input$yreg, intro.numericnames()[2]))
+  updateSelectizeInput(session, "xreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$xreg), input$xreg, intro.numericnames()[1]))
+  updateSelectizeInput(session, "yreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$yreg), input$yreg, intro.numericnames()[2]))
 })

### 2014-11-10 modules/regression/ui.R

@@ -1,12 +1,12 @@
 regression_ui <- tabPanel("Regression",
          column(4,
                 wellPanel(
-                  selectInput("xreg", "Independent Variable (x)", choices = numericNames(mpg), selected = numericNames(mpg)[4]),
-                  selectInput("yreg", "Dependent Variable (y)", choices = numericNames(mpg), selected = numericNames(mpg)[5]),
+                  selectizeInput("xreg", label = "Independent Variable (x)", choices = numericNames(mpg), selected = numericNames(mpg)[4]),
+                  selectizeInput("yreg", label = "Dependent Variable (y)", choices = numericNames(mpg), selected = numericNames(mpg)[5]),
                   
                   hr(),
+                  
                   tags$button("", id = "saveresid", type = "button", class = "btn action-button", list(icon("save"), "Save Residuals/Fitted"), onclick = "$('#side-nav :contains(\"Sources\")').highlight();")
-                  #actionButton("saveresid", "Save Residuals/Fitted", icon = icon("save"))
                 )
          ),
          

### 2014-11-10 modules/sources/observe.R

@@ -1,4 +1,3 @@
 observe({
     updateNumericInput(session, "randomsubrows", max = nrow(intro.data()))
-    updateCheckboxGroupInput(session, "tblvars", choices = names(intro.data()))
 })

### 2014-11-10 modules/sources/ui.R

@@ -3,7 +3,7 @@ sources_ui <- tabPanel("Sources",
                                                          wellPanel(
                                                              conditionalPanel(
                                                                  condition = "input.own == false",
-                                                                 selectInput("data", "Choose Dataset", c("MPG" = "mpg", "Air Quality" = "airquality", "Diamonds" = "diamonds"))
+                                                                 selectizeInput("data", label = "Choose Dataset", choices = c("MPG" = "mpg", "Air Quality" = "airquality", "Diamonds" = "diamonds"))
                                                              ),
                                                              conditionalPanel(
                                                                  condition = "input.own == true",

### 2014-11-10 modules/t_test/observe.R

@@ -1,4 +1,4 @@
     observe({
-        updateSelectInput(session, "group1", choices = intro.numericnames(), selected = ifelse(checkVariable(intro.data(), input$group1), input$group1, intro.numericnames()[1]))
-        updateSelectInput(session, "group2", choices = intro.numericnames(), selected = ifelse(checkVariable(intro.data(), input$group2), input$group2, intro.numericnames()[2]))
+        updateSelectizeInput(session, "group1", choices = intro.numericnames(), selected = ifelse(checkVariable(intro.data(), input$group1), input$group1, intro.numericnames()[1]))
+        updateSelectizeInput(session, "group2", choices = intro.numericnames(), selected = ifelse(checkVariable(intro.data(), input$group2), input$group2, intro.numericnames()[2]))
     })

### 2014-11-10 modules/t_test/ui.R

@@ -1,19 +1,19 @@
 t_test_ui <- tabPanel("T test",
                                                   column(4,
                                                          wellPanel(
-                                                             radioButtons("varts", "Type", choices=c("One Variable" = "onevart", "Two Variables" = "twovart")),
+                                                             selectizeInput("varts", label = "Type", choices=c("One Variable" = "onevart", "Two Variables" = "twovart")),
                                                              
                                                              hr(),
                                                              
-                                                             selectInput("group1", "Group 1 (x)", choices = numericNames(mpg)),
+                                                             selectizeInput("group1", label = "Group 1 (x)", choices = numericNames(mpg)),
                                                              conditionalPanel(
                                                                  condition = "input.varts == 'twovart'",
-                                                                 selectInput("group2", "Group 2 (y)", choices = numericNames(mpg))
+                                                                 selectizeInput("group2", "Group 2 (y)", choices = numericNames(mpg))
                                                              ),
                                                              
                                                              hr(),
                                                              
-                                                             selectInput("althyp", "Alternative Hypothesis", c("Two-Sided" = "two.sided", "Greater" = "greater", "Less" = "less")),
+                                                             selectizeInput("althyp", "Alternative Hypothesis", c("Two-Sided" = "two.sided", "Greater" = "greater", "Less" = "less")),
                                                              numericInput("hypval", "Hypothesized Value", value = 0),
                                                              sliderInput("conflevel", "Confidence Level", min=0.01, max=0.99, step=0.01, value=0.95)
                                                          )

### 2014-11-10 modules/transform/ui.R

@@ -1,8 +1,8 @@
 transform_ui <- tabPanel("Transform",
                                                   column(4,
                                                          wellPanel(                                          
-                                                             selectInput("trans", "Choose Transformation", choices = c("Power" = "power", "Categorical" = "categorical", "Numeric" = "numeric")),
-                                                             selectInput("var_trans", "Select Variable", choices = numericNames(mpg)),
+                                                             selectizeInput("trans", label = "Choose Transformation", choices = c("Power" = "power", "Categorical" = "categorical", "Numeric" = "numeric"), multiple = FALSE),
+                                                             selectizeInput("var_trans", label = "Select Variable", choices = numericNames(mpg), multiple = FALSE),
                                                              conditionalPanel(condition = "input.trans == 'power'",
                                                                               sliderInput("power", "Power", value = 1, min = -5, max = 5, step = 0.1)
                                                              ),

### 2014-11-10 global.R
Remove files from papers branch

@@ -1,38 +1,0 @@
-library(shiny)
-library(shinyAce)
-library(YaleToolkit)
-library(ggplot2)
-library(ggvis)
-library(dplyr)
-library(lubridate)
-library(gridExtra)
-library(R.utils)
-
-###
-### Global Helper Functions
-###
-numericNames <- function(data) {
-  vec <- as.character(subset(whatis(data), type == "numeric")$variable.name)
-  if (length(vec) == 0) vec <- ""
-  
-  return(vec)
-}
-
-categoricNames <- function(data) {
-  vec <- as.character(subset(whatis(data), type != "numeric")$variable.name)
-  if (length(vec) == 0) vec <- ""
-  
-  return(vec)
-}
-
-sourceDir <- function(path, type, local = FALSE, ...) { 
-  for (nm in list.files(path, pattern = paste0("\\.", type, "$"))) { 
-    source(file.path(path, nm), local=local) 
-  } 
-}
-
-checkVariable <- function(data, var) {
-    return(nchar(var) > 0 & var %in% names(data))
-}
-
-module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")

### 2014-11-10 modules/contingency/helper.R

@@ -1,24 +1,0 @@
-cont.table <- function(data, x, y, type, digits) {  
-    if (x %in% names(data) & y %in% names(data)) {
-        my.tbl <- table(data[,y], data[,x])
-        
-        totalsum <- sum(my.tbl)
-        
-        my.tbl <- rbind(my.tbl, Total = colSums(my.tbl, na.rm = TRUE))
-        my.tbl <- cbind(my.tbl, Total = c(rowSums(my.tbl[-(nrow(my.tbl)), ], na.rm = TRUE), totalsum))
-        
-        if (type == "totalpercs") my.tbl <- my.tbl / totalsum
-        if (type == "rowpercs") {
-            my.tbl <- t(apply(my.tbl[-nrow(my.tbl),], 1, function(row){row / row[ncol(my.tbl)]}))
-            #my.tbl <- rbind(my.tbl, Total = colSums(my.tbl, na.rm = TRUE))
-        }
-        if (type == "columnpercs") {
-            my.tbl <- apply(my.tbl[,-ncol(my.tbl)], 2, function(col){col / col[nrow(my.tbl)]})
-            #my.tbl <- cbind(my.tbl, Total = rowSums(my.tbl, na.rm = TRUE))
-        }
-        
-        return(format(my.tbl, digits = digits))
-    } else {
-        return(NULL)
-    }
-}

### 2014-11-10 modules/contingency/observe.R

@@ -1,4 +1,0 @@
-    observe({
-        updateSelectInput(session, "xcont", choices = intro.categoricnames(), selected = ifelse(checkVariable(intro.data(), input$xcont), input$xcont, intro.categoricnames()[1]))
-        updateSelectInput(session, "ycont", choices = intro.categoricnames(), selected = ifelse(checkVariable(intro.data(), input$ycont), input$ycont, intro.categoricnames()[2]))
-    })

### 2014-11-10 modules/contingency/output.R

@@ -1,3 +1,0 @@
-    output$conttable <- renderTable({
-        return(cont.table(intro.data(), input$xcont, input$ycont, input$conttype, my.digits()))
-    }, digits = my.digits())

### 2014-11-10 modules/contingency/reactive.R

@@ -1,4 +1,0 @@
-    my.digits <- reactive({
-        if (input$conttype %in% c("totalpercs", "rowpercs", "columnpercs")) return(3)
-        else return(0)
-    })

### 2014-11-10 modules/graphical/reactive.R

@@ -1,64 +1,0 @@
-x_choices <- reactive({
-    if (input$plottype %in% c("boxplot", "barchart", "paretochart", "mosaicplot")) intro.categoricnames()
-    else intro.numericnames()
-})
-    
-y_choices <- reactive({
-    if (input$plottype %in% c("mosaicplot")) intro.categoricnames()
-    else intro.numericnames()
-})
-
-x_selected <- reactive({
-    x_choices()[1]
-})
-
-y_selected <- reactive({
-    y_choices()[2]
-})
-
-intro.plot <- reactive({
-        if (is.null(intro.data())) return(NULL)
-        
-        mydat <- na.omit(intro.data())
-        
-        xvar <- if (checkVariable(intro.data(), input$x)) input$x else 1
-        yvar <- if (checkVariable(intro.data(), input$y)) input$y else 2
-
-        mydat$intro_x_cat <- factor(mydat[,xvar])
-        mydat$intro_x_num <- as.numeric(mydat[,xvar])
-        mydat$intro_y_cat <- factor(mydat[,yvar])
-        mydat$intro_y_num <- as.numeric(mydat[,yvar])
-        
-        if (input$plottype == "paretochart") mydat$intro_x_cat <- factor(mydat$intro_x_cat, levels = names(sort(table(mydat$intro_x_cat), decreasing = TRUE)))
-        
-        return(mydat)
-    })
-    
-    intro.quant <- reactive({
-        if (is.null(intro.plot())) return(NULL)
-        
-        mydat <- intro.plot()
-        
-        yy <- quantile(mydat$intro_x_num, na.rm = TRUE, c(0.25, 0.75))
-        xx <- qnorm(c(0.25, 0.75))
-        slope <- diff(yy) / diff(xx)
-        int <- yy[1] - slope * xx[1]
-        
-        mydat$intro_quant <- qnorm(seq(0, 1, by = (1/(length(mydat$intro_x_num) + 1)))[-c(1, (length(mydat$intro_x_num) + 2))])
-        mydat$intro_x_num <- sort(mydat$intro_x_num)
-        
-        return(mydat)
-    })
-    
-    input_binwidth <- reactive({
-        if (is.na(input$binwidth)) NULL
-        else input$binwidth
-    })
-    
-    input_xdomain <- reactive({
-        return(c(input$xmin, input$xmax))
-    })
-    
-    input_ydomain <- reactive({
-        return(c(input$ymin, input$ymax))
-    })

### 2014-11-10 modules/numerical/helper.R

@@ -1,2 +1,0 @@
-    q1 <- function(x) { return(quantile(x, .25)) }
-    q3 <- function(x) { return(quantile(x, .75)) }

### 2014-11-10 modules/numerical/observe.R

@@ -1,4 +1,0 @@
-    observe({
-        updateSelectInput(session, "grouping", choices = c("None" = "none", names(intro.data())))
-        updateCheckboxGroupInput(session, "tblvars", choices = names(intro.data()))
-    })

### 2014-11-10 modules/numerical/output.R

@@ -1,16 +1,0 @@
-    output$summary <- renderPrint({
-        if (is.null(input$tblvars)){
-            return(NULL)    
-        } else if (input$grouping == "none") {
-            return(summary(intro.data()[,input$tblvars]))
-        } else {
-            dat <- intro.data()[,input$tblvars]
-            dat$intro_grouping <- intro.data()[,input$grouping]
-            
-            dat.dplyr <- dat[,names(dat) %in% c("intro_grouping", numericNames(dat))] %>% group_by(intro_grouping) %>% summarise_each(funs(min, q1, median, q3, max, mean, sd))
-            dat.dplyr.df <- as.data.frame(dat.dplyr)
-            names(dat.dplyr.df)[1] <- input$grouping
-            
-            return(dat.dplyr.df)
-        }
-    })

### 2014-11-10 modules/numerical/ui.R

@@ -1,12 +1,0 @@
-numerical_ui <- tabPanel("Numerical",
-                                                  column(4,
-                                                         wellPanel(
-                                                             checkboxGroupInput("tblvars", "Select Variables", choices = names(mpg)),
-                                                             selectInput("grouping", "Select Grouping Variable", choices = names(mpg))
-                                                         )
-                                                  ),
-                                                  
-                                                  column(8,
-                                                         verbatimTextOutput("summary")
-                                                  )
-                                         )

### 2014-11-10 modules/regression/reactive.R

@@ -1,50 +1,0 @@
-intro.regression <- reactive({
-  if (is.null(intro.data())) return(NULL)
-  
-  lm.fit <- lm(intro.data()[,input$yreg] ~ intro.data()[,input$xreg], na.action = na.exclude)
-  
-  if (input$saveresid > oldsaveresid) {
-    curxreg <- input$xreg
-    curyreg <- input$yreg
-    
-    values$mydat <<- savefit(intro.data(), input$xreg, input$yreg, lm.fit)
-    
-    oldsaveresid <<- input$saveresid
-  }
-  
-  return(lm.fit)
-})
-
-reg.data <- reactive({
-  if (is.null(intro.regression())) return(NULL)
-  
-  reg.data <- intro.data()
-  reg.data$xreg <- reg.data[,input$xreg]
-  reg.data$yreg <- reg.data[,input$yreg]
-  
-  return(reg.data)
-})
-
-reg.resid1 <- reactive({
-  if (is.null(intro.regression())) return(NULL)
-  
-  mydat <- data.frame(residuals = resid(intro.regression()), x = intro.data()[as.numeric(names(resid(intro.regression()))),input$xreg])
-  
-  return(mydat)
-})
-
-reg.resid2 <- reactive({
-  if (is.null(intro.regression())) return(NULL)
-  
-  myresid <- resid(intro.regression())
-  
-  yy <- quantile(myresid, na.rm = TRUE, c(0.25, 0.75))
-  xx <- qnorm(c(0.25, 0.75))
-  slope <- diff(yy) / diff(xx)
-  int <- yy[1] - slope * xx[1]
-  
-  mydat <- data.frame(yy = qnorm(seq(0, 1, by = (1/(length(na.omit(myresid)) + 1)))[-c(1, (length(na.omit(myresid)) + 2))]),
-                      residuals = sort(myresid))
-  
-  return(mydat)
-})

### 2014-11-10 modules/sources/helper.R

@@ -1,49 +1,0 @@
-process_logical <- function(data, x) {
-    if (is.null(x)) {
-        return(data)
-    }
-    
-    relevant_cols <- names(data)[nchar(x) > 0]
-    if (length(relevant_cols) == 0) return(data)
-    
-    my_strs <- strsplit(gsub(",", " , ", x), split = ",")
-    new_strs <- my_strs[unlist(lapply(my_strs, length)) > 0]
-    
-    new_data <- data
-    if (length(new_strs) > 0) {
-        for (i in 1:length(new_strs)) {
-            test <- new_strs[[i]]
-            col <- relevant_cols[i]
-            
-            subset_str <- ""
-            if (length(test) == 1) {
-                if (is.na(as.numeric(test[1]))) {
-                    subset_str <- paste0("'", test[1], "' == ", col)
-                } else {
-                    subset_str <- paste(test[1], "==", col)
-                }
-            } else {
-                test[1] <- ifelse(test[1] == " ", -Inf, test[1])
-                test[2] <- ifelse(test[2] == " ", Inf, test[2])
-                if (is.na(test[2])) test[2] <- Inf
-                
-                subset_str <- paste(test[1], "<=", col, "&", col, "<=", test[2])
-            }          
-            
-            new_data <- eval(parse(text = paste0("subset(new_data, ", subset_str, ")")))
-        }
-    }
-    
-    return(new_data)
-}
-
-data.module <- function (inFile, dataset, own) {        
-    intro.data <- NULL
-    if (is.null(inFile) | !own) {
-        intro.data <- dataset
-    } else {
-        intro.data <- read.csv(inFile$datapath)
-    }
-    
-    intro.data
-}

### 2014-11-10 modules/sources/reactive.R

@@ -1,63 +1,0 @@
-intro.start <- reactive({
-    input$clearsubset
-
-    data.initial <- data.module(input$data_own, valid.datasets[[input$data]], input$own)
-    
-    if (values$firstrun) {
-        values$firstrun <- FALSE
-    }
-            
-    values$mydat <<- data.initial
-            
-    return(data.initial)
-})
-
-intro.data <- reactive({
-    if (is.null(intro.start())) return(NULL)
-    
-    data.subset <- process_logical(values$mydat, input$subs)
-    values$mydat <<- data.subset
-    values$mydat_rand <<- values$mydat
-    
-    if (input$randomsub & is.numeric(input$randomsubrows) & input$randomsubrows >= 1 & input$randomsubrows <= nrow(values$mydat)) {
-        values$mydat_rand <<- dplyr::sample_n(values$mydat, input$randomsubrows)
-    }
-    
-    if (input$savesubset > oldsavesub) {
-        values$mydat <<- values$mydat_rand
-        oldsavesub <<- input$savesubset
-    }
-    
-    return(values$mydat)
-})
-
-intro.numericnames <- reactive({
-    if (is.null(intro.data())) return(NULL)
-    
-    return(numericNames(intro.data()))
-})
-
-intro.categoricnames <- reactive({
-    if (is.null(intro.data())) return(NULL)
-    
-    return(categoricNames(intro.data()))
-}) 
-
-dt.options <- reactive({list(pageLength = 10,
-                             searching=0,
-                             destroy=1,
-                             headerCallback =  I(paste0("function(thead, data, start, end, display) {
-                                                        //color code the header items
-                                                        var col_types = [", 
-                                                        paste(paste0("'", ifelse(grepl("factor", whatis(intro.data())$type), "categorical", ifelse(grepl("character", whatis(intro.data())$type), "categorical", "numeric")),"'"),
-                                                              collapse=", "),
-                                                        "]
-                                                        var headers = $(thead).find('th');
-                                                        
-                                                        for(i = 0; i < col_types.length; i++) {
-                                                        if(col_types[i] == 'categorical') headers[i].style.color = '#95a5a6';
-                                                        else headers[i].style.color = '#3498db';
-                                                        }
-                                                        
[…]

### 2014-11-10 modules/sources/ui.R

@@ -1,35 +1,0 @@
-sources_ui <- tabPanel("Sources",
-                                                  column(4,
-                                                         wellPanel(
-                                                             conditionalPanel(
-                                                                 condition = "input.own == false",
-                                                                 selectInput("data", "Choose Dataset", c("MPG" = "mpg", "Air Quality" = "airquality", "Diamonds" = "diamonds"))
-                                                             ),
-                                                             conditionalPanel(
-                                                                 condition = "input.own == true",
-                                                                 fileInput('data_own', 'Choose CSV File',
-                                                                           accept=c('text/csv', 
-                                                                                    'text/comma-separated-values,text/plain', 
-                                                                                    '.csv'))
-                                                             ),
-                                                             checkboxInput("own", "Upload Dataset"),
-                                                             
-                                                             hr(),
-                                                             fluidRow(
-                                                                 column(6, checkboxInput("randomsub", "Random Subset")),
-                                                                 conditionalPanel(condition = "input.randomsub == true", 
-                                                                                  column(6, numericInput("randomsubrows", "Rows", value = 1, min = 1))
-                                                                 )
-                                                             ),
-                                                             tags$button("", id = "savesubset", type = "button", class = "btn action-button", onclick="var vals = []; var subsets = $('input[type = \"text\"][placeholder]'); for(i = 0; i < subsets.length; i++) {vals.push(subsets[i].value);}; Shiny.onInputChange(\"subs\", vals); $('#side-nav :contains(\"Sources\")').highlight();", list(icon("save"), "Save Subset")),
-                                                             br(), br(),
-                                                             tags$button("", id = "clearsubset", type = "button", class = "btn action-button", onclick="Shiny.onInputChange(\"subs\", null);", list(icon("eraser"), "Reset Data")),
-                                                             br(), br(),
-                                                             downloadButton("downloaddata", "Download Data")
-                                                         )
-                                                  ),
-                                                  
-                                                  column(8,
-                                                         dataTableOutput("data")
-                                                  )
-                                         )

### 2014-11-10 modules/t_test/observe.R

@@ -1,4 +1,0 @@
-    observe({
-        updateSelectInput(session, "group1", choices = intro.numericnames(), selected = ifelse(checkVariable(intro.data(), input$group1), input$group1, intro.numericnames()[1]))
-        updateSelectInput(session, "group2", choices = intro.numericnames(), selected = ifelse(checkVariable(intro.data(), input$group2), input$group2, intro.numericnames()[2]))
-    })

### 2014-11-10 modules/t_test/output.R

@@ -1,5 +1,0 @@
-    output$ttesttable <- renderText({
-        if (!(input$group1 %in% intro.numericnames())) return(NULL)
-        
-        return(ttesttable(intro.data(), input$group1, input$group2, input$varts == "twovart", input$conflevel, input$althyp, input$hypval))
-    })

### 2014-11-10 modules/transform/observe.R

@@ -1,4 +1,0 @@
-    ## Transform Observers
-    observe({
-        updateSelectInput(session, "var_trans", choices = (if (input$trans %in% c("power", "categorical")) intro.numericnames() else intro.categoricnames()))
-    })

### 2014-11-10 modules/transform/output.R

@@ -1,15 +1,0 @@
-    output$var_trans_text <- renderText({
-        if (is.null(intro.transform())) return(NULL)
-        
-        return(paste("Selected Variable:", input$var_trans))
-    })
-
-    intro.transform %>%
-        ggvis(x = ~var) %>%
-        layer_histograms() %>%
-        bind_shiny("var_plot")
-
-    intro.transform %>%
-        ggvis(x = ~trans_var) %>%
-        layer_histograms() %>%
-        bind_shiny("trans_plot")

### 2014-11-10 modules/transform/reactive.R

@@ -1,45 +1,0 @@
-    intro.transform.colname <- reactive({
-        if (is.null(intro.data())) return(NULL)
-        
-        curdata <- intro.data()
-        curtrans <- input$var_trans
-        colname <- ""
-        
-        if (input$trans == "power") {
-            colname <- paste(curtrans, sub("\\.", "", input$power), sep = "_")
-        } else if (input$trans %in% c("categorical", "numeric")) {
-            colname <- paste(curtrans, "trans", sep = "_")
-        }
-        
-        return(colname)
-    })
-
-    intro.transform <- reactive({
-        if (is.null(intro.data())) return(NULL)
-                        
-        curdata <- intro.data()
-        curtrans <- input$var_trans
-                
-        if (input$trans == "power" & curtrans %in% intro.numericnames()) {
-            trans_x <- if (input$power == 0) log(curdata[,curtrans]) else (curdata[,curtrans])^(input$power)
-            if (all(!is.infinite(trans_x))) {
-                curdata[, intro.transform.colname()] <- trans_x
-            }
-        } else if (input$trans %in% c("categorical", "numeric")) {
-            trans_x <- if (input$trans == "numeric") as.numeric(curdata[,curtrans]) else factor(curdata[,curtrans])
-            curdata[, intro.transform.colname()] <- trans_x
-        } else {
-            curdata[, intro.transform.colname()] <- curtrans
-        }
-                
-        if (input$savetrans > oldsavetrans) {
-            values$mydat <<- curdata
-            oldsavetrans <<- input$savetrans
-        }
-        
-        if (is.numeric(curdata[,curtrans])) curdata$var <- curdata[,curtrans] else curdata$var <- 0
-        if (is.numeric(curdata[,intro.transform.colname()])) curdata$trans_var <- curdata[,intro.transform.colname()] else curdata$trans_var <- 0
-        
-        return(curdata)
-    })
-

### 2014-11-10 modules/transform/static.R

@@ -1,1 +1,0 @@
-oldsavetrans <- 0

### 2014-11-10 scripts/other-helpers.js

@@ -1,18 +1,0 @@
-jQuery.fn.highlight = function() {
-  $(this).each(function() {
-        var el = $(this);
-        $("<div/>")
-          .width(el.outerWidth())
-          .height(el.outerHeight())
-          .css({
-              "position": "absolute",
-              "left": el.offset().left,
-              "top": el.offset().top,
-              "background-color": "#ffff99",
-              "opacity": ".7",
-              "z-index": "9999999"
-          }).appendTo('body').fadeOut(1000).queue(function () { $(this).remove(); });
-    });
-}
-
-//call with $('#side-nav :contains("Sources")').highlight()

### 2014-11-10 scripts/print.js

@@ -1,4 +1,0 @@
-function print_intRo() {
-  $('#codePrint').html(ace.edit('myEditor').getSession().getDocument().getAllLines().join('<br/>'))
-  window.print()
-}

### 2014-11-10 server.R

@@ -1,19 +1,0 @@
-###
-### Shiny Server definition
-###
-shinyServer(function(input, output, session) {
-    ## Maximum file upload size = 10MB
-    options(shiny.maxRequestSize=10*1024^2)
-    
-    ## Reactive values
-    values <- reactiveValues(firstrun = TRUE, mydat = NULL, mydat_rand = NULL)    
-    
-    ## Modules
-    types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
-    modules_tosource <- file.path("modules", apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
-
-    ## Source the modules
-    for (mod in modules_tosource) {
-        source(mod, local = TRUE)
-    }
-})

### 2014-11-10 ui.R

@@ -1,39 +1,0 @@
-addResourcePath(prefix="images", directoryPath="images/")
-
-## Source ui
-sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)
-
-## Generates the UI tabs
-mylist <- list()
-old_heading <- ""
-for (i in 1:nrow(module_info)) {
-    if (module_info[i,1] != old_heading) {
-        if (i != 1) mylist[[length(mylist) + 1]] <- "-----"
-        mylist[[length(mylist) + 1]] <- as.character(module_info[i,1])
-        old_heading <- module_info[i,1]
-    }
-    mylist[[length(mylist) + 1]] <- get(paste(module_info[i,2], "ui", sep = "_"))
-}
-
-shinyUI(
-    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
-               tabPanel(title="", icon=icon("home"),
-                        fluidRow(
-                            div(class='intRoPrint', h3('Results from intRo session:')),
-                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
-                        )     
-               ),
-               tabPanel(title="", value="http://gammarama.github.io/intRo", icon=icon('question-circle')),
-               tabPanel(title="", value="http://github.com/gammarama/intRo", icon=icon("github")),
-               navbarMenu("", icon=icon("envelope"),
-                          tabPanel("Eric Hare"),
-                          tabPanel("Andee Kaplan")),
-               #tabPanel(title="hide_me"),
-               #tabPanel(title="", icon=icon("print"), value = "javascript:print_intRo();"),
-               footer=tagList(includeScript("scripts/top-nav-links.js"),
-                              includeScript("scripts/print.js"),
-                              includeScript("http://code.jquery.com/color/jquery.color-2.1.2.min.js"),
-                              includeScript("scripts/other-helpers.js")
-               ),
-               tags$head(tags$link(rel="shortcut icon", href="images/icon.png"))
-    ))

### 2014-11-10 student-paper-competition-2014/paper.Rmd
started what is intro section

---
title: 'Introducing statistics with \texttt{intRo}'
author: "Andee Kaplan & Eric Hare"
date: "December 14, 2014"
output:
  pdf_document:
    number_sections: yes
    includes:
      in_header: preamble.tex
fontsize: 11pt
geometry: margin=1in
---

\begin{abstract}
Your abstract goes here...
...
\end{abstract}

\newpage

# Introduction

# What is \texttt{intRo}?
\texttt{intRo} is a web-based tool to accompany and introductory statistics class. It is meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class.

We have multiple philosophies that guided the creation of \texttt{intRo}, and this comes out in the design. Primarily, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics. Additionally, \texttt{intRo} is an \emph{extensible} tool, allowing for a user of \texttt{intRo} to tailor the tool for his or her own classroom needs.

## Easy
In order for an application to be useful to students in their first statistics course, it can't hog the spotlight. Statistics as a subject should shine through as a fun and interesting topic to study, rather than being bogged down by difficult to use software. With this in mind, in the development of \texttt{intRo} we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training. We used large, easy to click icons in the page header to help students find exactly what they need easily. With this easy to use framework, we also made the functionality available in \texttt{intRo} the minimal necessary for an introductory statistics course. Additional functionality is nice, but if it clutters the UI beyond the point of usability, it will frustrate an introductory student.

Beyong being simple, \texttt{intRo} is also consistent. The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules. Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code. In figure~\ref{fig:ui}, we highlight the five elements that comprise the intRo interface.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/ui_annotate.png}
\label{fig:ui}
\caption{The five elements that comprise the intRo application.}
\end{figure}

 - **Top Navigation** - The top navigation bar includes two sets of buttons. The left-aligned buttons are informational buttons. The first is a link to intRo. The second is a link to this documentation page. The third is a link to the GitHub repository where the code for \texttt{intRo} is housed. The final button is a link to our websites, which contain contact information if there are any questions or comments. The right-aligned buttons are \texttt{intRo} utilities. The first is a button to toggle the visibility of the code panel (5). The last is a button to print the active module results, and the associated code (if visible).
 - **Side Navigation** - The side navigation panel includes the modules which are detailed later in the tutorial.
 - **Options Panel** - The options panel includes module-specific options which the user can use to customize their results.
 - **Results Panel** - The results pane displays the result of the selected module and options.
 - **Code Panel** - The code panel displays the R code used to generate the results from the student's \texttt{intRo} session. The code panel is shown by default to facilitate an interest in coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.

There are multiple modules within \texttt{intRo}, which perform specific data analysis tasks that employ an easy to use point-and-click interface. Continue with the tutorial to read more about each one.

## Extensible

# In the classroom

# Reactive programming

# Conclusions and future work

# References

### 2014-11-10 student-paper-competition-2014/paper.Rmd
Merge in changes

@@
       in_header: preamble.tex
-fontsize: 11pt
-geometry: margin=1in
 ---
@@
 # In the classroom
+\texttt{intRo}'s ease of use lends itself to a natural presentation in an introductory statistics course.
+Statistics software typically used in introductory courses, such as JMP and Excel, have restrictive software licensing and platform requirements that can limit the software's accessibility to students.
+Even if these limitations can be overcome, these software programs have non-ideal interfaces for an exploratory data analysis.
+JMP, for instance, hides several basic regression functions such as a line of best fit behind small red arrows.
+Using R itself in the classroom is an appealing alternative, but suffers from some other problems.
+Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis.
+When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (This experience is made much easier with new tools like rmarkdown, but this again is a bit of a learning curve for the new statistics student.)
+
+We believe that \texttt{intRo} targets these key limitations.
+The interface is designed to be quick, seamless, and easy to understand.
+Because \texttt{intRo} is web-based, the statistical analysis can be performed from nearly any modern device.
+R underlies the analysis, but users need not be familiar with R or programming to perform it.
+The focal point of the interface is the output panel, allowing the student to remain focused on the results and limiting the need to fight with the interface itself to get work done.
+For those students who may have a natural inclination for programming, executable code is generated to allow a local reproduction of the analysis.
+
+The features of \texttt{intRo} were selected in order to apply to the most commonly used lesson plans in early statistics courses.
+The textbook "The Art and Science of Learning from Data" by Agresti and Franklin references many of the focal points of introductory statistics, and \texttt{intRo} supports performing a majority of them.
+Part one, titled "Gathering and Exploring Data", covers graphical summaries, measures of spread and center, and avoiding the misuse of graphics in situations they are not appropriate.
+\texttt{intRo} includes modules for both graphical and numerical summaries covering these topics.
+Furthermore, \texttt{intRo} limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{images/appropriate_graphics.png}
+\label{fig:ag}
+\caption{The plot and variable selection boxes in \texttt{intRo}'s graphical module.
+Because the selected plottype is histogram, the interface displays only a single variable selection box (x), and the variables selectable in this box are only numeric.}
+\end{figure}
 

### 2014-11-10 student-paper-competition-2014/paper.Rmd
Some more text for the education section

@@
 \texttt{intRo} includes modules for both graphical and numerical summaries covering these topics.
-Furthermore, \texttt{intRo} limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics.
+Furthermore, \texttt{intRo} limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics (Figure \ref{fig:ag}).
 
@@
 
+Part two of this text discusses probability and sampling distributions, which is outside the current realm of \texttt{intRo}, but part three focuses on inferential statistics.
+The inferential statistics include confidence intervals, hypothesis testing, and comparing two variables, all supported natively in \texttt{intRo}.
+Finally, part four of the text discusses the association between variables, including regression, functions that intRo performs with ease.
+Some of the more advanced topics of this section are currently unsupported by \texttt{intRo}, including multiple regression and nonparametric statistics.
+
+Statistics courses with a lab component can most benefit from the use of \texttt{intRo}.
+As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries.
+They simply upload a dataset and click the module appropriate to the section of the class.
+Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in.
+Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the \texttt{intRo} interface to obtain the results.
+For topics currently unsupported by \texttt{intRo}, the instructor can supplement the its use with other web-based modules that are publicly available.
+
 # Reactive programming

### 2014-11-10 student-paper-competition-2014/paper.Rmd
exciting section started

--- 2014-11-10 student-paper-competition-2014/paper.Rmd
@@
 Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code.
-In figure~\ref{fig:ui}, we highlight the five elements that comprise the intRo interface.
+In figure \ref{fig:ui}, we highlight the five elements that comprise the intRo interface.
 
-\begin{figure}
+\begin{figure}[ht!]
 \centering
-\includegraphics[width=\textwidth]{images/ui_annotate.png}
+\includegraphics[width=0.8\textwidth]{images/ui_annotate.png}
 \label{fig:ui}
@@
 The last is a button to print the active module results, and the associated code (if visible).
- - **Side Navigation** - The side navigation panel includes the modules which are detailed later in the tutorial.
+ - **Side Navigation** - The side navigation panel includes the modules that perform a specific data analysis task.
  - **Options Panel** - The options panel includes module-specific options which the user can use to customize their results.
  - **Results Panel** - The results pane displays the result of the selected module and options.
- - **Code Panel** - The code panel displays the R code used to generate the results from the student's \texttt{intRo} session.
+ - **Code Panel** - The code panel displays the \texttt{R} code used to generate the results from the student's \texttt{intRo} session.
 The code panel is shown by default to facilitate an interest in coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.
@@
 There are multiple modules within \texttt{intRo}, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
-Continue with the tutorial to read more about each one.
+Each module is placed within a higher level category to further help students navigate the tool.
+The default modules included in \texttt{intRo} are found in table \ref{tab:modules}, and more modules can be easy added by an instructor, as detailed in section \ref{sec:extensible}.
+
+\begin{table}
+\centering
+\begin{tabular}[ht!]{c|c|p{.6\textwidth}}
+Category & Module & Description \\
+\hline
+\multirow{2}{*}{Data} & Sources & \\
+& Transformation & \\
+\hline
+\multirow{2}{*}{Summaries} & Graphical & \\
+& Numerical & \\
+\hline
+\multirow{3}{*}{Statistics} & Contingency & \\
+& Regression & \\
+& T test & 
+\end{tabular}
+\label{tab:modules}
+\caption{A list of default modules that ship with \texttt{intRo} as well as brief descriptions of each.}
+\end{table}
+
+## Exciting
+In general, introduction to statistics is a class undergraduates dread.
+It's hard, the language is weird, and the software is hard to use.
+With \texttt{intRo}, we hope to flip this paradigm.
+\texttt{intRo} is a fun, easy to use tool that's available on the web, and is even usable on mobile devices.
+As a web-based application, this tool is immediately more familiar to students than a desktop application.
+With interactive plots from \texttt{ggvis}, \texttt{intRo} brings a `wow'-factor to learning introductory statistics.
+
+Once students get pulled in by an easy to use interface and fun interaction with their data, \texttt{intRo} has an ulterior motive: to get students excited about programming.
+By navigating about the user interface of \texttt{intRo}, students are actually creating a fully-executable \texttt{R} script that they can download and run locally.
+This code creation element of \texttt{intRo} is meant to generate excitement about programming in \texttt{R} and empower students to feel that they can generate code as well.
 
 ## Extensible
+\label{sec:extensible}
 
@@
 JMP, for instance, hides several basic regression functions such as a line of best fit behind small red arrows.
-Using R itself in the classroom is an appealing alternative, but suffers from some other problems.
+Using \texttt{R} itself in the classroom is an appealing alternative, but suffers from some other problems.
 Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis.
@@
 Because \texttt{intRo} is web-based, the statistical analysis can be performed from nearly any modern device.
-R underlies the analysis, but users need not be familiar with R or programming to perform it.
+\texttt{R} underlies the analysis, but users need not be familiar with \texttt{R} or programming to perform it.
 The focal point of the interface is the output panel, allowing the student to remain focused on the results and limiting the need to fight with the interface itself to get work done.

### 2014-11-17 modules/transform/reactive.R
Change power to numeric input with fine grained steps

@@ -21,7 +21,7 @@
         curtrans <- input$var_trans
                 
         if (input$trans == "power" & curtrans %in% intro.numericnames()) {
-            trans_x <- if (input$power == 0) log(curdata[,curtrans]) else (curdata[,curtrans])^(input$power)
+            trans_x <- if (!is.numeric(input$power) | is.null(input$power)) curdata[,curtrans] else if (input$power == 0) log(curdata[,curtrans]) else (curdata[,curtrans])^(input$power)
             if (all(!is.infinite(trans_x))) {
                 curdata[, intro.transform.colname()] <- trans_x
             }

### 2014-11-17 modules/transform/ui.R

@@ -4,7 +4,7 @@ transform_ui <- tabPanel("Transform",
                                                              selectizeInput("trans", label = "Choose Transformation", choices = c("Power" = "power", "Categorical" = "categorical", "Numeric" = "numeric"), multiple = FALSE),
                                                              selectizeInput("var_trans", label = "Select Variable", choices = numericNames(mpg), multiple = FALSE),
                                                              conditionalPanel(condition = "input.trans == 'power'",
-                                                                              sliderInput("power", "Power", value = 1, min = -5, max = 5, step = 0.1)
+                                                                              numericInput("power", "Power", value = 1, min = -100, max = 100, step = 0.01)
                                                              ),
                                                              
                                                              hr(),

### 2014-11-17 student-paper-competition-2014/paper.Rmd
filled out table, added refs, extensible

@@
 ---
-title: 'Introducing statistics with \texttt{intRo}'
+title: "Introducing statistics with \\texttt{intRo}"
 author: "Andee Kaplan & Eric Hare"
@@
   pdf_document:
-    number_sections: yes
     includes:
       in_header: preamble.tex
+    number_sections: yes
+bibliography: refs.bib
 ---
@@
 In order for an application to be useful to students in their first statistics course, it can't hog the spotlight.
-Statistics as a subject should shine through as a fun and interesting topic to study, rather than being bogged down by difficult to use software.
+Statistics as a subject should shine through as a fun and interesting topic to study, rather than being bogged down by difficult to learn software.
 With this in mind, in the development of \texttt{intRo} we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training.
@@
 Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code.
-In figure \ref{fig:ui}, we highlight the five elements that comprise the intRo interface.
+In figure \ref{fig:ui}, we highlight the five elements that comprise the \texttt{intRo} interface.
 
@@
 \includegraphics[width=0.8\textwidth]{images/ui_annotate.png}
+\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
 \label{fig:ui}
-\caption{The five elements that comprise the intRo application.}
 \end{figure}
 
- - **Top Navigation** - The top navigation bar includes two sets of buttons.
+ 1.
+**Top Navigation** - The top navigation bar includes two sets of buttons.
 The left-aligned buttons are informational buttons.
-The first is a link to intRo.
+The first is a link to \texttt{intRo}.
 The second is a link to this documentation page.
@@
 The last is a button to print the active module results, and the associated code (if visible).
- - **Side Navigation** - The side navigation panel includes the modules that perform a specific data analysis task.
- - **Options Panel** - The options panel includes module-specific options which the user can use to customize their results.
- - **Results Panel** - The results pane displays the result of the selected module and options.
- - **Code Panel** - The code panel displays the \texttt{R} code used to generate the results from the student's \texttt{intRo} session.
+ 2. **Side Navigation** - The side navigation panel includes the modules that perform a specific data analysis task.
+ 3. **Options Panel** - The options panel includes module-specific options which the user can use to customize their results.
+ 4. **Results Panel** - The results pane displays the result of the selected module and options.
+ 5.
+**Code Panel** - The code panel displays the \texttt{R} code used to generate the results from the student's \texttt{intRo} session.
 The code panel is shown by default to facilitate an interest in coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.
@@
 \centering
-\begin{tabular}[ht!]{c|c|p{.6\textwidth}}
+\begin{tabular}[ht!]{l|l|p{.65\textwidth}}
 Category & Module & Description \\
 \hline
-\multirow{2}{*}{Data} & Sources & \\
-& Transformation & \\
+\multirow{2}{*}{Data} & Sources & The data sources module allows the user to select or upload a dataset.
+Several datasets are pre-included, which can be selected via a drop-down menu.
+The user can check a box to upload a dataset, and then select the CSV file from his computer.
+Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
+From this module the user can select a subset of the data either by value or random selection.
+The dataset selected in the data module will be made available to all other modules of intRo.\\
+
+& Transformation & The data transformation modules allows transformation of numeric variables to categorical and categorical variable to numeric.
+Additionally, numeric variable can be transformed using a power transformation.
+The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.\\
 \hline
-\multirow{2}{*}{Summaries} & Graphical & \\
-& Numerical & \\
+\multirow{2}{*}{Summaries} & Graphical & The graphical summaries module provides several common graphical displays of variables in the dataset.
+First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
+The variable available to choose will only be those compatible with plot type selected.
+The plot visible on the right will automatically update when these values are selected.\\
+
+& Numerical & The numeric summaries module is another method of computing summaries of the data.
+The user is presented with a list of each variable from their dataset.
+The user can select one or more different variables, and will be presented with a table of some summary statistics.
+The summary statistics displayed vary depending on the type of the variables.
+For instance, for a numeric variable, a five number summary will be displayed.
+Character or factor variables will display counts of each unique value of that particular variable.
+Additionally, the user has the ability to display summaries by a grouping variable.\\
 \hline
-\multirow{3}{*}{Statistics} & Contingency & \\
-& Regression & \\
-& T test & 
+\multirow{3}{*}{Statistics} & Contingency & The contingency table module allows the user to compute a contingency table for two categorical variables.
+The results can either be displayed as counts, row percentages, column percentages, or total percentages.\\
+
+& Regression & The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
+The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
+Results of the regression will be displayed on the right.
+A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
+A user also has the ability to save residuals and predicted values to their dataset for further analysis.\\
+
+& T test & The T test module allows the user to perform either a one variable or two variable Student's T test.
+The user can select configuration options, including the direction of the test and the confidence level of the test.
+The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
 \end{tabular}
+\caption{A list of default modules that ship with \texttt{intRo} as well as brief descriptions of each.}
 \label{tab:modules}
-\caption{A list of default modules that ship with \texttt{intRo} as well as brief descriptions of each.}
 \end{table}
@@
 As a web-based application, this tool is immediately more familiar to students than a desktop application.
-With interactive plots from \texttt{ggvis}, \texttt{intRo} brings a `wow'-factor to learning introductory statistics.
+With interactive plots from \texttt{ggvis} [@ggvis], \texttt{intRo} brings a `wow'-factor to learning introductory statistics.
 
@@
 \label{sec:extensible}
+On the front end, user interaction with \texttt{intRo} is split into bitesize chunks that we call modules.
+This is true of the back end as well.
+Each module is a self contained set of \texttt{R} code that is dynamically added to the application at run time.
+What this means is that \texttt{intRo} is easily extended by the addition of modules within the framework underlying the application.
 
@@
 Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis.
-When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (This experience is made much easier with new tools like rmarkdown, but this again is a bit of a learning curve for the new statistics student.)
+When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (this experience is made much easier with new tools like \texttt{rmarkdown} [@rmarkdown], but this again is a bit of a learning curve for the new statistics student).
 
@@
 The features of \texttt{intRo} were selected in order to apply to the most commonly used lesson plans in early statistics courses.
-The textbook "The Art and Science of Learning from Data" by Agresti and Franklin references many of the focal points of introductory statistics, and \texttt{intRo} supports performing a majority of them.
+The textbook "The Art and Science of Learning from Data" by Agresti and Franklin [-@agresti2012statistics] references many of the focal points of introductory statistics, and \texttt{intRo} supports performing a majority of them.
 Part one, titled "Gathering and Exploring Data", covers graphical summaries, measures of spread and center, and avoiding the misuse of graphics in situations they are not appropriate.
@@
 \includegraphics[width=0.5\textwidth]{images/appropriate_graphics.png}
-\label{fig:ag}
 \caption{The plot and variable selection boxes in \texttt{intRo}'s graphical module.
 Because the selected plottype is histogram, the interface displays only a single variable selection box (x), and the variables selectable in this box are only numeric.}
+\label{fig:ag}
 \end{figure}
@@
 The inferential statistics include confidence intervals, hypothesis testing, and comparing two variables, all supported natively in \texttt{intRo}.
-Finally, part four of the text discusses the association between variables, including regression, functions that intRo performs with ease.
+Finally, part four of the text discusses the association between variables, including regression, functions that \texttt{intRo} performs with ease.
 Some of the more advanced topics of this section are currently unsupported by \texttt{intRo}, including multiple regression and nonparametric statistics.
@@
 Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the \texttt{intRo} interface to obtain the results.
-For topics currently unsupported by \texttt{intRo}, the instructor can supplement the its use with other web-based modules that are publicly available.
+For topics currently unsupported by \texttt{intRo}, the instructor can supplement its use with other web-based modules that are publicly available.
 

### 2014-11-17 modules/contingency/helper.R
Restructure for cat_and_eval

@@ -1,26 +1,28 @@
-cont.table <- function(data, x, y, type, digits) {  
-    if (x %in% names(data) & y %in% names(data)) {
-        my.tbl <- table(data[,y], data[,x])
+cont.table <- function(intro.data, x, y, type, digits) {  
+    if (x %in% names(intro.data) & y %in% names(intro.data)) {
+        cat_and_eval(paste0("
+            my.tbl <- table(intro.data[,'", y, "'], intro.data[,'", x, "'])
         
-        totalsum <- sum(my.tbl)
-        
-        my.tbl <- rbind(my.tbl, Total = colSums(my.tbl, na.rm = TRUE))
-        my.tbl <- cbind(my.tbl, Total = c(rowSums(my.tbl[-(nrow(my.tbl)), ], na.rm = TRUE), totalsum))
-        
-        if (type == "totalpercs") my.tbl <- my.tbl / totalsum
+            totalsum <- sum(my.tbl)
+            
+            my.tbl <- rbind(my.tbl, Total = colSums(my.tbl, na.rm = TRUE))
+            my.tbl <- cbind(my.tbl, Total = c(rowSums(my.tbl[-(nrow(my.tbl)), ], na.rm = TRUE), totalsum))
+        "), env = environment(), file = "code_Contingency.R")
+
+        if (type == "totalpercs") cat_and_eval("my.tbl <- my.tbl / totalsum", env = environment(), file = "code_Contingency.R", append = TRUE)
         if (type == "rowpercs") {
-            my.tbl <- t(apply(my.tbl[-nrow(my.tbl),], 1, function(row){row / row[ncol(my.tbl)]}))
-            #my.tbl <- rbind(my.tbl, Total = colSums(my.tbl, na.rm = TRUE))
+            cat_and_eval("my.tbl <- t(apply(my.tbl[-nrow(my.tbl),], 1, function(row){row / row[ncol(my.tbl)]}))", env = environment(), file = "code_Contingency.R", append = TRUE)
         }
         if (type == "columnpercs") {
-            my.tbl <- apply(my.tbl[,-ncol(my.tbl)], 2, function(col){col / col[nrow(my.tbl)]})
-            #my.tbl <- cbind(my.tbl, Total = rowSums(my.tbl, na.rm = TRUE))
+            cat_and_eval("my.tbl <- apply(my.tbl[,-ncol(my.tbl)], 2, function(col){col / col[nrow(my.tbl)]})", env = environment(), file = "code_Contingency.R", append = TRUE)
         }
         
[…]

### 2014-11-17 modules/graphical/output.R

@@ -1,48 +1,55 @@
-intro.plot %>%
+    cat_and_eval("p.histogram <- intro.plot %>%
         ggvis(x = ~intro_x_num, y = ~intro_y_num) %>%
         layer_histograms(width = input_binwidth) %>% 
-        scale_numeric("x", domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
-        scale_numeric("y", domain = input_ydomain, clamp = TRUE, nice = TRUE) %>%
-        bind_shiny("histogram")
+        scale_numeric('x', domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
+        scale_numeric('y', domain = input_ydomain, clamp = TRUE, nice = TRUE)", env = environment(), file = "code_histogram.R")
+    
+    p.histogram %>% bind_shiny("histogram")
 
-    intro.quant %>%
+    cat_and_eval("p.quantileplot <- intro.quant %>%
         ggvis(x = ~intro_quant, y = ~intro_x_num) %>%
         layer_points() %>% 
-        scale_numeric("x", domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
-        scale_numeric("y", domain = input_ydomain, clamp = TRUE, nice = TRUE) %>%
-        bind_shiny("quantileplot")
+        scale_numeric('x', domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
+        scale_numeric('y', domain = input_ydomain, clamp = TRUE, nice = TRUE)", env = environment(), file = "code_quantileplot.R")
     
-    intro.plot %>%
+    p.quantileplot %>% bind_shiny("quantileplot")
+    
+    cat_and_eval("p.scatterplot <- intro.plot %>%
         ggvis(x = ~intro_x_num, y = ~intro_y_num) %>%
         layer_points() %>% 
-        scale_numeric("x", domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
-        scale_numeric("y", domain = input_ydomain, clamp = TRUE, nice = TRUE) %>%
-        bind_shiny("scatterplot")
+        scale_numeric('x', domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
+        scale_numeric('y', domain = input_ydomain, clamp = TRUE, nice = TRUE)", env = environment(), file = "code_scatterplot.R")
     
-    intro.plot %>%
+    p.scatterplot %>% bind_shiny("scatterplot")
+    
[…]

### 2014-11-17 modules/graphical/reactive.R

@@ -16,49 +16,26 @@ y_selected <- reactive({
     y_choices()[2]
 })
 
-intro.plot <- reactive({
-        if (is.null(intro.data())) return(NULL)
-        
-        mydat <- na.omit(intro.data())
-        
-        xvar <- if (checkVariable(intro.data(), input$x)) input$x else 1
-        yvar <- if (checkVariable(intro.data(), input$y)) input$y else 2
-
-        mydat$intro_x_cat <- factor(mydat[,xvar])
-        mydat$intro_x_num <- as.numeric(mydat[,xvar])
-        mydat$intro_y_cat <- factor(mydat[,yvar])
-        mydat$intro_y_num <- as.numeric(mydat[,yvar])
-        
-        if (input$plottype == "paretochart") mydat$intro_x_cat <- factor(mydat$intro_x_cat, levels = names(sort(table(mydat$intro_x_cat), decreasing = TRUE)))
-        
-        return(mydat)
+    intro.plot <- reactive({
+            if (is.null(intro.data())) return(NULL)
+    
+            return(intro.plotdata(intro.data(), input$x, input$y, input$plottype))
     })
     
     intro.quant <- reactive({
         if (is.null(intro.plot())) return(NULL)
         
-        mydat <- intro.plot()
-        
-        yy <- quantile(mydat$intro_x_num, na.rm = TRUE, c(0.25, 0.75))
-        xx <- qnorm(c(0.25, 0.75))
-        slope <- diff(yy) / diff(xx)
-        int <- yy[1] - slope * xx[1]
-        
-        mydat$intro_quant <- qnorm(seq(0, 1, by = (1/(length(mydat$intro_x_num) + 1)))[-c(1, (length(mydat$intro_x_num) + 2))])
-        mydat$intro_x_num <- sort(mydat$intro_x_num)
-        
-        return(mydat)
+        return(intro.quantdata(intro.plot()))
     })
     
     input_binwidth <- reactive({
-        if (is.na(input$binwidth)) NULL
-        else input$binwidth
+        return(input_binwidthdata(input$binwidth))
     })
     
     input_xdomain <- reactive({
-        return(c(input$xmin, input$xmax))
+        return(intro.xdomaindata(input$xmin, input$xmax))
     })
     
     input_ydomain <- reactive({
-        return(c(input$ymin, input$ymax))
+        return(intro.ydomaindata(input$ymin, input$ymax))
     })

### 2014-11-17 modules/numerical/helper.R

@@ -1,9 +1,9 @@
-    generateSummary <- function(intro.data, tblvars, grouping) {
+    generateSummary <- function(intro.data, tblvars, grouping, save_result) {
         if (!is.null(tblvars)) {
             if (grouping == "none") {
-                cat_and_eval(paste0("summary(intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")])"))
+                cat_and_eval(paste0("summary(intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")])"), env = environment(), file = "code_Numerical.R", append = FALSE)
             } else {
-                cat_and_eval(paste0("dat <- intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")]; dat$intro_grouping <- intro.data[,'", grouping, "']; dat.dplyr <- dat[,names(dat) %in% c(\"intro_grouping\", numericNames(dat))] %>% group_by(intro_grouping) %>% summarise_each(funs(min, q1, median, q3, max, mean, sd)); dat.dplyr.df <- as.data.frame(dat.dplyr); names(dat.dplyr.df)[1] <- '", grouping, "'; dat.dplyr.df"))
+                cat_and_eval(paste0("dat <- intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")]; dat$intro_grouping <- intro.data[,'", grouping, "']; dat.dplyr <- dat[,names(dat) %in% c(\"intro_grouping\", numericNames(dat))] %>% group_by(intro_grouping) %>% summarise_each(funs(min, q1, median, q3, max, mean, sd)); dat.dplyr.df <- as.data.frame(dat.dplyr); names(dat.dplyr.df)[1] <- '", grouping, "'; dat.dplyr.df"), env = environment(), file = "code_Numerical.R")
             }
         }
     }

### 2014-11-17 modules/numerical/output.R

@@ -1,9 +1,3 @@
     output$summary <- renderPrint({
-        intro.inputs <- list("intro.data", input$tblvars, input$grouping)
-        get_code(generateSummary, intro.inputs)
-        
-        intro.inputs[[grep("intro.", intro.inputs)]] <- get(intro.inputs[[grep("intro.", intro.inputs)]])()
-        result <- do.call(generateSummary, intro.inputs)
-        
-        return(result)
+        return(generateSummary(intro.data(), input$tblvars, input$grouping, input$store_Numerical > values$store_value))
     })

### 2014-11-17 modules/t_test/helper.R

@@ -1,7 +1,7 @@
 ttesttable <- function (intro.data, x, y = NULL, twovar = FALSE, conflevel, althyp, hypval) {
     if (is.null(y) | y == "" | !twovar) {
-        cat_and_eval(paste0("t.test(x=intro.data[,'", x, "'], conf.level=", conflevel, ", alternative='", althyp, "', mu=", hypval, ")"), environment())
+        cat_and_eval(paste0("t.test(x=intro.data[,'", x, "'], conf.level=", conflevel, ", alternative='", althyp, "', mu=", hypval, ")"), env = environment(), file = "code_t_test.R")
     } else {
-        cat_and_eval(paste0("t.test(x=intro.data[,'", x, "'], y=intro.data[,'", y, "'], conf.level=", conflevel, ", alternative='", althyp, "', mu=", hypval, ")"), environment())
+        cat_and_eval(paste0("t.test(x=intro.data[,'", x, "'], y=intro.data[,'", y, "'], conf.level=", conflevel, ", alternative='", althyp, "', mu=", hypval, ")"), env = environment(), file = "code_t_test.R")
     }
 }

### 2014-11-17 server.R

@@ -2,14 +2,19 @@
 ### Shiny Server definition
 ###
 shinyServer(function(input, output, session) {
+    ## Remove last session's code
+    sapply(dir()[grep("code_", dir())], file.remove)
+    
     ## Maximum file upload size = 10MB
     options(shiny.maxRequestSize=10*1024^2)
     
     ## Reactive values
     values <- reactiveValues(mydat = NULL, mydat_rand = NULL)    
     
-    cat(paste(readLines("global.R"), collapse = "\n"), file = "test.R", append = TRUE)
-    cat("\n\n", file = "test.R", append = TRUE)
+    #cat(paste(readLines("global.R"), collapse = "\n"), file = "code_global.R")
+    cat("library(RCurl)\n\n", file = "code_All.R", append = TRUE)
+    cat("eval(parse(text = getURL('https://raw.githubusercontent.com/gammarama/intRo/master/global.R')))", file = "code_All.R", append = TRUE)
+    cat("\n\n", file = "code_All.R", append = TRUE)
     
     ## Modules
     types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")

### 2014-11-17 modules/graphical/helper.R
Fix xdomain and ydomain

@@ -30,16 +30,16 @@ intro.quantdata <- function(intro.plot) {
     return(intro.quant)
 }
 
-intro.xdomaindata <- function(xmin, xmax) {
+input_xdomaindata <- function(xmin, xmax) {
     cat_and_eval(paste0("input_xdomain <- c(", xmin, ", ", xmax, ")"), env = environment(), file = "code_Graphical.R", save_result = TRUE)
     
-    return(intro.xdomain)
+    return(input_xdomain)
 }
 
-intro.ydomaindata <- function(ymin, ymax) {
+input_ydomaindata <- function(ymin, ymax) {
     cat_and_eval(paste0("input_ydomain <- c(", ymin, ", ", ymax, ")"), env = environment(), file = "code_Graphical.R", save_result = TRUE)
     
-    return(intro.ydomain)
+    return(input_ydomain)
 }
 
 input_binwidthdata <- function(binwidth) {

### 2014-11-17 modules/graphical/reactive.R

@@ -33,9 +33,9 @@ y_selected <- reactive({
     })
     
     input_xdomain <- reactive({
-        return(intro.xdomaindata(input$xmin, input$xmax))
+        return(input_xdomaindata(input$xmin, input$xmax))
     })
     
     input_ydomain <- reactive({
-        return(intro.ydomaindata(input$ymin, input$ymax))
+        return(input_ydomaindata(input$ymin, input$ymax))
     })

### 2014-11-17 scripts/print.js
trying to print

@@ -1,4 +1,41 @@
-function print_intRo() {
-  $('#codePrint').html(ace.edit('myEditor').getSession().getDocument().getAllLines().join('<br/>'))
-  window.print()
+var outputBinding = new Shiny.OutputBinding();
+$.extend(outputBinding, {
+  find: function(scope) {
+    return $(scope).find('.print-results');
+  },
+  renderValue: function(el, data) {  
+    render_print(el, data);
+  }});
+Shiny.outputBindings.register(outputBinding);
+
+var inputBinding = new Shiny.InputBinding();
+$.extend(inputBinding, {
+  find: function(scope) {
+    return $(scope).find('.print-button');
+  },
+  getValue: function(el) {
+    return null;
+  },
+  subscribe: function(el, callback) {
+    $(el).on("click", function(e) {
+      callback();
+    });
+  },
+});
+Shiny.inputBindings.register(inputBinding);
+
+render_print = function(el, data) {
+  $(el).text(data);
+}
+
+print_intRo = function() {
+  $( ".print-results" ).dialog({
+    modal: true,
+    buttons: {
+      Print: function() {
+        //print stuff
+        $( this ).dialog( "close" );
+      }
+    }
+  });
 }

### 2014-11-17 server.R

@@ -24,4 +24,11 @@ shinyServer(function(input, output, session) {
     for (mod in modules_tosource) {
         source(mod, local = TRUE)
     }
+    
+    ## Printing
+    output$print_output <- reactive({ 
+      input[[paste0("store_", module_info$module)]]
+      loc <- render("code_All.R", output_dir = "www")
+      return(loc)      
+    })
 })

### 2014-11-17 ui.R

@@ -1,5 +1,11 @@
 addResourcePath(prefix="images", directoryPath="images/")
 
+## For printing
+dynDiv <- function(outputId, title) 
+{
+  div(id = outputId, class = "print-results", title = title)
+}
+
 ## Source ui
 sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)
 
@@ -19,6 +25,7 @@ shinyUI(
     navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
                tabPanel(title="", icon=icon("home"),
                         fluidRow(
+                            dynDiv(outputId = 'print_output', title="Print Results"),
                             div(class='intRoPrint', h3('Results from intRo session:')),
                             do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
                         )     
@@ -28,11 +35,14 @@ shinyUI(
                navbarMenu("", icon=icon("envelope"),
                           tabPanel("Eric Hare"),
                           tabPanel("Andee Kaplan")),
-               #tabPanel(title="hide_me"),
-               #tabPanel(title="", icon=icon("print"), value = "javascript:print_intRo();"),
+               tabPanel(title="hide_me"),
+               tabPanel(title="", icon=icon("print"), class="print-button", value = "javascript:print_intRo();"),
                footer=tagList(includeScript("scripts/top-nav-links.js"),
                               includeScript("scripts/print.js"),
-                              includeScript("scripts/other-helpers.js")
+                              includeScript("scripts/other-helpers.js"),
+                              includeCSS("http://code.jquery.com/ui/1.11.2/themes/smoothness/jquery-ui.css"),
+                              includeScript("http://code.jquery.com/jquery-1.10.2.js"),
+                              includeScript("http://code.jquery.com/ui/1.11.2/jquery-ui.js")
                ),
                tags$head(tags$link(rel="shortcut icon", href="images/icon.png"))
     ))

### 2014-11-17 global.R
Maybe less broken?

@@ -8,6 +8,7 @@ library(lubridate)
 library(gridExtra)
 library(R.utils)
 library(RCurl)
+library(rmarkdown)
 
 #cat("```{r, echo=FALSE}\nopts_chunk$set(echo=FALSE)\n```\n\n```{r, child="test.Rmd", message=FALSE, warning=FALSE}\n```", file = "outfile.Rmd")
 

### 2014-11-17 scripts/print.js

@@ -20,7 +20,7 @@ $.extend(inputBinding, {
     $(el).on("click", function(e) {
       callback();
     });
-  },
+  }
 });
 Shiny.inputBindings.register(inputBinding);
 
@@ -28,14 +28,3 @@ render_print = function(el, data) {
   $(el).text(data);
 }
 
-print_intRo = function() {
-  $( ".print-results" ).dialog({
-    modal: true,
-    buttons: {
-      Print: function() {
-        //print stuff
-        $( this ).dialog( "close" );
-      }
-    }
-  });
-}

### 2014-11-17 ui.R

@@ -26,7 +26,6 @@ shinyUI(
                tabPanel(title="", icon=icon("home"),
                         fluidRow(
                             dynDiv(outputId = 'print_output', title="Print Results"),
-                            div(class='intRoPrint', h3('Results from intRo session:')),
                             do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
                         )     
                ),
@@ -36,13 +35,12 @@ shinyUI(
                           tabPanel("Eric Hare"),
                           tabPanel("Andee Kaplan")),
                tabPanel(title="hide_me"),
-               tabPanel(title="", icon=icon("print"), class="print-button", value = "javascript:print_intRo();"),
+               tabPanel(title="", icon=icon("print"), class="print-button", value="javascript:return null;"),
                footer=tagList(includeScript("scripts/top-nav-links.js"),
                               includeScript("scripts/print.js"),
-                              includeScript("scripts/other-helpers.js"),
-                              includeCSS("http://code.jquery.com/ui/1.11.2/themes/smoothness/jquery-ui.css"),
-                              includeScript("http://code.jquery.com/jquery-1.10.2.js"),
-                              includeScript("http://code.jquery.com/ui/1.11.2/jquery-ui.js")
+                              includeScript("scripts/other-helpers.js")
+                              
+                              
                ),
                tags$head(tags$link(rel="shortcut icon", href="images/icon.png"))
     ))

### 2014-11-20 student-paper-competition-2014/paper.Rmd
more estensibility
worked on reactive programming

@@
 
-Beyong being simple, \texttt{intRo} is also consistent.
+Beyond being simple, \texttt{intRo} is also consistent.
 The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules.
@@
 What this means is that \texttt{intRo} is easily extended by the addition of modules within the framework underlying the application.
+This extensibility allows for \texttt{intRo} to be customized for an individual classroom setting, tailored to the needs of a particular course.
 
@@
 
+Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@reactive2013] As implemented by \texttt{shiny} [@shiny], results automatically update when users interact with the interface.
+
+\texttt{intRo} leverages reactive programming techniques as data is manipulated by the user, this carries throughout the entire session.
+In a typical \texttt{shiny} app, users interact with inputs, which in turn yield different results.
+Within \texttt{intRo}, the users are able to interact and manipulate the data underlying the entire application.
+This posed many challenges in the creation of \texttt{intRo}, namely timely save points according to the user's timeline, and reactive updating of variable lists tied to inputs across the entire application.
+
+In the creation of \texttt{intRo} we walked a fine line between giving the user flexibility and having realistic usability.
+At the same time, \texttt{intRo} was created as a consumer of another package, \texttt{shiny}, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability.
+For a tangible example, consider the graphical summaries module.
+We only allow variables of a type consistent with the selected plot to be displayed.
+This is a conscious decision that limits an \texttt{intRo} user's flexibility, while maximizing the usability (by minimizing crashes) of the application.
+On the flip side of this, \texttt{shiny} allows much higher flexibility.
+For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}.
+[Maybe add a figure here?] However, \texttt{shiny} does have limits in its flexibility based on the designers decisions for usability.
+One current example the slider element.
+This element allows for fixed width steps from its minimum to its maximum.
+The JavaScript library being utilized in \texttt{shiny} allows for arbitrary function calls to to generate these steps, however they much be written in plain JavaScript.
+This is an example of a decision made by the developers of \texttt{shiny} to limit functionality in favor of usability of their package.
+
 # Conclusions and future work

### 2014-11-25 student-paper-competition-2014/paper.Rmd
some rearranging and added thoughts

@@
 \includegraphics[width=0.8\textwidth]{images/ui_annotate.png}
-\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
+\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.
+\ak{[Don't forget to update this.]}}
 \label{fig:ui}
@@
 Each module is placed within a higher level category to further help students navigate the tool.
-The default modules included in \texttt{intRo} are found in table \ref{tab:modules}, and more modules can be easy added by an instructor, as detailed in section \ref{sec:extensible}.
+The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can be easy added by an instructor, as detailed in section \ref{sec:extensible}.
 
-\begin{table}
-\centering
-\begin{tabular}[ht!]{l|l|p{.65\textwidth}}
-Category & Module & Description \\
-\hline
-\multirow{2}{*}{Data} & Sources & The data sources module allows the user to select or upload a dataset.
+Within the data category there are two modules, sources and transformation.
+The data sources module allows the user to select or upload a dataset.
 Several datasets are pre-included, which can be selected via a drop-down menu.
@@
 From this module the user can select a subset of the data either by value or random selection.
-The dataset selected in the data module will be made available to all other modules of intRo.\\
+The dataset selected in the data module will be made available to all other modules of intRo.
+The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
+Additionally, numeric variable can be transformed using a power transformation.
+The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.
 
-& Transformation & The data transformation modules allows transformation of numeric variables to categorical and categorical variable to numeric.
-Additionally, numeric variable can be transformed using a power transformation.
-The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.\\
-\hline
-\multirow{2}{*}{Summaries} & Graphical & The graphical summaries module provides several common graphical displays of variables in the dataset.
+The graphical summaries module provides several common graphical displays of variables in the dataset.
 First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
 The variable available to choose will only be those compatible with plot type selected.
-The plot visible on the right will automatically update when these values are selected.\\
-
-& Numerical & The numeric summaries module is another method of computing summaries of the data.
+The plot visible on the right will automatically update when these values are selected.
+The numeric summaries module is another method of computing summaries of the data.
 The user is presented with a list of each variable from their dataset.
@@
 Character or factor variables will display counts of each unique value of that particular variable.
-Additionally, the user has the ability to display summaries by a grouping variable.\\
-\hline
-\multirow{3}{*}{Statistics} & Contingency & The contingency table module allows the user to compute a contingency table for two categorical variables.
-The results can either be displayed as counts, row percentages, column percentages, or total percentages.\\
+Additionally, the user has the ability to display summaries by a grouping variable.
 
-& Regression & The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
+Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in his first semester.
+The contingency table module allows the user to compute a contingency table for two categorical variables.
+The results can either be displayed as counts, row percentages, column percentages, or total percentages.
+The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
 The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
@@
 A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
-A user also has the ability to save residuals and predicted values to their dataset for further analysis.\\
-
-& T test & The T test module allows the user to perform either a one variable or two variable Student's T test.
+A user also has the ability to save residuals and predicted values to their dataset for further analysis.
+The T test module allows the user to perform either a one variable or two variable Student's T test.
 The user can select configuration options, including the direction of the test and the confidence level of the test.
 The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
-\end{tabular}
-\caption{A list of default modules that ship with \texttt{intRo} as well as brief descriptions of each.}
-\label{tab:modules}
-\end{table}
+
+\ak{[Add blurb about easy to read/consistent documentation]}
 
 ## Exciting
-In general, introduction to statistics is a class undergraduates dread.
-It's hard, the language is weird, and the software is hard to use.
+In general, introduction to statistics is a class undergraduate students dread.
+It's hard, the language is unfamiliar, and the software is hard to use.
 With \texttt{intRo}, we hope to flip this paradigm.
-\texttt{intRo} is a fun, easy to use tool that's available on the web, and is even usable on mobile devices.
+\texttt{intRo} is a fun, easy to use tool that's available on the web, and is also usable on mobile devices.
 As a web-based application, this tool is immediately more familiar to students than a desktop application.
@@
 Once students get pulled in by an easy to use interface and fun interaction with their data, \texttt{intRo} has an ulterior motive: to get students excited about programming.
-By navigating about the user interface of \texttt{intRo}, students are actually creating a fully-executable \texttt{R} script that they can download and run locally.
+By navigating about the user interface of \texttt{intRo}, students are actually creating a fully-executable \texttt{R} script that they can download and run locally as well as viewing the script change real-time within the application.
 This code creation element of \texttt{intRo} is meant to generate excitement about programming in \texttt{R} and empower students to feel that they can generate code as well.
@@
 This extensibility allows for \texttt{intRo} to be customized for an individual classroom setting, tailored to the needs of a particular course.
+
+For example, an instructor may wish to introduce hypothesis testing from a randomization framework rather than distributional.
+In this case, a non-parametric test could be added easily as a module to be used instead of the T test.
+This addition would be fairly simple to complete.
+The user would fork \texttt{intRo} from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `observe.R`, `output.R`, `reactive.R`, `static.R`, and `ui.R`.
+Each of these files is written to comply with the reactive nature available in a \texttt{shiny} application.
+The `ui.R` file contains all the user interface elements within the new module's tab.
+This will include any options in the options panel as well as results.
+The `helper.R` and `static.R` files refer to static functions and global variable that may be used within the module.
+The file `output.R` contains the reactive elements that tie output to the user interface within \texttt{shiny} and `observe.R` and `reactive.R` contain elements within an `observer` function or `reactive` function within the \texttt{shiny} framework.
+By placing each of these files within the `modules/nonparametric_test` folder and adding a single line to a navigation file indicating under which module grouping the non-parametric test module should go, a new module is available within \texttt{intRo} upon running the application.
 
@@
 
-Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@reactive2013] As implemented by \texttt{shiny} [@shiny], results automatically update when users interact with the interface.
+Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@bainomugisha2012survey] As implemented by \texttt{shiny} [@shiny], results automatically update when users interact with the interface.
 
@@
 This posed many challenges in the creation of \texttt{intRo}, namely timely save points according to the user's timeline, and reactive updating of variable lists tied to inputs across the entire application.
+
+\ak{[Add information about save buttons for individual pieces of code]}
 
@@
 For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}.
-[Maybe add a figure here?] However, \texttt{shiny} does have limits in its flexibility based on the designers decisions for usability.
+\ak{[Maybe add a figure here?, including example code for dynamically loading the ui]} However, \texttt{shiny} does have limits in its flexibility based on the designers decisions for usability.
 One current example the slider element.
 This element allows for fixed width steps from its minimum to its maximum.
-The JavaScript library being utilized in \texttt{shiny} allows for arbitrary function calls to to generate these steps, however they much be written in plain JavaScript.
+The JavaScript library being utilized in \texttt{shiny} allows for arbitrary function calls to to generate these steps, however they mustf be written in plain JavaScript.
 This is an example of a decision made by the developers of \texttt{shiny} to limit functionality in favor of usability of their package.
@@
 
+\ak{[Make sure we talk about the upcoming R package for module creation, mention also having intro in the classroom in the spring]}
+\ak{[Future challenges: server performance?]}
+
 # References

### 2014-11-26 modules/regression/helper.R
Regression is a real pain

@@ -1,8 +1,42 @@
+my.lm <- function (intro.data, x, y) {
+    cat_and_eval(paste0("intro.regression <- lm(intro.data[,'", y, "'] ~ intro.data[,'", x, "'], na.action = na.exclude)"), env = environment(), file = "code_regression.R", save_result = TRUE)
+    
+    return(intro.regression)
+}
+
+my_regdata <- function(intro.data, x, y) {
+    cat_and_eval("reg.data <- intro.data", , env = environment(), file = "code_regression.R", save_result = TRUE)
+    cat_and_eval(paste0("reg.data$xreg <- reg.data[,'", x, "']"), env = environment(), file = "code_regression.R", save_result = TRUE)
+    cat_and_eval(paste0("reg.data$yreg <- reg.data[,'", y, "']"), env = environment(), file = "code_regression.R", save_result = TRUE)
+    
+    return(reg.data)
+}
+
+my_regresid1 <- function(intro.data, intro.regression, x) {
+    cat_and_eval(paste0("reg.resid1 <- data.frame(residuals = resid(intro.regression), x = intro.data[as.numeric(names(resid(intro.regression))),'", x, "'])"), env = environment(), file = "code_regression.R", save_result = TRUE)
+    
+    return(reg.resid1)
+}
+
+my_regresid2 <- function(intro.regression) {
+    cat_and_eval("myresid <- resid(intro.regression)", env = environment(), file = "code_regression.R", save_result = TRUE)
+    
+    cat_and_eval("yy <- quantile(myresid, na.rm = TRUE, c(0.25, 0.75))", env = environment(), file = "code_regression.R", save_result = TRUE)
+    cat_and_eval("xx <- qnorm(c(0.25, 0.75))", env = environment(), file = "code_regression.R", save_result = TRUE)
+    cat_and_eval("slope <- diff(yy) / diff(xx)", env = environment(), file = "code_regression.R", save_result = TRUE)
+    cat_and_eval("int <- yy[1] - slope * xx[1]", env = environment(), file = "code_regression.R", save_result = TRUE)
+    
+    cat_and_eval("reg.resid2 <- data.frame(yy = qnorm(seq(0, 1, by = (1/(length(na.omit(myresid)) + 1)))[-c(1, (length(na.omit(myresid)) + 2))]),
+                        residuals = sort(myresid))", env = environment(), file = "code_regression.R", save_result = TRUE)
+    
+    return(reg.resid2)
+}
+
[…]

### 2014-11-26 modules/regression/observe.R

@@ -3,3 +3,7 @@ observe({
   updateSelectizeInput(session, "xreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$xreg), input$xreg, intro.numericnames()[1]))
   updateSelectizeInput(session, "yreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$yreg), input$yreg, intro.numericnames()[2]))
 })
+
+observeEvent(input$store_regression, {
+    cat(paste(readLines("code_regression.R"), collapse = "\n"), file = "code_All.R", append = TRUE)
+})

### 2014-11-26 modules/regression/output.R

@@ -1,9 +1,10 @@
 cat_and_eval("p.regplot <- reg.data %>%
   ggvis(x = ~xreg, y = ~yreg) %>%
   layer_points() %>%
-  layer_model_predictions(model = 'lm')", env = environment(), file = "code_regression.R")
+  layer_model_predictions(model = 'lm')", env = environment(), file = "code_regression.R", append = TRUE)
 
 p.regplot %>% bind_shiny("regplot")
+cat("p.regplot\n\n", file = "code_regression.R", append = TRUE)
 
 cat_and_eval("p.resplot1 <- reg.resid1 %>%
   ggvis(x = ~x, y = ~residuals) %>%
@@ -11,6 +12,7 @@ cat_and_eval("p.resplot1 <- reg.resid1 %>%
   set_options(width = 200, height = 200)", env = environment(), file = "code_regression.R", append = TRUE)
 
 p.resplot1 %>% bind_shiny("resplot1")
+cat("p.resplot1\n\n", file = "code_regression.R", append = TRUE)
 
 cat_and_eval("p.resplot2 <- reg.resid2 %>%
   ggvis(x = ~yy, y = ~residuals) %>%
@@ -18,6 +20,7 @@ cat_and_eval("p.resplot2 <- reg.resid2 %>%
   set_options(width = 200, height = 200)", env = environment(), file = "code_regression.R", append = TRUE)
 
 p.resplot2 %>% bind_shiny("resplot2")
+cat("p.resplot2\n\n", file = "code_regression.R", append = TRUE)
 
 cat_and_eval("p.resplot3 <- reg.resid1 %>%
   ggvis(x = ~residuals) %>%
@@ -25,6 +28,7 @@ cat_and_eval("p.resplot3 <- reg.resid1 %>%
   set_options(width = 200, height = 200)", env = environment(), file = "code_regression.R", append = TRUE)
 
 p.resplot3 %>% bind_shiny("resplot3")
+cat("p.resplot3\n\n", file = "code_regression.R", append = TRUE)
 
 output$regtable <- renderTable({
   if (!(input$xreg %in% intro.numericnames()) | !(input$yreg %in% intro.numericnames())) return(NULL)

### 2014-11-26 modules/regression/reactive.R

@@ -1,7 +1,7 @@
 intro.regression <- reactive({
   if (is.null(intro.data())) return(NULL)
   
-  lm.fit <- lm(intro.data()[,input$yreg] ~ intro.data()[,input$xreg], na.action = na.exclude)
+  lm.fit <- my.lm(intro.data(), input$xreg, input$yreg)
   
   if (input$saveresid > oldsaveresid) {
     curxreg <- input$xreg
@@ -18,33 +18,17 @@ intro.regression <- reactive({
 reg.data <- reactive({
   if (is.null(intro.regression())) return(NULL)
   
-  reg.data <- intro.data()
-  reg.data$xreg <- reg.data[,input$xreg]
-  reg.data$yreg <- reg.data[,input$yreg]
-  
-  return(reg.data)
+  return(my_regdata(intro.data(), input$xreg, input$yreg))
 })
 
 reg.resid1 <- reactive({
   if (is.null(intro.regression())) return(NULL)
-  
-  mydat <- data.frame(residuals = resid(intro.regression()), x = intro.data()[as.numeric(names(resid(intro.regression()))),input$xreg])
-  
-  return(mydat)
+    
+  return(my_regresid1(intro.data(), intro.regression(), input$xreg))
 })
 
 reg.resid2 <- reactive({
   if (is.null(intro.regression())) return(NULL)
   
-  myresid <- resid(intro.regression())
-  
-  yy <- quantile(myresid, na.rm = TRUE, c(0.25, 0.75))
-  xx <- qnorm(c(0.25, 0.75))
-  slope <- diff(yy) / diff(xx)
-  int <- yy[1] - slope * xx[1]
-  
-  mydat <- data.frame(yy = qnorm(seq(0, 1, by = (1/(length(na.omit(myresid)) + 1)))[-c(1, (length(na.omit(myresid)) + 2))]),
-                      residuals = sort(myresid))
-  
-  return(mydat)
+  return(my_regresid2(intro.regression()))
 })

### 2014-12-04 student-paper-competition-2014/paper.Rmd
Draft the abstract, intro, conclusion

@@
 \begin{abstract}
-Your abstract goes here...
-...
+\texttt{intRo} is a modern web-based application for performing basic data analysis and statistical routines.
+Leveraging the power of \texttt{R} and \texttt{Shiny}, \texttt{intRo} implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician.
+This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as \texttt{Excel} and \texttt{JMP}.
+\texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}
 \end{abstract}
@@
 # Introduction
+The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been a great development for the scientific community.
+However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student.
+This issue may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors would tend to use a tool like \texttt{R} in their own work.
+Still, signficant compromises must be made when using software like \texttt{JM}P, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.
+In teaching STAT 226 at Iowa State University, we witnessed profound struggles by students attempting to practice the introductory concepts discussed in class.
+Often, these struggles were directly related to the use of the software and not any sort of fundamental misunderstanding of the material.
+
+The release of \texttt{RStudio}'s \texttt{Shiny}, which allows for easily turning an R-based analysis into an interactive web application, led to the creation of an introductory statistics application which we call \texttt{intRo}.
+\texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.
 
@@
 
-\ak{[Make sure we talk about the upcoming R package for module creation, mention also having intro in the classroom in the spring]}
-\ak{[Future challenges: server performance?]}
+We believe \texttt{intRo} can be a powerful and effective tool for introductory statistics education.
+Its modular structure allows it to be flexible enough for many different applications and curriculums.
+Its ease-of-use allows the student to focus his or her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation.
+Reproducible code generated from each analysis can be used to spark an interest in \texttt{R} programming in those who might otherwise not be exposed to it.
+
+Despite the current functionality, there are some practical improvements in the works that will make \texttt{intRo} more accessible and useful.
+In particular, we have begun development on an \texttt{R} package which will allow modules to be created automatically.
+This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo}.
+This will vastly improve intRo's flexibility and allow it to be used in a wider range of curriculums, including more advanced statistics courses.
+We will also be trialing \texttt{intRo}'s use in a statistics course at Iowa State University in Spring Semester 2015.
+This will allow us to assess its usefulness relative to software used in the past, as well as areas for improvement.
+Challenges exist with regards to the wider adoption of \texttt{intRo}.
+For instance, we will need to monitor how well the server \texttt{intRo} is hosted on handles the load of dozens of students performing data analyses.
+If performance issues are noted, the infrastructure used may need to be expanded to handle current and future load.
 

### 2014-12-04 student-paper-competition-2014/paper.Rmd
small typos

@@
 title: "Introducing statistics with \\texttt{intRo}"
-author: "Andee Kaplan & Eric Hare"
+author: "Eric Hare & Andee Kaplan"
 date: "December 14, 2014"
@@
 This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as \texttt{Excel} and \texttt{JMP}.
-\texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}
+\texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}.
+Within this paper we introduce the application and explore the design decisions underlying \texttt{intRo} as well as highlight some challenges and advantages of reactive programming.
 \end{abstract}
@@
 This issue may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors would tend to use a tool like \texttt{R} in their own work.
-Still, signficant compromises must be made when using software like \texttt{JM}P, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.
+Still, signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.
 In teaching STAT 226 at Iowa State University, we witnessed profound struggles by students attempting to practice the introductory concepts discussed in class.
@@
 From this module the user can select a subset of the data either by value or random selection.
-The dataset selected in the data module will be made available to all other modules of intRo.
+The dataset selected in the data module will be made available to all other modules of \texttt{intRo}.
 The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
@@
 This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo}.
-This will vastly improve intRo's flexibility and allow it to be used in a wider range of curriculums, including more advanced statistics courses.
+This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curriculums, including more advanced statistics courses.
 We will also be trialing \texttt{intRo}'s use in a statistics course at Iowa State University in Spring Semester 2015.

### 2014-12-04 student-paper-competition-2014/paper.Rmd
Blurb about printing

@@
 This code creation element of \texttt{intRo} is meant to generate excitement about programming in \texttt{R} and empower students to feel that they can generate code as well.
+\texttt{intRo} uses \texttt{rmarkdown}'s [@rmarkdown] \texttt{render} function in order to print the results, by dymanically executing the user's \texttt{R} script.
+By default, the output will include the \texttt{R} code, but if the user elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
 

### 2014-12-04 student-paper-competition-2014/paper.Rmd
Blurb on store buttons

@@
 This posed many challenges in the creation of \texttt{intRo}, namely timely save points according to the user's timeline, and reactive updating of variable lists tied to inputs across the entire application.
-
-\ak{[Add information about save buttons for individual pieces of code]}
+Because the user may experiment with different configurations or select different variables, we did not want to store all actions taken in the intRo session.
+Rather, each module includes a button allowing the user to explicitly store the output visible in the results panel into the R script.
+This way, output is only stored when the user is satisfied, and the resulting output is not cluttered with unnecessary information.
 

### 2014-12-04 student-paper-competition-2014/paper.Rmd
more typos

--- 2014-12-04 student-paper-competition-2014/paper.Rmd
@@
 
-\ak{[Add blurb about easy to read/consistent documentation]}
+In addition to making \texttt{intRo} intuitive to use, we've also created a documentation website that is consistent with the interface of the application and hyperlinked within the application itself.
+This documentation covers all default modules and makes learning to use \texttt{intRo} even more painless.
 
@@
 This element allows for fixed width steps from its minimum to its maximum.
-The JavaScript library being utilized in \texttt{shiny} allows for arbitrary function calls to to generate these steps, however they mustf be written in plain JavaScript.
+The JavaScript library being utilized in \texttt{shiny} allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript.
 This is an example of a decision made by the developers of \texttt{shiny} to limit functionality in favor of usability of their package.

### 2014-12-05 student-paper-competition-2014/paper.Rmd
small chages to introduction

@@
 
-The release of \texttt{RStudio}'s \texttt{Shiny}, which allows for easily turning an R-based analysis into an interactive web application, led to the creation of an introductory statistics application which we call \texttt{intRo}.
+Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easy for an R-based analysis to be converted to an interactive web application.
+This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}.
+
 \texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.
@@
 # What is \texttt{intRo}?
-\texttt{intRo} is a web-based tool to accompany and introductory statistics class.
+\texttt{intRo} is a web-based tool to accompany an introductory statistics class.
 It is meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class.
 
-We have multiple philosophies that guided the creation of \texttt{intRo}, and this comes out in the design.
+We have three fundamental philosophies that guided the creation of \texttt{intRo}, and this comes out in the design.
 Primarily, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics.
@@
 
-For example, an instructor may wish to introduce hypothesis testing from a randomization framework rather than distributional.
+As an example of possible customization, an instructor may wish to introduce hypothesis testing from a randomization framework rather than distributional.
 In this case, a non-parametric test could be added easily as a module to be used instead of the T test.
@@
 The user would fork \texttt{intRo} from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `observe.R`, `output.R`, `reactive.R`, `static.R`, and `ui.R`.
-Each of these files is written to comply with the reactive nature available in a \texttt{shiny} application.
+Each of these files is written to comply with the reactive nature available in a \texttt{Shiny} application.
 The `ui.R` file contains all the user interface elements within the new module's tab.
@@
 The `helper.R` and `static.R` files refer to static functions and global variable that may be used within the module.
-The file `output.R` contains the reactive elements that tie output to the user interface within \texttt{shiny} and `observe.R` and `reactive.R` contain elements within an `observer` function or `reactive` function within the \texttt{shiny} framework.
+The file `output.R` contains the reactive elements that tie output to the user interface within \texttt{Shiny} and `observe.R` and `reactive.R` contain elements within an `observer` function or `reactive` function within the \texttt{Shiny} framework.
 By placing each of these files within the `modules/nonparametric_test` folder and adding a single line to a navigation file indicating under which module grouping the non-parametric test module should go, a new module is available within \texttt{intRo} upon running the application.
@@
 \texttt{intRo}'s ease of use lends itself to a natural presentation in an introductory statistics course.
-Statistics software typically used in introductory courses, such as JMP and Excel, have restrictive software licensing and platform requirements that can limit the software's accessibility to students.
+Statistics software typically used in introductory courses, such as \texttt{JMP} and \texttt{Excel}, have restrictive software licensing and platform requirements that can limit the software's accessibility to students.
 Even if these limitations can be overcome, these software programs have non-ideal interfaces for an exploratory data analysis.
-JMP, for instance, hides several basic regression functions such as a line of best fit behind small red arrows.
+\texttt{JMP}, for instance, hides several basic regression functions such as a line of best fit behind small red arrows.
 Using \texttt{R} itself in the classroom is an appealing alternative, but suffers from some other problems.
 Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis.
-When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (this experience is made much easier with new tools like \texttt{rmarkdown} [@rmarkdown], but this again is a bit of a learning curve for the new statistics student).
+When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (this experience is made much easier with new tools like \texttt{rmarkdown}, but this again is a bit of a learning curve for the new statistics student).
 
@@
 
-Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@bainomugisha2012survey] As implemented by \texttt{shiny} [@shiny], results automatically update when users interact with the interface.
+Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@bainomugisha2012survey] As implemented by \texttt{Shiny}, results automatically update when users interact with the interface.
 
 \texttt{intRo} leverages reactive programming techniques as data is manipulated by the user, this carries throughout the entire session.
-In a typical \texttt{shiny} app, users interact with inputs, which in turn yield different results.
+In a typical \texttt{Shiny} app, users interact with inputs, which in turn yield different results.
 Within \texttt{intRo}, the users are able to interact and manipulate the data underlying the entire application.
@@
 In the creation of \texttt{intRo} we walked a fine line between giving the user flexibility and having realistic usability.
-At the same time, \texttt{intRo} was created as a consumer of another package, \texttt{shiny}, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability.
+At the same time, \texttt{intRo} was created as a consumer of another package, \texttt{Shiny}, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability.
 For a tangible example, consider the graphical summaries module.
@@
 This is a conscious decision that limits an \texttt{intRo} user's flexibility, while maximizing the usability (by minimizing crashes) of the application.
-On the flip side of this, \texttt{shiny} allows much higher flexibility.
+On the flip side of this, \texttt{Shiny} allows much higher flexibility.
 For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}.
-\ak{[Maybe add a figure here?, including example code for dynamically loading the ui]} However, \texttt{shiny} does have limits in its flexibility based on the designers decisions for usability.
+\ak{[Maybe add a figure here?, including example code for dynamically loading the ui]} However, \texttt{Shiny} does have limits in its flexibility based on the designers decisions for usability.
 One current example the slider element.
 This element allows for fixed width steps from its minimum to its maximum.
-The JavaScript library being utilized in \texttt{shiny} allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript.
-This is an example of a decision made by the developers of \texttt{shiny} to limit functionality in favor of usability of their package.
+The JavaScript library being utilized in \texttt{Shiny} allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript.
+This is an example of a decision made by the developers of \texttt{Shiny} to limit functionality in favor of usability of their package.
 

### 2014-12-05 student-paper-competition-2014/paper.Rmd
Table instead of graphic for education section

@@
 \texttt{intRo} includes modules for both graphical and numerical summaries covering these topics.
-Furthermore, \texttt{intRo} limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics (Figure \ref{fig:ag}).
+Furthermore, \texttt{intRo} limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics (Table \ref{tab:plottypes}).
 
-\begin{figure}
+\begin{table}[ht!]
 \centering
-\includegraphics[width=0.5\textwidth]{images/appropriate_graphics.png}
-\caption{The plot and variable selection boxes in \texttt{intRo}'s graphical module.
-Because the selected plottype is histogram, the interface displays only a single variable selection box (x), and the variables selectable in this box are only numeric.}
-\label{fig:ag}
-\end{figure}
-
+\begin{tabular}{|c|c|c|}
+\hline
+Plot Type&X Variable Type Allowed&Y Variable Type Allowed\\
+\hline
+Histogram&Numeric&N/A\\
+Normal Quantile Plot&Numeric&N/A\\
+Scatterplot&Numeric&Numeric\\
+Line Chart&Numeric&Numeric\\
+Boxplot&Categorical&Numeric\\
+Bar Chart&Categorical&Numeric\\
+Pareto Chart&Categorical&Numeric\\
+Mosaic Plot&Categorical&Categorical\\
+\hline
+\end{tabular}
+\label{tab:plottypes}
+\caption{The plot and variable types allowed in \texttt{intRo}'s graphical module.
+For example, when the selected plottype is histogram, the interface displays only a single variable selection box (x), and the variables selectable in this box are only numeric.
+If a boxplot is selected, the x variable must be categorical and the y variable must be numeric, and the list of choices updates automatically to reflect this.}
+\end{table}
 Part two of this text discusses probability and sampling distributions, which is outside the current realm of \texttt{intRo}, but part three focuses on inferential statistics.

### 2014-12-05 student-paper-competition-2014/paper.Rmd
tweaked conclusion

--- 2014-12-05 student-paper-competition-2014/paper.Rmd
@@
 
+\ak{[Add some references about software discouraging people, and mention other new tools out there like swirl.]}
+
 # What is \texttt{intRo}?
@@
 Primarily, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics.
-Additionally, \texttt{intRo} is an \emph{extensible} tool, allowing for a user of \texttt{intRo} to tailor the tool for his or her own classroom needs.
+Additionally, \texttt{intRo} is an \emph{extensible} tool, allowing for a user of \texttt{intRo} to tailor the tool for her own classroom needs.
 
@@
 Several datasets are pre-included, which can be selected via a drop-down menu.
-The user can check a box to upload a dataset, and then select the CSV file from his computer.
+The user can check a box to upload a dataset, and then select the CSV file from her computer.
 Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
@@
 
-Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in his first semester.
+Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester.
 The contingency table module allows the user to compute a contingency table for two categorical variables.
@@
 Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the \texttt{intRo} interface to obtain the results.
-For topics currently unsupported by \texttt{intRo}, the instructor can supplement its use with other web-based modules that are publicly available.
+For topics currently unsupported by \texttt{intRo}, the instructor can supplement its use with other web-based modules that are publicly available or create their own modules within \texttt{intRo} as detailed in section \ref{sec:extensible}.
 
@@
 Its modular structure allows it to be flexible enough for many different applications and curriculums.
-Its ease-of-use allows the student to focus his or her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation.
+Its ease-of-use allows the student to focus her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation.
 Reproducible code generated from each analysis can be used to spark an interest in \texttt{R} programming in those who might otherwise not be exposed to it.
 
-Despite the current functionality, there are some practical improvements in the works that will make \texttt{intRo} more accessible and useful.
-In particular, we have begun development on an \texttt{R} package which will allow modules to be created automatically.
-This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo}.
-This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curriculums, including more advanced statistics courses.
-We will also be trialing \texttt{intRo}'s use in a statistics course at Iowa State University in Spring Semester 2015.
-This will allow us to assess its usefulness relative to software used in the past, as well as areas for improvement.
-Challenges exist with regards to the wider adoption of \texttt{intRo}.
-For instance, we will need to monitor how well the server \texttt{intRo} is hosted on handles the load of dozens of students performing data analyses.
-If performance issues are noted, the infrastructure used may need to be expanded to handle current and future load.
+In addition to the current functionality, there are some practical improvements in the works that will make \texttt{intRo} more useful to both students and instructors.
+In particular, we have begun development on an \texttt{R} package which will allow \texttt{intRo} modules to be created automatically from user written \texttt{R} code.
+This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo} as well as translate user code to \texttt{intRo} compatible code as populates the necessary files.
+This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
+
+Additionally,  \texttt{intRo} will be trialed in an introductory statistics course at Iowa State University in Spring 2015.
+This will allow us to assess its usefulness relative to software used in the past, as well as guage areas for improvement.
+Furthermore, we can determine the effectiveness of code printing on generating excitement from the students about programming in \texttt{R}.
+
+Challenges do exist with regards to the wider adoption of \texttt{intRo}.
+For instance, we will need to monitor how well the server hosting \texttt{intRo} handles the load of dozens of students performing data analyses at once.
+If performance issues are encountered, the infrastructure used may need to be expanded to handle current and future load.
+An unknown quantity will be how feasible it is to increase adoption of \texttt{intRo} across the department, as well as to other universities.
+
+Regardless, tools that focus on usability and extensibility in statistics education, such as \texttt{intRo}, are sure to encourage the next round of innovators to be interested and excited about statistical computing.
 

### 2014-12-05 student-paper-competition-2014/paper.Rmd
added flow chart for application creation

@@
 \includegraphics[width=0.8\textwidth]{images/ui_annotate.png}
-\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.
-\ak{[Don't forget to update this.]}}
+\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
 \label{fig:ui}
@@
 By placing each of these files within the `modules/nonparametric_test` folder and adding a single line to a navigation file indicating under which module grouping the non-parametric test module should go, a new module is available within \texttt{intRo} upon running the application.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=\textwidth]{images/app_creation_modules.png}
+\caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.}
+\label{fig:app_creation_modules}
+\end{figure}
+
+The process by which \texttt{intRo} is dynamically created upon running the application through module sourcing is detailed in figure~\ref{fig:app_creation_modules}.
+The key driver to populating \texttt{server.R} and \texttt{ui.R} is the file \texttt{modules.txt}, which defines the placement of each module.
+The interface is then created with the following statement.
+
+\footnotesize
+```{r ui, eval=FALSE}
+## Source ui
+module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
+sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)
+
+## Generates the UI tabs
+mylist <- list()
+old_heading <- ""
+for (i in 1:nrow(module_info)) {
+    if (module_info[i,1] != old_heading) {
+        if (i != 1) mylist[[length(mylist) + 1]] <- "-----"
+        mylist[[length(mylist) + 1]] <- as.character(module_info[i,1])
+        old_heading <- module_info[i,1]
+    }
+    mylist[[length(mylist) + 1]] <- get(paste(module_info[i,2], "ui", sep = "_"))
+}
+
+## Create the UI
+shinyUI(
+    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
+               tabPanel(title="", icon=icon("home"),
+                        fluidRow(
+                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
+                        )
+               ), ...
+    ))
+```
+\normalsize
+
+The key piece of code being the `do.call` statement loading the list of ui elements from the module's \texttt{ui.R} file. The server functions are then dynamically generated using a similar method.
+
+\footnotesize
+```{r server, eval=FALSE}
+shinyServer(function(input, output, session) {
+  ## Module info
+  module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
+
+  ## Modules
+  types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
+  modules_tosource <- file.path("modules", apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
+
+  ## Source the modules
+  for (mod in modules_tosource) {
+      source(mod, local = TRUE)
+  }
+}
+```
+\normalsize
+In this way, we were able to have \texttt{intRo} be fully extensible, its structure and functionality dependent entirely on the modules present within the application.
 

### 2014-12-05 student-paper-competition-2014/paper.Rmd
remove some comments

@@
 ---
-
+\thispagestyle{empty}
 \begin{abstract}
@@
 \newpage
+\setcounter{page}{1}
 
@@
 
-The process by which \texttt{intRo} is dynamically created upon running the application through module sourcing is detailed in figure~\ref{fig:app_creation_modules}.
+The process by which \texttt{intRo} is dynamically created upon running the application through module sourcing is detailed in figure \ref{fig:app_creation_modules}.
 The key driver to populating \texttt{server.R} and \texttt{ui.R} is the file \texttt{modules.txt}, which defines the placement of each module.
@@
 On the flip side of this, \texttt{Shiny} allows much higher flexibility.
-For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}.
-\ak{[Maybe add a figure here?, including example code for dynamically loading the ui]} However, \texttt{Shiny} does have limits in its flexibility based on the designers decisions for usability.
+For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}, as detailed in section \ref{sec:extensible}.
+However, \texttt{Shiny} does have limits in its flexibility based on the designers decisions for usability.
 One current example the slider element.

### 2014-12-09 modules/transform/reactive.R
Set mydir argument

@@ -27,7 +27,7 @@
                 cat_and_eval(paste0("intro.data[, '", intro.transform.colname(), "'] <- trans_x"),  mydir = userdir, file = "code_transform.R", append = TRUE)
             }
         } else if (input$trans %in% c("categorical", "numeric")) {
-            if (input$trans == "numeric") cat_and_eval(paste0("trans_x <- as.numeric(intro.data[,'", curtrans, "'])"),  mydir = userdir, file = "code_transform.R", append = TRUE) else cat_and_eval(paste0("trans_x <- factor(intro.data[,'", curtrans, "'])"), file = "code_transform.R", append = TRUE)
+            if (input$trans == "numeric") cat_and_eval(paste0("trans_x <- as.numeric(intro.data[,'", curtrans, "'])"),  mydir = userdir, file = "code_transform.R", append = TRUE) else cat_and_eval(paste0("trans_x <- factor(intro.data[,'", curtrans, "'])"), mydir = userdir, file = "code_transform.R", append = TRUE)
             cat_and_eval(paste0("intro.data[, '", intro.transform.colname(), "'] <- trans_x"),  mydir = userdir, file = "code_transform.R", append = TRUE)
         } else {
             cat_and_eval(paste0("intro.data[, '", intro.transform.colname(), "'] <- curtrans"),  mydir = userdir, file = "code_transform.R", append = TRUE)

### 2014-12-09 modules/regression/observe.R
Fix bug with transform

@@ -1,7 +1,7 @@
 ## Regression Observers
 observe({
   updateSelectizeInput(session, "xreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$xreg), input$xreg, intro.numericnames()[1]))
-  updateSelectizeInput(session, "yreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$yreg), input$yreg, intro.numericnames()[2]))
+  updateSelectizeInput(session, "yreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$yreg), input$yreg, intro.numericnames()[1]))
 })
 
 observeEvent(input$store_regression, {

### 2014-12-09 modules/transform/observe.R

@@ -1,4 +1,8 @@
     ## Transform Observers
     observe({
-        updateSelectInput(session, "var_trans", choices = (if (input$trans %in% c("power", "categorical")) intro.numericnames() else intro.categoricnames()))
+        updateSelectizeInput(session, "var_trans", choices = (if (input$trans %in% c("power", "categorical")) intro.numericnames() else intro.categoricnames()))
+    })
+    
+    observe({
+        updateSelectizeInput(session, "trans", choices = (if (numericNames(intro.data())[1] == "") c("Numeric" = "numeric") else if (categoricNames(intro.data())[1] == "") c("Power" = "power", "Categorical" = "categorical") else c("Power" = "power", "Categorical" = "categorical", "Numeric" = "numeric")))
     })

### 2014-12-09 server.R

@@ -19,7 +19,7 @@ shinyServer(function(input, output, session) {
     
     #cat(paste(readLines("global.R"), collapse = "\n"), file = "code_global.R")
     cat("library(RCurl)\n", file = file.path(userdir, "code_All.R"))
-    cat("eval(parse(text = getURL('https://raw.githubusercontent.com/gammarama/intRo/dev/global.R')))", file = file.path(userdir, "code_All.R"), append = TRUE)
+    cat("eval(parse(text = getURL('https://raw.githubusercontent.com/gammarama/intRo/master/global.R')))", file = file.path(userdir, "code_All.R"), append = TRUE)
     
     ## Modules
     types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")

### 2014-12-09 student-paper-competition-2014/paper.Rmd
Formatting for paper

@@
 \thispagestyle{empty}
+
 \begin{abstract}
@@
 
-\newpage
+\twocolumn
 \setcounter{page}{1}
@@
 
-\begin{figure}[ht!]
+\begin{figure*}[ht!]
 \centering
-\includegraphics[width=0.8\textwidth]{images/ui_annotate.png}
+\includegraphics[width=\linewidth]{images/ui_annotate.png}
 \caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
 \label{fig:ui}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[ht!]
+\begin{figure*}[ht!]
 \centering
-\includegraphics[width=\textwidth]{images/app_creation_modules.png}
+\includegraphics[width=\linewidth]{images/app_creation_modules.png}
 \caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.}
 \label{fig:app_creation_modules}
-\end{figure}
+\end{figure*}
 
@@
 \footnotesize
-```{r ui, eval=FALSE}
+```{r ui, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
 ## Source ui
@@
 \footnotesize
-```{r server, eval=FALSE}
+```{r server, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
 shinyServer(function(input, output, session) {
@@
   types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
-  modules_tosource <- file.path("modules", apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
+  modules_tosource <- file.path("modules", 
+                                apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
 
@@
   }
-}
+})
 ```

### 2014-12-11 student-paper-competition-2014/paper.Rmd
New powerpoint file and image updates

@@
 \centering
-\includegraphics[width=\linewidth]{images/ui_annotate.png}
+\includegraphics[width=\linewidth]{images/ui_annotate.pdf}
 \caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
@@
 \centering
-\includegraphics[width=\linewidth]{images/app_creation_modules.png}
+\includegraphics[width=\linewidth]{images/app_creation_modules.pdf}
 \caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.}
@@
 
-## Generates the UI tabs
-mylist <- list()
-old_heading <- ""
-for (i in 1:nrow(module_info)) {
-    if (module_info[i,1] != old_heading) {
-        if (i != 1) mylist[[length(mylist) + 1]] <- "-----"
-        mylist[[length(mylist) + 1]] <- as.character(module_info[i,1])
-        old_heading <- module_info[i,1]
-    }
-    mylist[[length(mylist) + 1]] <- get(paste(module_info[i,2], "ui", sep = "_"))
-}
-
+## mylist is a list containing the different ui module code
 ## Create the UI
@@
 
-\begin{table}[ht!]
+\begin{table*}[ht!]
 \centering
@@
 If a boxplot is selected, the x variable must be categorical and the y variable must be numeric, and the list of choices updates automatically to reflect this.}
-\end{table}
+\end{table*}
 Part two of this text discusses probability and sampling distributions, which is outside the current realm of \texttt{intRo}, but part three focuses on inferential statistics.

### 2014-12-11 student-paper-competition-2014/extra-text.Rmd
removed module info

--- 2014-12-11 student-paper-competition-2014/paper.Rmd
@@
 ---
-title: "Introducing statistics with \\texttt{intRo}"
-author: "Eric Hare & Andee Kaplan"
-date: "December 14, 2014"
-output:
-  pdf_document:
-    includes:
-      in_header: preamble.tex
-    number_sections: yes
-bibliography: refs.bib
+title: "Untitled"
+author: "Eric Hare and Andee Kaplan"
+date: "12/11/2014"
+output: html_document
 ---
-\thispagestyle{empty}
-
-\begin{abstract}
-\texttt{intRo} is a modern web-based application for performing basic data analysis and statistical routines.
-Leveraging the power of \texttt{R} and \texttt{Shiny}, \texttt{intRo} implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician.
-This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as \texttt{Excel} and \texttt{JMP}.
-\texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}.
-Within this paper we introduce the application and explore the design decisions underlying \texttt{intRo} as well as highlight some challenges and advantages of reactive programming.
-\end{abstract}
-
-\twocolumn
-\setcounter{page}{1}
-
-# Introduction
-The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been a great development for the scientific community.
-However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student.
-This issue may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors would tend to use a tool like \texttt{R} in their own work.
-Still, signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.
-In teaching STAT 226 at Iowa State University, we witnessed profound struggles by students attempting to practice the introductory concepts discussed in class.
-Often, these struggles were directly related to the use of the software and not any sort of fundamental misunderstanding of the material.
-
-Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easy for an R-based analysis to be converted to an interactive web application.
-This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}.
-
-\texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.
-
-\ak{[Add some references about software discouraging people, and mention other new tools out there like swirl.]}
-
-# What is \texttt{intRo}?
-\texttt{intRo} is a web-based tool to accompany an introductory statistics class.
-It is meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class.
-
-We have three fundamental philosophies that guided the creation of \texttt{intRo}, and this comes out in the design.
-Primarily, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics.
-Additionally, \texttt{intRo} is an \emph{extensible} tool, allowing for a user of \texttt{intRo} to tailor the tool for her own classroom needs.
-
-## Easy
-In order for an application to be useful to students in their first statistics course, it can't hog the spotlight.
-Statistics as a subject should shine through as a fun and interesting topic to study, rather than being bogged down by difficult to learn software.
-With this in mind, in the development of \texttt{intRo} we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training.
-We used large, easy to click icons in the page header to help students find exactly what they need easily.
-With this easy to use framework, we also made the functionality available in \texttt{intRo} the minimal necessary for an introductory statistics course.
-Additional functionality is nice, but if it clutters the UI beyond the point of usability, it will frustrate an introductory student.
-
-Beyond being simple, \texttt{intRo} is also consistent.
-The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules.
-Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code.
-In figure \ref{fig:ui}, we highlight the five elements that comprise the \texttt{intRo} interface.
-
-\begin{figure*}[ht!]
-\centering
-\includegraphics[width=\linewidth]{images/ui_annotate.pdf}
-\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
-\label{fig:ui}
-\end{figure*}
-
- 1.
-**Top Navigation** - The top navigation bar includes two sets of buttons.
-The left-aligned buttons are informational buttons.
-The first is a link to \texttt{intRo}.
-The second is a link to this documentation page.
-The third is a link to the GitHub repository where the code for \texttt{intRo} is housed.
-The final button is a link to our websites, which contain contact information if there are any questions or comments.
-The right-aligned buttons are \texttt{intRo} utilities.
-The first is a button to toggle the visibility of the code panel (5).
-The last is a button to print the active module results, and the associated code (if visible).
- 2. **Side Navigation** - The side navigation panel includes the modules that perform a specific data analysis task.
- 3. **Options Panel** - The options panel includes module-specific options which the user can use to customize their results.
- 4. **Results Panel** - The results pane displays the result of the selected module and options.
- 5.
-**Code Panel** - The code panel displays the \texttt{R} code used to generate the results from the student's \texttt{intRo} session.
-The code panel is shown by default to facilitate an interest in coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.
-
-There are multiple modules within \texttt{intRo}, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
-Each module is placed within a higher level category to further help students navigate the tool.
-The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can be easy added by an instructor, as detailed in section \ref{sec:extensible}.
 
@@
 The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
-
-In addition to making \texttt{intRo} intuitive to use, we've also created a documentation website that is consistent with the interface of the application and hyperlinked within the application itself.
-This documentation covers all default modules and makes learning to use \texttt{intRo} even more painless.
-
-## Exciting
-In general, introduction to statistics is a class undergraduate students dread.
-It's hard, the language is unfamiliar, and the software is hard to use.
-With \texttt{intRo}, we hope to flip this paradigm.
-\texttt{intRo} is a fun, easy to use tool that's available on the web, and is also usable on mobile devices.
-As a web-based application, this tool is immediately more familiar to students than a desktop application.
-With interactive plots from \texttt{ggvis} [@ggvis], \texttt{intRo} brings a `wow'-factor to learning introductory statistics.
-
-Once students get pulled in by an easy to use interface and fun interaction with their data, \texttt{intRo} has an ulterior motive: to get students excited about programming.
-By navigating about the user interface of \texttt{intRo}, students are actually creating a fully-executable \texttt{R} script that they can download and run locally as well as viewing the script change real-time within the application.
-This code creation element of \texttt{intRo} is meant to generate excitement about programming in \texttt{R} and empower students to feel that they can generate code as well.
-\texttt{intRo} uses \texttt{rmarkdown}'s [@rmarkdown] \texttt{render} function in order to print the results, by dymanically executing the user's \texttt{R} script.
-By default, the output will include the \texttt{R} code, but if the user elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
-
-## Extensible
-\label{sec:extensible}
-On the front end, user interaction with \texttt{intRo} is split into bitesize chunks that we call modules.
-This is true of the back end as well.
-Each module is a self contained set of \texttt{R} code that is dynamically added to the application at run time.
-What this means is that \texttt{intRo} is easily extended by the addition of modules within the framework underlying the application.
-This extensibility allows for \texttt{intRo} to be customized for an individual classroom setting, tailored to the needs of a particular course.
-
-As an example of possible customization, an instructor may wish to introduce hypothesis testing from a randomization framework rather than distributional.
-In this case, a non-parametric test could be added easily as a module to be used instead of the T test.
-This addition would be fairly simple to complete.
-The user would fork \texttt{intRo} from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `observe.R`, `output.R`, `reactive.R`, `static.R`, and `ui.R`.
-Each of these files is written to comply with the reactive nature available in a \texttt{Shiny} application.
-The `ui.R` file contains all the user interface elements within the new module's tab.
-This will include any options in the options panel as well as results.
-The `helper.R` and `static.R` files refer to static functions and global variable that may be used within the module.
-The file `output.R` contains the reactive elements that tie output to the user interface within \texttt{Shiny} and `observe.R` and `reactive.R` contain elements within an `observer` function or `reactive` function within the \texttt{Shiny} framework.
-By placing each of these files within the `modules/nonparametric_test` folder and adding a single line to a navigation file indicating under which module grouping the non-parametric test module should go, a new module is available within \texttt{intRo} upon running the application.
-
-\begin{figure*}[ht!]
-\centering
-\includegraphics[width=\linewidth]{images/app_creation_modules.pdf}
-\caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.}
-\label{fig:app_creation_modules}
-\end{figure*}
-
-The process by which \texttt{intRo} is dynamically created upon running the application through module sourcing is detailed in figure \ref{fig:app_creation_modules}.
-The key driver to populating \texttt{server.R} and \texttt{ui.R} is the file \texttt{modules.txt}, which defines the placement of each module.
-The interface is then created with the following statement.
-
-\footnotesize
-```{r ui, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
-## Source ui
-module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
-sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)
-
-## mylist is a list containing the different ui module code
-## Create the UI
-shinyUI(
-    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
-               tabPanel(title="", icon=icon("home"),
-                        fluidRow(
-                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
-                        )
-               ), ...
-    ))
-```
-\normalsize
-
-The key piece of code being the `do.call` statement loading the list of ui elements from the module's \texttt{ui.R} file. The server functions are then dynamically generated using a similar method.
-
-\footnotesize
-```{r server, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
-shinyServer(function(input, output, session) {
-  ## Module info
-  module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
-
-  ## Modules
-  types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
-  modules_tosource <- file.path("modules", 
-                                apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
-
-  ## Source the modules
-  for (mod in modules_tosource) {
-      source(mod, local = TRUE)
-  }
-})
-```
-\normalsize
-In this way, we were able to have \texttt{intRo} be fully extensible, its structure and functionality dependent entirely on the modules present within the application.
-
-# In the classroom
-\texttt{intRo}'s ease of use lends itself to a natural presentation in an introductory statistics course.
-Statistics software typically used in introductory courses, such as \texttt{JMP} and \texttt{Excel}, have restrictive software licensing and platform requirements that can limit the software's accessibility to students.
-Even if these limitations can be overcome, these software programs have non-ideal interfaces for an exploratory data analysis.
-\texttt{JMP}, for instance, hides several basic regression functions such as a line of best fit behind small red arrows.
-Using \texttt{R} itself in the classroom is an appealing alternative, but suffers from some other problems.
-Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis.
-When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (this experience is made much easier with new tools like \texttt{rmarkdown}, but this again is a bit of a learning curve for the new statistics student).
-
-We believe that \texttt{intRo} targets these key limitations.
-The interface is designed to be quick, seamless, and easy to understand.
-Because \texttt{intRo} is web-based, the statistical analysis can be performed from nearly any modern device.
-\texttt{R} underlies the analysis, but users need not be familiar with \texttt{R} or programming to perform it.
-The focal point of the interface is the output panel, allowing the student to remain focused on the results and limiting the need to fight with the interface itself to get work done.
-For those students who may have a natural inclination for programming, executable code is generated to allow a local reproduction of the analysis.
-
-The features of \texttt{intRo} were selected in order to apply to the most commonly used lesson plans in early statistics courses.
-The textbook "The Art and Science of Learning from Data" by Agresti and Franklin [-@agresti2012statistics] references many of the focal points of introductory statistics, and \texttt{intRo} supports performing a majority of them.
-Part one, titled "Gathering and Exploring Data", covers graphical summaries, measures of spread and center, and avoiding the misuse of graphics in situations they are not appropriate.
-\texttt{intRo} includes modules for both graphical and numerical summaries covering these topics.
-Furthermore, \texttt{intRo} limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics (Table \ref{tab:plottypes}).
-
-\begin{table*}[ht!]
-\centering
-\begin{tabular}{|c|c|c|}
-\hline
-Plot Type&X Variable Type Allowed&Y Variable Type Allowed\\
-\hline
-Histogram&Numeric&N/A\\
-Normal Quantile Plot&Numeric&N/A\\
-Scatterplot&Numeric&Numeric\\
-Line Chart&Numeric&Numeric\\
-Boxplot&Categorical&Numeric\\
-Bar Chart&Categorical&Numeric\\
-Pareto Chart&Categorical&Numeric\\
-Mosaic Plot&Categorical&Categorical\\
-\hline
-\end{tabular}
-\label{tab:plottypes}
-\caption{The plot and variable types allowed in \texttt{intRo}'s graphical module.
-For example, when the selected plottype is histogram, the interface displays only a single variable selection box (x), and the variables selectable in this box are only numeric.
-If a boxplot is selected, the x variable must be categorical and the y variable must be numeric, and the list of choices updates automatically to reflect this.}
-\end{table*}
-Part two of this text discusses probability and sampling distributions, which is outside the current realm of \texttt{intRo}, but part three focuses on inferential statistics.
-The inferential statistics include confidence intervals, hypothesis testing, and comparing two variables, all supported natively in \texttt{intRo}.
-Finally, part four of the text discusses the association between variables, including regression, functions that \texttt{intRo} performs with ease.
-Some of the more advanced topics of this section are currently unsupported by \texttt{intRo}, including multiple regression and nonparametric statistics.
-
-Statistics courses with a lab component can most benefit from the use of \texttt{intRo}.
-As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries.
-They simply upload a dataset and click the module appropriate to the section of the class.
-Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in.
-Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the \texttt{intRo} interface to obtain the results.
-For topics currently unsupported by \texttt{intRo}, the instructor can supplement its use with other web-based modules that are publicly available or create their own modules within \texttt{intRo} as detailed in section \ref{sec:extensible}.
-
-# Reactive programming
-
-Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@bainomugisha2012survey] As implemented by \texttt{Shiny}, results automatically update when users interact with the interface.
-
-\texttt{intRo} leverages reactive programming techniques as data is manipulated by the user, this carries throughout the entire session.
-In a typical \texttt{Shiny} app, users interact with inputs, which in turn yield different results.
-Within \texttt{intRo}, the users are able to interact and manipulate the data underlying the entire application.
-This posed many challenges in the creation of \texttt{intRo}, namely timely save points according to the user's timeline, and reactive updating of variable lists tied to inputs across the entire application.
-Because the user may experiment with different configurations or select different variables, we did not want to store all actions taken in the intRo session.
-Rather, each module includes a button allowing the user to explicitly store the output visible in the results panel into the R script.
-This way, output is only stored when the user is satisfied, and the resulting output is not cluttered with unnecessary information.
-
-In the creation of \texttt{intRo} we walked a fine line between giving the user flexibility and having realistic usability.
-At the same time, \texttt{intRo} was created as a consumer of another package, \texttt{Shiny}, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability.
-For a tangible example, consider the graphical summaries module.
-We only allow variables of a type consistent with the selected plot to be displayed.
-This is a conscious decision that limits an \texttt{intRo} user's flexibility, while maximizing the usability (by minimizing crashes) of the application.
-On the flip side of this, \texttt{Shiny} allows much higher flexibility.
-For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}, as detailed in section \ref{sec:extensible}.
-However, \texttt{Shiny} does have limits in its flexibility based on the designers decisions for usability.
-One current example the slider element.
-This element allows for fixed width steps from its minimum to its maximum.
-The JavaScript library being utilized in \texttt{Shiny} allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript.
-This is an example of a decision made by the developers of \texttt{Shiny} to limit functionality in favor of usability of their package.
-
-# Conclusions and future work
-
-We believe \texttt{intRo} can be a powerful and effective tool for introductory statistics education.
-Its modular structure allows it to be flexible enough for many different applications and curriculums.
-Its ease-of-use allows the student to focus her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation.
-Reproducible code generated from each analysis can be used to spark an interest in \texttt{R} programming in those who might otherwise not be exposed to it.
-
-In addition to the current functionality, there are some practical improvements in the works that will make \texttt{intRo} more useful to both students and instructors.
-In particular, we have begun development on an \texttt{R} package which will allow \texttt{intRo} modules to be created automatically from user written \texttt{R} code.
-This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo} as well as translate user code to \texttt{intRo} compatible code as populates the necessary files.
-This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
-
-Additionally,  \texttt{intRo} will be trialed in an introductory statistics course at Iowa State University in Spring 2015.
-This will allow us to assess its usefulness relative to software used in the past, as well as guage areas for improvement.
-Furthermore, we can determine the effectiveness of code printing on generating excitement from the students about programming in \texttt{R}.
-
-Challenges do exist with regards to the wider adoption of \texttt{intRo}.
-For instance, we will need to monitor how well the server hosting \texttt{intRo} handles the load of dozens of students performing data analyses at once.
-If performance issues are encountered, the infrastructure used may need to be expanded to handle current and future load.
-An unknown quantity will be how feasible it is to increase adoption of \texttt{intRo} across the department, as well as to other universities.
-
-Regardless, tools that focus on usability and extensibility in statistics education, such as \texttt{intRo}, are sure to encourage the next round of innovators to be interested and excited about statistical computing.
-
-# References

### 2014-12-11 student-paper-competition-2014/paper.Rmd
removed module info

--- 2014-12-09 student-paper-competition-2014/paper.Rmd
@@
 The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can be easy added by an instructor, as detailed in section \ref{sec:extensible}.
+The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.
 
-Within the data category there are two modules, sources and transformation.
-The data sources module allows the user to select or upload a dataset.
-Several datasets are pre-included, which can be selected via a drop-down menu.
-The user can check a box to upload a dataset, and then select the CSV file from her computer.
-Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
-From this module the user can select a subset of the data either by value or random selection.
-The dataset selected in the data module will be made available to all other modules of \texttt{intRo}.
-The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
-Additionally, numeric variable can be transformed using a power transformation.
-The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.
-
-The graphical summaries module provides several common graphical displays of variables in the dataset.
-First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
-The variable available to choose will only be those compatible with plot type selected.
-The plot visible on the right will automatically update when these values are selected.
-The numeric summaries module is another method of computing summaries of the data.
-The user is presented with a list of each variable from their dataset.
-The user can select one or more different variables, and will be presented with a table of some summary statistics.
-The summary statistics displayed vary depending on the type of the variables.
-For instance, for a numeric variable, a five number summary will be displayed.
-Character or factor variables will display counts of each unique value of that particular variable.
-Additionally, the user has the ability to display summaries by a grouping variable.
-
-Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester.
-The contingency table module allows the user to compute a contingency table for two categorical variables.
-The results can either be displayed as counts, row percentages, column percentages, or total percentages.
-The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
-The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
-Results of the regression will be displayed on the right.
-A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
-A user also has the ability to save residuals and predicted values to their dataset for further analysis.
-The T test module allows the user to perform either a one variable or two variable Student's T test.
-The user can select configuration options, including the direction of the test and the confidence level of the test.
-The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
-
-In addition to making \texttt{intRo} intuitive to use, we've also created a documentation website that is consistent with the interface of the application and hyperlinked within the application itself.
+In addition to making \texttt{intRo} intuitive to use, we've also created a documentation website that is consistent with the interface of the application.
+The documentation is available at http://gammarama.github.io/intRo/ and hyperlinked within the application itself.
 This documentation covers all default modules and makes learning to use \texttt{intRo} even more painless.

### 2014-12-11 student-paper-competition-2014/paper.Rmd
Talk about ggvis in further work

--- 2014-12-11 student-paper-competition-2014/paper.Rmd
@@
 This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
+Furthermore, we would like to utilize the expanded interactive capabilities of \texttt{ggvis} in order to make \texttt{intRo's} plots more engaging to students.
+One way to do this would be implementing linked plots, in which interactions with one plot are reflected in other plots that illustrate the same data.
+This would be particularly useful in the regression module so that students could explore observations with high influence and leverage.
 

### 2014-12-11 student-paper-competition-2014/paper.Rmd
Limitation of dataset size (further work)

@@
 An unknown quantity will be how feasible it is to increase adoption of \texttt{intRo} across the department, as well as to other universities.
+One limitation of \texttt{intRo} is that uploading a dataset beyond about 30,000 rows tends to be slow.
+Even once the data is successfully uploaded, the default modules produce results more slowly than with smaller datasets.
+This is a limitation that should be further investigated if and when \texttt{intRo} sees wider adoption.
 

### 2014-12-11 student-paper-competition-2014/paper.Rmd
Incorporate scrimshaw paper

@@
 Still, signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.
+
 In teaching STAT 226 at Iowa State University, we witnessed profound struggles by students attempting to practice the introductory concepts discussed in class.
-Often, these struggles were directly related to the use of the software and not any sort of fundamental misunderstanding of the material.
+Scrimshaw notes in his manuscript that "open-ended packages, like any others, may create obstacles to learning simply through their lack of user-friendliness in the sheer mechanics of operating them, rather than any intrinsic difficulty in the content...." [@scrimshaw2001computers] In our own experience teaching, students' struggles were often directly related to the use of the software and not any sort of fundamental misunderstanding of the material, in agreement with Scrimshaw's finding.
 

### 2014-12-11 student-paper-competition-2014/paper.Rmd
Blurb about the other software

@@
 
+Other software has recently been written in an attempt to spur student interest in introductory statistics and R programming.
+DataCamp's courses are user-friendly ways to learning basic \texttt{R} programming and data analysis techniques.
+\texttt{Swirl} is a similar interactive tool to make learning \texttt{R} more fun by learning \texttt{R} within \texttt{R} itself.
+Project MOSAIC has created a suite of tools to simplify the teaching of statistics in the form of an R package.
+The primary goal of DataCamp and Swirl is to teach \texttt{R} programming, rather than facilitate the learning of introductory statistics.
+Project MOSAIC's goal is to faciliate this learning, but using the package requires a knowledge of R programming that the introductory student may not have.
+
 Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easy for an R-based analysis to be converted to an interactive web application.
 This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}.
-
 \texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.

### 2014-12-11 student-paper-competition-2014/paper.Rmd
something something

@@
 Other software has recently been written in an attempt to spur student interest in introductory statistics and R programming.
-DataCamp's courses are user-friendly ways to learning basic \texttt{R} programming and data analysis techniques.
-\texttt{Swirl} is a similar interactive tool to make learning \texttt{R} more fun by learning \texttt{R} within \texttt{R} itself.
-Project MOSAIC has created a suite of tools to simplify the teaching of statistics in the form of an R package.
+DataCamp's [@datacamp] courses are user-friendly ways to learning basic \texttt{R} programming and data analysis techniques.
+\texttt{Swirl} [@swirl] is a similar interactive tool to make learning \texttt{R} more fun by learning \texttt{R} within \texttt{R} itself.
+Project MOSAIC [@mosaic] has created a suite of tools to simplify the teaching of statistics in the form of an R package.
 The primary goal of DataCamp and Swirl is to teach \texttt{R} programming, rather than facilitate the learning of introductory statistics.
@@
 \texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.
-
-\ak{[Add some references about software discouraging people, and mention other new tools out there like swirl.]}
 
@@
 \end{tabular}
-\label{tab:plottypes}
 \caption{The plot and variable types allowed in \texttt{intRo}'s graphical module.
@@
 If a boxplot is selected, the x variable must be categorical and the y variable must be numeric, and the list of choices updates automatically to reflect this.}
+\label{tab:plottypes}
 \end{table*}

### 2014-12-12 student-paper-competition-2014/paper.Rmd
Formatting of the document

@@
 bibliography: refs.bib
+geometry: margin=0.8in
 ---
@@
 # Introduction
+
 The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been a great development for the scientific community.
@@
 Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easy for an R-based analysis to be converted to an interactive web application.
-This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}.
+This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}, available at http://www.intro-stats.com.
 \texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.
@@
 ## Easy
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\linewidth]{images/user_experience.pdf}
+\caption{Typical user experience of generating a result in \texttt{intRo}. In this instance, a mosaic plot (and its corresponding code) was created and saved to the final compendium.}
+\label{fig:user_experience}
+\end{figure}
+
 In order for an application to be useful to students in their first statistics course, it can't hog the spotlight.
@@
 Additional functionality is nice, but if it clutters the UI beyond the point of usability, it will frustrate an introductory student.
+Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in \texttt{intRo}, and view the code producing the result.
 

### 2014-12-12 student-paper-competition-2014/paper.Rmd
Edits from reading through paper

@@
 \texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}.
-Within this paper we introduce the application and explore the design decisions underlying \texttt{intRo} as well as highlight some challenges and advantages of reactive programming.
+Within this paper, we introduce the application and explore the design decisions underlying \texttt{intRo}, as well as highlight some challenges and advantages of reactive programming.
 \end{abstract}
@@
 DataCamp's [@datacamp] courses are user-friendly ways to learning basic \texttt{R} programming and data analysis techniques.
-\texttt{Swirl} [@swirl] is a similar interactive tool to make learning \texttt{R} more fun by learning \texttt{R} within \texttt{R} itself.
+\texttt{Swirl} [@swirl] is a similar interactive tool to make learning \texttt{R} more fun by learning it within \texttt{R} itself.
 Project MOSAIC [@mosaic] has created a suite of tools to simplify the teaching of statistics in the form of an R package.
@@
 The first is a button to toggle the visibility of the code panel (5).
-The last is a button to print the active module results, and the associated code (if visible).
+The last is a button to print the stored module results, and the associated code (if visible).
  2. **Side Navigation** - The side navigation panel includes the modules that perform a specific data analysis task.
@@
 In addition to making \texttt{intRo} intuitive to use, we've also created a documentation website that is consistent with the interface of the application.
-The documentation is available at http://gammarama.github.io/intRo/ and hyperlinked within the application itself.
+The documentation is available at http://gammarama.github.io/intRo and hyperlinked within the application itself.
 This documentation covers all default modules and makes learning to use \texttt{intRo} even more painless.
@@
 In general, introduction to statistics is a class undergraduate students dread.
-It's hard, the language is unfamiliar, and the software is hard to use.
+It's hard, the language is unfamiliar, and the software is difficult to use.
 With \texttt{intRo}, we hope to flip this paradigm.
@@
 Each module is a self contained set of \texttt{R} code that is dynamically added to the application at run time.
-What this means is that \texttt{intRo} is easily extended by the addition of modules within the framework underlying the application.
+What this means is that \texttt{intRo} can be easily extended by the addition of modules within the framework underlying the application.
 This extensibility allows for \texttt{intRo} to be customized for an individual classroom setting, tailored to the needs of a particular course.
 
-As an example of possible customization, an instructor may wish to introduce hypothesis testing from a randomization framework rather than distributional.
-In this case, a non-parametric test could be added easily as a module to be used instead of the T test.
+As an example of possible customization, an instructor may wish to introduce hypothesis testing from a nonparametric framework rather than distributional.
+In this case, a wilcoxon rank sum test could be added easily as a module to be used instead of the T test.
 This addition would be fairly simple to complete.
@@
 \end{table*}
+
 Part two of this text discusses probability and sampling distributions, which is outside the current realm of \texttt{intRo}, but part three focuses on inferential statistics.
@@
 Finally, part four of the text discusses the association between variables, including regression, functions that \texttt{intRo} performs with ease.
-Some of the more advanced topics of this section are currently unsupported by \texttt{intRo}, including multiple regression and nonparametric statistics.
+Some of the more advanced topics of this section are currently unsupported by \texttt{intRo}, including multiple regression.
 
@@
 
-\texttt{intRo} leverages reactive programming techniques as data is manipulated by the user, this carries throughout the entire session.
+\texttt{intRo} leverages reactive programming techniques as data is manipulated by the user, and this is carried throughout the entire session.
 In a typical \texttt{Shiny} app, users interact with inputs, which in turn yield different results.
 Within \texttt{intRo}, the users are able to interact and manipulate the data underlying the entire application.
-This posed many challenges in the creation of \texttt{intRo}, namely timely save points according to the user's timeline, and reactive updating of variable lists tied to inputs across the entire application.
+This posed many challenges in the creation of \texttt{intRo}, namely timely save points according to the user's workflow, and reactive updating of variable lists tied to inputs across the entire application.
 Because the user may experiment with different configurations or select different variables, we did not want to store all actions taken in the intRo session.
@@
 For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}, as detailed in section \ref{sec:extensible}.
-However, \texttt{Shiny} does have limits in its flexibility based on the designers decisions for usability.
+However, \texttt{Shiny} does have limits on its flexibility based on the designers decisions for usability.
 One current example the slider element.

### 2014-12-12 student-paper-competition-2014/paper.Rmd
edits to paper

@@
 
-The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been a great development for the scientific community.
+The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly beenan important development for the scientific community.
 However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student.
@@
 
-In teaching STAT 226 at Iowa State University, we witnessed profound struggles by students attempting to practice the introductory concepts discussed in class.
-Scrimshaw notes in his manuscript that "open-ended packages, like any others, may create obstacles to learning simply through their lack of user-friendliness in the sheer mechanics of operating them, rather than any intrinsic difficulty in the content...." [@scrimshaw2001computers] In our own experience teaching, students' struggles were often directly related to the use of the software and not any sort of fundamental misunderstanding of the material, in agreement with Scrimshaw's finding.
+In teaching Introduction to Business Statistics at Iowa State University, we witnessed profound struggles by students attempting to practice the introductory concepts discussed in class using current software.
+Scrimshaw notes in his manuscript that "open-ended packages, like any others, may create obstacles to learning simply through their lack of user-friendliness in the sheer mechanics of operating them, rather than any intrinsic difficulty in the content...." [-@scrimshaw2001computers] In our own experience teaching, students' struggles were often directly related to the use of the software and not any sort of fundamental misunderstanding of the material, in agreement with Scrimshaw's finding.
 
-Other software has recently been written in an attempt to spur student interest in introductory statistics and R programming.
+Multiple software packages have recently been written in an attempt to spur student interest in introductory statistics and R programming.
 DataCamp's [@datacamp] courses are user-friendly ways to learning basic \texttt{R} programming and data analysis techniques.
@@
 
-Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easy for an R-based analysis to be converted to an interactive web application.
-This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}, available at http://www.intro-stats.com.
-\texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.
+Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easier for an R-based analysis to be converted to an interactive web application.
+This in turn led us to create an introductory statistics application which we call \texttt{intRo}, available at http://www.intro-stats.com.
+\texttt{intRo} offers a number of key advantages over traditional statistics software, including ease of use and an aim to foster student interest.
+Additionally, \texttt{intRo} stands apart from new tols in that it is a supplement to an existing class, fully usable by a beginning statistics student.
+These advantages will be discussed at length in this text.
 
@@
 Additional functionality is nice, but if it clutters the UI beyond the point of usability, it will frustrate an introductory student.
-Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in \texttt{intRo}, and view the code producing the result.
+Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in \texttt{intRo}, view the code producing the result, and save the results for later.
 
@@
 \end{figure*}
-
  1.
@@
 
-There are multiple modules within \texttt{intRo}, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
+There are seven default modules within \texttt{intRo}, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
 Each module is placed within a higher level category to further help students navigate the tool.
-The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can be easy added by an instructor, as detailed in section \ref{sec:extensible}.
+The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can easily be added by an instructor, as detailed in section \ref{sec:extensible}.
 The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.
@@
 In general, introduction to statistics is a class undergraduate students dread.
-It's hard, the language is unfamiliar, and the software is difficult to use.
+It's hard, the vocabulary is unfamiliar, and the software is difficult to use.
 With \texttt{intRo}, we hope to flip this paradigm.
@@
 The process by which \texttt{intRo} is dynamically created upon running the application through module sourcing is detailed in figure \ref{fig:app_creation_modules}.
-The key driver to populating \texttt{server.R} and \texttt{ui.R} is the file \texttt{modules.txt}, which defines the placement of each module.
+The key driver to populating \texttt{server.R} and \texttt{ui.R} is the navigation file \texttt{modules.txt}, which defines the placement of each module.
 The interface is then created with the following statement.
@@
 
-# Reactive programming
+# Reactive design choices
 
@@
 
-\texttt{intRo} leverages reactive programming techniques as data is manipulated by the user, and this is carried throughout the entire session.
-In a typical \texttt{Shiny} app, users interact with inputs, which in turn yield different results.
-Within \texttt{intRo}, the users are able to interact and manipulate the data underlying the entire application.
-This posed many challenges in the creation of \texttt{intRo}, namely timely save points according to the user's workflow, and reactive updating of variable lists tied to inputs across the entire application.
+\texttt{intRo} leverages the reactive programming nature of \texttt{Shiny}, and as such is designed around the idea of user input cascading through the entire application.
+In a typical \texttt{Shiny} application, users interact with inputs that act as parameters to function, which in turn yield different results.
+Within \texttt{intRo}, the users are able to interact with and manipulate the data underlying the entire application.
+This posed many challenges in the creation of \texttt{intRo} and drove design decisions, namely timely save points according to the user's workflow, and reactive updating of variable lists tied to inputs across the entire application.
 Because the user may experiment with different configurations or select different variables, we did not want to store all actions taken in the intRo session.
@@
 However, \texttt{Shiny} does have limits on its flexibility based on the designers decisions for usability.
-One current example the slider element.
+One current example is the slider element.
 This element allows for fixed width steps from its minimum to its maximum.
@@
 In particular, we have begun development on an \texttt{R} package which will allow \texttt{intRo} modules to be created automatically from user written \texttt{R} code.
-This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo} as well as translate user code to \texttt{intRo} compatible code as populates the necessary files.
+This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo} as well as translate user code to \texttt{intRo} compatible code and populate the necessary files.
 This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
-Furthermore, we would like to utilize the expanded interactive capabilities of \texttt{ggvis} in order to make \texttt{intRo's} plots more engaging to students.
+Additionally, we would like to utilize the expanded interactive capabilities of \texttt{ggvis} in order to make \texttt{intRo's} plots more engaging to students.
 One way to do this would be implementing linked plots, in which interactions with one plot are reflected in other plots that illustrate the same data.
@@
 
-Additionally,  \texttt{intRo} will be trialed in an introductory statistics course at Iowa State University in Spring 2015.
+\texttt{intRo} will be trialed in an introductory statistics course at Iowa State University in Spring 2015.
 This will allow us to assess its usefulness relative to software used in the past, as well as guage areas for improvement.

### 2014-12-16 hmwkch4.tex
first 6 homeworks plus datasets

\documentclass{article}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{epsfig, amsmath, amssymb}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\begin{document}

\hspace{2.5in} Name: (Print) \underline{\mbox{\hspace{3in}}}\\

\begin{center}
{\large STATISTICS 201 - Written Homework Chapter 4}\\[3mm]
\end{center}

\begin{enumerate}

\item One hundred middle-aged men and one hundred young adult men were rated on physical flexibility.  The scale went from 1, least flexible to 10, most flexible.  The table below shows a summary of these ratings.  For example, only 1 Young Adult Man had a flexibility rating of 1 while 2 Middle-Aged Men received this rating.

\begin{center}
\begin{tabular}{l|cccccccccc} \hline
Flexibility Rating & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10\\ \hline
Young Adult Men & 1 & 3 & 6 & 8 & 11 & 14 & 16 & 19 & 14 & 8\\
Middle-Aged Men & 2 & 6 & 10 & 14 & 17 & 18 & 14 & 11 & 6 & 2\\ \hline
\end{tabular}
\end{center}

Calculate the 5-number summary for the flexibility ratings for Young Adult Men and for Middle-Aged Men.  Make sure to clearly label the values and the group they belong to.

\newpage

\item A study examining the health risks of smoking measured the cholesterol levels of people who had smoked for at least 25 years and people of similar ages who had smoked for no more than 5 years and then stopped.  The cholesterol levels and smoking status of all 76 people can be found in the JMP data file {\bf Cholesterol.JMP}  in Blackboard.

\begin{enumerate}
\item Describe the Who for this data set.\\[0.5in]
\item Describe the What for this data set.\\[0.5in]
\item Use JMP to obtain the distribution of cholesterol levels for both groups (Smokers and Non-Smokers).  Print out this output and turn it in with your assignment.\\[1em]
\item Use the means and medians of the two distributions to discuss similarities and/or differences in the centers of the two distributions.\\[1in]
\item Use the ranges, IQRs, and standard deviations of the two distributions to discuss similarities and/or differences in the variability of the two distributions.\\[1in]
\item Doctors want patients to maintain a cholesterol level less than 200 points.  What percent of people in the Smokers group reached this goal?  What percent of people in the Ex-Smokers group reached this goal?  Why would it be important to report this information in the analysis of these data?\\[0.5em]

Percent in Smokers group less than 200 = \underline{\mbox{\hspace{1.5in}}}\\[1.5em]
Percent in Ex-Smokers group less than 200 = \underline{\mbox{\hspace{1.5in}}}\\[1em]

Why important to report?:
\end{enumerate}

\end{enumerate}

\end{document}

### 2014-12-16 hmwkch5.tex
first 6 homeworks plus datasets

\documentclass{article}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\begin{document}
\hspace{2in} Name: (Print) \underline{\mbox{\hspace{3.5in}}}\\

\begin{center}
{\large STATISTICS 201 - Written Homework Chapter 5}\\[3mm]
\end{center}

\begin{enumerate}

\item The GRE is a test to measure undergraduate achievement and is used as an admissions criterion to graduate school. The GRE is designed in such a way that scores on the test can be modeled with a normal model. From July 2006 to June 2007 the mean GRE Verbal score was 457 with a standard deviation of 121. Use the z-table to answer the following questions.

\begin{enumerate}
\item What proportion of all test takers will have a GRE Verbal score above 600 points?\\[1in]
\item What proportion of all test takers will have a GRE Verbal score below 375 points?\\[1in]
\item What proportion of all test takers will have a GRE Verbal score between 500 and 700 points?\\[1in]
\item What GRE Verbal score will 40\% of all test takers score below?\\[1in]
\item What GRE verbal score will 10\% of all test takers score above?
\end{enumerate}

\newpage

\item In a study to determine the nature of the relationship between heart rates and body temperature, body temperatures were collected from 130 individuals. The body temperatures, in degrees Fahrenheit, can be found in the JMP data file {\bf samplebodytemps.JMP} in Blackboard.

Use JMP to determine if the distribution of body temperatures can be modeled with a normal model. Make sure to turn in the output from JMP with your assignment. In your answer, make sure to use information from the JMP output that supports your conclusions.

\end{enumerate}

\end{document}

### 2014-12-16 hmwkch6.tex
first 6 homeworks plus datasets

\documentclass{article}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\begin{document}
\hspace{2in} Name: \underline{\mbox{\hspace{3.5in}}}\\

\begin{center}
{\large STATISTICS 201 - Written Homework Chapter 6}\\[3mm]
\end{center}

\begin{enumerate}
\item For the Somerset neighborhood 2009 Ames sales data tabulate the correlations between the SalePrice, LotArea, LivingArea and GarageArea. (See the textbook for an example of how to do this. Be sure to round numbers appropriately.)

\vspace{2.5in}

\item Which two variables have the strongest correlation? What is it?

\vspace{1in}

\item Make scatterplots of each of the pairs of variables (don't hand these in). Write a paragraph briefly describing the associations, and discussing whether correlation is  or isn't an appropriate summary.

\end{enumerate}

\end{document}

### 2014-12-16 global.R
Revamp the numeric module

@@ -10,6 +10,8 @@ library(R.utils)
 library(RCurl)
 library(rmarkdown)
 library(formatR)
+library(psych)
+library(Hmisc)
 
 #cat("```{r, echo=FALSE}\nopts_chunk$set(echo=FALSE)\n```\n\n```{r, child="test.Rmd", message=FALSE, warning=FALSE}\n```", file = "outfile.Rmd")
 

### 2014-12-16 modules/numerical/helper.R

@@ -1,9 +1,16 @@
-    generateSummary <- function(intro.data, tblvars, grouping, save_result) {
+    generateSummary <- function(intro.data, tblvars, grouping) {
         if (!is.null(tblvars)) {
             if (grouping == "none") {
-                cat_and_eval(paste0("summary(intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")])"),  mydir = userdir, env = environment(), file = "code_numerical.R", append = FALSE)
+                if (all(tblvars %in% categoricNames(intro.data))) {
+                    cat_and_eval(paste0("Hmisc::describe(intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")])"),  mydir = userdir, env = environment(), file = "code_numerical.R", append = FALSE)
+                } else {
+                    cat_and_eval(paste0("my.summary <- as.data.frame(psych::describe(intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")], skew = FALSE))[,c('n', 'mean', 'sd', 'min', 'median', 'max', 'range')]"), mydir = userdir, env = environment(), file = "code_numerical.R")
+                    cat_and_eval("my.summary[grep('\\\\*', rownames(my.summary)), c('mean', 'sd', 'min', 'median', 'max', 'range')] <- NA",  mydir = userdir, env = environment(), file = "code_numerical.R", append = TRUE)
+    
+                    cat_and_eval("my.summary",  mydir = userdir, env = environment(), file = "code_numerical.R", append = TRUE)
+                }
             } else {
-                cat_and_eval(paste0("dat <- intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")]; dat$intro_grouping <- intro.data[,'", grouping, "']; dat.dplyr <- dat[,names(dat) %in% c(\"intro_grouping\", numericNames(dat))] %>% group_by(intro_grouping) %>% summarise_each(funs(min, q1, median, q3, max, mean, sd)); dat.dplyr.df <- as.data.frame(dat.dplyr); names(dat.dplyr.df)[1] <- '", grouping, "'; dat.dplyr.df"),  mydir = userdir, env = environment(), file = "code_numerical.R")
[…]

### 2014-12-16 modules/numerical/observe.R

@@ -1,6 +1,9 @@
     observe({
         updateSelectizeInput(session, "grouping", choices = c("None" = "none", names(intro.data())))
-        updateSelectizeInput(session, "tblvars", choices = names(intro.data()))
+    })
+    
+    observe({
+        updateSelectizeInput(session, "tblvars", choices = (if (input$grouping == "none") names(intro.data()) else intro.numericnames()))
     })
     
     observeEvent(input$store_numerical, {

### 2014-12-16 modules/numerical/output.R

@@ -1,3 +1,3 @@
     output$summary <- renderPrint({
-        return(generateSummary(intro.data(), input$tblvars, input$grouping, input$store_numerical > values$store_value))
+        return(generateSummary(intro.data(), input$tblvars, input$grouping))
     })

### 2015-01-06 modules/contingency/observe.R
Get rid of files

@@ -1,8 +1,0 @@
-    observe({
-        updateSelectizeInput(session, "xcont", choices = intro.categoricnames(), selected = ifelse(checkVariable(intro.data(), input$xcont), input$xcont, intro.categoricnames()[1]))
-        updateSelectizeInput(session, "ycont", choices = intro.categoricnames(), selected = ifelse(checkVariable(intro.data(), input$ycont), input$ycont, intro.categoricnames()[2]))
-    })
-    
-    observeEvent(input$store_contingency, {
-        cat(paste0("\n\n", paste(readLines(file.path(userdir, "code_contingency.R")), collapse = "\n")), file = file.path(userdir, "code_All.R"), append = TRUE)
-    })

### 2015-01-06 modules/contingency/output.R

@@ -1,3 +1,0 @@
-    output$conttable <- renderTable({
-        return(cont.table(intro.data(), input$xcont, input$ycont, input$conttype, input$contdigits))
-    })

### 2015-01-06 modules/contingency/ui.R

@@ -1,23 +1,0 @@
-contingency_ui <- tabPanel("Contingency",
-                                                  column(4,
-                                                         wellPanel(
-                                                             selectizeInput("xcont", label = "X Variable (x)", choices = categoricNames(mpg), selected = categoricNames(mpg)[5]),
-                                                             selectizeInput("ycont", label = "Y Variable (y)", choices = categoricNames(mpg), selected = categoricNames(mpg)[6]),
-                                                             
-                                                             hr(),
-                                                             
-                                                             selectizeInput("conttype", label = "Type", choices = c("Counts" = "counts", "Row Proportions" = "rowpercs", "Column Proportions" = "columnpercs", "Total Proportions" = "totalpercs")),
-                                                             conditionalPanel(condition = "input.conttype != 'counts'",
-                                                                numericInput("contdigits", label = "Digits", value = 2, min = 1, step = 1)
-                                                             ),
-                                                             
-                                                             hr(),
-                                                             
-                                                             tags$button("", id = "store_contingency", type = "button", class = "btn action-button", list(icon("save"), "Store Contingency Result"), onclick = "$('a:has(> .fa-print, .fa-code)').highlight();")
-                                                         )
-                                                  ),
-                                                  
-                                                  column(8,
-                                                         tableOutput("conttable")
-                                                  )
-                                         )

### 2015-01-06 modules/graphical/observe.R

@@ -1,22 +1,0 @@
-observe({
-    updateSelectizeInput(session, "plottype", choices = if (intro.numericnames()[1] == "") c("Mosaic Plot" = "mosaicplot") else if (intro.categoricnames()[1] == "") c("Histogram" = "histogram", "Normal Quantile Plot" = "quantileplot", "Scatterplot" = "scatterplot", "Line Chart" = "linechart") else c("Histogram" = "histogram", "Normal Quantile Plot" = "quantileplot", "Scatterplot" = "scatterplot", "Line Chart" = "linechart", "Boxplot" = "boxplot", "Bar Chart" = "barchart", "Pareto Chart" = "paretochart", "Mosaic Plot" = "mosaicplot"))
-})
-
-observe({
-    input$plottype
-    updateSelectizeInput(session, "x", choices = x_choices(), selected = x_selected())
-    updateSelectizeInput(session, "y", choices = y_choices(), selected = y_selected())
-})
-
-observeEvent(input$store_graphical, {
-    cat(paste0("\n\n", paste(readLines(file.path(userdir, "code_graphical_reactive.R")), collapse = "\n")), file = file.path(userdir, "code_All.R"), append = TRUE)
-    
-    ## Fix this
-    cat(paste0("\n\ninput_xdomain <- c(", input$xmin, ", ", input$xmax, ")\n"), file = file.path(userdir, "code_All.R"), append = TRUE)
-    cat(paste0("input_ydomain <- c(", input$ymin, ", ", input$ymax, ")\n"), file = file.path(userdir, "code_All.R"), append = TRUE)
-    cat(paste0("input_binwidth <- ", ifelse(is.na(input$binwidth), "NULL", input$binwidth), "\n"), file = file.path(userdir, "code_All.R"), append = TRUE)
-    
-    cat("\n\n", file = file.path(userdir, paste0("code_", input$plottype, ".R")), append = TRUE)
-    cat(paste0("\n\n", paste(readLines(file.path(userdir, paste0("code_", input$plottype, ".R"))), collapse = "\n")), file = file.path(userdir, "code_All.R"), append = TRUE)
-    if (input$plottype != "mosaicplot") cat(paste0("p.", input$plottype), file = file.path(userdir, "code_All.R"), append = TRUE)
-})

### 2015-01-06 modules/graphical/output.R

@@ -1,58 +1,0 @@
-    cat_and_eval("p.histogram <- intro.plot %>%
-        ggvis(x = ~intro_x_num, y = ~intro_y_num) %>%
-        layer_histograms(width = input_binwidth) %>% 
-        scale_numeric('x', domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
-        scale_numeric('y', domain = input_ydomain, clamp = TRUE, nice = TRUE)",  mydir = userdir, env = environment(), file = "code_histogram.R")
-    
-    p.histogram %>% bind_shiny("histogram")
-
-    cat_and_eval("p.quantileplot <- intro.quant %>%
-        ggvis(x = ~intro_quant, y = ~intro_x_num, key := ~id) %>%
-        layer_points() %>% 
-        scale_numeric('x', domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
-        scale_numeric('y', domain = input_ydomain, clamp = TRUE, nice = TRUE) %>%
-        add_tooltip(function(df) {paste0('row id: ', df$id)}, 'hover')",  mydir = userdir, env = environment(), file = "code_quantileplot.R")
-    
-    p.quantileplot %>% bind_shiny("quantileplot")
-    
-    cat_and_eval("p.scatterplot <- intro.plot %>%
-        ggvis(x = ~intro_x_num, y = ~intro_y_num, key := ~id) %>%
-        layer_points() %>% 
-        scale_numeric('x', domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
-        scale_numeric('y', domain = input_ydomain, clamp = TRUE, nice = TRUE) %>%
-        add_tooltip(function(df) {paste0('row id: ', df$id)}, 'hover')",  mydir = userdir, env = environment(), file = "code_scatterplot.R")
-    
-    p.scatterplot %>% bind_shiny("scatterplot")
-    
-    cat_and_eval("p.linechart <- intro.plot %>%
-        ggvis(x = ~intro_x_num, y = ~intro_y_num) %>%
-        layer_lines() %>% 
-        scale_numeric('x', domain = input_xdomain, clamp = TRUE, nice = TRUE) %>%
-        scale_numeric('y', domain = input_ydomain, clamp = TRUE, nice = TRUE)",  mydir = userdir, env = environment(), file = "code_linechart.R")
-    
-    p.linechart %>% bind_shiny("linechart")
-    
[…]

### 2015-01-06 modules/graphical/reactive.R

@@ -1,41 +1,0 @@
-x_choices <- reactive({
-    if (input$plottype %in% c("boxplot", "barchart", "paretochart", "mosaicplot")) intro.categoricnames()
-    else intro.numericnames()
-})
-    
-y_choices <- reactive({
-    if (input$plottype %in% c("mosaicplot")) intro.categoricnames()
-    else intro.numericnames()
-})
-
-x_selected <- reactive({
-    x_choices()[1]
-})
-
-y_selected <- reactive({
-    y_choices()[2]
-})
-
-    intro.plot <- reactive({
-            if (is.null(intro.data())) return(NULL)
-    
-            return(intro.plotdata(intro.data(), input$x, input$y, input$plottype))
-    })
-    
-    intro.quant <- reactive({
-        if (is.null(intro.plot())) return(NULL)
-        
-        return(intro.quantdata(intro.plot()))
-    })
-    
-    input_binwidth <- reactive({        
-        return(input_binwidthdata(input$binwidth))
-    })
-    
-    input_xdomain <- reactive({        
-        return(input_xdomaindata(input$xmin, input$xmax))
-    })
-    
-    input_ydomain <- reactive({        
-        return(input_ydomaindata(input$ymin, input$ymax))
-    })

### 2015-01-06 modules/nonparametric/helper.R

@@ -1,3 +1,0 @@
-nonparametrictable <- function (intro.data, x, y, conflevel, althyp, hypval) {
-    cat_and_eval(paste0("wilcox.test(x=intro.data[,'", x, "'], y=intro.data[,'", y, "'], conf.level=", conflevel, ", alternative='", althyp, "', mu=", hypval, ")"),  mydir = userdir, env = environment(), file = "code_nonparametric.R")
-}

### 2015-01-06 modules/nonparametric/output.R

@@ -1,4 +1,0 @@
-    output$nonparametrictable <- renderText({
-        return(paste(capture.output(nonparametrictable(intro.data(), input$group1_non, input$group2_non, input$conflevel_non, input$althyp_non, input$hypval_non)), collapse = "\n"))
-    })
-

### 2015-01-06 modules/numerical/helper.R

@@ -1,13 +1,0 @@
-    generateSummary <- function(intro.data, tblvars, grouping, categoric_names) {
-        if (!is.null(tblvars)) {
-            if (grouping == "none") {
-                if (all(tblvars %in% categoric_names)) {
-                    cat_and_eval(paste0("describe(intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")])"), mydir = userdir, env = environment(), file = "code_numerical.R")
-                } else {
-                    cat_and_eval(paste0("my.summary(intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")])"), mydir = userdir, env = environment(), file = "code_numerical.R")
-                }
-            } else {
-                cat_and_eval(paste0("dat <- intro.data[,c(", paste(paste0("'", tblvars, "'"), collapse = ", "), ")]; dat$intro_grouping <- intro.data[,'", grouping, "']; dat.dplyr <- dat[,names(dat) %in% c(\"intro_grouping\", numericNames(dat))] %>% group_by(intro_grouping) %>% summarise_each(funs(min, q1, median, q3, max, mean, sd)); dat.dplyr.df <- as.data.frame(dat.dplyr); names(dat.dplyr.df)[1] <- '", grouping, "'; dat.dplyr.df"), mydir = userdir, env = environment(), file = "code_numerical.R")
-            }
-        }
-    }

### 2015-01-06 modules/numerical/output.R

@@ -1,3 +1,0 @@
-    output$summary <- renderPrint({
-        return(generateSummary(intro.data(), input$tblvars, input$grouping, intro.categoricnames()))
-    })

### 2015-01-06 modules/numerical/ui.R

@@ -1,16 +1,0 @@
-numerical_ui <- tabPanel("Numerical",
-                                                  column(4,
-                                                         wellPanel(
-                                                             selectizeInput("tblvars", label = "Select Variables", choices = names(mpg), multiple = TRUE),
-                                                             selectizeInput("grouping", label = "Select Grouping Variable", choices = names(mpg)),
-                                                             
-                                                             hr(),
-                                                             
-                                                             tags$button("", id = "store_numerical", type = "button", class = "btn action-button", list(icon("save"), "Store Numerical Result"), onclick = "$('a:has(> .fa-print, .fa-code)').highlight();")
-                                                         )
-                                                  ),
-                                                  
-                                                  column(8,
-                                                         verbatimTextOutput("summary")
-                                                  )
-                                         )

### 2015-01-06 modules/regression/observe.R

@@ -1,15 +1,0 @@
-## Regression Observers
-observe({
-  updateSelectizeInput(session, "xreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$xreg), input$xreg, intro.numericnames()[1]))
-  updateSelectizeInput(session, "yreg", choices = intro.numericnames(),  selected = ifelse(checkVariable(intro.data(), input$yreg), input$yreg, intro.numericnames()[1]))
-})
-
-observeEvent(input$store_regression, {
-    cat(paste0("\n\n", paste(readLines(file.path(userdir, "code_regression_reactives.R")), collapse = "\n")), file = file.path(userdir, "code_All.R"), append = TRUE)
-    cat(paste0("\n\n", paste(readLines(file.path(userdir, "code_regression.R")), collapse = "\n")), file = file.path(userdir, "code_All.R"), append = TRUE)
-    cat(paste0("\n\n", paste(readLines(file.path(userdir, "code_regression_ggvis.R")), collapse = "\n")), file = file.path(userdir, "code_All.R"), append = TRUE)
-})
-
-observeEvent(input$saveresid, {
-    values$mydat <<- savefit(intro.data(), intro.regression())
-})

### 2015-01-06 modules/regression/output.R

@@ -1,51 +1,0 @@
-cat_and_eval("p.regplot <- reg.data %>%
-  ggvis(x = ~xreg, y = ~yreg, key := ~test) %>%
-  layer_points() %>%
-  add_tooltip(function(df) {paste0('row id: ', df$test)}, 'hover')",  mydir = userdir, env = environment(), file = "code_regression_ggvis.R")
-
-p.regplot %>% bind_shiny("regplot")
-cat("p.regplot\n\n", file = file.path(userdir, "code_regression_ggvis.R"), append = TRUE)
-
-cat_and_eval("p.resplot1 <- reg.resid1 %>%
-  ggvis(x = ~x, y = ~residuals, key := ~id) %>%
-  layer_points() %>%
-  set_options(width = 200, height = 200) %>%
-  add_tooltip(function(df) {paste0('row id: ', df$id)}, 'hover')",  mydir = userdir, env = environment(), file = "code_regression_ggvis.R", append = TRUE)
-
-p.resplot1 %>% bind_shiny("resplot1")
-cat("p.resplot1\n\n", file = file.path(userdir, "code_regression_ggvis.R"), append = TRUE)
-
-cat_and_eval("p.resplot2 <- reg.resid2 %>%
-  ggvis(x = ~yy, y = ~residuals, key := ~id) %>%
-  layer_points() %>%
-  set_options(width = 200, height = 200) %>%
-  add_tooltip(function(df) {paste0('row id: ', df$id)}, 'hover')",  mydir = userdir, env = environment(), file = "code_regression_ggvis.R", append = TRUE)
-
-p.resplot2 %>% bind_shiny("resplot2")
-cat("p.resplot2\n\n", file = file.path(userdir, "code_regression_ggvis.R"), append = TRUE)
-
-cat_and_eval("p.resplot3 <- reg.resid1 %>%
-  ggvis(x = ~residuals) %>%
-  layer_histograms() %>%
-  set_options(width = 200, height = 200)",  mydir = userdir, env = environment(), file = "code_regression_ggvis.R", append = TRUE)
-
-p.resplot3 %>% bind_shiny("resplot3")
-cat("p.resplot3\n\n", file = file.path(userdir, "code_regression_ggvis.R"), append = TRUE)
-
-output$regtable <- renderTable({
-  if (!(input$xreg %in% intro.numericnames()) | !(input$yreg %in% intro.numericnames())) return(NULL)
-  
[…]

### 2015-01-06 modules/regression/ui.R

@@ -1,49 +1,0 @@
-regression_ui <- tabPanel("Regression",
-         column(4,
-                wellPanel(
-                  selectizeInput("xreg", label = "Independent Variable (x)", choices = numericNames(mpg), selected = numericNames(mpg)[4]),
-                  selectizeInput("yreg", label = "Dependent Variable (y)", choices = numericNames(mpg), selected = numericNames(mpg)[5]),
-                  
-                  hr(),
-                  
-                  tags$button("", id = "saveresid", type = "button", class = "btn action-button", list(icon("save"), "Save Residuals/Fitted"), onclick = "$('#side-nav :contains(\"Sources\")').highlight();"),
-
-                  hr(),
-                  
-                  tags$button("", id = "store_regression", type = "button", class = "btn action-button", list(icon("save"), "Store Regression Result"), onclick = "$('a:has(> .fa-print, .fa-code)').highlight();")
-                )
-         ),
-         
-         column(8,
-                conditionalPanel(condition = "input.xreg != ''", 
-                                 tags$b("Parameter Estimates"),
-                                 tableOutput("regtable"),
-                                 
-                                 hr(),
-                                 
-                                 tags$b("Correlation"),
-                                 textOutput("r"),
-                                 textOutput("r2"),
-                                 
-                                 hr(),
-                                 
-                                 tags$b("Plot of Fit"),
-                                 ggvisOutput("regplot"),
-                                 
-                                 hr(),
-                                 
-                                 tags$b("Residual Plots"),
-                                 fluidRow(
-                                     column(4, 
-                                            ggvisOutput("resplot1")
-                                     ),
-                                     column(4,
-                                            ggvisOutput("resplot2")
-                                     ),
-                                     column(4,
-                                            ggvisOutput("resplot3")
-                                     )
-                                 )           
-                )
-         )
-)

### 2015-01-06 modules/sources/helper.R

@@ -1,51 +1,0 @@
-process_logical <- function(data, x) {
-    if (is.null(x) || x == "c('')") {
-        return(data)
-    }
-    
-    relevant_cols <- names(data)[nchar(x) > 0]
-    if (length(relevant_cols) == 0) return(data)
-    
-    my_strs <- strsplit(gsub(",", " , ", x), split = ",")
-    new_strs <- my_strs[unlist(lapply(my_strs, length)) > 0]
-    
-    new_data <- data
-    if (length(new_strs) > 0) {
-        for (i in 1:length(new_strs)) {
-            test <- new_strs[[i]]
-            col <- relevant_cols[i]
-            
-            subset_str <- ""
-            if (length(test) == 1) {
-                if (is.na(as.numeric(test[1]))) {
-                    subset_str <- paste0("'", test[1], "' == ", col)
-                } else {
-                    subset_str <- paste(test[1], "==", col)
-                }
-            } else {
-                test[1] <- ifelse(test[1] == " ", -Inf, test[1])
-                test[2] <- ifelse(test[2] == " ", Inf, test[2])
-                if (is.na(test[2])) test[2] <- Inf
-                
-                subset_str <- paste(test[1], "<=", col, "&", col, "<=", test[2])
-            }          
-            
-            new_data <- eval(parse(text = paste0("subset(new_data, ", subset_str, ")")))
-        }
-    }
-    
-    return(new_data)
-}
-
-file.choose <- function () {
-    return(input$data_own[,"datapath"])
-}
-
-data.module <- function (inFile, dataset, own) {        
-    if (is.null(inFile) | !own) {
-        cat_and_eval(paste0("\n\nintro.data <- get('", dataset, "')"), file = "code_sources.R", mydir = userdir, append = FALSE, save_result = TRUE)
-    } else {
-        cat_and_eval(paste0("intro.data <- read.csv(file.choose())"), mydir = userdir, file = "code_sources.R", append = FALSE, save_result = TRUE)
-    }
-    return(intro.data)
-}

### 2015-01-06 modules/sources/observe.R

@@ -1,7 +1,0 @@
-observe({
-    updateNumericInput(session, "randomsubrows", max = nrow(intro.data()))
-})
-
-observeEvent(input$clearsubset, {
-    updateCheckboxInput(session, "randomsub", value = FALSE)
-})

### 2015-01-06 modules/sources/reactive.R

@@ -1,66 +1,0 @@
-intro.start <- reactive({
-    input$clearsubset
-    
-    data.initial <- data.module(input$data_own, input$data, input$own)
-    values$mydat <<- data.initial
-                
-    return(data.initial)
-})
-
-intro.data <- reactive({
-    if (is.null(intro.start())) return(NULL)
-        
-    intro.data <- values$mydat
-    intro.random <- values$mydat_rand
-    
-    cat_and_eval(paste0("\n\ndata.subset <- process_logical(intro.data, ", paste0("c(", paste(paste0("'", input$subs, "'"), collapse = ", "), ")"), ")"),  mydir = userdir, file = "code_sources.R")
-    cat_and_eval(paste0("intro.random <- data.subset"),  mydir = userdir, file = "code_sources.R", append = TRUE)
-        
-    values$mydat <<- data.subset
-    values$mydat_rand <<- values$mydat
-    
-    if (input$randomsub & is.numeric(input$randomsubrows) & input$randomsubrows >= 1 & input$randomsubrows <= nrow(values$mydat)) {
-        cat_and_eval(paste0("intro.random <- dplyr::sample_n(data.subset, ", input$randomsubrows, ")"),  mydir = userdir, file = "code_sources.R", append = TRUE)
-    }
-    
-    if (input$savesubset > oldsavesub) {                
-        cat_and_eval(paste0("intro.data <- intro.random"),  mydir = userdir, file = "code_sources.R", append = TRUE)
-        cat(paste(readLines(file.path(userdir, "code_sources.R")), collapse = "\n"), file = file.path(userdir, "code_All.R"), append = TRUE)
-        
-        values$mydat <<- values$mydat_rand
-        oldsavesub <<- input$savesubset
-    }
-    
-    return(intro.data)
-})
-
-intro.numericnames <- reactive({
-    if (is.null(intro.data())) return(NULL)
-    
-    return(numericNames(intro.data()))
-})
-
-intro.categoricnames <- reactive({
-    if (is.null(intro.data())) return(NULL)
-    
-    return(categoricNames(intro.data()))
-}) 
-
-dt.options <- reactive({list(pageLength = 10,
-                             searching=0,
-                             destroy=1,
-                             headerCallback =  I(paste0("function(thead, data, start, end, display) {
-                                                        //color code the header items
-                                                        var col_types = [", 
[…]

### 2015-01-06 modules/t_test/helper.R

@@ -1,7 +1,0 @@
-ttesttable <- function (intro.data, x, y = NULL, twovar = FALSE, conflevel, althyp, hypval) {
-    if (is.null(y) | y == "" | !twovar) {
-        cat_and_eval(paste0("t.test(x=intro.data[,'", x, "'], conf.level=", conflevel, ", alternative='", althyp, "', mu=", hypval, ")"),  mydir = userdir, env = environment(), file = "code_t_test.R")
-    } else {
-        cat_and_eval(paste0("t.test(x=intro.data[,'", x, "'], y=intro.data[,'", y, "'], conf.level=", conflevel, ", alternative='", althyp, "', mu=", hypval, ")"),  mydir = userdir, env = environment(), file = "code_t_test.R")
-    }
-}

### 2015-01-06 modules/transform/observe.R

@@ -1,8 +1,0 @@
-    ## Transform Observers
-    observe({
-        updateSelectizeInput(session, "var_trans", choices = (if (input$trans %in% c("power", "categorical")) intro.numericnames() else intro.categoricnames()))
-    })
-    
-    observe({
-        updateSelectizeInput(session, "trans", choices = (if (numericNames(intro.data())[1] == "") c("Numeric" = "numeric") else if (categoricNames(intro.data())[1] == "") c("Power" = "power", "Categorical" = "categorical") else c("Power" = "power", "Categorical" = "categorical", "Numeric" = "numeric")))
-    })

### 2015-01-06 modules/transform/output.R

@@ -1,17 +1,0 @@
-    output$var_trans_text <- renderText({
-        if (is.null(intro.transform())) return(NULL)
-        
-        return(paste("Selected Variable:", input$var_trans))
-    })
-
-    intro.transform %>%
-        ggvis(x = ~var) %>%
-        layer_histograms() %>%
-        set_options(width = 600, height = 250) %>%
-        bind_shiny("var_plot")
-
-    intro.transform %>%
-        ggvis(x = ~trans_var) %>%
-        layer_histograms() %>%
-        set_options(width = 600, height = 250) %>%
-        bind_shiny("trans_plot")

### 2015-01-06 modules/transform/reactive.R

@@ -1,48 +1,0 @@
-    intro.transform.colname <- reactive({
-        if (is.null(intro.data())) return(NULL)
-        
-        curdata <- intro.data()
-        curtrans <- input$var_trans
-        colname <- ""
-        
-        if (input$trans == "power") {
-            colname <- paste(curtrans, sub("\\.", "", input$power), sep = "_")
-        } else if (input$trans %in% c("categorical", "numeric")) {
-            colname <- paste(curtrans, "trans", sep = "_")
-        }
-        
-        return(colname)
-    })
-
-    intro.transform <- reactive({
-        if (is.null(intro.data())) return(NULL)
-                        
-        intro.data <- intro.data()
-        
-        cat_and_eval(paste0("curtrans <- '", input$var_trans, "'"),  mydir = userdir, env = environment(), file = "code_transform.R")
-                
-        if (input$trans == "power" & curtrans %in% intro.numericnames()) {
-            if (!is.numeric(input$power) | is.null(input$power)) cat_and_eval(paste0("trans_x <- intro.data[,'", curtrans, "']"),  mydir = userdir, file = "code_transform.R", append = TRUE) else if (input$power == 0) cat_and_eval(paste0("trans_x <- log(intro.data[,'", curtrans, "'])"),  mydir = userdir, file = "code_transform.R", append = TRUE) else cat_and_eval(paste0("trans_x <- (intro.data[,'", curtrans, "'])^", input$power),  mydir = userdir, file = "code_transform.R", append = TRUE)
-            if (all(!is.infinite(trans_x))) {
-                cat_and_eval(paste0("intro.data[, '", intro.transform.colname(), "'] <- trans_x"),  mydir = userdir, file = "code_transform.R", append = TRUE)
-            }
-        } else if (input$trans %in% c("categorical", "numeric")) {
-            if (input$trans == "numeric") cat_and_eval(paste0("trans_x <- as.numeric(intro.data[,'", curtrans, "'])"),  mydir = userdir, file = "code_transform.R", append = TRUE) else cat_and_eval(paste0("trans_x <- factor(intro.data[,'", curtrans, "'])"), mydir = userdir, file = "code_transform.R", append = TRUE)
-            cat_and_eval(paste0("intro.data[, '", intro.transform.colname(), "'] <- trans_x"),  mydir = userdir, file = "code_transform.R", append = TRUE)
-        } else {
[…]

### 2015-01-06 scripts/print.js

@@ -1,29 +1,0 @@
-var content;
-
-function code_clicked() {
-    Shiny.onInputChange("code_clicked", $('a:has(> .fa-code)').parent().hasClass("active"));
-}
-
-function print_clicked() {
-    code_clicked();
-    Shiny.onInputChange("print_clicked", true);
-}
-
-Shiny.addCustomMessageHandler("renderFinished",
-  function(file) {
-    var frame = $('.print_results')
-    //frame.contents().find("html").html(file);
-    frame.attr('src', 'code_All.html');
-
-    content = frame.contents().find("body").html();
-    check_print(frame)
-  })
-
-function check_print(frame) {
-  if (frame.contents().find("body").html() == content) {
-    setTimeout(check_print, 500, frame);
-  } else {
-    $('.print_results').get(0).contentWindow.print();
-    Shiny.onInputChange("print_clicked", false);
-  }    
-}

### 2015-01-06 server.R

@@ -1,64 +1,0 @@
-###
-### Shiny Server definition
-###
-shinyServer(function(input, output, session) {
-    
-    ## Module info
-    module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
-    
-    ## Get directory ready for code printing
-    userdir <- file.path(tempdir(), tempfile())
-    dir.create(userdir, recursive = TRUE)
-    sapply(file.path(userdir, dir(userdir)[grep("code_", dir(userdir))]), file.remove)
-    
-    ## Maximum file upload size = 10MB
-    options(shiny.maxRequestSize=10*1024^2)
-    
-    ## Reactive values
-    values <- reactiveValues(mydat = NULL, mydat_rand = NULL)
-    
-    #cat(paste(readLines("global.R"), collapse = "\n"), file = "code_global.R")
-    cat("library(RCurl)\n", file = file.path(userdir, "code_All.R"))
-    cat("eval(parse(text = getURL('https://raw.githubusercontent.com/gammarama/intRo/master/global.R')))", file = file.path(userdir, "code_All.R"), append = TRUE)
-    
-    ## Modules
-    types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
-    modules_tosource <- file.path("modules", apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
-
-    ## Source the modules
-    for (mod in modules_tosource) {
-        source(mod, local = TRUE)
-    }
-    
-    ## Check for file update every 5 seconds
-    code <- reactiveFileReader(500, session, file.path(userdir, "code_All.R"), clean_readlines)
-    
-    ## Code Viewing
-    observe({    
-      updateAceEditor(session, "myEditor", value=paste(code(), collapse="\n"))  
-    })
-    
-    
-    ## Printing
-    observe({ 
-      if(is.null(input)) return
-      if(length(input$print_clicked) > 0) {
-        file <- NULL
-        if(input$print_clicked) {
-          oldwd <- getwd()
-          
-          if(!input$code_clicked) {
-            file <- render(file.path(userdir, "code_All.R"), 
-                           html_document(css = file.path(oldwd, "www/hide_code.css")),
-                           output_dir = file.path(oldwd, "www"))            
-          } else {
-            file <- render(file.path(userdir, "code_All.R"), 
-                           output_dir = file.path(oldwd, "www"))
-          }
-          
-          
[…]

### 2015-01-06 ui.R

@@ -1,54 +1,0 @@
-addResourcePath(prefix="images", directoryPath="images/")
-
-## For printing
-dynFrame <- function(outputId) 
-{
-  HTML(paste0("<iframe class = 'print_results' src='about:blank' id = '", outputId, "'></iframe>"))
-}
-
-## Source ui
-module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
-sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)
-
-## Generates the UI tabs
-mylist <- list()
-old_heading <- ""
-for (i in 1:nrow(module_info)) {
-    if (module_info[i,1] != old_heading) {
-        if (i != 1) mylist[[length(mylist) + 1]] <- "-----"
-        mylist[[length(mylist) + 1]] <- as.character(module_info[i,1])
-        old_heading <- module_info[i,1]
-    }
-    mylist[[length(mylist) + 1]] <- get(paste(module_info[i,2], "ui", sep = "_"))
-}
-
-shinyUI(
-    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
-               tabPanel(title="", icon=icon("home"),
-                        fluidRow(
-                            dynFrame(outputId = 'print_output'),
-                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
-                        ),
-                        hr(),                        
-                        fluidRow(
-                          column(12,
-                            aceEditor("myEditor", "", mode="r", readOnly=TRUE, theme="chrome")
-                          )
-                        )
-               ),
-               tabPanel(title="", value="http://gammarama.github.io/intRo", icon=icon('question-circle')),
-               tabPanel(title="", value="http://github.com/gammarama/intRo", icon=icon("github")),
-               navbarMenu("", icon=icon("envelope"),
-                          tabPanel("Eric Hare"),
-                          tabPanel("Andee Kaplan")),
-               tabPanel(title="hide_me"),
-               tabPanel(title="", icon=icon('code'), value = "javascript:$('#myEditor').slideToggle(); $('.fa-code').parent().parent().toggleClass('active'); code_clicked();"),
-               tabPanel(title="", icon=icon("print"), value="javascript: $(this).addClass('print_button'); print_clicked();"),
-               footer=tagList(includeScript("scripts/top-nav-links.js"),
-                              includeScript("scripts/print.js"),
-                              includeScript("scripts/other-helpers.js")
[…]

### 2015-01-06 departmental-seminar/seminar-slides-markdown.Rmd
Markdown presentation

<style>
.reveal section del,
.reveal h1 del,
.reveal h3 del {
  color: #1abc9c;
}

.reveal, .reveal h3 {
  color: #2c3e50;
}

.reveal section code {
  font-size: 24px;
}

.reveal table {
  font-size: 14px;
  border-collapse: collapse;
  float: left;
}

/*** section background ***/
.section .reveal .state-background {
   background: #2c3e50;
}

.section .reveal a {
  color: #ecf0f1;
}

/*** navigation color ***/
.reveal .controls div.navigate-left,
.reveal .controls div.navigate-left.enabled {
  border-right-color: #1abc9c;
}

.reveal .controls div.navigate-right,
.reveal .controls div.navigate-right.enabled {
  border-left-color: #1abc9c;
}

.reveal .controls div.navigate-up,
.reveal .controls div.navigate-up.enabled {
  border-bottom-color: #1abc9c;
}

.reveal .controls div.navigate-down,
.reveal .controls div.navigate-down.enabled {
  border-top-color: #1abc9c;
}

.reveal .controls div.navigate-left.enabled:hover {
  border-right-color: #148f77;
}

.reveal .controls div.navigate-right.enabled:hover {
  border-left-color: #148f77;
}

.reveal .controls div.navigate-up.enabled:hover {
  border-bottom-color: #148f77;
}

.reveal .controls div.navigate-down.enabled:hover {
  border-top-color: #148f77;
}

.reveal .progress span {
  background: #1abc9c;
}

.reveal blockquote {
  background: #1abc9c;
  color: white;
}
</style>

---
title: "intRo"
author: "Andee Kaplan & Eric Hare"
date: "January 12, 2015"
output: ioslides_presentation
font-family: Helvetica
---

Statistical Analysis Software for Teaching

## Introduction
type: section

## Do we really need another statistical software package?
Short answer: ~~yes~~

 - `R` is great, but requires students to have some knowledge/interest in programming
 - JMP is powerful, but confusing to students
   - Licenses
   - Inconsistent UI
 - Students fight with software rather than focusing on the goal: statistics

## What is intRo?
 - A modern ~~web-based~~ application for performing basic data analysis and statistical routines
 - Built using `R` and `Shiny`
 - Powerful and extensible modular structure
 - Simple enough for the novice statistician
 - Meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education

## Easy
- Focused on aspects of the user interface (UI) and output that make it easy to pick up without training
- Minimal necessary functionality for an introductory statistics course
- Organized around specific tasks a student may perform in the process of a data analysis

<img src=images/user_experience.png height="300"></img>

## Exciting
 - Fun, easy to use (available on the web)
 - Interactive plots using `ggvis`

~~Ulterior motive~~: get students excited about programming
 - By navigating about the user interface of intRo, students are creating a fully-executable `R` script that they can download and run locally
 - Viewing their script change real-time within the application

## Extensible
 - User interaction with intRo is split into bitesize chunks that we call *modules*
 - Each module is a self contained set of `R` code that is dynamically added to the application at run time
 - `intRo` can be easily extended by the addition of modules within the frame-work underlying the application
 - Allows `intRo` to be tailored to the needs of a particular course

## First look
type: section

http://intro-stats.com

## Classroom
type: section

## intRo in the wild
 - `intRo` is currently being used in STAT 201
  - Homework assignments and labs will use `intRo` to emphasize class concepts
 - `intRo` is a supplement, not the focus of the class

## Homework 1
 - *Data*: OECD PISA Results in Focus report to assess the workforce readiness of 15-year old students
 - *Goal*: Explore graphically and numerically how the students perceive academics and math
 - `intRo` *Tasks*:
   1. Make a summary of the Enjoy.Maths variable. What proportion of these students agree or strongly agree with the statement "I enjoy mathematics"?
  1. Make a barchart of the variable Enjoy.Maths.
  1. Give the conditional distribution of ``Strongly agree'' given the Country.
  2. Obtain a contingency table and mosaic plot of the relationship between Country and Enjoy.Maths

## Our turn
type: section

## Design decisions
type: section

## Consistent UI
Elements of `intRo`: 1) top navigation, 2) side navigation, 3) options
panel, 4) results panel, and 5) code panel.

<img src="images/ui_annotate.png" height="500"></img>

## Consistent UI (Cont'd)
 - Large, easy to click icons in the page header to help students find
exactly what they need easily
 - Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code
 - Documentation website that is consistent with
the interface of the application; covers all default modules and makes learning to use intRo even more painless

## Modularity
![app creation](images/app_creation_modules.png)

## Modularity (Cont'd)
The interface is created with the following statement

```r
## Source ui
module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)

## mylist is a list containing the different ui module code
## Create the UI
shinyUI(
    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
               tabPanel(title="", icon=icon("home"),
                        fluidRow(
                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
                        )
               ), ...
    ))
```

## Modularity (Cont'd)
The server functions are dynamically generated using a similar method

```r
shinyServer(function(input, output, session) {
  ## Module info
  module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")

  ## Modules
  types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
  modules_tosource <- file.path("modules", 
                                apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))

  ## Source the modules
  for (mod in modules_tosource) {
      source(mod, local = TRUE)
  }
})
```

## Challenges
type: section

## Tradeoffs
There were a number of challenges we encountered in striving to strike the balance of functionality and ease-of-use:

* Storing the user's results
* Presenting only relevant options to the user
* Producing clean and executable R code

## Storing results
* Users must be able to print results, but only the results they want
* Reactivity makes this somewhat challenging

~~Solution~~: Store Buttons

![](images/store.png)

## Storing results (Cont'd)
* Each module contains, at the bottom, a Store button
* When Store is pressed, code is stored in code panel
    * Additional challenge: All options selected must also be stored!
* When user goes to print results, only this particular code is executed

## Presenting choices
* intRo is flexible, but must also be simple for the introductory student
* Despite increasing the complexity of the code, we decided to only show options appropriate for the given selections.

~~Example~~:

<img src="images/graphical.png" height="300">

## Presenting choices (Cont'd)
<img src="images/graphical_table.png">

## Producing clean code
Code generated must be:

* Executable on the server
* Executable on the user's machine
* Clean

## Producing clean code (Cont'd)
But how do we handle uploaded data?

```{r, eval=FALSE}
file.choose <- function () {
    return(input$data_own[,"datapath"])
}
```

Now we can provide to the user *and* the server:

```{r, eval=FALSE}
intro.data <- read.csv(file.choose())
```

## Future work
type: section

## What's next?
* Module creation package
* Server load
* Use in other classes

## Module creation package
*Modularity* is a key feature of intRo, but module creation is currently:
* Undocumented
* Entirely manual
* Unnecessarily lengthy

~~Idea~~: R package (with associated Shiny app?) to automate creation of intRo modules

## Server load
Limited testing has been done to assess how intRo handles heavy user load
* Dozens of students at accessing app at once
* Simultaneous larger computations like printing results

More server resources may need to be devoted to handle this, particularly if intRo is more widely adopted

## More classes
* We would like to see intRo used in more classes
* Module creation package will allow it to suit more curricula
* We welcome collaborators interested in extending intRo to submit pull requests on GitHub (http://www.github.com/gammarama/intRo)
    * Additional modules
    * Improvements and bug fixes

## Special thanks
Dr. Cook, for guiding many of the features and interface decisions, and helping test the functionality in preparation for STAT 201

## Any questions?
type: section

Thank you!

Contact:<br/>
ajkaplan@iastate.edu<br/>
erichare@iastate.edu

### 2015-01-06 departmental-seminar/seminar-slides-markdown.Rmd
working with rmarkdown instead

@@
-<style>
-.reveal section del,
-.reveal h1 del,
-.reveal h3 del {
-  color: #1abc9c;
-}
-
-.reveal, .reveal h3 {
-  color: #2c3e50;
-}
-
-.reveal section code {
-  font-size: 24px;
-}
-
-.reveal table {
-  font-size: 14px;
-  border-collapse: collapse;
-  float: left;
-}
-
-/*** section background ***/
-.section .reveal .state-background {
-   background: #2c3e50;
-}
-
-.section .reveal a {
-  color: #ecf0f1;
-}
-
-/*** navigation color ***/
-.reveal .controls div.navigate-left,
-.reveal .controls div.navigate-left.enabled {
-  border-right-color: #1abc9c;
-}
-
-.reveal .controls div.navigate-right,
-.reveal .controls div.navigate-right.enabled {
-  border-left-color: #1abc9c;
-}
-
-.reveal .controls div.navigate-up,
-.reveal .controls div.navigate-up.enabled {
-  border-bottom-color: #1abc9c;
-}
-
-.reveal .controls div.navigate-down,
-.reveal .controls div.navigate-down.enabled {
-  border-top-color: #1abc9c;
-}
-
-.reveal .controls div.navigate-left.enabled:hover {
-  border-right-color: #148f77;
-}
-
-.reveal .controls div.navigate-right.enabled:hover {
-  border-left-color: #148f77;
-}
-
-.reveal .controls div.navigate-up.enabled:hover {
-  border-bottom-color: #148f77;
-}
-
-.reveal .controls div.navigate-down.enabled:hover {
-  border-top-color: #148f77;
-}
-
-.reveal .progress span {
-  background: #1abc9c;
-}
-
-.reveal blockquote {
-  background: #1abc9c;
-  color: white;
-}
-</style>
-
 ---
-title: "intRo"
+title: "`intRo`"
+subtitle: "Statistical Analysis Software for Teaching"
 author: "Andee Kaplan & Eric Hare"
@@
 font-family: Helvetica
+ioslides_presentation:
+    logo: images/icon.png
+    widescreen: true
 ---
 
-Statistical Analysis Software for Teaching
-
-## Introduction
-type: section
+# Introduction
 
 ## Do we really need another statistical software package?
-Short answer: ~~yes~~
+Short answer: **yes**
 
@@
 
-## What is intRo?
- - A modern ~~web-based~~ application for performing basic data analysis and statistical routines
+## What is `intRo`?
+ - A modern **web-based** application for performing basic data analysis and statistical routines
  - Built using `R` and `Shiny`
@@
 
-<img src=images/user_experience.png height="300"></img>
+<img src=images/user_experience.png height="250"></img>
 
@@
 
-~~Ulterior motive~~: get students excited about programming
- - By navigating about the user interface of intRo, students are creating a fully-executable `R` script that they can download and run locally
+**Ulterior motive**: get students excited about programming
+ - By navigating about the user interface of `intRo`, students are creating a fully-executable `R` script that they can download and run locally
  - Viewing their script change real-time within the application
@@
 ## Extensible
- - User interaction with intRo is split into bitesize chunks that we call *modules*
+ - User interaction with `intRo` is split into bitesize chunks that we call *modules*
  - Each module is a self contained set of `R` code that is dynamically added to the application at run time
@@
 
-## First look
-type: section
+# First look | http://intro-stats.com
 
-http://intro-stats.com
+# Classroom
 
-## Classroom
-type: section
-
-## intRo in the wild
+## `intRo` in the wild
  - `intRo` is currently being used in STAT 201
@@
   1. Make a barchart of the variable Enjoy.Maths.
-  1. Give the conditional distribution of ``Strongly agree'' given the Country.
+  1. Give the conditional distribution of "Strongly agree'' given the Country.
   2. Obtain a contingency table and mosaic plot of the relationship between Country and Enjoy.Maths
 
-## Our turn
-type: section
+# Our turn
 
-## Design decisions
-type: section
+# Design decisions
 
@@
 
-<img src="images/ui_annotate.png" height="500"></img>
+<img src="images/ui_annotate.png" height="400"></img>
 
@@
  - Documentation website that is consistent with
-the interface of the application; covers all default modules and makes learning to use intRo even more painless
+the interface of the application; covers all default modules and makes learning to use `intRo` even more painless
 
 ## Modularity
-![app creation](images/app_creation_modules.png)
+<img src="images/app_creation_modules.png" height="500"></img>
 
@@
                         fluidRow(
+                        ### <b>
                             do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
+                        ### </b>
                         )
@@
 
-## Challenges
-type: section
+# Challenges
 
@@
 
-* Storing the user's results
-* Presenting only relevant options to the user
-* Producing clean and executable R code
+ - Storing the user's results
+ - Presenting only relevant options to the user
+ - Producing clean and executable R code
 
 ## Storing results
-* Users must be able to print results, but only the results they want
-* Reactivity makes this somewhat challenging
+ - Users must be able to print results, but only the results they want
+ - Reactivity makes this somewhat challenging
 
-~~Solution~~: Store Buttons
+**Solution**: Store Buttons
 
-![](images/store.png)
+<img src="images/store.png" height="300"></img>
 
 ## Storing results (Cont'd)
-* Each module contains, at the bottom, a Store button
-* When Store is pressed, code is stored in code panel
-    * Additional challenge: All options selected must also be stored!
-* When user goes to print results, only this particular code is executed
+ - Each module contains, at the bottom, a Store button
+ - When Store is pressed, code is stored in code panel
+     - Additional challenge: All options selected must also be stored!
+ - When user goes to print results, only this particular code is executed
 
 ## Presenting choices
-* intRo is flexible, but must also be simple for the introductory student
-* Despite increasing the complexity of the code, we decided to only show options appropriate for the given selections.
+ - `intRo` is flexible, but must also be simple for the introductory student
+ - Despite increasing the complexity of the code, we decided to only show options appropriate for the given selections.
 
-~~Example~~:
+**Example**:
 
@@
 ## Presenting choices (Cont'd)
-<img src="images/graphical_table.png">
+<img src="images/graphical_table.png" height="275">
 
@@
 
-* Executable on the server
-* Executable on the user's machine
-* Clean
+ - Executable on the server
+ - Executable on the user's machine
+ - Clean
 
@@
 
-## Future work
-type: section
+# Future work
 
 ## What's next?
-* Module creation package
-* Server load
-* Use in other classes
+ - Module creation package
+ - Server load
+ - Use in other classes
 
 ## Module creation package
-*Modularity* is a key feature of intRo, but module creation is currently:
-* Undocumented
-* Entirely manual
-* Unnecessarily lengthy
+*Modularity* is a key feature of `intRo`, but module creation is currently:
+ - Undocumented
+ - Entirely manual
+ - Unnecessarily lengthy
 
-~~Idea~~: R package (with associated Shiny app?) to automate creation of intRo modules
+**Idea**: R package (with associated Shiny app?) to automate creation of `intRo` modules
 
 ## Server load
-Limited testing has been done to assess how intRo handles heavy user load
-* Dozens of students at accessing app at once
-* Simultaneous larger computations like printing results
+Limited testing has been done to assess how `intRo` handles heavy user load
+ - Dozens of students at accessing app at once
+ - Simultaneous larger computations like printing results
 
-More server resources may need to be devoted to handle this, particularly if intRo is more widely adopted
+More server resources may need to be devoted to handle this, particularly if `intRo` is more widely adopted
 
 ## More classes
-* We would like to see intRo used in more classes
-* Module creation package will allow it to suit more curricula
-* We welcome collaborators interested in extending intRo to submit pull requests on GitHub (http://www.github.com/gammarama/intRo)
-    * Additional modules
-    * Improvements and bug fixes
+ - We would like to see `intRo` used in more classes
+ - Module creation package will allow it to suit more curricula
+ - We welcome collaborators interested in extending `intRo` to submit pull requests on GitHub (http://www.github.com/gammarama/intRo)
+     - Additional modules
+     - Improvements and bug fixes
 
@@
 
-## Any questions?
-type: section
-
-Thank you!
+# Any questions? | Thank you!
 

### 2015-01-06 departmental-seminar/seminar-slides-markdown.Rmd
Format the code better

@@
 
-```r
+```{r, tidy=FALSE, eval=FALSE}
 ## Source ui
-module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
-sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)
-
-## mylist is a list containing the different ui module code
+module_info <- read.table("modules/modules.txt", header = TRUE, 
+                          sep=",")
+sapply(file.path("modules", dir("modules")[dir("modules") 
+                      != "modules.txt"], "ui.R"), source)
 ## Create the UI
-shinyUI(
-    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
+shinyUI(navbarPage("intRo", theme = "bootstrap.min.css",
                tabPanel(title="", icon=icon("home"),
@@
                         ### <b>
-                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
+                          do.call(navlistPanel, 
+                              c(list(id = "side-nav", 
+                              widths = c(2, 10)), mylist))
                         ### </b>
@@
 
-```r
+```{r, eval=FALSE, tidy=FALSE}
 shinyServer(function(input, output, session) {
   ## Module info
-  module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
-
+  module_info <- read.table("modules/modules.txt", 
+                            header = TRUE, sep=",")
   ## Modules
-  types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
+  types <- c("helper.R", "static.R", "observe.R", 
+             "reactive.R", "output.R")
   modules_tosource <- file.path("modules", 
-                                apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
-
+                                apply(expand.grid(
+                                module_info$module, types)
+                                , 1, paste, collapse = "/"))
   ## Source the modules
-  for (mod in modules_tosource) {
-      source(mod, local = TRUE)
-  }
+  for (mod in modules_tosource) { source(mod, local = TRUE) }
 })

### 2015-01-06 departmental-seminar/seminar-slides-markdown.Rmd
code highlighting

@@
   ## Source the modules
-  for (mod in modules_tosource) { source(mod, local = TRUE) }
+  for (mod in modules_tosource) {
+    ### <b>
+    source(mod, local = TRUE) 
+    ### </b>
+  }
 })

### 2015-01-06 departmental-seminar/index.Rmd
first draft of slides

--- 2015-01-06 departmental-seminar/seminar-slides-markdown.Rmd
@@
 ---
-title: "`intRo`"
+title: "intRo"
 subtitle: "Statistical Analysis Software for Teaching"
@@
  - Simple enough for the novice statistician
- - Meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education
+ - Assists in the learning of statistics rather than acting as a stand-alone deliverer of statistics education
 
@@
 **Ulterior motive**: get students excited about programming
+
  - By navigating about the user interface of `intRo`, students are creating a fully-executable `R` script that they can download and run locally
@@
  - `intRo` can be easily extended by the addition of modules within the frame-work underlying the application
- - Allows `intRo` to be tailored to the needs of a particular course
+ - Allows instructors/collaborators to tailor `intRo` to the needs of a particular course
 
@@
 ## Tradeoffs
-There were a number of challenges we encountered in striving to strike the balance of functionality and ease-of-use:
+We encountered a number of challenges in striving to strike the balance of functionality and ease-of-use:
 
@@
 ## Presenting choices (Cont'd)
-<img src="images/graphical_table.png" height="275">
+<img src="images/graphical_table.png">
 
@@
 *Modularity* is a key feature of `intRo`, but module creation is currently:
+
  - Undocumented
@@
 Limited testing has been done to assess how `intRo` handles heavy user load
+
  - Dozens of students at accessing app at once
@@
 ## More classes
- - We would like to see `intRo` used in more classes
+ - We would like work with instructors to get `intRo` into more classes
  - Module creation package will allow it to suit more curricula

### 2015-01-09 presentations/departmental-seminar/index.Rmd
updates from @dicook

--- 2015-01-06 departmental-seminar/index.Rmd
@@
  - `R` is great, but requires students to have some knowledge/interest in programming
- - JMP is powerful, but confusing to students
+ - JMP is powerful, but too big
       - Licenses
       - Inconsistent UI
- - Students fight with software rather than focusing on the goal: statistics
+ - Want students to focus on data analysis rather than fight with software
 
 ## What is `intRo`?
- - A modern **web-based** application for performing basic data analysis and statistical routines
+ - A simple **web-based** application for performing basic data analysis and statistical routines
  - Built using `R` and `Shiny`
- - Powerful and extensible modular structure
- - Simple enough for the novice statistician
+ - Extensible modular structure
+ - Designed for a first statistics class student
  - Assists in the learning of statistics rather than acting as a stand-alone deliverer of statistics education
@@
 ## `intRo` in the wild
- - `intRo` is currently being used in STAT 201
+ - `intRo` will be used this semester in STAT 201
       - Homework assignments and labs will use `intRo` to emphasize class concepts

### 2015-01-11 presentations/departmental-seminar/index.Rmd
New slides again

@@
  - `R` is great, but requires students to have some knowledge/interest in programming
- - JMP is powerful, but too big
-      - Licenses
-      - Inconsistent UI
+ - `JMP` is powerful, but too big
+      - Licenses and inconsistent UI
+ - New tools recently released to spark an interest in `R`
+      - `Swirl` and DataCamp teach `R` programming
+      - Project MOSAIC facilitates learning, but assumes knowledge of `R`
  - Want students to focus on data analysis rather than fight with software
@@
  - `intRo` *Tasks*:
+      1. Make a barchart of the variable Enjoy.Maths.
       1. Make a summary of the Enjoy.Maths variable. What proportion of these students agree or strongly agree with the statement "I enjoy mathematics"?
-      1. Make a barchart of the variable Enjoy.Maths.
-      1. Give the conditional distribution of "Strongly agree'' given the Country.
-      2. Obtain a contingency table and mosaic plot of the relationship between Country and Enjoy.Maths
+      2. Obtain a contingency table and mosaic plot of the relationship between Country and Enjoy.Maths. Print contingency tables with the **Count** and **Column Proportion** values.
 

### 2015-02-06 math-siam/index.Rmd
changes for math-siam talk

--- 2015-01-11 presentations/departmental-seminar/index.Rmd
@@
 author: "Andee Kaplan & Eric Hare"
-date: "January 12, 2015"
+date: "February 6, 2015"
 output: ioslides_presentation
@@
 ## `intRo` in the wild
- - `intRo` will be used this semester in STAT 201
+ - `intRo` will be used this semester in STAT 201 and STAT 104
       - Homework assignments and labs will use `intRo` to emphasize class concepts

### 2015-02-16 lessonplans/hmwkch3.tex
homeworks to be done using intRo

\documentclass{article}[11pt]
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\begin{document}

\hspace{2in} Name: (Print) \underline{\mbox{\hspace{3.5in}}}\\

\begin{center}
{\Large STATISTICS 201 - Written Homework Chapter 3}\\[3mm]
\end{center}

\begin{enumerate}

\item Michael Jordan is highly renowned basketball player who played for 15 seasons in the NBA.  The following table contains the average points per game scored by Jordan in each of his 15 seasons.

\begin{center}
\begin{tabular}{|l|c||l|c||l|c|} \hline
Year & Avg Pts/Game & Year & Avg Pts/Game & Year & Avg Pts/Game 
\\ \hline
1984 & 28 & 1989 & 34 & 1995 & 30 \\
1985 & 23 & 1990 & 32 & 1996 & 30 \\
1986 & 37 & 1991 & 30 & 1997 & 29 \\
1987 & 35 & 1992 & 33 & 2001 & 23 \\
1988 & 33 & 1994 & 27 & 2002 & 20 \\ \hline
\end{tabular}
\end{center}

\begin{enumerate}
\item Make a stem and leaf display of the average number of points per game that Michael Jordan scored per season.  Do this by hand, use a split stem, and ensure that the leaves are sorted numerically. Describe the shape of this distribution. \\[2.5in]

\item Make a line plot of average number of points per game by year. Describe the pattern of points relative to time. \\[2.5in]

\end{enumerate}

\newpage
\item For this part of the assignment you will be using data collected on IA high school quarterbacks for the 2012 season.

\begin{enumerate}
\item Describe the Who for this data.\\[0.35in]
\item Use intRo to obtain the a histogram of yards thrown, with a suitable bin width.  Describe the shape of the distribution of this variable. Print out the output and turn it in with this assignment.
\item Give the five number summary for this variable.\\[0.2in]
\item Calculate the range and IQR for this variable.\\[0.1in]
\item What percentage of the quarterbacks  threw less than 366 yards? \\[0.1in]
\item What percentage of the quarterbacks  threw more than 3396 yards?  \\[0.1in]
\item Give the mean and standard deviation for this variable.\\[0.1in]
\item Explain why there is a difference of approximately 70 yards in the mean and median for this variable.\\[1in]
\item Which numerical summaries are the most appropriate for this distribution - the 5-number summary or the mean and standard deviation?  Circle the appropriate answer below.  Then explain your answer.

\hspace{1in} 5-number summary \hspace{1in} mean and standard deviation \\[0.5em]

Explain your answer:\\[0.5in]
\item Explore transformations of this variable to make it more symmetric. Circle which is best of the choices below:

\hspace{1in} 0.5 \hspace{1in} 0.25 \hspace{1in} 0

Print a histogram of your transformed data. Compute what the transformed value of 3000 yards would be.

\end{enumerate}

\end{enumerate}

\end{document}

### 2015-02-16 lessonplans/hmwkch3sol.tex
homeworks to be done using intRo

\documentclass{article}[11pt]
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\begin{document}
\begin{center}
{\Large STATISTICS 201 - Written Homework Chapter 3 Solutions}\\[3mm]
\end{center}

\noindent 
This assignment is worth a total of 35 points.

\begin{enumerate}

\item (8 pts) Michael Jordan is highly renowned basketball player who played for 15 seasons in the NBA.  The following table contains the average points per game scored by Jordan in each of his 15 seasons.

\begin{center}
\begin{tabular}{|l|c||l|c||l|c|} \hline
Year & Avg Pts/Game & Year & Avg Pts/Game & Year & Avg Pts/Game 
\\ \hline
1984 & 28 & 1989 & 34 & 1995 & 30 \\
1985 & 23 & 1990 & 32 & 1996 & 30 \\
1986 & 37 & 1991 & 30 & 1997 & 29 \\
1987 & 35 & 1992 & 33 & 2001 & 23 \\
1988 & 33 & 1994 & 27 & 2002 & 20 \\ \hline
\end{tabular}
\end{center}

\begin{enumerate}
\item (5 pts) Make a stem and leaf display of the average number of points per game that Michael Jordan scored per season.  Use a split stem. \\
\begin{center}
\begin{tabular}{c|cccccccc}
2 & 0 & 3 & 3 \\ 
2 & 7 & 8 & 9 \\
3 & 0 & 0 & 0 & 2 & 3 & 3 & 4 \\
3 & 5 & 7 \\ 
\end{tabular}
\end{center}

\item (3 pts) Using your stem and leaf display, describe the shape of this distribution.

Unimodal, skewed left and no apparent outliers.
\end{enumerate}

\item (27 pts) For this part of the assignment you will be using data collected on the population of each state in the United States in 1975.  The state populations can be found in the JMP data file {\bf StatePop1975.JMP} in the Chapter 3 Learning Module in the Unit 1 content area in Blackboard. (Note: These data are recorded by the thousands of people, so a value of 850 should be read as 850,000 people.)

\begin{enumerate}
\item (1 pt) Describe the Who for this quantitative variable.

The 50 States

\item (5 pts) Use JMP to obtain the distribution of the state populations.  Change the display to a Horizontal Layout, add a Count Axis to the histogram, and a Stem-and-Leaf display to the output.  Print out the output and turn it in with this assignment.

See the attached JMP output.

\item (2 pts) What percentage of the 50 states had between 800 thousand and 900 thousand residents in 1975?

3/50 = 6\%

\item (2 pts) What percentage of the 50 states had greater than 10 million residents in 1975?

6/50 = 12\%

\item (3 pts) Describe the shape of the distribution of this variable.

Unimodal, Skewed Right, two possible outliers

\item (5 pts) Give the five number summary for this variable.

The 5 number summary in this case is: Min = 365,000; Q1 = 1,026,250; Median = 2,838,500; Q3 = 5,064,000; Max = 21,198,000.

\item (2 pts) Calculate the range and IQR for this variable.

Range = Max - Min = 21,198,000 - 365,000 = 20,833,000 \hspace{0.25in} IQR = Q3 - Q1 = 5,064,000 - 1,026,250 = 4,037,750.

\item (2 pts) Give the mean and standard deviation for this variable.

$\overline{y} = $4,246,420; $s = $4,464,490.

\item (2 pts) Explain why there is a difference of approximately 1.4 million people in the mean and median for this variable.

This is because the plot is skewed right, which causes the mean to be dragged upward, but leaves the median unaffected.

\item (3 pts) Which numerical summaries are the most appropriate for this distribution - the 5-number summary or the mean and standard deviation?  Circle the appropriate answer below.  Then explain your answer.

The correct answer is: The 5-number summary

Explain your answer: The 5-number summary is more appropriate because the distribution is heavily skewed right and has outliers.

\end{enumerate}

\end{enumerate}

\end{document}

### 2015-02-16 student-paper-competition-2014/paper.Rmd
homeworks to be done using intRo

--- 2014-12-12 student-paper-competition-2014/paper.Rmd
@@
 ---
-title: "Introducing statistics with \\texttt{intRo}"
+title: "Introducing statistics with \texttt{intRo}"
 author: "Eric Hare & Andee Kaplan"
@@
     number_sections: yes
+  html_document: default
 bibliography: refs.bib
-geometry: margin=0.8in
 ---
 \thispagestyle{empty}
-
 \begin{abstract}
@@
 
-\twocolumn
+\newpage
 \setcounter{page}{1}
@@
 # Introduction
-
 The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been a great development for the scientific community.
@@
 Still, signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.
-
 In teaching STAT 226 at Iowa State University, we witnessed profound struggles by students attempting to practice the introductory concepts discussed in class.
-Scrimshaw notes in his manuscript that "open-ended packages, like any others, may create obstacles to learning simply through their lack of user-friendliness in the sheer mechanics of operating them, rather than any intrinsic difficulty in the content...." [@scrimshaw2001computers] In our own experience teaching, students' struggles were often directly related to the use of the software and not any sort of fundamental misunderstanding of the material, in agreement with Scrimshaw's finding.
-
-Other software has recently been written in an attempt to spur student interest in introductory statistics and R programming.
-DataCamp's [@datacamp] courses are user-friendly ways to learning basic \texttt{R} programming and data analysis techniques.
-\texttt{Swirl} [@swirl] is a similar interactive tool to make learning \texttt{R} more fun by learning \texttt{R} within \texttt{R} itself.
-Project MOSAIC [@mosaic] has created a suite of tools to simplify the teaching of statistics in the form of an R package.
-The primary goal of DataCamp and Swirl is to teach \texttt{R} programming, rather than facilitate the learning of introductory statistics.
-Project MOSAIC's goal is to faciliate this learning, but using the package requires a knowledge of R programming that the introductory student may not have.
+Often, these struggles were directly related to the use of the software and not any sort of fundamental misunderstanding of the material.
 
 Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easy for an R-based analysis to be converted to an interactive web application.
-This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}, available at http://www.intro-stats.com.
+This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}.
+
 \texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.
+
+\ak{[Add some references about software discouraging people, and mention other new tools out there like swirl.]}
 
@@
 ## Easy
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\linewidth]{images/user_experience.pdf}
-\caption{Typical user experience of generating a result in \texttt{intRo}. In this instance, a mosaic plot (and its corresponding code) was created and saved to the final compendium.}
-\label{fig:user_experience}
-\end{figure}
-
 In order for an application to be useful to students in their first statistics course, it can't hog the spotlight.
@@
 Additional functionality is nice, but if it clutters the UI beyond the point of usability, it will frustrate an introductory student.
-Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in \texttt{intRo}, and view the code producing the result.
 
@@
 
-\begin{figure*}[ht!]
+\begin{figure}[ht!]
 \centering
-\includegraphics[width=\linewidth]{images/ui_annotate.pdf}
+\includegraphics[width=0.8\textwidth]{images/ui_annotate.png}
 \caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
 \label{fig:ui}
-\end{figure*}
+\end{figure}
 
@@
 The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can be easy added by an instructor, as detailed in section \ref{sec:extensible}.
-The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.
 
-In addition to making \texttt{intRo} intuitive to use, we've also created a documentation website that is consistent with the interface of the application.
-The documentation is available at http://gammarama.github.io/intRo/ and hyperlinked within the application itself.
+Within the data category there are two modules, sources and transformation.
+The data sources module allows the user to select or upload a dataset.
+Several datasets are pre-included, which can be selected via a drop-down menu.
+The user can check a box to upload a dataset, and then select the CSV file from her computer.
+Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
+From this module the user can select a subset of the data either by value or random selection.
+The dataset selected in the data module will be made available to all other modules of \texttt{intRo}.
+The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
+Additionally, numeric variable can be transformed using a power transformation.
+The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.
+
+The graphical summaries module provides several common graphical displays of variables in the dataset.
+First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
+The variable available to choose will only be those compatible with plot type selected.
+The plot visible on the right will automatically update when these values are selected.
+The numeric summaries module is another method of computing summaries of the data.
+The user is presented with a list of each variable from their dataset.
+The user can select one or more different variables, and will be presented with a table of some summary statistics.
+The summary statistics displayed vary depending on the type of the variables.
+For instance, for a numeric variable, a five number summary will be displayed.
+Character or factor variables will display counts of each unique value of that particular variable.
+Additionally, the user has the ability to display summaries by a grouping variable.
+
+Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester.
+The contingency table module allows the user to compute a contingency table for two categorical variables.
+The results can either be displayed as counts, row percentages, column percentages, or total percentages.
+The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
+The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
+Results of the regression will be displayed on the right.
+A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
+A user also has the ability to save residuals and predicted values to their dataset for further analysis.
+The T test module allows the user to perform either a one variable or two variable Student's T test.
+The user can select configuration options, including the direction of the test and the confidence level of the test.
+The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
+
+In addition to making \texttt{intRo} intuitive to use, we've also created a documentation website that is consistent with the interface of the application and hyperlinked within the application itself.
 This documentation covers all default modules and makes learning to use \texttt{intRo} even more painless.
@@
 
-\begin{figure*}[ht!]
+\begin{figure}[ht!]
 \centering
-\includegraphics[width=\linewidth]{images/app_creation_modules.pdf}
+\includegraphics[width=\textwidth]{images/app_creation_modules.png}
 \caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.}
 \label{fig:app_creation_modules}
-\end{figure*}
+\end{figure}
 
@@
 \footnotesize
-```{r ui, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
+```{r ui, eval=FALSE}
 ## Source ui
@@
 
-## mylist is a list containing the different ui module code
+## Generates the UI tabs
+mylist <- list()
+old_heading <- ""
+for (i in 1:nrow(module_info)) {
+    if (module_info[i,1] != old_heading) {
+        if (i != 1) mylist[[length(mylist) + 1]] <- "-----"
+        mylist[[length(mylist) + 1]] <- as.character(module_info[i,1])
+        old_heading <- module_info[i,1]
+    }
+    mylist[[length(mylist) + 1]] <- get(paste(module_info[i,2], "ui", sep = "_"))
+}
+
 ## Create the UI
@@
 \footnotesize
-```{r server, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
+```{r server, eval=FALSE}
 shinyServer(function(input, output, session) {
@@
   types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
-  modules_tosource <- file.path("modules", 
-                                apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
+  modules_tosource <- file.path("modules", apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
 
@@
   }
-})
+}
 ```
@@
 
-\begin{table*}[ht!]
+\begin{table}[ht!]
 \centering
@@
 \end{tabular}
+\label{tab:plottypes}
 \caption{The plot and variable types allowed in \texttt{intRo}'s graphical module.
@@
 If a boxplot is selected, the x variable must be categorical and the y variable must be numeric, and the list of choices updates automatically to reflect this.}
-\label{tab:plottypes}
-\end{table*}
+\end{table}
 Part two of this text discusses probability and sampling distributions, which is outside the current realm of \texttt{intRo}, but part three focuses on inferential statistics.
@@
 This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
-Furthermore, we would like to utilize the expanded interactive capabilities of \texttt{ggvis} in order to make \texttt{intRo's} plots more engaging to students.
-One way to do this would be implementing linked plots, in which interactions with one plot are reflected in other plots that illustrate the same data.
-This would be particularly useful in the regression module so that students could explore observations with high influence and leverage.
 
@@
 An unknown quantity will be how feasible it is to increase adoption of \texttt{intRo} across the department, as well as to other universities.
-One limitation of \texttt{intRo} is that uploading a dataset beyond about 30,000 rows tends to be slow.
-Even once the data is successfully uploaded, the default modules produce results more slowly than with smaller datasets.
-This is a limitation that should be further investigated if and when \texttt{intRo} sees wider adoption.
 

### 2015-02-16 lessonplans/hmwkch5.tex
populating the first few homeworks

--- 2014-12-16 hmwkch5.tex
@@
 \begin{enumerate}
-\item What proportion of all test takers will have a GRE Verbal score above 600 points?\\[1in]
-\item What proportion of all test takers will have a GRE Verbal score below 375 points?\\[1in]
-\item What proportion of all test takers will have a GRE Verbal score between 500 and 700 points?\\[1in]
-\item What GRE Verbal score will 40\% of all test takers score below?\\[1in]
+\item What proportion of all test takers will have a GRE Verbal score above 600 points?\\%[1in]
+\item What proportion of all test takers will have a GRE Verbal score below 375 points?\\%[1in]
+\item What proportion of all test takers will have a GRE Verbal score between 500 and 700 points?\\%[1in]
+\item What GRE Verbal score will 40\% of all test takers score below?\\%[1in]
 \item What GRE verbal score will 10\% of all test takers score above?
@@
 
-\newpage
+%\newpage
 
 \item In a study to determine the nature of the relationship between heart rates and body temperature, body temperatures were collected from 130 individuals.
-The body temperatures, in degrees Fahrenheit, can be found in the JMP data file {\bf samplebodytemps.JMP} in Blackboard.
+The body temperatures, in degrees Fahrenheit, can be found in the  data file {\bf samplebodytemps.csv} in Blackboard.
 
-Use JMP to determine if the distribution of body temperatures can be modeled with a normal model.
-Make sure to turn in the output from JMP with your assignment.
-In your answer, make sure to use information from the JMP output that supports your conclusions.
+Use the intRo software to determine if the distribution of body temperatures can be modeled with a normal model.
+Make sure to turn in the output from intRo with your assignment.
+In your answer, make sure to use information from the computer output that supports your conclusions.
 

### 2015-02-16 lessonplans/hmwkch5sol.tex
populating the first few homeworks

\documentclass{article}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\usepackage{graphicx}

\begin{document}

\begin{center}
{\large STATISTICS 201 - Written Homework Chapter 5}\\[6mm]
\end{center}

\noindent
This assignment is worth a total of 30 points.

\begin{enumerate}

\item (19 pts) The GRE is a test to measure undergraduate achievement and is used as an admissions criterion to graduate school. The GRE is designed in such a way that scores on the test can be modeled with a normal model. From July 2006 to June 2007 the mean GRE Verbal score was 457 with a standard deviation of 121. Use the z-table to answer the following questions.

\begin{enumerate}
\item (3 pts) What proportion of all test takers will have a GRE Verbal score above 600 points?

The standardized value of 600 points is $z = \frac{600-457}{121} = 1.18182$ rounded to $z = 1.18$.

The percentile from the normal table for $z = 1.18$ is $0.8810$. Since we want the proportion above 600 points, the answer is $1-0.8810 = 0.1190$.

\item (3 pts) What proportion of all test takers will have a GRE Verbal score below 375 points?

The standardized value of 375 points is $z = \frac{375-457}{121} = -0.677686$, rounded to $z = -0.68$.

The percentile from the normal table for $z = -0.68$ is $0.2483$. Since we want the proportion below 375 points, the answer is $0.2483$.

\item (5 pts) What proportion of all test takers will have a GRE Verbal score between 500 and 700 points?

The standardized value of 500 points is $z = \frac{500-457}{121} = 0.355372$, rounded to $z = 0.36$; the standardized value of 700 points is $z = \frac{700-457}{121} = 2.00826$, rounded to $z = 2.01$.

The percentile from the normal table for $z = 0.36$ is 0.6406 and for $z = 2.01$ is 0.9778. Since we want the proportion between 500 and 700 points, the answer is $0.9778 - 0.6406 = 0.3372$.

\item (4 pts) What GRE Verbal score will 40\% of all test takers score below?

The score where 40\% of test takers are below is the same as the 40th percentile. From the normal table, the percentile closest to 0.40 is 0.4013 when $z = -0.25$, so we will use $z = -0.25$.

We are looking for the score that is 0.25 standard deviations below the mean, this value is $457 + (-0.25*121) = 426.75$.

\item (4 pts) What GRE verbal score will 10\% of all test takers score above?

The score where 10\% of test takers are above is the same score where 90\% are below. So we need to find the 90th percentile. From the normal table, the percentile 0.8997 when $z = 1.28$ is closest to 0.9, so we will use $z = 1.28$.  We are looking for the score that is 1.28 standard deviations above the mean, this value is $457 + (1.28*121) = 611.88$.

\end{enumerate}

\newpage

\item (11 pts) IIn a study to determine the nature of the relationship between heart rates and body temperature, body temperatures were collected from 130 individuals. The body temperatures, in degrees Fahrenheit, can be found in the  data file {\bf samplebodytemps.csv} in Blackboard.

Use the intRo software to determine if the distribution of body temperatures can be modeled with a normal model. Make sure to turn in the output from intRo with your assignment. In your answer, make sure to use information from the computer output that supports your conclusions.

%\begin{tabular}{p{3in}p{2.5in}}
(5 pts) Computer output

%\includegraphics[width=2.5in]{hwk5jmpoutput.pdf}

%&
%\vspace{0.5in}
(6 pts)  We see that the distribution of body temperatures is unimodal and symmetric, with a general bell-shape to the histogram.  The points in the normal quantile plot generally follow the straight line in the plot. Due to the shape of the histogram and the straight line pattern in the normal quantile plot, it appears that the distribution of body temperatures in the sample can be modeled with the normal model.

%\end{tabular}

\end{enumerate}

\end{document}

### 2015-02-16 lessonplans/hmwkch6-7.tex
populating the first few homeworks

--- 2014-12-16 hmwkch6.tex
@@
 
-\vspace{2.5in}
+%\vspace{2.5in}
 
+\begin{enumerate}
 \item Which two variables have the strongest correlation? What is it?
 
-\vspace{1in}
+%\vspace{1in}
 
@@
 Write a paragraph briefly describing the associations, and discussing whether correlation is  or isn't an appropriate summary.
+\end{enumerate}
 
+\item Fit a regression model for SalePrice against LivingArea.
+
+\begin{enumerate}
+\item Write out the model. 
+\item Print the scatterplot and sketch the linear model on this plot. 
+\item What proportion of variation in house price does the living area explain?
+\item
+Predict the average sales price for a house having 1500 sq ft of living space.
+\item There is a (small) problem with the fit. Diagnose this using the residual plots, and write a paragraph on how you would correct it, and re-fit the model. 
+\end{enumerate}
 \end{enumerate}

### 2015-02-16 student-paper-competition-2014/paper.Rmd

--- 2014-12-12 student-paper-competition-2014/paper.Rmd
@@
     number_sections: yes
+  html_document: default
 bibliography: refs.bib

### 2015-02-16 lessonplans/hmwkch2.tex
hwk 2, not sure if having a folder under papers is best, or whether should be shifted back to separated branch

\documentclass{article}
\usepackage{hyperref}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\parindent 0pt
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9in \textwidth 6.5in

\begin{document}

\hspace{2in} Name: (Print) \underline{\mbox{\hspace{3.5in}}}\\

\begin{center}
{\Large STATISTICS 201 - Written Homework Chapter 2}\\[3mm]
\end{center}

\noindent
The OECD PISA Results in Focus report describes the survey as ``the world's global metric for quality, equity and efficiency in school education". The goal of the PISA survey is to assess the workforce readiness of 15-year old students. Nearly 500,000 students were tested across 65 countries and economies. Students were examined on how well they can apply the knowledge they learned in school to applications outside of school. The  reported scores range between 0-1000. Information about the students, parents, and schools is also collected. The students completed a questionnaire providing information about themselves, their homes, their schools, and a variety of psychological views regarding factors they believe affect their performance in school. School principals responded to a questionnaire covering their school system and learning experiences for their students. In some countries, parents completed a questionnaire requesting information about their perceptions regarding the school system, expectations for their child, and their involvement in their child's schooling.

You have a sample of this data for three countries. The variables are responses of the students to questions about how they perceive academics and math, and how many books they have in the household. Students taking the PISA tests were asked if their view on the following questions:

\begin{itemize} \itemsep 0in
\item I enjoy mathematics (Enjoy.Maths)
\item Math is important for future study (Math.Important.Study)
\item Math is important for getting a job (Math.Helps.Job)
\item My friends do well in math (Friends.Do.Well)
\end{itemize}

to which they could answer "Strongly disagree", "Disagree", "Agree", "Strongly agree". They also estimated the number of books in the household, 0=0-10, 1=11-25, 2=26-100, 3=101-200, 4=201-500, 5=more than 500.

\begin{enumerate}
\item Identify the Who for this data set.\\[0.5in]

\item Identify the What for this data set.\\[0.5in]

\item Use the Intro Stats app (\url{http://www.intro-stats.com}) to make a barchart of the variable Enjoy.Maths.  Print out the output and turn it in with this assignment.\\[1em]

\item Make a summary of the Enjoy.Maths variable. What proportion of these students agree or strongly agree with the statement ``I enjoy mathematics"?\\[0.5in]

\item Use the Intro Stats app to obtain a contingency table and mosaic plot of the relationship between Country (X, Factor) and Enjoy.Maths (Y, Response).  
Print contingency tables with the {\bf Count} and {\bf Column \%} values.  Print out the output and turn it in with this assignment. \\[1em]

\item Explain the meaning of the number {\bf 52} from the contingency table of counts.\\[0.5in]

\item Give the conditional distribution of ``Strongly agree'' given the Country. \\[1in]

\item Compare this conditional distribution to the overall marginal distribution of ``Strongly Agree". \\[1in]

\item Based on the contingency table and mosaic plot, does it appear there is an association between whether students report enjoying math and the country in which they attend school?  Circle the appropriate answer below.  Then given an explanation for your answer.\\[0.5em]

\hspace{1in} Yes, there is an association \hspace{1in} No, there is not an association \\[0.5em]

Explain your answer:

\end{enumerate}

\end{document}

### 2015-02-16 lessonplans/hmwkch2sol.tex
hwk 2, not sure if having a folder under papers is best, or whether should be shifted back to separated branch

\documentclass{article}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\parindent 0pt
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9in \textwidth 6.5in

\usepackage{graphicx}

\begin{document}
\begin{center}
{\Large STATISTICS 201 - Written Homework Chapter 2 Solutions}\\[3mm]
\end{center}

\noindent
This assignment is worth a total of 25 points.\\

\noindent
A study investigated the relationship between different treatments (Tree Treatment) on the lifespan of 24 trees infected with a particular disease.  The lifespan of a tree (Tree Lifespan) was categorized as follows: Trees that died in under one year, Trees that lived for between one and four years and Trees that lived longer than four years.  The data are located in the file {\bf InfectedTrees.JMP} in the Chapter 2 Learning Module in the Unit 1 content area in Blackboard.

\begin{enumerate}
\item (2 pts) Identify the Who for this data set.

The Who in this study was the 24 infected trees.

\item (2 pts) Identify the What for this data set.

The What in this study was the two variables collected for each tree: Tree Treatment and Tree Lifespan.

\item (5 pts) Use JMP to obtain the distribution of the variable Tree Lifespan.  Print out the output and turn it in with this assignment.

\centerline{\includegraphics[width=4.8in]{hmwkch2jmpoutput.pdf}}

\item (1 pt) What proportion of these trees died in under one year?

The proportion is $10/24 = 0.417$

\item (5 pts) Use JMP to obtain a contingency table and mosaic plot of the relationship between Tree Treatment (X, Factor) and Tree Lifespan (Y, Response).  
Change the contingency table so that it includes only the {\bf Count} and {\bf Row \%} values.  Print out the output and turn it in with this assignment.

See the attached JMP output.

\item (2 pts) Explain the meaning of the number {\bf 4} from the contingency table.

4 is the number of trees that received Treatment C and lived between 1 and 4 years.

%\newpage

\item (3 pts) Give the conditional distribution of Tree Lifespan given the tree received treatment B.

Of the trees that received Treatment B: 37.5\% lived for less than one year, 37.5\% lived between one and four years and 25\% lived for longer than four years.

\item (2 pts) Compare this conditional distribution to the overall marginal distribution of Tree Lifespan.

The proportion of trees that lived less than one year was slightly lower among trees that received treatment B than the overall proportion.  Also, the proportion of trees that lived longer than four years was slightly higher among the trees that received treatment B than the overall proportion.

\item (3 pts) Based on the contingency table and mosaic plot, does it appear there is an association between Tree Treatment and Tree Lifespan?  Circle the appropriate answer below.  Then given an explanation for your answer.

Yes, there is an association

Explain your answer:  The conditional distributions of Tree Lifespan given the different treatments are noticeably different, both in the contingency table and in the mosaic plot.  Since the conditional distributions are different, there is an association between the two variables.
\end{enumerate}

\end{document}

### 2015-02-21 lessonplans/hmwkch2sol.tex
removing directory that was inadvertently created

@@
 \documentclass{article}
+\usepackage{hyperref}
+\usepackage{graphicx}
 \oddsidemargin 0in \evensidemargin 0in \topmargin 0in
@@
 
-\usepackage{graphicx}
+\begin{document}
 
-\begin{document}
 \begin{center}
-{\Large STATISTICS 201 - Written Homework Chapter 2 Solutions}\\[3mm]
+{\Large STATISTICS 201 - Written Homework Chapter 2 Solution}\\[3mm]
 \end{center}
@@
 \noindent 
-This assignment is worth a total of 25 points.\\
+This assignment is worth a total of 35 points.
 
 \noindent
-A study investigated the relationship between different treatments (Tree Treatment) on the lifespan of 24 trees infected with a particular disease.
-The lifespan of a tree (Tree Lifespan) was categorized as follows: Trees that died in under one year, Trees that lived for between one and four years and Trees that lived longer than four years.
-The data are located in the file {\bf InfectedTrees.JMP} in the Chapter 2 Learning Module in the Unit 1 content area in Blackboard.
+The OECD PISA Results in Focus report describes the survey as ``the world's global metric for quality, equity and efficiency in school education".
+The goal of the PISA survey is to assess the workforce readiness of 15-year old students.
+Nearly 500,000 students were tested across 65 countries and economies.
+Students were examined on how well they can apply the knowledge they learned in school to applications outside of school.
+The  reported scores range between 0-1000.
+Information about the students, parents, and schools is also collected.
+The students completed a questionnaire providing information about themselves, their homes, their schools, and a variety of psychological views regarding factors they believe affect their performance in school.
+School principals responded to a questionnaire covering their school system and learning experiences for their students.
+In some countries, parents completed a questionnaire requesting information about their perceptions regarding the school system, expectations for their child, and their involvement in their child's schooling.
+
+You have a sample of this data for three countries.
+The variables are responses of the students to questions about how they perceive academics and math, and how many books they have in the household.
+Students taking the PISA tests were asked if their view on the following questions:
+
+\begin{itemize} \itemsep 0in
+\item I enjoy mathematics (Enjoy.Maths)
+\item Math is important for future study (Math.Important.Study)
+\item Math is important for getting a job (Math.Helps.Job)
+\item My friends do well in math (Friends.Do.Well)
+\end{itemize}
+
+to which they could answer "Strongly disagree", "Disagree", "Agree", "Strongly agree".
+They also estimated the number of books in the household, 0=0-10, 1=11-25, 2=26-100, 3=101-200, 4=201-500, 5=more than 500.
 
 \begin{enumerate}
-\item (2 pts) Identify the Who for this data set.
+\item {\em (5pts)} Identify the Who for this data set.
 
-The Who in this study was the 24 infected trees.
+{\em The students who sat the PISA tests.}
 
-\item (2 pts) Identify the What for this data set.
+\item {\em (5pts)} Identify the What for this data set.
 
-The What in this study was the two variables collected for each tree: Tree Treatment and Tree Lifespan.
+{\em Country, Enjoy.Maths, Math.Important.Study, Math.Helps.Job, Friends.Do.Well}
 
-\item (5 pts) Use JMP to obtain the distribution of the variable Tree Lifespan.  Print out the output and turn it in with this assignment.
+\item Use the Intro Stats app (\url{http://www.intro-stats.com}) to make a barchart of the variable Enjoy.Maths.  
+{\em (5pts)} Print out the output and turn it in with this assignment.
 
-\centerline{\includegraphics[width=4.8in]{hmwkch2jmpoutput.pdf}}
+\centerline{\includegraphics[width=2.5in]{barchart.pdf}}
 
-\item (1 pt) What proportion of these trees died in under one year?
+\item {\em (5pts)} Make a summary of the Enjoy.Maths variable. What proportion of these students agree or strongly agree with the statement ``I enjoy mathematics"?
 
-The proportion is $10/24 = 0.417$
+\begin{tabular}{lrr}
+Category & Count & Proportion \\\hline
+Strongly disagree & 2074 & 0.131\\
+Disagree & 5776 & 0.365 \\
+Agree & 5820 &0.368 \\
+Strongly Agree & 2161 & 0.137\\
+\end{tabular}
 
-\item (5 pts) Use JMP to obtain a contingency table and mosaic plot of the relationship between Tree Treatment (X, Factor) and Tree Lifespan (Y, Response).  
-Change the contingency table so that it includes only the {\bf Count} and {\bf Row \%} values.  Print out the output and turn it in with this assignment.
+{\em 0.50 say they agree or strongly agree.}
 
-See the attached JMP output.
+\item Use the Intro Stats app to obtain a contingency table and mosaic plot of the relationship between Country (X, Factor) and Enjoy.Maths (Y, Response).  
+{\em (5pts)} Print contingency tables with the {\bf Count} and {\bf Column \%} values.  Print out the output and turn it in with this assignment. \\
 
-\item (2 pts) Explain the meaning of the number {\bf 4} from the contingency table.
+\centerline{\includegraphics[width=5in]{contingency.pdf}}
 
-4 is the number of trees that received Treatment C and lived between 1 and 4 years.
+\centerline{\includegraphics[width=4in]{mosaic.pdf}}
 
-%\newpage
+\item {\em (2pts)} Explain the meaning of the number {\bf 52} from the contingency table of counts.
 
-\item (3 pts) Give the conditional distribution of Tree Lifespan given the tree received treatment B.
+{\em This is the number of students in Indonesia who Strongly disagreed with the statement ``I enjoy mathematics".}
 
-Of the trees that received Treatment B: 37.5\% lived for less than one year, 37.5\% lived between one and four years and 25\% lived for longer than four years.
+\item {\em (3pts)} Give the conditional distribution of ``Strongly agree'' given the Country.
 
-\item (2 pts) Compare this conditional distribution to the overall marginal distribution of Tree Lifespan.
+{\em Conditional on country the values for strongly disagree are 0.123,  0.193, 0.117.}
 
-The proportion of trees that lived less than one year was slightly lower among trees that received treatment B than the overall proportion.
-Also, the proportion of trees that lived longer than four years was slightly higher among the trees that received treatment B than the overall proportion.
+\item {\em (2pts)} Compare this conditional distribution to the overall marginal distribution of ``Strongly Agree".
 
-\item (3 pts) Based on the contingency table and mosaic plot, does it appear there is an association between Tree Treatment and Tree Lifespan?
+{\em Overall the proportions that strongly agree is 0.137. Indonesia has a bigger proportion that agree than we would expect if all countries were the same.}
+
+\item {\em (3pts)} Based on the contingency table and mosaic plot, does it appear there is an association between whether students report enjoying math and the country in which they attend school?
 Circle the appropriate answer below.
@@
 
-Yes, there is an association
+{\em Yes, there is an association}
 
-Explain your answer:  The conditional distributions of Tree Lifespan given the different treatments are noticeably different, both in the contingency table and in the mosaic plot.
-Since the conditional distributions are different, there is an association between the two variables.
+Explain your answer:
+
+{\em The proportions for the different responses are quite different from one country to another, which indicates the association.
+Students in Indonesia tend to more often report that they agree with the statement ``I enjoy mathematics", than either Colombia or the USA.
+}
+
 \end{enumerate}

### 2015-02-21 hmwkch3.tex
refresh of material

--- 2015-02-16 lessonplans/hmwkch3.tex
@@
 Describe the shape of this distribution.
-\\[2.5in]
+\\
 
-\item Make a line plot of average number of points per game by year. Describe the pattern of points relative to time. \\[2.5in]
+\item Make a line plot of average number of points per game by year. Describe the pattern of points relative to time. \\
 
@@
 \begin{enumerate}
-\item Describe the Who for this data.\\[0.35in]
-\item Use intRo to obtain the a histogram of yards thrown, with a suitable bin width.
+\item Describe the Who for this data.\\
+\item Use intRo to obtain the a histogram of yards thrown (Yards), with a suitable bin width.
 Describe the shape of the distribution of this variable.
 Print out the output and turn it in with this assignment.
-\item Give the five number summary for this variable.\\[0.2in]
-\item Calculate the range and IQR for this variable.\\[0.1in]
-\item What percentage of the quarterbacks  threw less than 366 yards? \\[0.1in]
-\item What percentage of the quarterbacks  threw more than 3396 yards?  \\[0.1in]
-\item Give the mean and standard deviation for this variable.\\[0.1in]
-\item Explain why there is a difference of approximately 70 yards in the mean and median for this variable.\\[1in]
-\item Which numerical summaries are the most appropriate for this distribution - the 5-number summary or the mean and standard deviation?
-Circle the appropriate answer below.
-Then explain your answer.
+\item Give the five number summary for this variable.\\
+\item Calculate the range and IQR for this variable.\\
+\item What percentage of the quarterbacks  threw less than 366 yards? \\
+\item What percentage of the quarterbacks  threw more than 3396 yards?  \\
+\item Give the mean and standard deviation for this variable.\\
+\item Explain why there is a difference of approximately 70 yards in the mean and median for this variable.\\
+\item Which numerical summaries are the most appropriate for this distribution - the 5-number summary or the mean and standard deviation?  Circle the appropriate answer below.  Explain your answer.
 
-\hspace{1in} 5-number summary \hspace{1in} mean and standard deviation \\[0.5em]
+\hspace{1in} 5-number summary \hspace{1in} or   \hspace{1in} mean and standard deviation \\
 
-Explain your answer:\\[0.5in]
-\item Explore transformations of this variable to make it more symmetric. Circle which is best of the choices below:
+\item Explore transformations of this variable to make it more symmetric. Which is best of the choices below:
 

### 2015-02-21 hmwkch3sol.tex
up to date with homeworks, i need to put lab assignments here too

\documentclass{article}[11pt]
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\usepackage{graphicx}

\begin{document}

\hspace{2in} Name: (Print) \underline{\mbox{\hspace{3.5in}}}\\

\begin{center}
{\Large STATISTICS 201 - Written Homework Chapter 3}\\[3mm]
\end{center}

\begin{enumerate}

\item Michael Jordan is highly renowned basketball player who played for 15 seasons in the NBA.  The following table contains the average points per game scored by Jordan in each of his 15 seasons.

\begin{center}
\begin{tabular}{|l|c||l|c||l|c|} \hline
Year & Avg Pts/Game & Year & Avg Pts/Game & Year & Avg Pts/Game 
\\ \hline
1984 & 28 & 1989 & 34 & 1995 & 30 \\
1985 & 23 & 1990 & 32 & 1996 & 30 \\
1986 & 37 & 1991 & 30 & 1997 & 29 \\
1987 & 35 & 1992 & 33 & 2001 & 23 \\
1988 & 33 & 1994 & 27 & 2002 & 20 \\ \hline
\end{tabular}
\end{center}

\begin{enumerate}
\item {\em (3 pts)} Make a stem and leaf display of the average number of points per game that Michael Jordan scored per season.  Do this by hand, use a split stem, and ensure that the leaves are sorted numerically. Describe the shape of this distribution. \\

\includegraphics[width=4in]{hwk3-stuff.pdf}

\item  {\em (2 pts)} Make a line plot of average number of points per game by year. Describe the pattern of points relative to time. \\

\includegraphics[width=4in]{hwk3-stuff2.pdf}

\end{enumerate}

\newpage
\item For this part of the assignment you will be using data collected on IA high school quarterbacks for the 2012 season.

\begin{enumerate}
\item {\em (2 pts)} Describe the Who for this data.\\

{\em The IA high school quarterbacks who played in 2012.}

\item {\em (3 pts)} Use intRo to obtain the a histogram of yards thrown (Yards), with a suitable bin width.  Describe the shape of the distribution of this variable. Print out the output and turn it in with this assignment.

\includegraphics[width=4in]{hwk3-intRo1.pdf}

{\em The distribution is right-skewed, unimodal, centered at around 500-600 yards, with a few outliers at 2500 and 3500 yards.}

\item {\em (2 pts)} Give the five number summary for this variable.\\

{\em 29, 366, 663.5, 958.5, 3396}

\item {\em (2 pts)} Calculate the range and IQR for this variable.\\

{\em Range=3396-29=3367, IQR=592.5.}

\item {\em (2 pts)} What percentage of the quarterbacks  threw less than 366 yards? \\

{\em  25\%, 366 is $Q_1$.}

\item {\em (2 pts)} What percentage of the quarterbacks  threw more than 3396 yards?  \\

{\em 0\%, None}

\item {\em (2 pts)}Give the mean and standard deviation for this variable.\\

{\em $\bar{y}=735.0, s=482.5$}

\item {\em (3 pts)} Explain why there is a difference of approximately 70 yards in the mean and median for this variable.\\

{\em The mean is larger than the median because the distribution is right-skewed, and outliers with large values.}

\item {\em (2 pts)} Which numerical summaries are the most appropriate for this distribution - the 5-number summary or the mean and standard deviation?  Explain your answer.

\hspace{1in} {\em 5-number summary}  \\

{\em Because the data is not symmetric and unimodal, the mean and standard deviation do not give a good summary of the center and spread of the numbers. }\\

\item {\em (3 pts)} Explore transformations of this variable to make it more symmetric. Which is best of the choices below:

\hspace{1in} {\em 0.5, because the other two make the distribution left-skewed.}

Print a histogram of your transformed data. Compute what the transformed value of 3000 yards would be.

\includegraphics[width=4in]{hwk3-intRo2.pdf}

{\em There is still one outlier after the transformation. This is Blake Gimbel, who was the quarterback for Marshalltown and was widely celebrated in the news as being the bets quarterback in the state. These numbers would suggest that it is a valid praise. Although, if we also considered pass success and touchdowns some other quarterbacks have better statistics.}

{\em $\sqrt{3000} = 54.8$}

\end{enumerate}

\end{enumerate}

\end{document}

### 2015-02-21 hmwkch5sol.tex
up to date with homeworks, i need to put lab assignments here too

--- 2015-02-16 lessonplans/hmwkch5sol.tex
@@
 
-%\includegraphics[width=2.5in]{hwk5jmpoutput.pdf}
+\centerline{\includegraphics[width=3in]{hmwkch5-histogram.pdf}\includegraphics[width=3in]{hmwkch5-normprob.pdf}}
+
+\includegraphics[width=6in]{hmwkch5-summary.pdf}
 
@@
 The points in the normal quantile plot generally follow the straight line in the plot.
-Due to the shape of the histogram and the straight line pattern in the normal quantile plot, it appears that the distribution of body temperatures in the sample can be modeled with the normal model.
+Due to the shape of the histogram and the straight line pattern in the normal quantile plot.
+We can also count the number of observations that fall within 1, 2, and 3 standard deviations to compare against the 68-95-99.7 rule.
+These are 90,
+123, 129 which corresponds to 69-95-99.2, extremely close to the ideal.
+
+It appears that the distribution of body temperatures in the sample can be modeled with the normal model.
 

### 2015-02-21 hmwkch6-7sol.tex
up to date with homeworks, i need to put lab assignments here too

\documentclass{article}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\usepackage{graphicx}

\begin{document}
\hspace{2in} Name: \underline{SOLUTION}\\

\begin{center}
{\large STATISTICS 201 - Written Homework Chapter 6}\\[3mm]

\end{center}

\begin{enumerate}
\item {\em (3pts)} For the Somerset neighborhood 2009 Ames sales data tabulate the correlations between the SalePrice, LotArea, LivingArea and GarageArea. (See the textbook for an example of how to do this. Be sure to round numbers appropriately.)

%\includegraphics[width=5in]{somerset-cor.png}

\begin{tabular}{l|rrrr}
 & SalePrice & LotArea & LivingArea & GarageArea \\\hline
SalePrice & 1.00 & 0.54 & 0.63 & 0.55 \\
LotArea & & 1.00 & 0.31 & 0.62 \\
LivingArea & & & 1.00 & 0.50 \\
\end{tabular}

\item {\em (2pts)} Which two variables have the strongest correlation? What is it?

{\em SalePrice and LivingArea}

\item {\em (5pts)} Make scatterplots of each of the pairs of variables (don't hand these in). Write a paragraph briefly describing the associations, and discussing whether correlation is  or isn't an appropriate summary.

{\em Based on the scatterplots, we can say that generally there is positive association between all of these variables. If the lot area is bigger then so is the living area and garage area, and also the sale price of the home. There are some patterns between some of the variables which would make correlation not the best summary of the association. For example, between sale price and lot area we can see clusters, and within these clusters the association varies a little. There are also some outliers which may affect the correlation, for example, one house that has a sale price close to \$500000, much higher than all of the others in the data set. }

\end{enumerate}

\end{document}

### 2015-03-02 hmwkch6-7sol.tex
Chapter 6-7 homework

@@
 \begin{enumerate}
-\item {\em (3pts)} For the Somerset neighborhood 2009 Ames sales data tabulate the correlations between the SalePrice, LotArea, LivingArea and GarageArea.
+\item {\em (5 pts)} For the Somerset neighborhood 2009 Ames sales data tabulate the correlations between the SalePrice, LotArea, LivingArea and GarageArea.
 (See the textbook for an example of how to do this.
@@
 
-\item {\em (2pts)} Which two variables have the strongest correlation? What is it?
+\begin{enumerate}
+\item {\em (2 pts)} Which two variables have the strongest correlation? What is it?
 
@@
 
-\item {\em (5pts)} Make scatterplots of each of the pairs of variables (don't hand these in).
+\item {\em (3 pts)} Make scatterplots of each of the pairs of variables (don't hand these in).
 Write a paragraph briefly describing the associations, and discussing whether correlation is  or isn't an appropriate summary.
@@
 
+\item Fit a regression model for SalePrice against LivingArea.
+
+\begin{enumerate}
+\item {\em (2 pts)} Write out the model.
+
+\[
+\hat{y} = 68548.0 + 96.9x
+\]
+
+\item {\em (2 pts)} Print the scatterplot and sketch the linear model on this plot.
+
+{\em Predict values for $x=1200$ gives $\hat{y}=184843.7$ and $x=2000$ gives $\hat{y}=262374.1$  respectively. Plot these two points and draw the line between. }
+
+\centerline{\includegraphics[width=4in]{hmwkch6-7-scatplot.png}}
+
+\item {\em (2 pts)} What proportion of variation in house price does the living area explain?
+
+\[
+R^2 = 0.40
+\]
+
+\item
+{\em (2 pts)} Predict the average sales price for a house having 1500 sq ft of living space.
+
+\[
+68548.0 + 96.9\times 1500 = 213917.6
+\]
+
+\item {\em (2 pts)} There is a (small) problem with the fit. Diagnose this using the residual plots, and write a paragraph on how you would correct it, and re-fit the model.
+
+{\em There is a large house (3000 sq ft) that sold for a relatively low price of  approximately \$285,000.
+It is clear from the residual plot that it pulls the regression line away from the rest of the points, so that the trend is not accurately captured by the model.
+It would be a good idea to remove this house from the data and re-fit the model, to examine the influence of this point.
+If the slope and intercept of the model changes by a lot then this house should be considered to have some special qualities which differ from the other houses in the data base, and removed from the analysis.}
+
+\end{enumerate}
+
+\end{enumerate}
+
 \end{document}

### 2015-07-15 jsm-2015/eric/index.Rmd
Cut out parts that wont be said

--- 2015-01-11 presentations/departmental-seminar/index.Rmd
@@
 
-# First look | http://intro-stats.com
-
 # Classroom
@@
 ## `intRo` in the wild
- - `intRo` will be used this semester in STAT 201
-      - Homework assignments and labs will use `intRo` to emphasize class concepts
+ - `intRo` was used last semester at Iowa State University in STAT 201
+      - Homework assignments and labs used `intRo` to emphasize class concepts
  - `intRo` is a supplement, not the focus of the class
@@
 
-# Our turn
+# Live Demo
 
@@
 
-## Consistent UI (Cont'd)
- - Large, easy to click icons in the page header to help students find
-exactly what they need easily
- - Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code
- - Documentation website that is consistent with
-the interface of the application; covers all default modules and makes learning to use `intRo` even more painless
-
 ## Modularity
 <img src="images/app_creation_modules.png" height="500"></img>
-
-## Modularity (Cont'd)
-The interface is created with the following statement
-
-```{r, tidy=FALSE, eval=FALSE}
-## Source ui
-module_info <- read.table("modules/modules.txt", header = TRUE, 
-                          sep=",")
-sapply(file.path("modules", dir("modules")[dir("modules") 
-                      != "modules.txt"], "ui.R"), source)
-## Create the UI
-shinyUI(navbarPage("intRo", theme = "bootstrap.min.css",
-               tabPanel(title="", icon=icon("home"),
-                        fluidRow(
-                        ### <b>
-                          do.call(navlistPanel, 
-                              c(list(id = "side-nav", 
-                              widths = c(2, 10)), mylist))
-                        ### </b>
-                        )
-               ), ...
-))
-```
-
-## Modularity (Cont'd)
-The server functions are dynamically generated using a similar method
-
-```{r, eval=FALSE, tidy=FALSE}
-shinyServer(function(input, output, session) {
-  ## Module info
-  module_info <- read.table("modules/modules.txt", 
-                            header = TRUE, sep=",")
-  ## Modules
-  types <- c("helper.R", "static.R", "observe.R", 
-             "reactive.R", "output.R")
-  modules_tosource <- file.path("modules", 
-                                apply(expand.grid(
-                                module_info$module, types)
-                                , 1, paste, collapse = "/"))
-  ## Source the modules
-  for (mod in modules_tosource) {
-    ### <b>
-    source(mod, local = TRUE) 
-    ### </b>
-  }
-})
-```
-
-# Challenges
-
-## Tradeoffs
-We encountered a number of challenges in striving to strike the balance of functionality and ease-of-use:
-
- - Storing the user's results
- - Presenting only relevant options to the user
- - Producing clean and executable R code
-
-## Storing results
- - Users must be able to print results, but only the results they want
- - Reactivity makes this somewhat challenging
-
-**Solution**: Store Buttons
-
-<img src="images/store.png" height="300"></img>
-
-## Storing results (Cont'd)
- - Each module contains, at the bottom, a Store button
- - When Store is pressed, code is stored in code panel
-     - Additional challenge: All options selected must also be stored!
- - When user goes to print results, only this particular code is executed
-
-## Presenting choices
- - `intRo` is flexible, but must also be simple for the introductory student
- - Despite increasing the complexity of the code, we decided to only show options appropriate for the given selections.
-
-**Example**:
-
-<img src="images/graphical.png" height="300">
-
-## Presenting choices (Cont'd)
-<img src="images/graphical_table.png">
-
-## Producing clean code
-Code generated must be:
-
- - Executable on the server
- - Executable on the user's machine
- - Clean
-
-## Producing clean code (Cont'd)
-But how do we handle uploaded data?
-
-```{r, eval=FALSE}
-file.choose <- function () {
-    return(input$data_own[,"datapath"])
-}
-```
-
-Now we can provide to the user *and* the server:
-
-```{r, eval=FALSE}
-intro.data <- read.csv(file.choose())
-```
 

### 2015-08-10 presentations/jsm-2015/andee/index.Rmd
Presentations added to gh-pages

--- 2015-01-11 presentations/departmental-seminar/index.Rmd
@@
 author: "Andee Kaplan & Eric Hare"
-date: "January 12, 2015"
+date: "August 11, 2015"
 output: ioslides_presentation
@@
 # First look | http://intro-stats.com
-
-# Classroom
-
-## `intRo` in the wild
- - `intRo` will be used this semester in STAT 201
-      - Homework assignments and labs will use `intRo` to emphasize class concepts
- - `intRo` is a supplement, not the focus of the class
-
-## Homework 1
- - *Data*: OECD PISA Results in Focus report to assess the workforce readiness of 15-year old students
- - *Goal*: Explore graphically and numerically how the students perceive academics and math
- - `intRo` *Tasks*:
-      1. Make a barchart of the variable Enjoy.Maths.
-      1. Make a summary of the Enjoy.Maths variable. What proportion of these students agree or strongly agree with the statement "I enjoy mathematics"?
-      2. Obtain a contingency table and mosaic plot of the relationship between Country and Enjoy.Maths. Print contingency tables with the **Count** and **Column Proportion** values.
-
-# Our turn
 
@@
 
- - Dozens of students at accessing app at once
+ - Dozens of students accessing app at once
  - Simultaneous larger computations like printing results

### 2015-08-10 presentations/jsm-2015/eric/index.Rmd
Presentations added to gh-pages

--- 2015-07-15 jsm-2015/eric/index.Rmd
@@
 author: "Andee Kaplan & Eric Hare"
-date: "January 12, 2015"
+date: "August 12, 2015"
 output: ioslides_presentation

### 2015-08-11 jsm-2015/andee/index.Rmd
small changes

--- 2015-08-10 presentations/jsm-2015/andee/index.Rmd
@@
 subtitle: "Statistical Analysis Software for Teaching"
-author: "Andee Kaplan & Eric Hare"
-date: "August 11, 2015"
+author: "Andee Kaplan & Eric Hare <br/> Iowa State University"
+date: "August 11, 2015 <br> http://j.mp/intRo-JSM2015"
 output: ioslides_presentation
@@
 
-## Special thanks
-Dr. Cook, for guiding many of the features and interface decisions, and helping test the functionality in preparation for STAT 201
-
 # Any questions? | Thank you!

### 2015-08-12 jsm-2015/eric/index.Rmd
Blank space
Some small changes

@@
 subtitle: "Statistical Analysis Software for Teaching"
-author: "Andee Kaplan & Eric Hare"
-date: "January 12, 2015"
+author: "Eric Hare & Andee Kaplan"
+date: "August 12, 2015"
 output: ioslides_presentation
@@
 
-## Do we really need another statistical software package?
-Short answer: **yes**
-
- - `R` is great, but requires students to have some knowledge/interest in programming
- - `JMP` is powerful, but too big
-      - Licenses and inconsistent UI
- - New tools recently released to spark an interest in `R`
-      - `Swirl` and DataCamp teach `R` programming
-      - Project MOSAIC facilitates learning, but assumes knowledge of `R`
- - Want students to focus on data analysis rather than fight with software
-
 ## What is `intRo`?
@@
  - Assists in the learning of statistics rather than acting as a stand-alone deliverer of statistics education
+
+## Do we really need another statistical software package?
+Short answer: **yes**
+
+ - `R` is great, but requires students to have some knowledge/interest in programming
+ - `JMP` is powerful, but lots of functionality for intro stats
+      - Software licensing issues
+ - New tools recently released to spark an interest in `R`
+      - `Swirl` and DataCamp teach `R` programming
+      - Project MOSAIC facilitates learning, but assumes knowledge of `R`
+ - Want students to focus on data analysis rather than fight with software
 
@@
 # Live Demo
+Feel free to try intRo yourself at
+
+http://www.intro-stats.com
 

### 2015-11-09 jcgs-paper-2016/paper.Rmd
Start on JCGS Paper

--- 2015-02-16 student-paper-competition-2014/paper.Rmd
@@
 ---
-title: "Introducing statistics with \texttt{intRo}"
-author: "Eric Hare & Andee Kaplan"
-date: "December 14, 2014"
+title: Introducing statistics with \texttt{intRo}
+author: Eric Hare & Andee Kaplan
 output:
@@
 bibliography: refs.bib
-geometry: margin=0.8in
 ---
-\thispagestyle{empty}
 
@@
 \end{abstract}
-
-\twocolumn
-\setcounter{page}{1}
 
@@
 
+# Modules
+Within the data category there are two modules, sources and transformation.
+The data sources module allows the user to select or upload a dataset.
+Several datasets are pre-included, which can be selected via a drop-down menu.
+The user can check a box to upload a dataset, and then select the CSV file from her computer.
+Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
+From this module the user can select a subset of the data either by value or random selection.
+The dataset selected in the data module will be made available to all other modules of \texttt{intRo}.
+The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
+Additionally, numeric variable can be transformed using a power transformation.
+The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.
+
+The graphical summaries module provides several common graphical displays of variables in the dataset.
+First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
+The variable available to choose will only be those compatible with plot type selected.
+The plot visible on the right will automatically update when these values are selected.
+The numeric summaries module is another method of computing summaries of the data.
+The user is presented with a list of each variable from their dataset.
+The user can select one or more different variables, and will be presented with a table of some summary statistics.
+The summary statistics displayed vary depending on the type of the variables.
+For instance, for a numeric variable, a five number summary will be displayed.
+Character or factor variables will display counts of each unique value of that particular variable.
+Additionally, the user has the ability to display summaries by a grouping variable.
+
+Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester.
+The contingency table module allows the user to compute a contingency table for two categorical variables.
+The results can either be displayed as counts, row percentages, column percentages, or total percentages.
+The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
+The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
+Results of the regression will be displayed on the right.
+A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
+A user also has the ability to save residuals and predicted values to their dataset for further analysis.
+The T test module allows the user to perform either a one variable or two variable Student's T test.
+The user can select configuration options, including the direction of the test and the confidence level of the test.
+The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
+
 # In the classroom

### 2015-11-12 jcgs-paper-2016/paper.Rmd
TODO items for the paper

@@
 # Introduction
+
+TODO: Settle on licensing for intRo
 
@@
 # What is \texttt{intRo}?
+
+TODO: Talk about the package (briefly)
+
 \texttt{intRo} is a web-based tool to accompany an introductory statistics class.
@@
 
-# Modules
-Within the data category there are two modules, sources and transformation.
-The data sources module allows the user to select or upload a dataset.
-Several datasets are pre-included, which can be selected via a drop-down menu.
-The user can check a box to upload a dataset, and then select the CSV file from her computer.
-Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
-From this module the user can select a subset of the data either by value or random selection.
-The dataset selected in the data module will be made available to all other modules of \texttt{intRo}.
-The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
-Additionally, numeric variable can be transformed using a power transformation.
-The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.
-
-The graphical summaries module provides several common graphical displays of variables in the dataset.
-First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
-The variable available to choose will only be those compatible with plot type selected.
-The plot visible on the right will automatically update when these values are selected.
-The numeric summaries module is another method of computing summaries of the data.
-The user is presented with a list of each variable from their dataset.
-The user can select one or more different variables, and will be presented with a table of some summary statistics.
-The summary statistics displayed vary depending on the type of the variables.
-For instance, for a numeric variable, a five number summary will be displayed.
-Character or factor variables will display counts of each unique value of that particular variable.
-Additionally, the user has the ability to display summaries by a grouping variable.
-
-Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester.
-The contingency table module allows the user to compute a contingency table for two categorical variables.
-The results can either be displayed as counts, row percentages, column percentages, or total percentages.
-The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
-The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
-Results of the regression will be displayed on the right.
-A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
-A user also has the ability to save residuals and predicted values to their dataset for further analysis.
-The T test module allows the user to perform either a one variable or two variable Student's T test.
-The user can select configuration options, including the direction of the test and the confidence level of the test.
-The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
-
 # In the classroom
@@
 
-# Reactive design choices
+# Using intRo
+
+# Details
+
+## Modules
+TODO: Condense this into either a single paragraph, list, and/or table - Move to Extensible section
+TODO: Rewrite this as the details behind the modularity. Add information on writing your own modules
+
+Within the data category there are two modules, sources and transformation.
+The data sources module allows the user to select or upload a dataset.
+Several datasets are pre-included, which can be selected via a drop-down menu.
+The user can check a box to upload a dataset, and then select the CSV file from her computer.
+Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
+From this module the user can select a subset of the data either by value or random selection.
+The dataset selected in the data module will be made available to all other modules of \texttt{intRo}.
+The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
+Additionally, numeric variable can be transformed using a power transformation.
+The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.
+
+The graphical summaries module provides several common graphical displays of variables in the dataset.
+First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
+The variable available to choose will only be those compatible with plot type selected.
+The plot visible on the right will automatically update when these values are selected.
+The numeric summaries module is another method of computing summaries of the data.
+The user is presented with a list of each variable from their dataset.
+The user can select one or more different variables, and will be presented with a table of some summary statistics.
+The summary statistics displayed vary depending on the type of the variables.
+For instance, for a numeric variable, a five number summary will be displayed.
+Character or factor variables will display counts of each unique value of that particular variable.
+Additionally, the user has the ability to display summaries by a grouping variable.
+
+Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester.
+The contingency table module allows the user to compute a contingency table for two categorical variables.
+The results can either be displayed as counts, row percentages, column percentages, or total percentages.
+The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
+The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
+Results of the regression will be displayed on the right.
+A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
+A user also has the ability to save residuals and predicted values to their dataset for further analysis.
+The T test module allows the user to perform either a one variable or two variable Student's T test.
+The user can select configuration options, including the direction of the test and the confidence level of the test.
+The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
+
+## Reactive design choices
 
@@
 This is an example of a decision made by the developers of \texttt{Shiny} to limit functionality in favor of usability of their package.
+
+## Reproducibility
 

### 2015-11-23 jcgs-paper-2016/paper.Rmd
Blurb on intRo package

@@
 
-TODO: Settle on licensing for intRo
-
-The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly beenan important development for the scientific community.
+The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been an important development for the scientific community.
 However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student.
@@
 
-TODO: Talk about the package (briefly)
-
 \texttt{intRo} is a web-based tool to accompany an introductory statistics class.
 It is meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class.
+An accompanying R package, titled \texttt{intRo} and available on GitHub, assists in the downloading, running, and deploying of \texttt{intRo} instances.
 

### 2015-11-23 jcgs-paper-2016/paper.Rmd
Section on using intRo package

@@
 # Using intRo
+\texttt{intRo} can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package \texttt{intRo}.
+Currently, the package is only available and GitHub, and can be installed using the \texttt{devtools} package as follows:
+
+```{r, eval=FALSE}
+devtools::install_github("gammarama/intRo")
+```
+
+After installing the \texttt{intRo} package, the first function one should call is `download_intRo`.
+`download_intRo` takes as an argument a directory in which to store the application.
+By default, it selects the working directory of the R session.
+This function clones the application branch of the \texttt{intRo} repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
+
+Running `download_intRo` will produce an \texttt{intRo} folder in the specified folder.
+It can then be ran as any \texttt{Shiny} application, using \texttt{Shiny}'s `runApp` command.
+However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
+`run_intRo` takes as argument the path to the folder containing the \texttt{intRo} application.
+It also takes several more optional arguments:
+
+* `enabled_modules`: A character vector containing the modules to enable
+* `theme`: A string representing a \texttt{shinythemes} theme to use
+* `...`: Additional arguments passed to \texttt{Shiny}'s `runApp` function
+
+The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download \texttt{intRo} to my working directory, execute an \texttt{intRo} session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
+The series of calls to do so would be as follows:
+
+```{r, eval=FALSE}
+download_intRo()
+run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), theme = "cerulean")
+```
+
+Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
+
+If the intent is to use a specific instance of \texttt{intRo} where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of \texttt{intRo} to a publicly accessible URL.
+The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the \texttt{shinyapps} package.
+Once the shinyapps package is installed and configured, `deploy_intRo` will upload \texttt{intRo} as an application on the user's account.
+The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
+It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
+It also takes `...` as additional arguments to be passed into the `deployApp` routine.
+For example, if we wished to deploy the instance of \texttt{intRo} we ran previously, we would call it like so:
+
+```{r, eval=FALSE}
+deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), theme = "cerulean")
+```
+
+Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
 

### 2015-11-23 jcgs-paper-2016/paper.Rmd
Move around the modularity stuff

@@
 On the front end, user interaction with \texttt{intRo} is split into bitesize chunks that we call modules.
-This is true of the back end as well.
+Modules are self-contained pieces of functionality which implement common statistical procedures that are often used in introductory statistics classes.
+
+Within the data category there are two modules, sources and transformation.
+The data sources module allows the user to select or upload a dataset.
+Several datasets are pre-included, which can be selected via a drop-down menu.
+The user can check a box to upload a dataset, and then select the CSV file from her computer.
+Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
+From this module the user can select a subset of the data either by value or random selection.
+The dataset selected in the data module will be made available to all other modules of \texttt{intRo}.
+The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
+Additionally, numeric variable can be transformed using a power transformation.
+The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.
+
+The graphical summaries module provides several common graphical displays of variables in the dataset.
+First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
+The variable available to choose will only be those compatible with plot type selected.
+The plot visible on the right will automatically update when these values are selected.
+The numeric summaries module is another method of computing summaries of the data.
+The user is presented with a list of each variable from their dataset.
+The user can select one or more different variables, and will be presented with a table of some summary statistics.
+The summary statistics displayed vary depending on the type of the variables.
+For instance, for a numeric variable, a five number summary will be displayed.
+Character or factor variables will display counts of each unique value of that particular variable.
+Additionally, the user has the ability to display summaries by a grouping variable.
+
+Within the set of statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester.
+The contingency table module allows the user to compute a contingency table for two categorical variables.
+The results can either be displayed as counts, row percentages, column percentages, or total percentages.
+The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
+The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
+Results of the regression will be displayed on the right.
+A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
+A user also has the ability to save residuals and predicted values to their dataset for further analysis.
+The T test module allows the user to perform either a one variable or two variable Student's T test.
+The user can select configuration options, including the direction of the test and the confidence level of the test.
+The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
+
 Each module is a self contained set of \texttt{R} code that is dynamically added to the application at run time.
@@
 ## Modules
-TODO: Condense this into either a single paragraph, list, and/or table - Move to Extensible section
 TODO: Rewrite this as the details behind the modularity. Add information on writing your own modules
-
-Within the data category there are two modules, sources and transformation.
-The data sources module allows the user to select or upload a dataset.
-Several datasets are pre-included, which can be selected via a drop-down menu.
-The user can check a box to upload a dataset, and then select the CSV file from her computer.
-Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
-From this module the user can select a subset of the data either by value or random selection.
-The dataset selected in the data module will be made available to all other modules of \texttt{intRo}.
-The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
-Additionally, numeric variable can be transformed using a power transformation.
-The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.
-
-The graphical summaries module provides several common graphical displays of variables in the dataset.
-First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
-The variable available to choose will only be those compatible with plot type selected.
-The plot visible on the right will automatically update when these values are selected.
-The numeric summaries module is another method of computing summaries of the data.
-The user is presented with a list of each variable from their dataset.
-The user can select one or more different variables, and will be presented with a table of some summary statistics.
-The summary statistics displayed vary depending on the type of the variables.
-For instance, for a numeric variable, a five number summary will be displayed.
-Character or factor variables will display counts of each unique value of that particular variable.
-Additionally, the user has the ability to display summaries by a grouping variable.
-
-Within the set statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester.
-The contingency table module allows the user to compute a contingency table for two categorical variables.
-The results can either be displayed as counts, row percentages, column percentages, or total percentages.
-The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
-The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
-Results of the regression will be displayed on the right.
-A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
-A user also has the ability to save residuals and predicted values to their dataset for further analysis.
-The T test module allows the user to perform either a one variable or two variable Student's T test.
-The user can select configuration options, including the direction of the test and the confidence level of the test.
-The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
 

### 2015-11-23 jcgs-paper-2016/paper.Rmd
Discuss reproducibility more

@@
 
+## Reproducibility
+Though they can make common actions quicker and interacting with software much simplier, many Graphical User Interfaces for scientific software packages come with a major drawback.
+That is, the actions taken by the user are typically not reproducible.
+Code scripting, on the other hand, allows us to share a file containing code, and another user can execute the code to reproduce the same results.
+In terms of the scientific process, where results can and should be scrutinized by other parties, the utility of GUIs are greatly limited.
+
+We have designed \texttt{intRo} modules to reproduce actions taken by the user so that they can obtain \texttt{R} code associated with each module interacted with.
+When a user chooses to store the results associated with a particular module, R code yielding the visible results is automatically added to the code window at the bottom.
+This framework yields several advantages:
+
+1.
+This eases a user who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
+Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many inexperienced users may have.
+
+2. An analysis created by \texttt{intRo} can be reproduced in an R session to easily assess and extend the results.
+
+3. "Printing" the results of an \texttt{intRo} analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
+
+We designed the modules in such a way that the code printed to the console is the same as the code executed by the server, which makes the process of writing an \texttt{intRo} module significantly more seamless.
+The workhorse function for this reproducibility framework is called `interpolate`, which is based on a function proposed by Hadley Wickham[ref].
+The idea of interpolation is that the call can be captured, printed, and evaluated, with all the appropriate arguments substituted into the call automatically.
+This means that depending on the options selected in \texttt{intRo}, the corresponding arguments to functions are appropriately set for reproduction in a standard R session.
+
+```{r, eval=FALSE}
+interpolate <- function(code, ..., mydir, `_env` = parent.frame(), file = "code_All.R", 
+                        append = FALSE, save_result = FALSE, eval = TRUE) {
+    stopifnot(inherits(code, "formula"), length(code) == 2)
+
+    args <- lapply(list(...), as_call)
+    expr <- methods::substituteDirect(as_call(code), args)
+
+    cat(paste0(as.character(expr)[2], "\n"), file = file.path(mydir, file), append = append)
+
+    if (save_result) cat(paste0(paste(readLines(file.path(mydir, file)), collapse = "\n"),
+                                "\n"), file = file.path(mydir, "code_All.R"), append = TRUE)
+    if (eval) eval(expr, `_env`)
+}
+```
+
 ## Reactive design choices
@@
 This is an example of a decision made by the developers of \texttt{Shiny} to limit functionality in favor of usability of their package.
-
-## Reproducibility
 

### 2015-11-23 jcgs-paper-2016/paper.Rmd
jcgs template

--- 2015-11-23 jcgs-paper-2016/paper.Rmd
@@
 ---
-title: Introducing statistics with \texttt{intRo}
-author: Eric Hare & Andee Kaplan
+title: "Introducing statistics with `intRo`"
+date: "November 23, 2015"
+author:
+  - name: Eric Hare
+    affiliation: Iowa State University
+    department: Department of Statistics
+  - name: Andee Kaplan
+    affiliation: Iowa State University
+    department: Department of Statistics
 output:
   pdf_document:
+    fig_caption: yes
     includes:
@@
     number_sections: yes
-  html_document: default
+    template: jcgs_article.tex
+abstract: |
+  `intRo` is a modern web-based application for performing basic data analysis and statistical routines.
+Leveraging the power of `R` and `Shiny`, `intRo` implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician.
+This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as `Excel` and `JMP`.
+`intRo` is currently deployed at the URL \url{http://www.intro-stats.com}.
+Within this paper, we introduce the application and explore the design decisions underlying `intRo`, as well as highlight some challenges and advantages of reactive programming.
+keywords: blah, blah blah
 bibliography: refs.bib
+blind: false
 ---
-
-\begin{abstract}
-\texttt{intRo} is a modern web-based application for performing basic data analysis and statistical routines.
-Leveraging the power of \texttt{R} and \texttt{Shiny}, \texttt{intRo} implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician.
-This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as \texttt{Excel} and \texttt{JMP}.
-\texttt{intRo} is currently deployed at the URL \url{http://www.intro-stats.com}.
-Within this paper, we introduce the application and explore the design decisions underlying \texttt{intRo}, as well as highlight some challenges and advantages of reactive programming.
-\end{abstract}
 
@@
 
-The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been an important development for the scientific community.
+TODO: Settle on licensing for intRo
+
+The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly beenan important development for the scientific community.
 However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student.
@@
 
+TODO: Talk about the package (briefly)
+
 \texttt{intRo} is a web-based tool to accompany an introductory statistics class.
 It is meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class.
-An accompanying R package, titled \texttt{intRo} and available on GitHub, assists in the downloading, running, and deploying of \texttt{intRo} instances.
 
@@
 # Using intRo
-\texttt{intRo} can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package \texttt{intRo}.
-Currently, the package is only available and GitHub, and can be installed using the \texttt{devtools} package as follows:
-
-```{r, eval=FALSE}
-devtools::install_github("gammarama/intRo")
-```
-
-After installing the \texttt{intRo} package, the first function one should call is `download_intRo`.
-`download_intRo` takes as an argument a directory in which to store the application.
-By default, it selects the working directory of the R session.
-This function clones the application branch of the \texttt{intRo} repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
-
-Running `download_intRo` will produce an \texttt{intRo} folder in the specified folder.
-It can then be ran as any \texttt{Shiny} application, using \texttt{Shiny}'s `runApp` command.
-However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
-`run_intRo` takes as argument the path to the folder containing the \texttt{intRo} application.
-It also takes several more optional arguments:
-
-* `enabled_modules`: A character vector containing the modules to enable
-* `theme`: A string representing a \texttt{shinythemes} theme to use
-* `...`: Additional arguments passed to \texttt{Shiny}'s `runApp` function
-
-The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download \texttt{intRo} to my working directory, execute an \texttt{intRo} session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
-The series of calls to do so would be as follows:
-
-```{r, eval=FALSE}
-download_intRo()
-run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), theme = "cerulean")
-```
-
-Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
-
-If the intent is to use a specific instance of \texttt{intRo} where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of \texttt{intRo} to a publicly accessible URL.
-The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the \texttt{shinyapps} package.
-Once the shinyapps package is installed and configured, `deploy_intRo` will upload \texttt{intRo} as an application on the user's account.
-The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
-It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
-It also takes `...` as additional arguments to be passed into the `deployApp` routine.
-For example, if we wished to deploy the instance of \texttt{intRo} we ran previously, we would call it like so:
-
-```{r, eval=FALSE}
-deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), theme = "cerulean")
-```
-
-Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
 

### 2016-01-08 jcgs-paper-2016/paper.Rmd
Start on the modules section

@@
 download_intRo()
-run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), theme = "cerulean")
+run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+          theme = "cerulean")
 ```
@@
 ```{r, eval=FALSE}
-deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), theme = "cerulean")
+deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+             theme = "cerulean")
 ```
@@
 ## Modules
-TODO: Rewrite this as the details behind the modularity. Add information on writing your own modules
+An \texttt{intRo} module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality.
+\texttt{intRo} modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the \texttt{intRo} framework.
+In this section, we detail the structure of these modules, and the process of creating and deploying new modules.
+
+A module consists of the following scripts:
+
+* *helper.R* - R code that performs some statistical analysis or transformation. This would typically be in the form of a function, and similar to any standard R script.
+* *libraries.R* - Code to load any libraries which are not part of core R.
+* *observe.R* - Shiny observer code typically used to update choices of an input box.
+* *output.R* - Shiny output code defining the results of the analysis that should be displayed to the user.
+* *reactive.R* - Shiny reactives, typically containing data that depend on inputs.
+* *ui.R* - Shiny user interface definition, including the placement of the inputs and outputs.
+
+The modules provided with \texttt{intRo} are contained in the modules folder.
+The top level directory in the modules folder defines the category of the module (currently `data`, `summaries`, or `statistics`).
+Within each of these categories is a folder named according to the name of the module.
+This folder houses the previously defined scripts.
+As an example, we will walk through the contents of the `contingency` module.
+
+Since the `contingency` module performs a statistical analysis, it is part of the statistics category, and hence can be found in the \texttt{intRo} repository at **modules/statistics/contingency**.
+Let's first examine *helper.R*:
+
+```{r, eval=FALSE}
+cont.table <- function(intro.data, x, y, type, digits) {  
+    interpolate(
+        ~ (my.tbl <- cbind(rbind(table(df$y, df$x), 
+                      Total = colSums(table(df$y, df$x))), 
+                Total = c(rowSums(table(df$y, df$x))[-(length(df$x))], sum(table(df$y, df$x))))),
+        df = quote(intro.data),
+        x = x,
+        y = y,
+        mydir = userdir,
+        `_env` = environment(),
+        file = "code_contingency.R"
+    )
+
+    if (type == "totalpercs") interpolate(~(my.tbl <- my.tbl / sum(table(df$y, df$x))), 
+                                          df = quote(intro.data), x = x, y = y,  
+                                          mydir = userdir, `_env` = environment(), 
+                                          file = "code_contingency.R", append = TRUE)
+    if (type == "rowpercs") interpolate(~(my.tbl <- t(apply(my.tbl[-nrow(my.tbl),], 1, 
+                                                            function(row){row / row[ncol(my.tbl)]}))),  
+                                        mydir = userdir,`_env` = environment(), 
+                                        file = "code_contingency.R", append = TRUE)
+    if (type == "columnpercs") interpolate(~(my.tbl <- apply(my.tbl[,-ncol(my.tbl)], 2, 
+                                                             function(col){col / col[nrow(my.tbl)]})),  
+                                           mydir = userdir, `_env` = environment(), 
+                                           file = "code_contingency.R", append = TRUE)
+
+    if (type %in% c("totalpercs", "rowpercs", "columnpercs")) interpolate(~(new.digits <- digits), 
+                                                                          digits = digits, mydir = userdir,
+                                                                          `_env` = environment(), 
+                                                                          file = "code_contingency.R", 
+                                                                          append = TRUE)
+    else interpolate(~(new.digits <- 0),  mydir = userdir, `_env` = environment(), 
+                     file = "code_contingency.R", append = TRUE)
+
+    interpolate(~(format(round(my.tbl, digits = new.digits))), mydir = userdir, `_env` = environment(), 
+                file = "code_contingency.R", append = TRUE)
+}
+```
 

### 2016-01-08 jcgs-paper-2016/paper.Rmd
Continue drafting modules section

@@
                       Total = colSums(table(df$y, df$x))), 
-                Total = c(rowSums(table(df$y, df$x))[-(length(df$x))], sum(table(df$y, df$x))))),
+                Total = c(rowSums(table(df$y, df$x))[-(length(df$x))], 
+                          sum(table(df$y, df$x))))),
         df = quote(intro.data),
@@
 
-    if (type == "totalpercs") interpolate(~(my.tbl <- my.tbl / sum(table(df$y, df$x))), 
+    if (type == "totalpercs") interpolate(~(my.tbl <- my.tbl / 
+                                          sum(table(df$y, df$x))), 
                                           df = quote(intro.data), x = x, y = y,  
-                                          mydir = userdir, `_env` = environment(), 
-                                          file = "code_contingency.R", append = TRUE)
-    if (type == "rowpercs") interpolate(~(my.tbl <- t(apply(my.tbl[-nrow(my.tbl),], 1, 
-                                                            function(row){row / row[ncol(my.tbl)]}))),  
-                                        mydir = userdir,`_env` = environment(), 
-                                        file = "code_contingency.R", append = TRUE)
-    if (type == "columnpercs") interpolate(~(my.tbl <- apply(my.tbl[,-ncol(my.tbl)], 2, 
-                                                             function(col){col / col[nrow(my.tbl)]})),  
-                                           mydir = userdir, `_env` = environment(), 
-                                           file = "code_contingency.R", append = TRUE)
+                                          mydir = userdir, 
+                                          `_env` = environment(), 
+                                          file = "code_contingency.R", 
+                                          append = TRUE)
+    if (type == "rowpercs") interpolate(~(my.tbl <- t(apply(
+        my.tbl[-nrow(my.tbl),], 1, function(row){row / row[ncol(my.tbl)]}))),  
+                                        mydir = userdir,
+                                        `_env` = environment(), 
+                                        file = "code_contingency.R", 
+                                        append = TRUE)
+    if (type == "columnpercs") interpolate(~(my.tbl <- apply(
+        my.tbl[,-ncol(my.tbl)], 2, function(col){col / col[nrow(my.tbl)]})),  
+                                           mydir = userdir, 
+                                           `_env` = environment(), 
+                                           file = "code_contingency.R", 
+                                           append = TRUE)
 
-    if (type %in% c("totalpercs", "rowpercs", "columnpercs")) interpolate(~(new.digits <- digits), 
-                                                                          digits = digits, mydir = userdir,
+    if (type %in% c("totalpercs", "rowpercs", "columnpercs")) 
+        interpolate(~(new.digits <- digits), 
+                    digits = digits, 
+                    mydir = userdir,
                     `_env` = environment(), 
@@
 
-    interpolate(~(format(round(my.tbl, digits = new.digits))), mydir = userdir, `_env` = environment(), 
-                file = "code_contingency.R", append = TRUE)
+    interpolate(~(format(round(my.tbl, digits = new.digits))), 
+                mydir = userdir, 
+                `_env` = environment(), 
+                file = "code_contingency.R",
+                append = TRUE)
 }
 ```
+
+This script is most immediately similar to standard R code.
+In this case, a function `cont.table` is created which, depending on the values of the parameters, ultimately returns a contingency table.
+One important difference from a typical R script is that each call in the script is wrapped in a function called `interpolate`.
+`interpolate` both executes the given R code on the server, and also writes the code executed to the script window at the bottom of \texttt{intRo}.
+More details on the reasons and choices behind `interpolate` can be found in the next section.
+
+Because all the code needed to implement a contingency table is found in the `base` and `stats` package, the `libraries.R` file is empty for the `contingency` module.
+`observe.R`, which defines the Shiny observers needed, is reproduced below:
+
+```{r, eval=FALSE}
+observe({
+    xselect <- ifelse(checkVariable(intro.data(), input$xcont),
+                      input$xcont, 
+                      intro.categoricnames()[1])
+    yselect <- ifelse(checkVariable(intro.data(), input$ycont), 
+                      input$ycont, 
+                      intro.categoricnames()[2])
+
+    updateSelectizeInput(session, 
+                         "xcont", 
+                         choices = setdiff(intro.categoricnames(), yselect), 
+                         selected = xselect)
+    updateSelectizeInput(session, 
+                         "ycont", 
+                         choices = setdiff(intro.categoricnames(), xselect), 
+                         selected = yselect)
+})
+
+observeEvent(input$store_contingency, {
+    cat(paste0("\n\n", paste(readLines(
+        file.path(userdir, "code_contingency.R")), 
+        collapse = "\n")), 
+        file = file.path(userdir, "code_All.R"), 
+        append = TRUE)
+})
+```
+
+Shiny observers are created to ensure that the choices of variable for the `contingency` module are only categorical variables.
+This is accomplished by utilizing the global reactive `intro.categoricnames()`, which returns a character vector containing the variables in the current dataset that are categorical.
+Finally, there is an event observer to store code generated from the module into the overall code script upon clicking the store button.
+The presence of this observer code and the definition of the button in the user interface are enforced, and must be present in any \texttt{intRo} module.
+
+The `output.R` and `reactive.R` code is very simple:
+
+```{r, eval=FALSE}
+output$conttable <- renderTable({
+    return(intro.contingency())
+})
+```
+
+```{r, eval=FALSE}
+intro.contingency <- reactive({
+    if (is.null(intro.data())) return(NULL)
+    if (input$xcont == input$ycont || 
+        length(categoricNames(intro.data())) < 2) return(NULL)
+
+    my.conttable <- cont.table(intro.data(), 
+                               input$xcont, input$ycont, 
+                               input$conttype, input$contdigits)
+
+    return(my.conttable)
+})
+```
+
+`intro.contingency` is defined as a reactive object.
+Some small error checks are performed at the beginning to ensure that the data is present and available, and that two different categorical variables are selected.
+Assuming those conditions hold, a call to the previously defined `helper.R` function is made with the appropriate parameters, and the results are returned.
+The `output.R` script then simply uses Shiny's `renderTable` function to display the resulting table.
+
+Finally, the `ui.R` code is shown below:
+
+```{r, eval=FALSE}
+contingency_ui <- tabPanel("Contingency",
+    column(4,
+         wellPanel(
+             selectizeInput("xcont", label = "X Variable (x)", 
+                            choices = categoricNames(mpg), 
+                            selected = categoricNames(mpg)[5]),
+             selectizeInput("ycont", label = "Y Variable (y)", 
+                            choices = categoricNames(mpg), 
+                            selected = categoricNames(mpg)[6]),
+
+             hr(),
+
+             selectizeInput("conttype", label = "Type", 
+                            choices = c("Counts" = "counts", 
+                                        "Row Proportions" = "rowpercs", 
+                                        "Column Proportions" = "columnpercs", 
+                                        "Total Proportions" = "totalpercs")),
+             conditionalPanel(condition = "input.conttype != 'counts'",
+                numericInput("contdigits", label = "Digits", 
+                             value = 2, min = 1, step = 1)
+             ),
+
+             hr(),
+
+             tags$button("", id = "store_contingency", type = "button", 
+                         class = "btn action-button", 
+                         list(icon("save"), "Store Contingency Result"), 
+                         onclick = "$('#top-nav a:has(> .fa-print, 
+                         .fa-code, .fa-download)').highlight();")
+         )
+    ),
+
+    column(8,
+         tableOutput("conttable")
+    )
+)
+```
+
+This script defines all the inputs and outputs that the user will see.
+The only requirements from \texttt{intRo}'s perspective are (1) that there exist a store button at the bottom of the middle panel for storing the results of the analysis in the code script, and (2) that configuration options appear in the width 4 column in the middle, and output appears in the width 8 column on the right.
+The remaining input and output definitions depend on the statistical analysis or transformation being performed.
+
+Although the structure of an \texttt{intRo} module is relatively straightforward, producing the code needed in a more seamless fashion would certainly help open up the creation of such modules to a wider audience.
+As we discuss in the conclusions and future work section, providing an \texttt{intRo} module creation tool to abstract away some of the less common coding paradigms, like the use of `interpolate`, is an important effort that will continue to be pursued.
 
@@
 ```{r, eval=FALSE}
-interpolate <- function(code, ..., mydir, `_env` = parent.frame(), file = "code_All.R", 
-                        append = FALSE, save_result = FALSE, eval = TRUE) {
+interpolate <- function(code, ..., mydir, `_env` = parent.frame(), 
+                        file = "code_All.R", append = FALSE, 
+                        save_result = FALSE, eval = TRUE) {
     stopifnot(inherits(code, "formula"), length(code) == 2)
@@
 
-    cat(paste0(as.character(expr)[2], "\n"), file = file.path(mydir, file), append = append)
+    cat(paste0(as.character(expr)[2], "\n"), 
+        file = file.path(mydir, file), append = append)
 
-    if (save_result) cat(paste0(paste(readLines(file.path(mydir, file)), collapse = "\n"),
-                                "\n"), file = file.path(mydir, "code_All.R"), append = TRUE)
+    if (save_result) cat(paste0(paste(readLines(
+        file.path(mydir, file)), collapse = "\n"), "\n"), 
+        file = file.path(mydir, "code_All.R"), append = TRUE)
     if (eval) eval(expr, `_env`)

### 2016-01-08 modules/statistics/contingency/reactive.R
Reference the correct inputs in contingency

@@ -1,6 +1,6 @@
 intro.contingency <- reactive({
     if (is.null(intro.data())) return(NULL)
-    if (input$xreg == input$yreg || length(categoricNames(intro.data())) < 2) return(NULL)
+    if (input$xcont == input$ycont || length(categoricNames(intro.data())) < 2) return(NULL)
     
     my.conttable <- cont.table(intro.data(), input$xcont, input$ycont, input$conttype, input$contdigits)
     

### 2016-01-20 ui.R
Switch to using .click

@@ -57,7 +57,7 @@ shinyUI(
                           tabPanel("Andee Kaplan")),
                tabPanel(title = "hide_me"),
                tabPanel(title = "", icon = icon('code', "fa-2x"), value = "javascript:$('#myEditor').slideToggle(); $('.fa-code').parent().parent().toggleClass('active'); code_clicked();"),
-               tabPanel(title = "", icon = icon('download', "fa-2x"), value = "javascript:$('#top-nav i.fa-download').parent().attr('href', $('#mydownload').attr('href'));"),
+               tabPanel(title = "", icon = icon('download', "fa-2x"), value = "javascript:document.getElementById('mydownload').click();"),
                tabPanel(title = "", icon = icon("print", "fa-2x"), value = "javascript: $(this).addClass('print_button'); print_clicked();"),
                footer = tagList(includeScript("scripts/top-nav-links.js"),
                               includeScript("scripts/print.js"),

### 2016-01-20 jcgs-paper-2016/paper.Rmd
Remove date

@@
 title: "Introducing statistics with `intRo`"
-date: "November 23, 2015"
 author:

### 2016-01-20 jcgs-paper-2016/paper.Rmd
Clean up the output of helper.R

@@
 
-    if (type == "totalpercs") interpolate(~(my.tbl <- my.tbl / 
-                                          sum(table(df$y, df$x))), 
-                                          df = quote(intro.data), x = x, y = y,  
-                                          mydir = userdir, 
-                                          `_env` = environment(), 
-                                          file = "code_contingency.R", 
-                                          append = TRUE)
-    if (type == "rowpercs") interpolate(~(my.tbl <- t(apply(
-        my.tbl[-nrow(my.tbl),], 1, function(row){row / row[ncol(my.tbl)]}))),  
-                                        mydir = userdir,
-                                        `_env` = environment(), 
-                                        file = "code_contingency.R", 
-                                        append = TRUE)
-    if (type == "columnpercs") interpolate(~(my.tbl <- apply(
-        my.tbl[,-ncol(my.tbl)], 2, function(col){col / col[nrow(my.tbl)]})),  
-                                           mydir = userdir, 
-                                           `_env` = environment(), 
-                                           file = "code_contingency.R", 
-                                           append = TRUE)
-
-    if (type %in% c("totalpercs", "rowpercs", "columnpercs")) 
-        interpolate(~(new.digits <- digits), 
-                    digits = digits, 
-                    mydir = userdir,
-                    `_env` = environment(), 
-                    file = "code_contingency.R", 
-                    append = TRUE)
-    else interpolate(~(new.digits <- 0),  mydir = userdir, `_env` = environment(), 
-                     file = "code_contingency.R", append = TRUE)
+    # Additional formatting code suppressed for readability
+    ...
 
@@
 `interpolate` both executes the given R code on the server, and also writes the code executed to the script window at the bottom of \texttt{intRo}.
-More details on the reasons and choices behind `interpolate` can be found in the next section.
+More details on the reasons and choices behind `interpolate` can be found in section \ref{reproducibility}.
 

### 2016-01-20 jcgs-paper-2016/paper.Rmd
added a couple sentences

@@
 
-Shiny observers are created to ensure that the choices of variable for the `contingency` module are only categorical variables.
+Shiny observers are a class of reactive objects within the Shiny paradigm which do not return a value [@shiny].
+For further discussion of reactivity, see section \ref{reactive-design-choices}.
+In this example, observers are created to ensure that the choices of variable for the `contingency` module are only categorical variables.
 This is accomplished by utilizing the global reactive `intro.categoricnames()`, which returns a character vector containing the variables in the current dataset that are categorical.
@@
                             selected = categoricNames(mpg)[5]),
-             selectizeInput("ycont", label = "Y Variable (y)", 
-                            choices = categoricNames(mpg), 
-                            selected = categoricNames(mpg)[6]),
 
-             hr(),
-
-             selectizeInput("conttype", label = "Type", 
-                            choices = c("Counts" = "counts", 
-                                        "Row Proportions" = "rowpercs", 
-                                        "Column Proportions" = "columnpercs", 
-                                        "Total Proportions" = "totalpercs")),
-             conditionalPanel(condition = "input.conttype != 'counts'",
-                numericInput("contdigits", label = "Digits", 
-                             value = 2, min = 1, step = 1)
-             ),
+             # Additional inputs suppressed for readability
+             ...
 

### 2016-01-20 jcgs-paper-2016/paper.Rmd
Some work on the interpolation section

@@
 Within this paper, we introduce the application and explore the design decisions underlying `intRo`, as well as highlight some challenges and advantages of reactive programming.
-keywords: blah, blah blah
+keywords: R, Shiny, Statistics, Teaching, Interactive
 bibliography: refs.bib
@@
 This addition would be fairly simple to complete.
-The user would fork \texttt{intRo} from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `observe.R`, `output.R`, `reactive.R`, `static.R`, and `ui.R`.
+The user would fork \texttt{intRo} from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `libraries.R`, `observe.R`, `output.R`, `reactive.R`, and `ui.R`.
 Each of these files is written to comply with the reactive nature available in a \texttt{Shiny} application.
@@
 By placing each of these files within the `modules/nonparametric_test` folder and adding a single line to a navigation file indicating under which module grouping the non-parametric test module should go, a new module is available within \texttt{intRo} upon running the application.
+Further details on the creation of new modules can be found in section \ref{modules}.
 
@@
 We designed the modules in such a way that the code printed to the console is the same as the code executed by the server, which makes the process of writing an \texttt{intRo} module significantly more seamless.
-The workhorse function for this reproducibility framework is called `interpolate`, which is based on a function proposed by Hadley Wickham[ref].
-The idea of interpolation is that the call can be captured, printed, and evaluated, with all the appropriate arguments substituted into the call automatically.
+The workhorse function for this reproducibility framework is called `interpolate`, which is based on a function proposed by Hadley Wickham [-@interpolate].
+The idea of interpolation is that by using non-standard evaluation, the call can be captured, printed, and evaluated, with all the appropriate arguments substituted into the call automatically.
 This means that depending on the options selected in \texttt{intRo}, the corresponding arguments to functions are appropriately set for reproduction in a standard R session.
-
-```{r, eval=FALSE}
-interpolate <- function(code, ..., mydir, `_env` = parent.frame(), 
-                        file = "code_All.R", append = FALSE, 
-                        save_result = FALSE, eval = TRUE) {
-    stopifnot(inherits(code, "formula"), length(code) == 2)
-
-    args <- lapply(list(...), as_call)
-    expr <- methods::substituteDirect(as_call(code), args)
-
-    cat(paste0(as.character(expr)[2], "\n"), 
-        file = file.path(mydir, file), append = append)
-
-    if (save_result) cat(paste0(paste(readLines(
-        file.path(mydir, file)), collapse = "\n"), "\n"), 
-        file = file.path(mydir, "code_All.R"), append = TRUE)
-    if (eval) eval(expr, `_env`)
-}
-```
 
@@
 
-\texttt{intRo} will be trialed in an introductory statistics course at Iowa State University in Spring 2015.
-This will allow us to assess its usefulness relative to software used in the past, as well as guage areas for improvement.
+We hope to use \texttt{intRo} in courses to collect feedback regarding the ease of use and functionality.
+This will allow us to assess its usefulness relative to software used in the past, as well as gauge areas for improvement.
 Furthermore, we can determine the effectiveness of code printing on generating excitement from the students about programming in \texttt{R}.

### 2016-01-20 jcgs-paper-2016/paper.Rmd
Remove unneeded text about module customization

@@
 Each of these files is written to comply with the reactive nature available in a \texttt{Shiny} application.
-The `ui.R` file contains all the user interface elements within the new module's tab.
-This will include any options in the options panel as well as results.
-The `helper.R` and `static.R` files refer to static functions and global variable that may be used within the module.
-The file `output.R` contains the reactive elements that tie output to the user interface within \texttt{Shiny} and `observe.R` and `reactive.R` contain elements within an `observer` function or `reactive` function within the \texttt{Shiny} framework.
-By placing each of these files within the `modules/nonparametric_test` folder and adding a single line to a navigation file indicating under which module grouping the non-parametric test module should go, a new module is available within \texttt{intRo} upon running the application.
 Further details on the creation of new modules can be found in section \ref{modules}.

### 2016-01-24 jcgs-paper-2016/paper.Rmd
Fix some typos

@@
 \texttt{intRo} offers a number of key advantages over traditional statistics software, including ease of use and an aim to foster student interest.
-Additionally, \texttt{intRo} stands apart from new tols in that it is a supplement to an existing class, fully usable by a beginning statistics student.
+Additionally, \texttt{intRo} stands apart from new tools in that it is a supplement to an existing class, fully usable by a beginning statistics student.
 These advantages will be discussed at length in this text.
@@
 \texttt{intRo} is a web-based tool to accompany an introductory statistics class.
-It is meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class.
+It is meant to assist in the learning of statistics rather than act as a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class.
 An accompanying R package, titled \texttt{intRo} and available on GitHub, assists in the downloading, running, and deploying of \texttt{intRo} instances.
 
-We have three fundamental philosophies that guided the creation of \texttt{intRo}, and this comes out in the design.
-Primarily, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics.
-Additionally, \texttt{intRo} is an \emph{extensible} tool, allowing for a user of \texttt{intRo} to tailor the tool for her own classroom needs.
+We have three fundamental philosophies that guided the creation of \texttt{intRo}.
+In particular, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics.
+Additionally, \texttt{intRo} is an \emph{extensible} tool, allowing for a user of \texttt{intRo} to tailor the tool for his or her own classroom needs.
 
@@
 \includegraphics[width=\linewidth]{images/user_experience.pdf}
-\caption{Typical user experience of generating a result in \texttt{intRo}. In this instance, a mosaic plot (and its corresponding code) was created and saved to the final compendium.}
+\caption{A typical user experience of generating a result in \texttt{intRo}. In this instance, a mosaic plot (and its corresponding code) was created and saved to the final compendium.}
 \label{fig:user_experience}
@@
 Each module is placed within a higher level category to further help students navigate the tool.
-The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can easily be added by an instructor, as detailed in section \ref{sec:extensible}.
+The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can easily be added by an instructor, as detailed in section \ref{sec:extensible} and section \ref{sec:modules}.
 The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.

### 2016-01-24 jcgs-paper-2016/paper.Rmd
Fix a broken reference

@@
 Each module is placed within a higher level category to further help students navigate the tool.
-The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can easily be added by an instructor, as detailed in section \ref{sec:extensible} and section \ref{sec:modules}.
+The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can easily be added by an instructor, as detailed in section \ref{sec:extensible} and section \ref{modules}.
 The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.

### 2016-01-24 jcgs-paper-2016/paper.Rmd
Unneeded reference removed

@@
 Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the \texttt{intRo} interface to obtain the results.
-For topics currently unsupported by \texttt{intRo}, the instructor can supplement its use with other web-based modules that are publicly available or create their own modules within \texttt{intRo} as detailed in section \ref{sec:extensible}.
+For topics currently unsupported by \texttt{intRo}, the instructor can supplement its use with other web-based modules that are publicly available or create their own modules within \texttt{intRo} as detailed in section \ref{modules}.
 
@@
 On the flip side of this, \texttt{Shiny} allows much higher flexibility.
-For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}, as detailed in section \ref{sec:extensible}.
+For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}.
 However, \texttt{Shiny} does have limits on its flexibility based on the designers decisions for usability.

### 2016-02-03 jcgs-paper-2016/paper.Rmd
Font size change

@@
     template: jcgs_article.tex
+fontsize: 12pt
 abstract: |

### 2016-02-05 jcgs-paper-2016/paper.Rmd
Paper edits

--- 2016-01-24 jcgs-paper-2016/paper.Rmd
@@
 ---
-title: "Introducing statistics with `intRo`"
+title: "Introductory statistics with `intRo`"
 author:
@@
     template: jcgs_article.tex
+fontsize: 12pt
 abstract: |
-  `intRo` is a modern web-based application for performing basic data analysis and statistical routines.
-Leveraging the power of `R` and `Shiny`, `intRo` implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician.
+  `intRo` is a web-based application for performing basic data analysis and statistical routines.
+Leveraging the power of `R` and `Shiny`, `intRo` implements common statistical functions in an extensible modular structure, while including a point-and-click interface for the novice statistician.
 This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as `Excel` and `JMP`.
@@
 Within this paper, we introduce the application and explore the design decisions underlying `intRo`, as well as highlight some challenges and advantages of reactive programming.
-keywords: R, Shiny, Statistics, Teaching, Interactive
+keywords: Graphical User Interface, Interactivity, Reproducibility, Reactive Programming, Statistical Software, Programming Paradigms
 bibliography: refs.bib
@@
 The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been an important development for the scientific community.
-However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student.
-This issue may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors would tend to use a tool like \texttt{R} in their own work.
-Still, signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.
+However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student.
+This additional time commitment may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors tend to use a tool like \texttt{R} in their own work.
+This brings along signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses and unsupported desktop platforms.
 
-In teaching Introduction to Business Statistics at Iowa State University, we witnessed profound struggles by students attempting to practice the introductory concepts discussed in class using current software.
-Scrimshaw notes in his manuscript that "open-ended packages, like any others, may create obstacles to learning simply through their lack of user-friendliness in the sheer mechanics of operating them, rather than any intrinsic difficulty in the content...." [-@scrimshaw2001computers] In our own experience teaching, students' struggles were often directly related to the use of the software and not any sort of fundamental misunderstanding of the material, in agreement with Scrimshaw's finding.
+In teaching Introduction to Business Statistics at Iowa State University, we witnessed profound struggles by students attempting to practice introductory concepts discussed in class using current software.
+Scrimshaw [-@scrimshaw2001computers] notes in his manuscript that "open-ended packages, like any others, may create obstacles to learning simply through their lack of user-friendliness in the sheer mechanics of operating them, rather than any intrinsic difficulty in the content...." In our own experience teaching, students' struggles were often directly related to the use of the software and not any sort of fundamental misunderstanding of the material, in agreement with Scrimshaw's finding.
 
@@
 \texttt{intRo} is a web-based tool to accompany an introductory statistics class.
-It is meant to assist in the learning of statistics rather than act as a stand-alone deliverer of statistics education, with the intention of being used in conjunction with a guided class.
+It is meant to assist in the learning of statistics rather than as a stand-alone delivery of statistics education, with the intention of being used in conjunction with a guided class.
 An accompanying R package, titled \texttt{intRo} and available on GitHub, assists in the downloading, running, and deploying of \texttt{intRo} instances.
 
-We have three fundamental philosophies that guided the creation of \texttt{intRo}.
+Three fundamental philosophies that guided the creation of \texttt{intRo}.
 In particular, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics.
@@
 
-## Easy
+In the development of \texttt{intRo}, we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training.
+We used large, easy to click icons in the page header to help students find what they need more easily.
+We also made the functionality available the minimal necessary for an introductory statistics course.
+Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in \texttt{intRo}.
+In this instance, a user clicks on the Graphical tab to create a mosaic plot.
+The user sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.
 
@@
 \includegraphics[width=\linewidth]{images/user_experience.pdf}
-\caption{A typical user experience of generating a result in \texttt{intRo}. In this instance, a mosaic plot (and its corresponding code) was created and saved to the final compendium.}
+\caption{A typical user experience of generating a result in \texttt{intRo}.
+In this instance, a user clicks on the Graphical tab to create a mosaic plot.
+The user sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.}
 \label{fig:user_experience}
@@
 
-In order for an application to be useful to students in their first statistics course, it can't hog the spotlight.
-Statistics as a subject should shine through as a fun and interesting topic to study, rather than being bogged down by difficult to learn software.
-With this in mind, in the development of \texttt{intRo} we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training.
-We used large, easy to click icons in the page header to help students find exactly what they need easily.
-With this easy to use framework, we also made the functionality available in \texttt{intRo} the minimal necessary for an introductory statistics course.
-Additional functionality is nice, but if it clutters the UI beyond the point of usability, it will frustrate an introductory student.
-Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in \texttt{intRo}, view the code producing the result, and save the results for later.
-
 Beyond being simple, \texttt{intRo} is also consistent.
 The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules.
-Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code.
-In figure \ref{fig:ui}, we highlight the five elements that comprise the \texttt{intRo} interface.
+To the user, a module is simply a page of statistics functionality that maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code.
+Figure \ref{fig:ui} highlights the five elements that comprise the \texttt{intRo} interface.
 
@@
 
-# In the classroom
+# Using intRo
+
+## intRo Package
+
+\texttt{intRo} can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package \texttt{intRo}.
+Currently, the package is only available and GitHub, and can be installed using the \texttt{devtools} package as follows:
+
+```{r, eval=FALSE}
+devtools::install_github("gammarama/intRo")
+```
+
+After installing the \texttt{intRo} package, the first function one should call is `download_intRo`.
+`download_intRo` takes as an argument a directory in which to store the application.
+By default, it selects the working directory of the R session.
+This function clones the application branch of the \texttt{intRo} repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
+
+Running `download_intRo` will produce an \texttt{intRo} folder in the specified folder.
+It can then be ran as any \texttt{Shiny} application, using \texttt{Shiny}'s `runApp` command.
+However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
+`run_intRo` takes as argument the path to the folder containing the \texttt{intRo} application.
+It also takes several more optional arguments:
+
+* `enabled_modules`: A character vector containing the modules to enable
+* `theme`: A string representing a \texttt{shinythemes} theme to use
+* `...`: Additional arguments passed to \texttt{Shiny}'s `runApp` function
+
+The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download \texttt{intRo} to my working directory, execute an \texttt{intRo} session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
+The series of calls to do so would be as follows:
+
+```{r, eval=FALSE}
+download_intRo()
+run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+          theme = "cerulean")
+```
+
+Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
+
+If the intent is to use a specific instance of \texttt{intRo} where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of \texttt{intRo} to a publicly accessible URL.
+The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the \texttt{shinyapps} package.
+Once the shinyapps package is installed and configured, `deploy_intRo` will upload \texttt{intRo} as an application on the user's account.
+The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
+It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
+It also takes `...` as additional arguments to be passed into the `deployApp` routine.
+For example, if we wished to deploy the instance of \texttt{intRo} we ran previously, we would call it like so:
+
+```{r, eval=FALSE}
+deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+             theme = "cerulean")
+```
+
+Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
+
+## In the classroom
+
 \texttt{intRo}'s ease of use lends itself to a natural presentation in an introductory statistics course.
@@
 
-# Using intRo
-\texttt{intRo} can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package \texttt{intRo}.
-Currently, the package is only available and GitHub, and can be installed using the \texttt{devtools} package as follows:
-
-```{r, eval=FALSE}
-devtools::install_github("gammarama/intRo")
-```
-
-After installing the \texttt{intRo} package, the first function one should call is `download_intRo`.
-`download_intRo` takes as an argument a directory in which to store the application.
-By default, it selects the working directory of the R session.
-This function clones the application branch of the \texttt{intRo} repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
-
-Running `download_intRo` will produce an \texttt{intRo} folder in the specified folder.
-It can then be ran as any \texttt{Shiny} application, using \texttt{Shiny}'s `runApp` command.
-However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
-`run_intRo` takes as argument the path to the folder containing the \texttt{intRo} application.
-It also takes several more optional arguments:
-
-* `enabled_modules`: A character vector containing the modules to enable
-* `theme`: A string representing a \texttt{shinythemes} theme to use
-* `...`: Additional arguments passed to \texttt{Shiny}'s `runApp` function
-
-The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download \texttt{intRo} to my working directory, execute an \texttt{intRo} session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
-The series of calls to do so would be as follows:
-
-```{r, eval=FALSE}
-download_intRo()
-run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
-          theme = "cerulean")
-```
-
-Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
-
-If the intent is to use a specific instance of \texttt{intRo} where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of \texttt{intRo} to a publicly accessible URL.
-The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the \texttt{shinyapps} package.
-Once the shinyapps package is installed and configured, `deploy_intRo` will upload \texttt{intRo} as an application on the user's account.
-The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
-It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
-It also takes `...` as additional arguments to be passed into the `deployApp` routine.
-For example, if we wished to deploy the instance of \texttt{intRo} we ran previously, we would call it like so:
-
-```{r, eval=FALSE}
-deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
-             theme = "cerulean")
-```
-
-Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
-
-# Details
+# Understand intRo
 

### 2016-02-11 jcgs-paper-2016/paper.Rmd
Some small changes

@@
 Each module is placed within a higher level category to further help students navigate the tool.
-The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can easily be added by an instructor, as detailed in section \ref{sec:extensible} and section \ref{modules}.
+The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can easily be added by an instructor, as detailed in section \ref{modules}.
 The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.
@@
 
-## Exciting
-In general, introduction to statistics is a class undergraduate students dread.
-It's hard, the vocabulary is unfamiliar, and the software is difficult to use.
-With \texttt{intRo}, we hope to flip this paradigm.
-\texttt{intRo} is a fun, easy to use tool that's available on the web, and is also usable on mobile devices.
-As a web-based application, this tool is immediately more familiar to students than a desktop application.
-With interactive plots from \texttt{ggvis} [@ggvis], \texttt{intRo} brings a `wow'-factor to learning introductory statistics.
-
-Once students get pulled in by an easy to use interface and fun interaction with their data, \texttt{intRo} has an ulterior motive: to get students excited about programming.
+\texttt{intRo} has an ulterior motive as well: to get students excited about programming.
 By navigating about the user interface of \texttt{intRo}, students are actually creating a fully-executable \texttt{R} script that they can download and run locally as well as viewing the script change real-time within the application.
@@
 
-## Extensible
-\label{sec:extensible}
 On the front end, user interaction with \texttt{intRo} is split into bitesize chunks that we call modules.
@@
 
-# Understand intRo
+# Understanding intRo
 

### 2016-02-11 jcgs-paper-2016/paper.Rmd
Finish heike's edits

@@
  1.
-**Top Navigation** - The top navigation bar includes two sets of buttons.
+**Top Navigation** - The top navigation bar includes two sets of clickable icons.
 The left-aligned buttons are informational buttons.
 The first is a link to \texttt{intRo}.
-The second is a link to this documentation page.
+The second is a link to the documentation page.
 The third is a link to the GitHub repository where the code for \texttt{intRo} is housed.
@@
 The right-aligned buttons are \texttt{intRo} utilities.
-The first is a button to toggle the visibility of the code panel (5).
-The last is a button to print the stored module results, and the associated code (if visible).
- 2. **Side Navigation** - The side navigation panel includes the modules that perform a specific data analysis task.
- 3. **Options Panel** - The options panel includes module-specific options which the user can use to customize their results.
+The first is a link to toggle the visibility of the code panel (5).
+The middle icon downloads an rmarkdown [@rmarkdown] document of the analysis performed.
+The last is a link to print the stored module results, and the associated code (if visible).
+ 2. **Side Navigation** - The side navigation panel includes a list of data analysis tasks.
+ 3. **Options Panel** - The options panel includes task-specific options which the user can use to customize their results.
  4. **Results Panel** - The results pane displays the result of the selected module and options.
@@
 **Code Panel** - The code panel displays the \texttt{R} code used to generate the results from the student's \texttt{intRo} session.
-The code panel is shown by default to facilitate an interest in coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.
+The code panel is shown by default to facilitate a transition to coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.
 
-There are seven default modules within \texttt{intRo}, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
-Each module is placed within a higher level category to further help students navigate the tool.
-The default modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics - and more modules can easily be added by an instructor, as detailed in section \ref{modules}.
+The modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics.
+Under each of these categories, there are seven default modules, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
+More modules can easily be added by an instructor, as detailed in section \ref{modules}.
 The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.
 
-In addition to making \texttt{intRo} intuitive to use, we've also created a documentation website that is consistent with the interface of the application.
+We've also created a documentation website that is consistent with the interface of the application.
 The documentation is available at http://gammarama.github.io/intRo and hyperlinked within the application itself.
-This documentation covers all default modules and makes learning to use \texttt{intRo} even more painless.
 
@@
 This code creation element of \texttt{intRo} is meant to generate excitement about programming in \texttt{R} and empower students to feel that they can generate code as well.
-\texttt{intRo} uses \texttt{rmarkdown}'s [@rmarkdown] \texttt{render} function in order to print the results, by dymanically executing the user's \texttt{R} script.
+\texttt{intRo} uses \texttt{rmarkdown}'s \texttt{render} function in order to print the results, by dymanically executing the user's \texttt{R} script.
 By default, the output will include the \texttt{R} code, but if the user elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.

### 2016-02-18 jcgs-paper-2016/paper.Rmd
Typo typo

@@
 This additional time commitment may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors tend to use a tool like \texttt{R} in their own work.
-This brings along signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses and unsupported desktop platforms.
+Still, some compromises must be made when using software like \texttt{JMP}, including dealing with software licenses and unsupported desktop platforms.
 
@@
 This in turn led us to create an introductory statistics application which we call \texttt{intRo}, available at http://www.intro-stats.com.
-\texttt{intRo} offers a number of key advantages over traditional statistics software, including ease of use and an aim to foster student interest.
+\texttt{intRo} offers a number of key advantages over traditional statistics software, including ease of use and an aim to foster student interest in coding.
 Additionally, \texttt{intRo} stands apart from new tools in that it is a supplement to an existing class, fully usable by a beginning statistics student.

### 2016-02-19 jcgs-paper-2016/paper.Rmd
No more texttt

@@
 
-The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been an important development for the scientific community.
-However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student.
-This additional time commitment may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors tend to use a tool like \texttt{R} in their own work.
-Still, some compromises must be made when using software like \texttt{JMP}, including dealing with software licenses and unsupported desktop platforms.
+The widespread adoption of R as a tool for statistical analysis has undoubtedly been an important development for the scientific community.
+However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student.
+This additional time commitment may explain why introductory courses often utilize software applications such as JMP, even though instructors tend to use a tool like R in their own work.
+Still, some compromises must be made when using software like JMP, including dealing with software licenses and unsupported desktop platforms.
 
@@
 Multiple software packages have recently been written in an attempt to spur student interest in introductory statistics and R programming.
-DataCamp's [@datacamp] courses are user-friendly ways to learning basic \texttt{R} programming and data analysis techniques.
-\texttt{Swirl} [@swirl] is a similar interactive tool to make learning \texttt{R} more fun by learning it within \texttt{R} itself.
+DataCamp's [@datacamp] courses are user-friendly ways to learning basic R programming and data analysis techniques.
+Swirl [@swirl] is a similar interactive tool to make learning R more fun by learning it within R itself.
 Project MOSAIC [@mosaic] has created a suite of tools to simplify the teaching of statistics in the form of an R package.
-The primary goal of DataCamp and Swirl is to teach \texttt{R} programming, rather than facilitate the learning of introductory statistics.
+The primary goal of DataCamp and Swirl is to teach R programming, rather than facilitate the learning of introductory statistics.
 Project MOSAIC's goal is to faciliate this learning, but using the package requires a knowledge of R programming that the introductory student may not have.
 
-Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easier for an R-based analysis to be converted to an interactive web application.
-This in turn led us to create an introductory statistics application which we call \texttt{intRo}, available at http://www.intro-stats.com.
-\texttt{intRo} offers a number of key advantages over traditional statistics software, including ease of use and an aim to foster student interest in coding.
-Additionally, \texttt{intRo} stands apart from new tools in that it is a supplement to an existing class, fully usable by a beginning statistics student.
+Upon the release of RStudio's Shiny [@shiny] it became easier for an R-based analysis to be converted to an interactive web application.
+This in turn led us to create an introductory statistics application which we call `intRo`, available at http://www.intro-stats.com.
+`intRo` offers a number of key advantages over traditional statistics software, including ease of use and an aim to foster student interest in coding.
+Additionally, `intRo` stands apart from new tools in that it is a supplement to an existing class, fully usable by a beginning statistics student.
 These advantages will be discussed at length in this text.
 
-# What is \texttt{intRo}?
+# What is `intRo`?
 
-\texttt{intRo} is a web-based tool to accompany an introductory statistics class.
+`intRo` is a web-based tool to accompany an introductory statistics class.
 It is meant to assist in the learning of statistics rather than as a stand-alone delivery of statistics education, with the intention of being used in conjunction with a guided class.
-An accompanying R package, titled \texttt{intRo} and available on GitHub, assists in the downloading, running, and deploying of \texttt{intRo} instances.
+An accompanying R package, titled `intRo` and available on GitHub, assists in the downloading, running, and deploying of `intRo` instances.
 
-Three fundamental philosophies that guided the creation of \texttt{intRo}.
-In particular, \texttt{intRo} is \emph{easy} to use and can be an \emph{exciting} part of learning statistics.
-Additionally, \texttt{intRo} is an \emph{extensible} tool, allowing for a user of \texttt{intRo} to tailor the tool for his or her own classroom needs.
+Three fundamental philosophies that guided the creation of `intRo`.
+In particular, `intRo` is \emph{easy} to use and can be an \emph{exciting} part of learning statistics.
+Additionally, `intRo` is an \emph{extensible} tool, allowing for a user of `intRo` to tailor the tool for his or her own classroom needs.
 
-In the development of \texttt{intRo}, we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training.
+In the development of `intRo`, we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training.
 We used large, easy to click icons in the page header to help students find what they need more easily.
 We also made the functionality available the minimal necessary for an introductory statistics course.
-Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in \texttt{intRo}.
+Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in `intRo`.
 In this instance, a user clicks on the Graphical tab to create a mosaic plot.
@@
 \includegraphics[width=\linewidth]{images/user_experience.pdf}
-\caption{A typical user experience of generating a result in \texttt{intRo}.
+\caption{A typical user experience of generating a result in `intRo`.
 In this instance, a user clicks on the Graphical tab to create a mosaic plot.
@@
 
-Beyond being simple, \texttt{intRo} is also consistent.
+Beyond being simple, `intRo` is also consistent.
 The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules.
 To the user, a module is simply a page of statistics functionality that maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code.
-Figure \ref{fig:ui} highlights the five elements that comprise the \texttt{intRo} interface.
+Figure \ref{fig:ui} highlights the five elements that comprise the `intRo` interface.
 
@@
 \includegraphics[width=\linewidth]{images/ui_annotate.pdf}
-\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
+\caption{The five elements that comprise the `intRo` application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
 \label{fig:ui}
@@
 The left-aligned buttons are informational buttons.
-The first is a link to \texttt{intRo}.
+The first is a link to `intRo`.
 The second is a link to the documentation page.
-The third is a link to the GitHub repository where the code for \texttt{intRo} is housed.
+The third is a link to the GitHub repository where the code for `intRo` is housed.
 The final button is a link to our websites, which contain contact information if there are any questions or comments.
-The right-aligned buttons are \texttt{intRo} utilities.
+The right-aligned buttons are `intRo` utilities.
 The first is a link to toggle the visibility of the code panel (5).
@@
  5.
-**Code Panel** - The code panel displays the \texttt{R} code used to generate the results from the student's \texttt{intRo} session.
+**Code Panel** - The code panel displays the R code used to generate the results from the student's `intRo` session.
 The code panel is shown by default to facilitate a transition to coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.
 
-The modules included in \texttt{intRo} are split into three higher level categories - data, summaries, and statistics.
+The modules included in `intRo` are split into three higher level categories - data, summaries, and statistics.
 Under each of these categories, there are seven default modules, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
@@
 
-\texttt{intRo} has an ulterior motive as well: to get students excited about programming.
-By navigating about the user interface of \texttt{intRo}, students are actually creating a fully-executable \texttt{R} script that they can download and run locally as well as viewing the script change real-time within the application.
-This code creation element of \texttt{intRo} is meant to generate excitement about programming in \texttt{R} and empower students to feel that they can generate code as well.
-\texttt{intRo} uses \texttt{rmarkdown}'s \texttt{render} function in order to print the results, by dymanically executing the user's \texttt{R} script.
-By default, the output will include the \texttt{R} code, but if the user elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
+`intRo` has an ulterior motive as well: to get students excited about programming.
+By navigating about the user interface of `intRo`, students are actually creating a fully-executable R script that they can download and run locally as well as viewing the script change real-time within the application.
+This code creation element of `intRo` is meant to generate excitement about programming in R and empower students to feel that they can generate code as well.
+`intRo` uses rmarkdown's render function in order to print the results, by dymanically executing the user's R script.
+By default, the output will include the R code, but if the user elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
 
-On the front end, user interaction with \texttt{intRo} is split into bitesize chunks that we call modules.
+On the front end, user interaction with `intRo` is split into bitesize chunks that we call modules.
 Modules are self-contained pieces of functionality which implement common statistical procedures that are often used in introductory statistics classes.
@@
 From this module the user can select a subset of the data either by value or random selection.
-The dataset selected in the data module will be made available to all other modules of \texttt{intRo}.
+The dataset selected in the data module will be made available to all other modules of `intRo`.
 The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
 Additionally, numeric variable can be transformed using a power transformation.
-The transformed variables can be saved to the dataset for use elsewhere in \texttt{intRo}.
+The transformed variables can be saved to the dataset for use elsewhere in `intRo`.
 
@@
 
-Each module is a self contained set of \texttt{R} code that is dynamically added to the application at run time.
-What this means is that \texttt{intRo} can be easily extended by the addition of modules within the framework underlying the application.
-This extensibility allows for \texttt{intRo} to be customized for an individual classroom setting, tailored to the needs of a particular course.
+Each module is a self contained set of R code that is dynamically added to the application at run time.
+What this means is that `intRo` can be easily extended by the addition of modules within the framework underlying the application.
+This extensibility allows for `intRo` to be customized for an individual classroom setting, tailored to the needs of a particular course.
 
@@
 This addition would be fairly simple to complete.
-The user would fork \texttt{intRo} from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `libraries.R`, `observe.R`, `output.R`, `reactive.R`, and `ui.R`.
-Each of these files is written to comply with the reactive nature available in a \texttt{Shiny} application.
+The user would fork `intRo` from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `libraries.R`, `observe.R`, `output.R`, `reactive.R`, and `ui.R`.
+Each of these files is written to comply with the reactive nature available in a Shiny application.
 Further details on the creation of new modules can be found in section \ref{modules}.
@@
 \includegraphics[width=\linewidth]{images/app_creation_modules.pdf}
-\caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.}
+\caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within `intRo`.}
 \label{fig:app_creation_modules}
@@
 
-The process by which \texttt{intRo} is dynamically created upon running the application through module sourcing is detailed in figure \ref{fig:app_creation_modules}.
+The process by which `intRo` is dynamically created upon running the application through module sourcing is detailed in figure \ref{fig:app_creation_modules}.
 The key driver to populating \texttt{server.R} and \texttt{ui.R} is the navigation file \texttt{modules.txt}, which defines the placement of each module.
@@
 \normalsize
-In this way, we were able to have \texttt{intRo} be fully extensible, its structure and functionality dependent entirely on the modules present within the application.
+In this way, we were able to have `intRo` be fully extensible, its structure and functionality dependent entirely on the modules present within the application.
 
@@
 
-\texttt{intRo} can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package \texttt{intRo}.
-Currently, the package is only available and GitHub, and can be installed using the \texttt{devtools} package as follows:
+`intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.
+Currently, the package is only available and GitHub, and can be installed using the devtools package as follows:
 
@@
 
-After installing the \texttt{intRo} package, the first function one should call is `download_intRo`.
+After installing the `intRo` package, the first function one should call is `download_intRo`.
 `download_intRo` takes as an argument a directory in which to store the application.
 By default, it selects the working directory of the R session.
-This function clones the application branch of the \texttt{intRo} repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
+This function clones the application branch of the `intRo` repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
 
-Running `download_intRo` will produce an \texttt{intRo} folder in the specified folder.
-It can then be ran as any \texttt{Shiny} application, using \texttt{Shiny}'s `runApp` command.
+Running `download_intRo` will produce an `intRo` folder in the specified folder.
+It can then be ran as any Shiny application, using Shiny's `runApp` command.
 However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
-`run_intRo` takes as argument the path to the folder containing the \texttt{intRo} application.
+`run_intRo` takes as argument the path to the folder containing the `intRo` application.
 It also takes several more optional arguments:
@@
 * `enabled_modules`: A character vector containing the modules to enable
-* `theme`: A string representing a \texttt{shinythemes} theme to use
-* `...`: Additional arguments passed to \texttt{Shiny}'s `runApp` function
+* `theme`: A string representing a shinythemes theme to use
+* `...`: Additional arguments passed to Shiny's `runApp` function
 
-The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download \texttt{intRo} to my working directory, execute an \texttt{intRo} session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
+The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download `intRo` to my working directory, execute an `intRo` session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
 The series of calls to do so would be as follows:
@@
 
-If the intent is to use a specific instance of \texttt{intRo} where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of \texttt{intRo} to a publicly accessible URL.
-The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the \texttt{shinyapps} package.
-Once the shinyapps package is installed and configured, `deploy_intRo` will upload \texttt{intRo} as an application on the user's account.
+If the intent is to use a specific instance of `intRo` where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
+The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package.
+Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the user's account.
 The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
@@
 It also takes `...` as additional arguments to be passed into the `deployApp` routine.
-For example, if we wished to deploy the instance of \texttt{intRo} we ran previously, we would call it like so:
+For example, if we wished to deploy the instance of `intRo` we ran previously, we would call it like so:
 
@@
 
-\texttt{intRo}'s ease of use lends itself to a natural presentation in an introductory statistics course.
-Statistics software typically used in introductory courses, such as \texttt{JMP} and \texttt{Excel}, have restrictive software licensing and platform requirements that can limit the software's accessibility to students.
+`intRo`'s ease of use lends itself to a natural presentation in an introductory statistics course.
+Statistics software typically used in introductory courses, such as JMP and Excel, have restrictive software licensing and platform requirements that can limit the software's accessibility to students.
 Even if these limitations can be overcome, these software programs have non-ideal interfaces for an exploratory data analysis.
-\texttt{JMP}, for instance, hides several basic regression functions such as a line of best fit behind small red arrows.
-Using \texttt{R} itself in the classroom is an appealing alternative, but suffers from some other problems.
+JMP, for instance, hides several basic regression functions such as a line of best fit behind small red arrows.
+Using R itself in the classroom is an appealing alternative, but suffers from some other problems.
 Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis.
-When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (this experience is made much easier with new tools like \texttt{rmarkdown}, but this again is a bit of a learning curve for the new statistics student).
+When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (this experience is made much easier with new tools like rmarkdown, but this again is a bit of a learning curve for the new statistics student).
 
-We believe that \texttt{intRo} targets these key limitations.
+We believe that `intRo` targets these key limitations.
 The interface is designed to be quick, seamless, and easy to understand.
-Because \texttt{intRo} is web-based, the statistical analysis can be performed from nearly any modern device.
-\texttt{R} underlies the analysis, but users need not be familiar with \texttt{R} or programming to perform it.
+Because `intRo` is web-based, the statistical analysis can be performed from nearly any modern device.
+R underlies the analysis, but users need not be familiar with R or programming to perform it.
 The focal point of the interface is the output panel, allowing the student to remain focused on the results and limiting the need to fight with the interface itself to get work done.
@@
 
-The features of \texttt{intRo} were selected in order to apply to the most commonly used lesson plans in early statistics courses.
-The textbook "The Art and Science of Learning from Data" by Agresti and Franklin [-@agresti2012statistics] references many of the focal points of introductory statistics, and \texttt{intRo} supports performing a majority of them.
+The features of `intRo` were selected in order to apply to the most commonly used lesson plans in early statistics courses.
+The textbook "The Art and Science of Learning from Data" by Agresti and Franklin [-@agresti2012statistics] references many of the focal points of introductory statistics, and `intRo` supports performing a majority of them.
 Part one, titled "Gathering and Exploring Data", covers graphical summaries, measures of spread and center, and avoiding the misuse of graphics in situations they are not appropriate.
-\texttt{intRo} includes modules for both graphical and numerical summaries covering these topics.
-Furthermore, \texttt{intRo} limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics (Table \ref{tab:plottypes}).
+`intRo` includes modules for both graphical and numerical summaries covering these topics.
+Furthermore, `intRo` limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics (Table \ref{tab:plottypes}).
 
@@
 \end{tabular}
-\caption{The plot and variable types allowed in \texttt{intRo}'s graphical module.
+\caption{The plot and variable types allowed in `intRo`'s graphical module.
 For example, when the selected plottype is histogram, the interface displays only a single variable selection box (x), and the variables selectable in this box are only numeric.
@@
 
-Part two of this text discusses probability and sampling distributions, which is outside the current realm of \texttt{intRo}, but part three focuses on inferential statistics.
-The inferential statistics include confidence intervals, hypothesis testing, and comparing two variables, all supported natively in \texttt{intRo}.
-Finally, part four of the text discusses the association between variables, including regression, functions that \texttt{intRo} performs with ease.
-Some of the more advanced topics of this section are currently unsupported by \texttt{intRo}, including multiple regression.
+Part two of this text discusses probability and sampling distributions, which is outside the current realm of `intRo`, but part three focuses on inferential statistics.
+The inferential statistics include confidence intervals, hypothesis testing, and comparing two variables, all supported natively in `intRo`.
+Finally, part four of the text discusses the association between variables, including regression, functions that `intRo` performs with ease.
+Some of the more advanced topics of this section are currently unsupported by `intRo`, including multiple regression.
 
-Statistics courses with a lab component can most benefit from the use of \texttt{intRo}.
+Statistics courses with a lab component can most benefit from the use of `intRo`.
 As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries.
@@
 Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in.
-Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the \texttt{intRo} interface to obtain the results.
-For topics currently unsupported by \texttt{intRo}, the instructor can supplement its use with other web-based modules that are publicly available or create their own modules within \texttt{intRo} as detailed in section \ref{modules}.
+Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the `intRo` interface to obtain the results.
+For topics currently unsupported by `intRo`, the instructor can supplement its use with other web-based modules that are publicly available or create their own modules within `intRo` as detailed in section \ref{modules}.
 
@@
 ## Modules
-An \texttt{intRo} module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality.
-\texttt{intRo} modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the \texttt{intRo} framework.
+An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality.
+`intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework.
 In this section, we detail the structure of these modules, and the process of creating and deploying new modules.
@@
 
-The modules provided with \texttt{intRo} are contained in the modules folder.
+The modules provided with `intRo` are contained in the modules folder.
 The top level directory in the modules folder defines the category of the module (currently `data`, `summaries`, or `statistics`).
@@
 
-Since the `contingency` module performs a statistical analysis, it is part of the statistics category, and hence can be found in the \texttt{intRo} repository at **modules/statistics/contingency**.
+Since the `contingency` module performs a statistical analysis, it is part of the statistics category, and hence can be found in the `intRo` repository at **modules/statistics/contingency**.
 Let's first examine *helper.R*:
@@
 One important difference from a typical R script is that each call in the script is wrapped in a function called `interpolate`.
-`interpolate` both executes the given R code on the server, and also writes the code executed to the script window at the bottom of \texttt{intRo}.
+`interpolate` both executes the given R code on the server, and also writes the code executed to the script window at the bottom of `intRo`.
 More details on the reasons and choices behind `interpolate` can be found in section \ref{reproducibility}.
@@
 Finally, there is an event observer to store code generated from the module into the overall code script upon clicking the store button.
-The presence of this observer code and the definition of the button in the user interface are enforced, and must be present in any \texttt{intRo} module.
+The presence of this observer code and the definition of the button in the user interface are enforced, and must be present in any `intRo` module.
 
@@
 This script defines all the inputs and outputs that the user will see.
-The only requirements from \texttt{intRo}'s perspective are (1) that there exist a store button at the bottom of the middle panel for storing the results of the analysis in the code script, and (2) that configuration options appear in the width 4 column in the middle, and output appears in the width 8 column on the right.
+The only requirements from `intRo`'s perspective are (1) that there exist a store button at the bottom of the middle panel for storing the results of the analysis in the code script, and (2) that configuration options appear in the width 4 column in the middle, and output appears in the width 8 column on the right.
 The remaining input and output definitions depend on the statistical analysis or transformation being performed.
 
-Although the structure of an \texttt{intRo} module is relatively straightforward, producing the code needed in a more seamless fashion would certainly help open up the creation of such modules to a wider audience.
-As we discuss in the conclusions and future work section, providing an \texttt{intRo} module creation tool to abstract away some of the less common coding paradigms, like the use of `interpolate`, is an important effort that will continue to be pursued.
+Although the structure of an `intRo` module is relatively straightforward, producing the code needed in a more seamless fashion would certainly help open up the creation of such modules to a wider audience.
+As we discuss in the conclusions and future work section, providing an `intRo` module creation tool to abstract away some of the less common coding paradigms, like the use of `interpolate`, is an important effort that will continue to be pursued.
 
@@
 
-We have designed \texttt{intRo} modules to reproduce actions taken by the user so that they can obtain \texttt{R} code associated with each module interacted with.
+We have designed `intRo` modules to reproduce actions taken by the user so that they can obtain R code associated with each module interacted with.
 When a user chooses to store the results associated with a particular module, R code yielding the visible results is automatically added to the code window at the bottom.
@@
 
-2. An analysis created by \texttt{intRo} can be reproduced in an R session to easily assess and extend the results.
+2. An analysis created by `intRo` can be reproduced in an R session to easily assess and extend the results.
 
-3. "Printing" the results of an \texttt{intRo} analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
+3. "Printing" the results of an `intRo` analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
 
-We designed the modules in such a way that the code printed to the console is the same as the code executed by the server, which makes the process of writing an \texttt{intRo} module significantly more seamless.
+We designed the modules in such a way that the code printed to the console is the same as the code executed by the server, which makes the process of writing an `intRo` module significantly more seamless.
 The workhorse function for this reproducibility framework is called `interpolate`, which is based on a function proposed by Hadley Wickham [-@interpolate].
 The idea of interpolation is that by using non-standard evaluation, the call can be captured, printed, and evaluated, with all the appropriate arguments substituted into the call automatically.
-This means that depending on the options selected in \texttt{intRo}, the corresponding arguments to functions are appropriately set for reproduction in a standard R session.
+This means that depending on the options selected in `intRo`, the corresponding arguments to functions are appropriately set for reproduction in a standard R session.
 
@@
 
-Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@bainomugisha2012survey] As implemented by \texttt{Shiny}, results automatically update when users interact with the interface.
+Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@bainomugisha2012survey] As implemented by Shiny, results automatically update when users interact with the interface.
 
-\texttt{intRo} leverages the reactive programming nature of \texttt{Shiny}, and as such is designed around the idea of user input cascading through the entire application.
-In a typical \texttt{Shiny} application, users interact with inputs that act as parameters to function, which in turn yield different results.
-Within \texttt{intRo}, the users are able to interact with and manipulate the data underlying the entire application.
-This posed many challenges in the creation of \texttt{intRo} and drove design decisions, namely timely save points according to the user's workflow, and reactive updating of variable lists tied to inputs across the entire application.
+intRo leverages the reactive programming nature of Shiny, and as such is designed around the idea of user input cascading through the entire application.
+In a typical Shiny application, users interact with inputs that act as parameters to function, which in turn yield different results.
+Within `intRo`, the users are able to interact with and manipulate the data underlying the entire application.
+This posed many challenges in the creation of `intRo` and drove design decisions, namely timely save points according to the user's workflow, and reactive updating of variable lists tied to inputs across the entire application.
 Because the user may experiment with different configurations or select different variables, we did not want to store all actions taken in the intRo session.
@@
 
-In the creation of \texttt{intRo} we walked a fine line between giving the user flexibility and having realistic usability.
-At the same time, \texttt{intRo} was created as a consumer of another package, \texttt{Shiny}, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability.
+In the creation of `intRo` we walked a fine line between giving the user flexibility and having realistic usability.
+At the same time, `intRo` was created as a consumer of another package, Shiny, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability.
 For a tangible example, consider the graphical summaries module.
 We only allow variables of a type consistent with the selected plot to be displayed.
-This is a conscious decision that limits an \texttt{intRo} user's flexibility, while maximizing the usability (by minimizing crashes) of the application.
-On the flip side of this, \texttt{Shiny} allows much higher flexibility.
-For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within \texttt{intRo}.
-However, \texttt{Shiny} does have limits on its flexibility based on the designers decisions for usability.
+This is a conscious decision that limits an `intRo` user's flexibility, while maximizing the usability (by minimizing crashes) of the application.
+On the flip side of this, Shiny allows much higher flexibility.
+For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within `intRo`.
+However, Shiny does have limits on its flexibility based on the designers decisions for usability.
 One current example is the slider element.
 This element allows for fixed width steps from its minimum to its maximum.
-The JavaScript library being utilized in \texttt{Shiny} allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript.
-This is an example of a decision made by the developers of \texttt{Shiny} to limit functionality in favor of usability of their package.
+The JavaScript library being utilized in Shiny allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript.
+This is an example of a decision made by the developers of Shiny to limit functionality in favor of usability of their package.
 
@@
 
-We believe \texttt{intRo} can be a powerful and effective tool for introductory statistics education.
+We believe `intRo` can be a powerful and effective tool for introductory statistics education.
 Its modular structure allows it to be flexible enough for many different applications and curriculums.
 Its ease-of-use allows the student to focus her attention on the statistics task at hand, rather than struggling with software licenses and confusing interface navigation.
-Reproducible code generated from each analysis can be used to spark an interest in \texttt{R} programming in those who might otherwise not be exposed to it.
+Reproducible code generated from each analysis can be used to spark an interest in R programming in those who might otherwise not be exposed to it.
 
-In addition to the current functionality, there are some practical improvements in the works that will make \texttt{intRo} more useful to both students and instructors.
-In particular, we have begun development on an \texttt{R} package which will allow \texttt{intRo} modules to be created automatically from user written \texttt{R} code.
-This package will generate the necessary file structure to allow the module's incorporation into \texttt{intRo} as well as translate user code to \texttt{intRo} compatible code and populate the necessary files.
-This will vastly improve \texttt{intRo}'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
-Additionally, we would like to utilize the expanded interactive capabilities of \texttt{ggvis} in order to make \texttt{intRo's} plots more engaging to students.
+In addition to the current functionality, there are some practical improvements in the works that will make `intRo` more useful to both students and instructors.
+In particular, we have begun development on an R package which will allow `intRo` modules to be created automatically from user written R code.
+This package will generate the necessary file structure to allow the module's incorporation into `intRo` as well as translate user code to `intRo` compatible code and populate the necessary files.
+This will vastly improve `intRo`'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
+Additionally, we would like to utilize the expanded interactive capabilities of ggvis in order to make \texttt{intRo's} plots more engaging to students.
 One way to do this would be implementing linked plots, in which interactions with one plot are reflected in other plots that illustrate the same data.
@@
 
-We hope to use \texttt{intRo} in courses to collect feedback regarding the ease of use and functionality.
+We hope to use `intRo` in courses to collect feedback regarding the ease of use and functionality.
 This will allow us to assess its usefulness relative to software used in the past, as well as gauge areas for improvement.
-Furthermore, we can determine the effectiveness of code printing on generating excitement from the students about programming in \texttt{R}.
+Furthermore, we can determine the effectiveness of code printing on generating excitement from the students about programming in R.
 
-Challenges do exist with regards to the wider adoption of \texttt{intRo}.
-For instance, we will need to monitor how well the server hosting \texttt{intRo} handles the load of dozens of students performing data analyses at once.
+Challenges do exist with regards to the wider adoption of `intRo`.
+For instance, we will need to monitor how well the server hosting `intRo` handles the load of dozens of students performing data analyses at once.
 If performance issues are encountered, the infrastructure used may need to be expanded to handle current and future load.
-An unknown quantity will be how feasible it is to increase adoption of \texttt{intRo} across the department, as well as to other universities.
-One limitation of \texttt{intRo} is that uploading a dataset beyond about 30,000 rows tends to be slow.
+An unknown quantity will be how feasible it is to increase adoption of `intRo` across the department, as well as to other universities.
+One limitation of `intRo` is that uploading a dataset beyond about 30,000 rows tends to be slow.
 Even once the data is successfully uploaded, the default modules produce results more slowly than with smaller datasets.
-This is a limitation that should be further investigated if and when \texttt{intRo} sees wider adoption.
+This is a limitation that should be further investigated if and when `intRo` sees wider adoption.
 
-Regardless, tools that focus on usability and extensibility in statistics education, such as \texttt{intRo}, are sure to encourage the next round of innovators to be interested and excited about statistical computing.
+Regardless, tools that focus on usability and extensibility in statistics education, such as `intRo`, are sure to encourage the next round of innovators to be interested and excited about statistical computing.
 

### 2016-02-19 jcgs-paper-2016/paper.Rmd
Sections

@@
 These advantages will be discussed at length in this text.
+The paper is structured as follows: Section 2 introduces the application, its features and its usability.
+Section 3 describes the process of creating a classroom instance of `intRo` tailored to the individual instructor.
+Section 4 provides technical details on the implementation of `intRo`.
+Finally, Section 5 discusses some future possibilities and limitations of the software.
 

### 2016-02-19 jcgs-paper-2016/paper.Rmd
references + section titles

--- 2016-02-19 jcgs-paper-2016/paper.Rmd
@@
 These advantages will be discussed at length in this text.
+The paper is structured as follows: section \ref{what-is-intro} introduces the application, its features and its usability.
+Section \ref{using-intro} describes the process of creating a classroom instance of `intRo` tailored to the individual instructor.
+Section \ref{understanding-intro} provides technical details on the implementation of `intRo`.
+Finally, Section \ref{conclusions-and-future-work} discusses some future possibilities and limitations of the software.
 
@@
 
-# Using intRo
+# Using `intRo`
 
-## intRo Package
+## `intRo` Package
 
@@
 
-# Understanding intRo
+# Understanding `intRo`
 

### 2016-02-19 jcgs-paper-2016/paper.Rmd
Final changes

@@
 Section \ref{understanding-intro} provides technical details on the implementation of `intRo`.
-Finally, Section \ref{conclusions-and-future-work} discusses some future possibilities and limitations of the software.
+Finally, section \ref{conclusions-and-future-work} discusses some future possibilities and limitations of the software.
 
@@
 
+Although users can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to customize their own instance.
+
 ## `intRo` Package
@@
 `intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.
-Currently, the package is only available and GitHub, and can be installed using the devtools package as follows:
+Currently, the package is only available on GitHub, and can be installed using the devtools package as follows:
 
@@
 
+To fully understand `intRo`, it is important to expand upon three underlying concepts.
+The first is the idea of modularity in the context of a Shiny application.
+The second is the idea of reproducibility in the context of a graphical user interface.
+Finally, the third is the idea of reactive programming in the context of a web application.
+
 ## Modules
+
 An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality.
@@
                          "xcont", 
-                         choices = setdiff(intro.categoricnames(), yselect), 
+                         choices = setdiff(intro.categoricnames(), 
+                                           yselect), 
                          selected = xselect)
@@
                          "ycont", 
-                         choices = setdiff(intro.categoricnames(), xselect), 
+                         choices = setdiff(intro.categoricnames(), 
+                                           xselect), 
                          selected = yselect)

### 2016-02-19 jcgs-paper-2016/paper.Rmd
More texttt

@@
   `intRo` is a web-based application for performing basic data analysis and statistical routines.
-Leveraging the power of `R` and `Shiny`, `intRo` implements common statistical functions in an extensible modular structure, while including a point-and-click interface for the novice statistician.
-This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as `Excel` and `JMP`.
+Leveraging the power of R and Shiny, `intRo` implements common statistical functions in an extensible modular structure, while including a point-and-click interface for the novice statistician.
+This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as Excel and JMP.
 `intRo` is currently deployed at the URL \url{http://www.intro-stats.com}.
 Within this paper, we introduce the application and explore the design decisions underlying `intRo`, as well as highlight some challenges and advantages of reactive programming.
-keywords: Graphical User Interface, Interactivity, Reproducibility, Reactive Programming, Statistical Software, Programming Paradigms
+keywords: Graphical User Interface, Interactivity, Programming Paradigms, Reactive Programming, Reproducibility, Statistical Software
 bibliography: refs.bib
@@
 
-## `intRo` Package
+## `intRo` package
 
@@
 
+# Supplementary material
+All code and documents related to this manuscript are available at \url{https://github.com/gammarama/intRo}.
+
 # References

### 2016-05-26 jcgs-paper-2016/paper.Rmd
Start restructuring and removing a lot

@@
 Within this paper, we introduce the application and explore the design decisions underlying `intRo`, as well as highlight some challenges and advantages of reactive programming.
-keywords: Graphical User Interface, Interactivity, Programming Paradigms, Reactive Programming, Reproducibility, Statistical Software
+keywords: Interactivity, Programming Paradigms, Reactive Programming, Reproducibility, Statistical Software
 bibliography: refs.bib
@@
 By default, the output will include the R code, but if the user elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
+This reproducibility framework has a number of key advantages:
+
+1.
+This eases a user who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
+Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many inexperienced users may have.
+
+2. An analysis created by `intRo` can be reproduced in an R session to easily assess and extend the results.
+
+3. "Printing" the results of an `intRo` analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
 
@@
 
-Each module is a self contained set of R code that is dynamically added to the application at run time.
-What this means is that `intRo` can be easily extended by the addition of modules within the framework underlying the application.
-This extensibility allows for `intRo` to be customized for an individual classroom setting, tailored to the needs of a particular course.
+# Understanding `intRo`
 
-As an example of possible customization, an instructor may wish to introduce hypothesis testing from a nonparametric framework rather than distributional.
-In this case, a wilcoxon rank sum test could be added easily as a module to be used instead of the T test.
-This addition would be fairly simple to complete.
-The user would fork `intRo` from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `libraries.R`, `observe.R`, `output.R`, `reactive.R`, and `ui.R`.
-Each of these files is written to comply with the reactive nature available in a Shiny application.
-Further details on the creation of new modules can be found in section \ref{modules}.
+To fully understand `intRo`, it is important to expand upon two underlying concepts.
+The first is the idea of modularity in the context of a Shiny application.
+The second is the idea of reactive programming in the context of a web application.
+
+## Modules
+
+An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality.
+`intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework.
+In this section, we detail the structure of these modules, and the process of creating and deploying new modules.
 
@@
 
-The process by which `intRo` is dynamically created upon running the application through module sourcing is detailed in figure \ref{fig:app_creation_modules}.
+`intRo`'s user interface and functionality is dynamically generated depending on the set of modules enabled.
+This process is detailed in figure \ref{fig:app_creation_modules}.
 The key driver to populating \texttt{server.R} and \texttt{ui.R} is the navigation file \texttt{modules.txt}, which defines the placement of each module.
@@
 In this way, we were able to have `intRo` be fully extensible, its structure and functionality dependent entirely on the modules present within the application.
-
-# Using `intRo`
-
-Although users can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to customize their own instance.
-
-## `intRo` package
-
-`intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.
-Currently, the package is only available on GitHub, and can be installed using the devtools package as follows:
-
-```{r, eval=FALSE}
-devtools::install_github("gammarama/intRo")
-```
-
-After installing the `intRo` package, the first function one should call is `download_intRo`.
-`download_intRo` takes as an argument a directory in which to store the application.
-By default, it selects the working directory of the R session.
-This function clones the application branch of the `intRo` repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
-
-Running `download_intRo` will produce an `intRo` folder in the specified folder.
-It can then be ran as any Shiny application, using Shiny's `runApp` command.
-However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
-`run_intRo` takes as argument the path to the folder containing the `intRo` application.
-It also takes several more optional arguments:
-
-* `enabled_modules`: A character vector containing the modules to enable
-* `theme`: A string representing a shinythemes theme to use
-* `...`: Additional arguments passed to Shiny's `runApp` function
-
-The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download `intRo` to my working directory, execute an `intRo` session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
-The series of calls to do so would be as follows:
-
-```{r, eval=FALSE}
-download_intRo()
-run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
-          theme = "cerulean")
-```
-
-Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
-
-If the intent is to use a specific instance of `intRo` where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
-The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package.
-Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the user's account.
-The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
-It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
-It also takes `...` as additional arguments to be passed into the `deployApp` routine.
-For example, if we wished to deploy the instance of `intRo` we ran previously, we would call it like so:
-
-```{r, eval=FALSE}
-deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
-             theme = "cerulean")
-```
-
-Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
-
-## In the classroom
-
-`intRo`'s ease of use lends itself to a natural presentation in an introductory statistics course.
-Statistics software typically used in introductory courses, such as JMP and Excel, have restrictive software licensing and platform requirements that can limit the software's accessibility to students.
-Even if these limitations can be overcome, these software programs have non-ideal interfaces for an exploratory data analysis.
-JMP, for instance, hides several basic regression functions such as a line of best fit behind small red arrows.
-Using R itself in the classroom is an appealing alternative, but suffers from some other problems.
-Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis.
-When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (this experience is made much easier with new tools like rmarkdown, but this again is a bit of a learning curve for the new statistics student).
-
-We believe that `intRo` targets these key limitations.
-The interface is designed to be quick, seamless, and easy to understand.
-Because `intRo` is web-based, the statistical analysis can be performed from nearly any modern device.
-R underlies the analysis, but users need not be familiar with R or programming to perform it.
-The focal point of the interface is the output panel, allowing the student to remain focused on the results and limiting the need to fight with the interface itself to get work done.
-For those students who may have a natural inclination for programming, executable code is generated to allow a local reproduction of the analysis.
-
-The features of `intRo` were selected in order to apply to the most commonly used lesson plans in early statistics courses.
-The textbook "The Art and Science of Learning from Data" by Agresti and Franklin [-@agresti2012statistics] references many of the focal points of introductory statistics, and `intRo` supports performing a majority of them.
-Part one, titled "Gathering and Exploring Data", covers graphical summaries, measures of spread and center, and avoiding the misuse of graphics in situations they are not appropriate.
-`intRo` includes modules for both graphical and numerical summaries covering these topics.
-Furthermore, `intRo` limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics (Table \ref{tab:plottypes}).
-
-\begin{table*}[ht!]
-\centering
-\begin{tabular}{|c|c|c|}
-\hline
-Plot Type&X Variable Type Allowed&Y Variable Type Allowed\\
-\hline
-Histogram&Numeric&N/A\\
-Normal Quantile Plot&Numeric&N/A\\
-Scatterplot&Numeric&Numeric\\
-Line Chart&Numeric&Numeric\\
-Boxplot&Categorical&Numeric\\
-Bar Chart&Categorical&Numeric\\
-Pareto Chart&Categorical&Numeric\\
-Mosaic Plot&Categorical&Categorical\\
-\hline
-\end{tabular}
-\caption{The plot and variable types allowed in `intRo`'s graphical module.
-For example, when the selected plottype is histogram, the interface displays only a single variable selection box (x), and the variables selectable in this box are only numeric.
-If a boxplot is selected, the x variable must be categorical and the y variable must be numeric, and the list of choices updates automatically to reflect this.}
-\label{tab:plottypes}
-\end{table*}
-
-Part two of this text discusses probability and sampling distributions, which is outside the current realm of `intRo`, but part three focuses on inferential statistics.
-The inferential statistics include confidence intervals, hypothesis testing, and comparing two variables, all supported natively in `intRo`.
-Finally, part four of the text discusses the association between variables, including regression, functions that `intRo` performs with ease.
-Some of the more advanced topics of this section are currently unsupported by `intRo`, including multiple regression.
-
-Statistics courses with a lab component can most benefit from the use of `intRo`.
-As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries.
-They simply upload a dataset and click the module appropriate to the section of the class.
-Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in.
-Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the `intRo` interface to obtain the results.
-For topics currently unsupported by `intRo`, the instructor can supplement its use with other web-based modules that are publicly available or create their own modules within `intRo` as detailed in section \ref{modules}.
-
-# Understanding `intRo`
-
-To fully understand `intRo`, it is important to expand upon three underlying concepts.
-The first is the idea of modularity in the context of a Shiny application.
-The second is the idea of reproducibility in the context of a graphical user interface.
-Finally, the third is the idea of reactive programming in the context of a web application.
-
-## Modules
-
-An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality.
-`intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework.
-In this section, we detail the structure of these modules, and the process of creating and deploying new modules.
 
@@
 In this case, a function `cont.table` is created which, depending on the values of the parameters, ultimately returns a contingency table.
-One important difference from a typical R script is that each call in the script is wrapped in a function called `interpolate`.
+One important difference from a typical R script is that each call in the script is wrapped in a function called `interpolate` [@interpolate].
 `interpolate` both executes the given R code on the server, and also writes the code executed to the script window at the bottom of `intRo`.
-More details on the reasons and choices behind `interpolate` can be found in section \ref{reproducibility}.
 
@@
 
-## Reproducibility
-Though they can make common actions quicker and interacting with software much simplier, many Graphical User Interfaces for scientific software packages come with a major drawback.
-That is, the actions taken by the user are typically not reproducible.
-Code scripting, on the other hand, allows us to share a file containing code, and another user can execute the code to reproduce the same results.
-In terms of the scientific process, where results can and should be scrutinized by other parties, the utility of GUIs are greatly limited.
-
-We have designed `intRo` modules to reproduce actions taken by the user so that they can obtain R code associated with each module interacted with.
-When a user chooses to store the results associated with a particular module, R code yielding the visible results is automatically added to the code window at the bottom.
-This framework yields several advantages:
-
-1.
-This eases a user who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
-Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many inexperienced users may have.
-
-2. An analysis created by `intRo` can be reproduced in an R session to easily assess and extend the results.
-
-3. "Printing" the results of an `intRo` analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
-
-We designed the modules in such a way that the code printed to the console is the same as the code executed by the server, which makes the process of writing an `intRo` module significantly more seamless.
-The workhorse function for this reproducibility framework is called `interpolate`, which is based on a function proposed by Hadley Wickham [-@interpolate].
-The idea of interpolation is that by using non-standard evaluation, the call can be captured, printed, and evaluated, with all the appropriate arguments substituted into the call automatically.
-This means that depending on the options selected in `intRo`, the corresponding arguments to functions are appropriately set for reproduction in a standard R session.
-
 ## Reactive design choices
@@
 This is an example of a decision made by the developers of Shiny to limit functionality in favor of usability of their package.
+
+# Using and Deploying `intRo`
+
+Although users can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and their own instance, perhaps with new modules or modified theming or functionality.
+`intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.
+Currently, the package is only available on GitHub, and can be installed using the devtools package as follows:
+
+```{r, eval=FALSE}
+devtools::install_github("gammarama/intRo")
+```
+
+After installing the `intRo` package, the first function one should call is `download_intRo`.
+`download_intRo` takes as an argument a directory in which to store the application.
+By default, it selects the working directory of the R session.
+This function clones the application branch of the `intRo` repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
+
+Running `download_intRo` will produce an `intRo` folder in the specified folder.
+It can then be ran as any Shiny application, using Shiny's `runApp` command.
+However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
+`run_intRo` takes as argument the path to the folder containing the `intRo` application.
+It also takes several more optional arguments:
+
+* `enabled_modules`: A character vector containing the modules to enable
+* `theme`: A string representing a shinythemes theme to use
+* `...`: Additional arguments passed to Shiny's `runApp` function
+
+The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download `intRo` to my working directory, execute an `intRo` session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
+The series of calls to do so would be as follows:
+
+```{r, eval=FALSE}
+download_intRo()
+run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+          theme = "cerulean")
+```
+
+Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
+
+If the intent is to use a specific instance of `intRo` where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
+The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package.
+Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the user's ShinyApps.io account.
+The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
+It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
+It also takes `...` as additional arguments to be passed into the `deployApp` routine.
+For example, if we wished to deploy the instance of `intRo` we ran previously, we would call it like so:
+
+```{r, eval=FALSE}
+deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+             theme = "cerulean")
+```
+
+Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
 
@@
 # Supplementary material
+
 All code and documents related to this manuscript are available at \url{https://github.com/gammarama/intRo}.

### 2016-05-26 jcgs-paper-2016/response.Rmd
Include paper rev1

---
title: "Response to reviewers"
author: "Eric Hare, Andee Kaplan"
date: "May 26, 2016"
output: html_document
---

# Reviewer 1

This is an intelligent paper describing interesting software, but it (and the associated software) suffer from three problems:

(1) The paper seems to me inappropriate for the JCGS. Because it focuses purely on software, I think that (once the other issues are addressed) it would be more appropriate for the Journal of Statistical Software or the R Journal.

(2) Although the software looks promising, unless I've missed something, it currently has very limited functionality -- indeed, not enough to sustain most introductory statistics courses. This isn't a permanent limitation, however, since the authors have created interesting infrastructure for extending the software. In its current state, the software doesn't quite look ready for prime time.

(3) The authors ignore other (and at present, much more capable) GUIs for R -- the R Commander and Deducer. Since these are the most direct competitors to intRo -- more so than, e.g., Excel and JMP -- it seems obvious that the authors should address similarities, differences, advantages, and disadvantages. One obvious difference is that these other GUIs run on individual computers rather than being web-based, and it would also make sense to address more generally the advantages and disadvantages of the two approaches. (For example, in my tests of intRo, I noticed slow response, even with very small data sets. Of course, it's hard to know to what to attribute that -- it could as easily have been a network issue as slowness at the server end.)

# Reviewer 2

The “Article Assessment” questions on the report form don’t work for a paper like this.
The authors report on some very impressive work. The simplicity of conception and execution is one of its biggest strengths. Their motivations are compelling, as is their system design and execution but I think the story-telling can be improved considerably.

I think the authors need to decide, “What sort of paper am I writing?” Included in the manuscript are elements of “a paper about teaching introductory statistics using a software system”, elements of “a paper about designing and building a simple extensible system”, and “a paper about how to use and extend a particular system”.

I think the most compelling story for the Journal of Computational and Graphical Statistics is the designing and building of a simple online system for intro stat that is easy to learn, use and extend – and teaches R as a side effect. This would leave room for a sister publication in something like the Journal of Statistics Education which actually focusses on teaching aspects.

If you buy this conception, the teaching-related aspects should be shortened, shaved of much detail and form the basis of an introductory motivational section that leads into a design brief and a plan of attack (using shiny etc). There needs to be high-level discussion of how the system is conceived and put together accompanied by high-level diagrams before descending into code fragments (much detailed code-fragment discussion could go to appendices). The paper just needs to convey the sense of how things are done. Actual details of how to customise and extend the system are provided in the system’s documentation.

Detailed comments follow. (Perhaps unfortunately, many draw on what I have said above.):
Everywhere – distinguish between user=student (who uses what is served up) and user=teacher (who customises what is served up)

p. 2: Another online analysis system implemented in shiny is iNZight Lite at http://lite.docker.stat.auckland.ac.nz/. It is more ambitious and therefore more complex. Also it does not have the “exposing the R code” capability.

p 3 bottom: Remove parentheses around “(and its corresponding code)” because this is actually one of the key features. R Commander should enter the discussion around here (even though it isn’t online)

P 5: Fig 2 shows a simple design that should be easy for casual users (as intro students really are) to get their heads around. Not all that keen on the use of “statistics” as a category heading in Panel 2 because it reinforces the ghettoizing of the word conflicting with a message that says everything you are doing here is statistics.

P 6 top: When you look at the actual R code generated it can be quite complex with its pipe structures etc which I would have thought does rather limit its introductory value.

P 6-7: discussion of particular modules should be in a “for example” vein. There is no need to justify what particular features are and are not included. (That would be in a teaching paper.)

P 7 top: Being able to choose only plot types that are “compatible” with the variables used does, however, preclude learning experiences of the form “this thing you want to do is not appropriate for this type of data.”

P 8: I think this descends to details too quickly. The big features are designing for extensibility, with conceptions about switching modules on and off and adding modules, and the interface as presented being assembled on the fly when the system starts up.

P 9-10 looks like appendix stuff to me.

Section 3 discusses downloading onto a local machine before customising the hosted online system which should be the focus (again conception and issues involved before any details).

P 11 near bottom: “as an application on the user’s account” puzzled me -- needs to say shinyapps.io account
Section 3.2 mainly belongs in up-front motivation. Table 1 is an implementation mechanism.
P 14 top: “currently unsupported” justifications are more relevant to a teaching paper than a software design paper.
P 14 Last para Section 3: could go into motivational discussion.
Section 4.1 is a big part of what I see as the real story but descends to details too quickly.
Most of Section 4.2 looks to me like more motivation, except the interpolate stuff which is an implementation mechanism
Section 4.3: more “main story”.

### 2016-05-26 jcgs-paper-2016/response.Rmd
Continue response

@@
 
+**We appreciate the very kind comments and hope that this revision has improved the story being told.**
+
 I think the authors need to decide, “What sort of paper am I writing?” Included in the manuscript are elements of “a paper about teaching introductory statistics using a software system”, elements of “a paper about designing and building a simple extensible system”, and “a paper about how to use and extend a particular system”.
-
 I think the most compelling story for the Journal of Computational and Graphical Statistics is the designing and building of a simple online system for intro stat that is easy to learn, use and extend – and teaches R as a side effect.
 This would leave room for a sister publication in something like the Journal of Statistics Education which actually focusses on teaching aspects.
+
+**We have taken care to significantly shift the focus away from the teaching aspects, and towards the extensible software system aspects.
+In particular, we have removed the lengthy teaching discussion and moved elements of it to the motivational section.
+We have also trimmed much of the content involving supported and unsupported functionality.
+Specific changes are noted in the responses to the further points raised below.**
 
@@
 
-Detailed comments follow. (Perhaps unfortunately, many draw on what I have said above.):
 Everywhere – distinguish between user=student (who uses what is served up) and user=teacher (who customises what is served up)
@@
 
-p 3 bottom: Remove parentheses around “(and its corresponding code)” because this is actually one of the key features. R Commander should enter the discussion around here (even though it isn’t online)
+p 3 bottom: Remove parentheses around “(and its corresponding code)” because this is actually one of the key features.
+
+**Done**
+
+R Commander should enter the discussion around here (even though it isn’t online)
 
@@
 
+**We appreciate the comment.
+In teaching R Workshops at Iowa State, we observed that among the majority of the introductory R students we encountered, the pipe structure was often preferred to using traditional nested function calls.
+The pipe structure is also the standard by which `ggvis` plots are created, which we are using as the backend to our graphical module.
+It would also be a relatively simple task for an instructor to modify or remove the piping structure in a modified version of `intRo` for use in a particular class.**
+
 P 6-7: discussion of particular modules should be in a “for example” vein. There is no need to justify what particular features are and are not included. (That would be in a teaching paper.)
+
+**We have removed the in-depth discussion about the particular modules, and instead more vaguely described the functionality.
+We have also made sure to indicate that the full documentation is available online, which covers the supported functionality of every provided module.**
 
@@
 
+**We have also gone back and forth on this issue.
+The reasoning behind choosing only plot types compatible is that the user will see the different plot types available as they select different variables.
+So while the teaching moment could take place post-selection (by, for instance, producing an error if a plot type is used with incompatible variable types), we chose to have the teaching moment be a step earlier, where the student may see the list and wonder why the plot type they thought they wanted is not actually an option.
+Further, by doing a transformation of a variable, the plot types will automatically appear if they are compatible.**
+
 P 8: I think this descends to details too quickly.
 The big features are designing for extensibility, with conceptions about switching modules on and off and adding modules, and the interface as presented being assembled on the fly when the system starts up.
+
+**We agree.
+We have restructured this portion of the paper to merely discuss the motivation behind an extensible system, with mentions of disabling and enabling modules.
+We have moved the details to the Understanding intRo section with the details on modules.**
 
@@
 
+**We have moved this section to after the discussion on customizing intRo. It now comes just before the conclusion.**
+
 P 11 near bottom: “as an application on the user’s account” puzzled me -- needs to say shinyapps.io account
+
+**Done**
+
 Section 3.2 mainly belongs in up-front motivation. Table 1 is an implementation mechanism.
+
+**We have merged remnants of Section 3.2 into Section 2 / motivation.
+We have also drastically cut down the content to focus this paper more on the extensible systems side, rather than the teaching side.
+As such, we have also removed Table 1.**
+
 P 14 top: “currently unsupported” justifications are more relevant to a teaching paper than a software design paper.
+
+**We have removed this comment.**
+
 P 14 Last para Section 3: could go into motivational discussion.
+
+**We have moved this paragraph to the motivational discussion.**
+
 Section 4.1 is a big part of what I see as the real story but descends to details too quickly.
+
 Most of Section 4.2 looks to me like more motivation, except the interpolate stuff which is an implementation mechanism
+
+**We have merged most of the content from the former Section 4.2 (on reproducibility) with Section 2, and removed a discussion on interpolation.**
+
 Section 4.3: more “main story”.
+
+**We agree and have emphasized this section as one of the two primary pieces to the main story, the other being modularity.**

### 2016-05-26 jcgs-paper-2016/paper.Rmd
Continue response

@@
 These advantages will be discussed at length in this text.
-The paper is structured as follows: section \ref{what-is-intro} introduces the application, its features and its usability.
-Section \ref{using-intro} describes the process of creating a classroom instance of `intRo` tailored to the individual instructor.
-Section \ref{understanding-intro} provides technical details on the implementation of `intRo`.
+The paper is structured as follows: section \ref{what-is-intro} introduces the application, its features and its usability, and provides motivations for the design decisions underlying it.
+Section \ref{understanding-intro} provides technical details on the design decisions of `intRo`, in particular, its modular and extensible structure.
+Section \ref{deploying-intro-instances} describes the process of deploying a classroom instance of `intRo` tailored to the individual instructor.
 Finally, section \ref{conclusions-and-future-work} discusses some future possibilities and limitations of the software.
@@
 In this instance, a user clicks on the Graphical tab to create a mosaic plot.
-The user sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.
+The user sees the plot, and elects to click the save button to store the plot and its corresponding code to the final compendium.
 
@@
 The documentation is available at http://gammarama.github.io/intRo and hyperlinked within the application itself.
+The documentation describes the specific functionality provided by the default `intRo` modules.
+
+Statistics courses with a lab component can most benefit from the use of `intRo`.
+As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries.
+They simply upload a dataset and click the module appropriate to the section of the class.
+Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in.
+Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the `intRo` interface to obtain the results.
 
@@
 Modules are self-contained pieces of functionality which implement common statistical procedures that are often used in introductory statistics classes.
-
-Within the data category there are two modules, sources and transformation.
-The data sources module allows the user to select or upload a dataset.
-Several datasets are pre-included, which can be selected via a drop-down menu.
-The user can check a box to upload a dataset, and then select the CSV file from her computer.
-Upon selection or successful upload of a dataset, a table view of the data will be visible on the right.
-From this module the user can select a subset of the data either by value or random selection.
-The dataset selected in the data module will be made available to all other modules of `intRo`.
-The data transformation module allows transformation of numeric variables to categorical and categorical variable to numeric.
-Additionally, numeric variable can be transformed using a power transformation.
-The transformed variables can be saved to the dataset for use elsewhere in `intRo`.
-
-The graphical summaries module provides several common graphical displays of variables in the dataset.
-First, the user selects the plot desired, and either a single variable or two variables from the dataset, depending on plot type.
-The variable available to choose will only be those compatible with plot type selected.
-The plot visible on the right will automatically update when these values are selected.
-The numeric summaries module is another method of computing summaries of the data.
-The user is presented with a list of each variable from their dataset.
-The user can select one or more different variables, and will be presented with a table of some summary statistics.
-The summary statistics displayed vary depending on the type of the variables.
-For instance, for a numeric variable, a five number summary will be displayed.
-Character or factor variables will display counts of each unique value of that particular variable.
-Additionally, the user has the ability to display summaries by a grouping variable.
-
-Within the set of statistics modules, there are three simple statistics tasks that an introductory statistics student will learn in her first semester.
-The contingency table module allows the user to compute a contingency table for two categorical variables.
-The results can either be displayed as counts, row percentages, column percentages, or total percentages.
-The simple linear regression module allows the user to compute a simple linear regression of a dependent variable (y) on an independent variable (x).
-The module will automatically filter any variables that are non-numeric so that the user may only select variables that are suitable for a simple linear regression.
-Results of the regression will be displayed on the right.
-A table of parameter estimates, a scatterplot of the data along with a line of best fit, The R-Squared value, and residual diagnostic plots will all be displayed.
-A user also has the ability to save residuals and predicted values to their dataset for further analysis.
-The T test module allows the user to perform either a one variable or two variable Student's T test.
-The user can select configuration options, including the direction of the test and the confidence level of the test.
-The results, including the t statistic, degrees of freedom, p-value, and a confidence interval, will be displayed in a text box in the results panel.
+These modules form the core of `intRo` and are discussed at length in the next section.
 
@@
 
-# Using and Deploying `intRo`
+# Deploying `intRo` Instances
 

### 2016-05-26 jcgs-paper-2016/response.Rmd
More response updates

@@
 In its current state, the software doesn't quite look ready for prime time.
+
+**Our intent was not for `intRo` to sustain an introductory statistics course, but rather to supplement it.
+In other words, limited functionality was actually a design goal from the beginning so as not to overwhelm a student seeing statistics for the first time.
+While we agree that it is likely there will be course topics that `intRo` doesn't support, if additional functionality is desired by the instructor, more modules can be added to support this.**
 
@@
 In teaching R Workshops at Iowa State, we observed that among the majority of the introductory R students we encountered, the pipe structure was often preferred to using traditional nested function calls.
+TODO: modify this to show that pipe structure is used more frequently.
 The pipe structure is also the standard by which `ggvis` plots are created, which we are using as the backend to our graphical module.
@@
 So while the teaching moment could take place post-selection (by, for instance, producing an error if a plot type is used with incompatible variable types), we chose to have the teaching moment be a step earlier, where the student may see the list and wonder why the plot type they thought they wanted is not actually an option.
+We felt that for a lot of introductory students, seeing a red error message is a frustrating moment and can discourage them from having any sort of autonomy within the software.
 Further, by doing a transformation of a variable, the plot types will automatically appear if they are compatible.**

### 2016-05-26 jcgs-paper-2016/paper.Rmd
More response updates

@@
 
-Although users can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and their own instance, perhaps with new modules or modified theming or functionality.
+Although users can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and deploy their own instance, perhaps with new modules or modified theming or functionality.
 `intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.

### 2016-06-06 jcgs-paper-2016/paper-original.Rmd
Rename to original

--- 2016-05-26 jcgs-paper-2016/paper.Rmd
@@
 By default, the output will include the R code, but if the user elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
-This reproducibility framework has a number of key advantages:
-
-1.
-This eases a user who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
-Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many inexperienced users may have.
-
-2. An analysis created by `intRo` can be reproduced in an R session to easily assess and extend the results.
-
-3. "Printing" the results of an `intRo` analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
 
@@
 
-# Understanding `intRo`
+Each module is a self contained set of R code that is dynamically added to the application at run time.
+What this means is that `intRo` can be easily extended by the addition of modules within the framework underlying the application.
+This extensibility allows for `intRo` to be customized for an individual classroom setting, tailored to the needs of a particular course.
 
-To fully understand `intRo`, it is important to expand upon two underlying concepts.
-The first is the idea of modularity in the context of a Shiny application.
-The second is the idea of reactive programming in the context of a web application.
-
-## Modules
-
-An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality.
-`intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework.
-In this section, we detail the structure of these modules, and the process of creating and deploying new modules.
+As an example of possible customization, an instructor may wish to introduce hypothesis testing from a nonparametric framework rather than distributional.
+In this case, a wilcoxon rank sum test could be added easily as a module to be used instead of the T test.
+This addition would be fairly simple to complete.
+The user would fork `intRo` from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `libraries.R`, `observe.R`, `output.R`, `reactive.R`, and `ui.R`.
+Each of these files is written to comply with the reactive nature available in a Shiny application.
+Further details on the creation of new modules can be found in section \ref{modules}.
 
@@
 
-`intRo`'s user interface and functionality is dynamically generated depending on the set of modules enabled.
-This process is detailed in figure \ref{fig:app_creation_modules}.
+The process by which `intRo` is dynamically created upon running the application through module sourcing is detailed in figure \ref{fig:app_creation_modules}.
 The key driver to populating \texttt{server.R} and \texttt{ui.R} is the navigation file \texttt{modules.txt}, which defines the placement of each module.
@@
 In this way, we were able to have `intRo` be fully extensible, its structure and functionality dependent entirely on the modules present within the application.
+
+# Using `intRo`
+
+Although users can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to customize their own instance.
+
+## `intRo` package
+
+`intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.
+Currently, the package is only available on GitHub, and can be installed using the devtools package as follows:
+
+```{r, eval=FALSE}
+devtools::install_github("gammarama/intRo")
+```
+
+After installing the `intRo` package, the first function one should call is `download_intRo`.
+`download_intRo` takes as an argument a directory in which to store the application.
+By default, it selects the working directory of the R session.
+This function clones the application branch of the `intRo` repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
+
+Running `download_intRo` will produce an `intRo` folder in the specified folder.
+It can then be ran as any Shiny application, using Shiny's `runApp` command.
+However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
+`run_intRo` takes as argument the path to the folder containing the `intRo` application.
+It also takes several more optional arguments:
+
+* `enabled_modules`: A character vector containing the modules to enable
+* `theme`: A string representing a shinythemes theme to use
+* `...`: Additional arguments passed to Shiny's `runApp` function
+
+The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download `intRo` to my working directory, execute an `intRo` session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
+The series of calls to do so would be as follows:
+
+```{r, eval=FALSE}
+download_intRo()
+run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+          theme = "cerulean")
+```
+
+Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
+
+If the intent is to use a specific instance of `intRo` where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
+The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package.
+Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the user's account.
+The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
+It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
+It also takes `...` as additional arguments to be passed into the `deployApp` routine.
+For example, if we wished to deploy the instance of `intRo` we ran previously, we would call it like so:
+
+```{r, eval=FALSE}
+deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+             theme = "cerulean")
+```
+
+Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
+
+## In the classroom
+
+`intRo`'s ease of use lends itself to a natural presentation in an introductory statistics course.
+Statistics software typically used in introductory courses, such as JMP and Excel, have restrictive software licensing and platform requirements that can limit the software's accessibility to students.
+Even if these limitations can be overcome, these software programs have non-ideal interfaces for an exploratory data analysis.
+JMP, for instance, hides several basic regression functions such as a line of best fit behind small red arrows.
+Using R itself in the classroom is an appealing alternative, but suffers from some other problems.
+Many students in introductory statistics courses have not been exposed to programming languages and concepts, which could require a substantial overhead in order to begin a basic statistical analysis.
+When results need to be submitted as homework, getting the output into a document can be a frustrating and tedious experience (this experience is made much easier with new tools like rmarkdown, but this again is a bit of a learning curve for the new statistics student).
+
+We believe that `intRo` targets these key limitations.
+The interface is designed to be quick, seamless, and easy to understand.
+Because `intRo` is web-based, the statistical analysis can be performed from nearly any modern device.
+R underlies the analysis, but users need not be familiar with R or programming to perform it.
+The focal point of the interface is the output panel, allowing the student to remain focused on the results and limiting the need to fight with the interface itself to get work done.
+For those students who may have a natural inclination for programming, executable code is generated to allow a local reproduction of the analysis.
+
+The features of `intRo` were selected in order to apply to the most commonly used lesson plans in early statistics courses.
+The textbook "The Art and Science of Learning from Data" by Agresti and Franklin [-@agresti2012statistics] references many of the focal points of introductory statistics, and `intRo` supports performing a majority of them.
+Part one, titled "Gathering and Exploring Data", covers graphical summaries, measures of spread and center, and avoiding the misuse of graphics in situations they are not appropriate.
+`intRo` includes modules for both graphical and numerical summaries covering these topics.
+Furthermore, `intRo` limits the display of variables to ones that are appropriate for a chosen plot type, helping to reinforce the appropriateness of particular graphics (Table \ref{tab:plottypes}).
+
+\begin{table*}[ht!]
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+Plot Type&X Variable Type Allowed&Y Variable Type Allowed\\
+\hline
+Histogram&Numeric&N/A\\
+Normal Quantile Plot&Numeric&N/A\\
+Scatterplot&Numeric&Numeric\\
+Line Chart&Numeric&Numeric\\
+Boxplot&Categorical&Numeric\\
+Bar Chart&Categorical&Numeric\\
+Pareto Chart&Categorical&Numeric\\
+Mosaic Plot&Categorical&Categorical\\
+\hline
+\end{tabular}
+\caption{The plot and variable types allowed in `intRo`'s graphical module.
+For example, when the selected plottype is histogram, the interface displays only a single variable selection box (x), and the variables selectable in this box are only numeric.
+If a boxplot is selected, the x variable must be categorical and the y variable must be numeric, and the list of choices updates automatically to reflect this.}
+\label{tab:plottypes}
+\end{table*}
+
+Part two of this text discusses probability and sampling distributions, which is outside the current realm of `intRo`, but part three focuses on inferential statistics.
+The inferential statistics include confidence intervals, hypothesis testing, and comparing two variables, all supported natively in `intRo`.
+Finally, part four of the text discusses the association between variables, including regression, functions that `intRo` performs with ease.
+Some of the more advanced topics of this section are currently unsupported by `intRo`, including multiple regression.
+
+Statistics courses with a lab component can most benefit from the use of `intRo`.
+As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries.
+They simply upload a dataset and click the module appropriate to the section of the class.
+Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in.
+Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the `intRo` interface to obtain the results.
+For topics currently unsupported by `intRo`, the instructor can supplement its use with other web-based modules that are publicly available or create their own modules within `intRo` as detailed in section \ref{modules}.
+
+# Understanding `intRo`
+
+To fully understand `intRo`, it is important to expand upon three underlying concepts.
+The first is the idea of modularity in the context of a Shiny application.
+The second is the idea of reproducibility in the context of a graphical user interface.
+Finally, the third is the idea of reactive programming in the context of a web application.
+
+## Modules
+
+An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality.
+`intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework.
+In this section, we detail the structure of these modules, and the process of creating and deploying new modules.
 
@@
 In this case, a function `cont.table` is created which, depending on the values of the parameters, ultimately returns a contingency table.
-One important difference from a typical R script is that each call in the script is wrapped in a function called `interpolate` [@interpolate].
+One important difference from a typical R script is that each call in the script is wrapped in a function called `interpolate`.
 `interpolate` both executes the given R code on the server, and also writes the code executed to the script window at the bottom of `intRo`.
+More details on the reasons and choices behind `interpolate` can be found in section \ref{reproducibility}.
 
@@
 
+## Reproducibility
+Though they can make common actions quicker and interacting with software much simplier, many Graphical User Interfaces for scientific software packages come with a major drawback.
+That is, the actions taken by the user are typically not reproducible.
+Code scripting, on the other hand, allows us to share a file containing code, and another user can execute the code to reproduce the same results.
+In terms of the scientific process, where results can and should be scrutinized by other parties, the utility of GUIs are greatly limited.
+
+We have designed `intRo` modules to reproduce actions taken by the user so that they can obtain R code associated with each module interacted with.
+When a user chooses to store the results associated with a particular module, R code yielding the visible results is automatically added to the code window at the bottom.
+This framework yields several advantages:
+
+1.
+This eases a user who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
+Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many inexperienced users may have.
+
+2. An analysis created by `intRo` can be reproduced in an R session to easily assess and extend the results.
+
+3. "Printing" the results of an `intRo` analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
+
+We designed the modules in such a way that the code printed to the console is the same as the code executed by the server, which makes the process of writing an `intRo` module significantly more seamless.
+The workhorse function for this reproducibility framework is called `interpolate`, which is based on a function proposed by Hadley Wickham [-@interpolate].
+The idea of interpolation is that by using non-standard evaluation, the call can be captured, printed, and evaluated, with all the appropriate arguments substituted into the call automatically.
+This means that depending on the options selected in `intRo`, the corresponding arguments to functions are appropriately set for reproduction in a standard R session.
+
 ## Reactive design choices
@@
 This is an example of a decision made by the developers of Shiny to limit functionality in favor of usability of their package.
-
-# Using and Deploying `intRo`
-
-Although users can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and their own instance, perhaps with new modules or modified theming or functionality.
-`intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.
-Currently, the package is only available on GitHub, and can be installed using the devtools package as follows:
-
-```{r, eval=FALSE}
-devtools::install_github("gammarama/intRo")
-```
-
-After installing the `intRo` package, the first function one should call is `download_intRo`.
-`download_intRo` takes as an argument a directory in which to store the application.
-By default, it selects the working directory of the R session.
-This function clones the application branch of the `intRo` repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
-
-Running `download_intRo` will produce an `intRo` folder in the specified folder.
-It can then be ran as any Shiny application, using Shiny's `runApp` command.
-However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
-`run_intRo` takes as argument the path to the folder containing the `intRo` application.
-It also takes several more optional arguments:
-
-* `enabled_modules`: A character vector containing the modules to enable
-* `theme`: A string representing a shinythemes theme to use
-* `...`: Additional arguments passed to Shiny's `runApp` function
-
-The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download `intRo` to my working directory, execute an `intRo` session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
-The series of calls to do so would be as follows:
-
-```{r, eval=FALSE}
-download_intRo()
-run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
-          theme = "cerulean")
-```
-
-Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
-
-If the intent is to use a specific instance of `intRo` where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
-The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package.
-Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the user's ShinyApps.io account.
-The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
-It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
-It also takes `...` as additional arguments to be passed into the `deployApp` routine.
-For example, if we wished to deploy the instance of `intRo` we ran previously, we would call it like so:
-
-```{r, eval=FALSE}
-deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
-             theme = "cerulean")
-```
-
-Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
 
@@
 # Supplementary material
-
 All code and documents related to this manuscript are available at \url{https://github.com/gammarama/intRo}.

### 2016-06-06 jcgs-paper-2016/paper.Rmd
Update response + references

@@
 
-Multiple software packages have recently been written in an attempt to spur student interest in introductory statistics and R programming.
+Multiple software packages have recently been written in an attempt to spur interest in introductory statistics and R programming.
 DataCamp's [@datacamp] courses are user-friendly ways to learning basic R programming and data analysis techniques.
@@
 Project MOSAIC's goal is to faciliate this learning, but using the package requires a knowledge of R programming that the introductory student may not have.
+iNZight Lite [@inzight] also attempts to expose students to data analysis without requiring programming knowledge.
+While this tool has a lot of functionality and a very broad scope, it does not include reproducible R code, and therefore has less of a focus on spurring interest in coding for students.
+R Commander [@fox2005] and Deducer [@fellows2012] provides a graphical front-end to many statistical functions in R, but is not web-based and thus requires local installation and configuration.
 

### 2016-06-06 R/set_options.R
Bump package version, switch to inference instead of statistics

@@ -4,6 +4,7 @@
 #' @param enabled_modules The modules to enable
 #' @param theme The shinythemes theme to use
 #' @param google_analytics The google analytics tracking ID to use
+#' @importFrom utils installed.packages
 #' @examples
 #' \dontrun{
 #'     deploy_intRo()
@@ -13,9 +14,9 @@ set_options <- function(path, enabled_modules, theme, google_analytics) {
     if (is.null(enabled_modules)) enabled_modules <- c("data/transform", 
                                                        "summaries/graphical",
                                                        "summaries/numerical",
-                                                       "statistics/contingency",
-                                                       "statistics/regression",
-                                                       "statistics/t_test")
+                                                       "inference/contingency",
+                                                       "inference/regression",
+                                                       "inference/t_test")
     if (is.null(theme)) theme <- "united"
     if (is.null(google_analytics)) google_analytics <- "intRo_tracking_code"
     

### 2016-06-06 jcgs-paper-2016/cover_letter.Rmd
cover letter template

---
author: 
 - Eric Hare
 - PhD Student 
 - Department of Statistics 
 - Iowa State University
date: "`r format(Sys.time(), '%B %d, %Y')`"
opening: To whom it may concern,
closing: Best regards,
address: 
 - Editor
 - Journal of Computational and Graphical Statistics
return-address:
 - Eric Hare
 - Department of Statistics
 - Iowa State University
 - Ames, IA 50011
output:
  pdf_document:
    template: template-letter.tex
geometry: margin=1in
---

Thank you in advance for reviewing our manuscript "Introductory statistics with `intRo`", which I have submitted for consideration of publication in the Journal of Computational and Graphical Statistics.

This paper details an interactive web-based statistical software application intended for use in an introductory statistics classroom. We introduce the tool as well as an accompanying R package to ease the process of deploying intRo within a curriculum. Additionally, the paper describes the process for extending the functionality in intRo in order to personalize the software.

Thank you for your consideration of our submission.

### 2016-06-06 jcgs-paper-2016/response.Rmd
Update response + references
Update response and paper based on meeting

@@
 
-This is an intelligent paper describing interesting software, but it (and the associated software) suffer from three problems:
-
 (1) The paper seems to me inappropriate for the JCGS.
 Because it focuses purely on software, I think that (once the other issues are addressed) it would be more appropriate for the Journal of Statistical Software or the R Journal.
+
+**We appreciate the reviewer's perspective on the context in which our paper fits.
+While we are focusing on a software system, the aim of JCGS includes to "extend the use of computational and graphical methods in statistics and data analysis".
+It also contains "more on numerical graphical displays and methods, and perceptions." While this is an educational piece of software, this paper focuses on more general computational concepts of reproducibility, reactivity, and modularity, in the context of bringing statistical content to an end user.
+As such, we feel it does have a place within JCGS.**
 
@@
 In other words, limited functionality was actually a design goal from the beginning so as not to overwhelm a student seeing statistics for the first time.
-While we agree that it is likely there will be course topics that `intRo` doesn't support, if additional functionality is desired by the instructor, more modules can be added to support this.**
+While we agree with the reviewer that it is likely there will be course topics that `intRo` doesn't support, if additional functionality is desired by the instructor, more modules can be added to support this.
+We have extended the discussion of modularity in intRo to further emphasize this idea.**
 
@@
 Of course, it's hard to know to what to attribute that -- it could as easily have been a network issue as slowness at the server end.)
+
+**We have included references and a brief discussion on some other GUIs for R, including R Commander and Deducer.
+We have described some of the differences in our approach to providing a front-end.
+As for the slow-response time in intRo, this can certainly occur in a couple of different scenarios.
+One would be if the student is working with a large dataset, as mentioned in Section 5.
+A second reason for slowdown would be if the student's internet connection is performing slowly.
+This is an unfortunate limitation of web-based interactive software.
+However, we have detailed the process by which a student may download and run an intRo instance locally to counteract this possibility.**
 
@@
 This would leave room for a sister publication in something like the Journal of Statistics Education which actually focusses on teaching aspects.
+If you buy this conception, the teaching-related aspects should be shortened, shaved of much detail and form the basis of an introductory motivational section that leads into a design brief and a plan of attack (using shiny etc).
+There needs to be high-level discussion of how the system is conceived and put together accompanied by high-level diagrams before descending into code fragments (much detailed code-fragment discussion could go to appendices).
+The paper just needs to convey the sense of how things are done.
+Actual details of how to customise and extend the system are provided in the system’s documentation.
 
@@
 
-If you buy this conception, the teaching-related aspects should be shortened, shaved of much detail and form the basis of an introductory motivational section that leads into a design brief and a plan of attack (using shiny etc).
-There needs to be high-level discussion of how the system is conceived and put together accompanied by high-level diagrams before descending into code fragments (much detailed code-fragment discussion could go to appendices).
-The paper just needs to convey the sense of how things are done.
-Actual details of how to customise and extend the system are provided in the system’s documentation.
+Everywhere – distinguish between user=student (who uses what is served up) and user=teacher (who customises what is served up)
 
-Everywhere – distinguish between user=student (who uses what is served up) and user=teacher (who customises what is served up)
+**We went through and replaced instances of "user" with either "student" or "teacher" as appropriate.**
 
@@
 Also it does not have the “exposing the R code” capability.
+
+**We have included a mention of iNZight Lite, along with the shared goals and some differences with intRo.**
 
@@
 
+**We have included R Commander in the discussion.**
+
 P 5: Fig 2 shows a simple design that should be easy for casual users (as intro students really are) to get their heads around.
 Not all that keen on the use of “statistics” as a category heading in Panel 2 because it reinforces the ghettoizing of the word conflicting with a message that says everything you are doing here is statistics.
+
+**We have updated the application (and the screenshot) to refer to the contingency, regression, and t-test modules category heading as "inference" rather than "statistics", to help emphasize that numerical and graphical summaries are also statistics.**
 
@@
 
+**We have move the implementation details of dynamic UI generation to the appendix.**
+
 Section 3 discusses downloading onto a local machine before customising the hosted online system which should be the focus (again conception and issues involved before any details).
@@
 
+**We have eased the transition into this section (now Section 3.1) by talking more generally about the benefits behind a modular structure.
+We also have made sure to clarify that our concept of modules is not to be confused with Shiny's new modules feature**
+
 Most of Section 4.2 looks to me like more motivation, except the interpolate stuff which is an implementation mechanism

### 2016-06-06 jcgs-paper-2016/paper.Rmd
Update response and paper based on meeting

@@
 In particular, `intRo` is \emph{easy} to use and can be an \emph{exciting} part of learning statistics.
-Additionally, `intRo` is an \emph{extensible} tool, allowing for a user of `intRo` to tailor the tool for his or her own classroom needs.
+Additionally, `intRo` is an \emph{extensible} tool, allowing for a course instructor using `intRo` to tailor the tool for his or her own classroom needs.
 
@@
 Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in `intRo`.
-In this instance, a user clicks on the Graphical tab to create a mosaic plot.
-The user sees the plot, and elects to click the save button to store the plot and its corresponding code to the final compendium.
+In this instance, a student clicks on the Graphical tab to create a mosaic plot.
+The student sees the plot, and elects to click the save button to store the plot and its corresponding code to the final compendium.
 
@@
 \includegraphics[width=\linewidth]{images/user_experience.pdf}
-\caption{A typical user experience of generating a result in `intRo`.
-In this instance, a user clicks on the Graphical tab to create a mosaic plot.
-The user sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.}
+\caption{A typical student experience of generating a result in \texttt{intRo}.
+In this instance, a student clicks on the Graphical tab to create a mosaic plot.
+The student sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.}
 \label{fig:user_experience}
@@
 The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules.
-To the user, a module is simply a page of statistics functionality that maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code.
+To the student, a module is simply a page of statistics functionality that maintains a consistent layout, helping the student to become familiar with the location of the options, the results, and the code.
 Figure \ref{fig:ui} highlights the five elements that comprise the `intRo` interface.
@@
 \includegraphics[width=\linewidth]{images/ui_annotate.pdf}
-\caption{The five elements that comprise the `intRo` application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
+\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
 \label{fig:ui}
@@
  2. **Side Navigation** - The side navigation panel includes a list of data analysis tasks.
- 3. **Options Panel** - The options panel includes task-specific options which the user can use to customize their results.
+ 3. **Options Panel** - The options panel includes task-specific options which the student can use to customize their results.
  4. **Results Panel** - The results pane displays the result of the selected module and options.
@@
 Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in.
-Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the `intRo` interface to obtain the results.
+Early labs can walk the student through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the student can easily navigate the `intRo` interface to obtain the results.
 
@@
 This code creation element of `intRo` is meant to generate excitement about programming in R and empower students to feel that they can generate code as well.
-`intRo` uses rmarkdown's render function in order to print the results, by dymanically executing the user's R script.
-By default, the output will include the R code, but if the user elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
+`intRo` uses rmarkdown's render function in order to print the results, by dymanically executing the student's R script.
+By default, the output will include the R code, but if the student elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
 This reproducibility framework has a number of key advantages:
@@
 1.
-This eases a user who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
-Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many inexperienced users may have.
+This eases a student who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
+Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many students may have.
 
@@
 `intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework.
-In this section, we detail the structure of these modules, and the process of creating and deploying new modules.
+A high-level diagram of this process is given in figure \ref{fig:app_creation_modules}.
+Based on the modules contained within the `modules` folder, the interface and functionality of the app is dynamically generated at runtime.
 
@@
 \includegraphics[width=\linewidth]{images/app_creation_modules.pdf}
-\caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within `intRo`.}
+\caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.}
 \label{fig:app_creation_modules}
@@
 
-`intRo`'s user interface and functionality is dynamically generated depending on the set of modules enabled.
-This process is detailed in figure \ref{fig:app_creation_modules}.
-The key driver to populating \texttt{server.R} and \texttt{ui.R} is the navigation file \texttt{modules.txt}, which defines the placement of each module.
-The interface is then created with the following statement.
+Section \ref{dynamic-ui-generation} provides some technical details on how we implemented this.
+For the rest of this section, we focus on the structure and development of the modules themselves, to aid in the process of creating and deploying new modules.
 
-\footnotesize
-```{r ui, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
-## Source ui
-module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
-sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)
+Modularity was a design decision we focused on from the start of `intRo's` development.
+There are some practical benefits to thinking of related statistics and data science functionality in terms of modules:
 
-## mylist is a list containing the different ui module code
-## Create the UI
-shinyUI(
-    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
-               tabPanel(title="", icon=icon("home"),
-                        fluidRow(
-                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
-                        )
-               ), ...
-    ))
-```
-\normalsize
+1.
+Course instructors can easily enable or disable functionality depending on the exact needs of the particular class being taught.
+Because this is done at run-time, including new modules is as simple as downloading and placing it within 'intRo's` modules folder, or removing existing modules from that folder.
+2. Errors can be more easily isolated to specific components. For instance, if an error is encountered, simply disabling the module can provide a temporary workaround while the issue is identified.
+3. Modularity helps to organize the different pieces of code into functionality chunks that make it easier for developers to maintain.
 
-The key piece of code being the `do.call` statement loading the list of ui elements from the module's \texttt{ui.R} file. The server functions are then dynamically generated using a similar method.
+`intRo` modules are not to be confused with Shiny modules [@shinymodules].
+Shiny modules are a recent feature added to Shiny which allows the bundling of inputs and outputs into a single set of functionality.
+They are more general and suitable for any application.
+`intRo` modules are specifically for statistics functionality and work within the `intRo` application only.
 
-\footnotesize
-```{r server, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
-shinyServer(function(input, output, session) {
-  ## Module info
-  module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
-
-  ## Modules
-  types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
-  modules_tosource <- file.path("modules", 
-                                apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))
-
-  ## Source the modules
-  for (mod in modules_tosource) {
-      source(mod, local = TRUE)
-  }
-})
-```
-\normalsize
-In this way, we were able to have `intRo` be fully extensible, its structure and functionality dependent entirely on the modules present within the application.
-
-A module consists of the following scripts:
+An `intRo` module consists of the following scripts:
 
@@
 * *observe.R* - Shiny observer code typically used to update choices of an input box.
-* *output.R* - Shiny output code defining the results of the analysis that should be displayed to the user.
+* *output.R* - Shiny output code defining the results of the analysis that should be displayed to the student.
 * *reactive.R* - Shiny reactives, typically containing data that depend on inputs.
@@
 
-This script defines all the inputs and outputs that the user will see.
-The only requirements from `intRo`'s perspective are (1) that there exist a store button at the bottom of the middle panel for storing the results of the analysis in the code script, and (2) that configuration options appear in the width 4 column in the middle, and output appears in the width 8 column on the right.
+This script defines all the inputs and outputs that the student will see.
+The only requirements from `intRo's` perspective are (1) that there exist a store button at the bottom of the middle panel for storing the results of the analysis in the code script, and (2) that configuration options appear in the width 4 column in the middle, and output appears in the width 8 column on the right.
 The remaining input and output definitions depend on the statistical analysis or transformation being performed.
@@
 In a typical Shiny application, users interact with inputs that act as parameters to function, which in turn yield different results.
-Within `intRo`, the users are able to interact with and manipulate the data underlying the entire application.
-This posed many challenges in the creation of `intRo` and drove design decisions, namely timely save points according to the user's workflow, and reactive updating of variable lists tied to inputs across the entire application.
-Because the user may experiment with different configurations or select different variables, we did not want to store all actions taken in the intRo session.
-Rather, each module includes a button allowing the user to explicitly store the output visible in the results panel into the R script.
-This way, output is only stored when the user is satisfied, and the resulting output is not cluttered with unnecessary information.
+Within `intRo`, the students are able to interact with and manipulate the data underlying the entire application.
+This posed many challenges in the creation of `intRo` and drove design decisions, namely timely save points according to the student's workflow, and reactive updating of variable lists tied to inputs across the entire application.
+Because the student may experiment with different configurations or select different variables, we did not want to store all actions taken in the intRo session.
+Rather, each module includes a button allowing the student to explicitly store the output visible in the results panel into the R script.
+This way, output is only stored when the student is satisfied, and the resulting output is not cluttered with unnecessary information.
 
-In the creation of `intRo` we walked a fine line between giving the user flexibility and having realistic usability.
+In the creation of `intRo` we walked a fine line between giving the student flexibility and having realistic usability.
 At the same time, `intRo` was created as a consumer of another package, Shiny, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability.
@@
 
-Although users can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and deploy their own instance, perhaps with new modules or modified theming or functionality.
+Although students can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and deploy their own instance, perhaps with new modules or modified theming or functionality.
 `intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.
@@
 
-If the intent is to use a specific instance of `intRo` where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
+If the intent is to use a specific instance of `intRo` where many students will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
 The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package.
-Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the user's ShinyApps.io account.
+Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the instructor's ShinyApps.io account.
 The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
@@
 This package will generate the necessary file structure to allow the module's incorporation into `intRo` as well as translate user code to `intRo` compatible code and populate the necessary files.
-This will vastly improve `intRo`'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
+This will vastly improve `intRo's` flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
 Additionally, we would like to utilize the expanded interactive capabilities of ggvis in order to make \texttt{intRo's} plots more engaging to students.
@@
 
+# Appendix
+
+## Dynamic UI Generation
+
+`intRo's` user interface and functionality is dynamically generated depending on the set of modules enabled.
+The key driver to populating \texttt{server.R} and \texttt{ui.R} is the modules folder, the directory structure of which defines the placement of each module.
+The interface is then created with the following statement.
+
+\footnotesize
+```{r ui, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
+## Source ui
+mylist <- list()
+old_heading <- ""
+for (i in seq_along(modules)) {
+    my.module <- strsplit(modules[i], "/")[[1]]
+    if (my.module[1] != old_heading) {
+        mylist[[length(mylist) + 1]] <- Hmisc::capitalize(my.module[1])
+        old_heading <- my.module[1]
+    }
+    mylist[[length(mylist) + 1]] <- get(paste(my.module[2], "ui", sep = "_"))
+}
+
+## mylist is a list containing the different ui module code
+## Create the UI
+shinyUI(
+    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
+               tabPanel(title="", icon=icon("home"),
+                        fluidRow(
+                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
+                        )
+               ), ...
+    ))
+```
+\normalsize
+
+The key piece of code being the `do.call` statement loading the list of ui elements from the module's \texttt{ui.R} file. The server functions are then dynamically generated using a similar method.
+
+\footnotesize
+```{r server, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
+shinyServer(function(input, output, session) {
+    types <- c("helper.R", "observe.R", "reactive.R", "output.R")
+
+    modules_tosource <- file.path("modules", apply(expand.grid(modules, types), 1, paste, collapse = "/"))
+
+    for (mod in modules_tosource) {
+        source(mod, local = TRUE)
+    }
+})
+```
+\normalsize
+In this way, we were able to have `intRo` be fully extensible, its structure and functionality dependent entirely on the modules present within the application.
+
 # References

### 2016-06-14 jcgs-paper-2016/paper.Rmd
some comments from AK before meeting

@@
 Still, some compromises must be made when using software like JMP, including dealing with software licenses and unsupported desktop platforms.
+**AK: This is heavy on the JMP**
 
@@
 iNZight Lite [@inzight] also attempts to expose students to data analysis without requiring programming knowledge.
-While this tool has a lot of functionality and a very broad scope, it does not include reproducible R code, and therefore has less of a focus on spurring interest in coding for students.
-R Commander [@fox2005] and Deducer [@fellows2012] provides a graphical front-end to many statistical functions in R, but is not web-based and thus requires local installation and configuration.
+While this tool has a lot of functionality and a very broad scope, it does not include reproducible R code, and therefore has less of a focus on spurring interest in coding for students **AK: this seems a little out of place, did we even mention the interest thing yet?**.
+R Commander [@fox2005] and Deducer [@fellows2012] provide a graphical front-end to many statistical functions in R, but are not web-based and thus require local installation and configuration.
 
@@
 
-The modules included in `intRo` are split into three higher level categories - data, summaries, and statistics.
+The modules included in `intRo` are split into three higher level categories - data, summaries, and inference.
 Under each of these categories, there are seven default modules, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
@@
 Based on the modules contained within the `modules` folder, the interface and functionality of the app is dynamically generated at runtime.
+**AK: We need a more detailed description of this image.**
 
@@
 Modularity was a design decision we focused on from the start of `intRo's` development.
-There are some practical benefits to thinking of related statistics and data science functionality in terms of modules:
+There are some practical benefits to thinking of related statistics and data science functionality in terms of modules: **AK: We have too many lists in this paper.**
 

### 2016-06-15 jcgs-paper-2016/paper.Rmd
more comments

@@
 Course instructors can easily enable or disable functionality depending on the exact needs of the particular class being taught.
-Because this is done at run-time, including new modules is as simple as downloading and placing it within 'intRo's` modules folder, or removing existing modules from that folder.
+Because this is done at run-time, including new modules is as simple as downloading and placing it within `intRo's` modules folder, or removing existing modules from that folder.
 2. Errors can be more easily isolated to specific components. For instance, if an error is encountered, simply disabling the module can provide a temporary workaround while the issue is identified.
@@
 The modules provided with `intRo` are contained in the modules folder.
-The top level directory in the modules folder defines the category of the module (currently `data`, `summaries`, or `statistics`).
+The top level directory in the modules folder defines the category of the module (currently `data`, `summaries`, or `inference`).
 Within each of these categories is a folder named according to the name of the module.
@@
 
-Since the `contingency` module performs a statistical analysis, it is part of the statistics category, and hence can be found in the `intRo` repository at **modules/statistics/contingency**.
+**AK: Should we walk through a difference models? Since this is not inference, it is confusing**
+Since the `contingency` module performs a statistical analysis, it is part of the inference category, and hence can be found in the `intRo` repository at **modules/inference/contingency**.
 Let's first examine *helper.R*:
@@
 
-Reactive programming is a programming paradigm that ``tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies.'' [@bainomugisha2012survey] As implemented by Shiny, results automatically update when users interact with the interface.
+Reactive programming is a programming paradigm that "tackles issues posed by event-driven applications by providing abstractions to express programs as reactions to external events and having the language automatically manage the flow of time (by conceptually supporting simultaneity), and data and computation dependencies" [@bainomugisha2012survey].
+As implemented by Shiny, results automatically update when users interact with the interface.
 
@@
 This is an example of a decision made by the developers of Shiny to limit functionality in favor of usability of their package.
+
+**AK: I feel like this section could use a little more beef, but I'm not sure how**
 
@@
 
-We believe `intRo` can be a powerful and effective tool for introductory statistics education.
+We believe **AK: we believe is weak language, consider removing** `intRo` can be a powerful and effective tool for introductory statistics education.
 Its modular structure allows it to be flexible enough for many different applications and curriculums.
@@
 If performance issues are encountered, the infrastructure used may need to be expanded to handle current and future load.
-An unknown quantity will be how feasible it is to increase adoption of `intRo` across the department, as well as to other universities.
+An unknown quantity will be how feasible it is to increase adoption of `intRo` across the department **AK: not sure about "the department"**, as well as to other universities.
 One limitation of `intRo` is that uploading a dataset beyond about 30,000 rows tends to be slow.

### 2016-06-15 modules/inference/nonparametric/helper.R
clean up nonparametric module

@@ -1,4 +1,4 @@
-nonparametrictable <- function (intro.data, x, y, conflevel, althyp, hypval) {
+nonparametrictest <- function (intro.data, x, y, conflevel, althyp, hypval) {
     interpolate(~(wilcox.test(x = df$x, y = df$y, conf.level = conf, alternative = althyp, mu = hypval)),
                   df = quote(intro.data),
                   x = x,

### 2016-06-15 modules/inference/nonparametric/output.R

@@ -1,4 +1,4 @@
-    output$nonparametrictable <- renderText({
-        return(paste(capture.output(nonparametrictable(intro.data(), input$group1_non, input$group2_non, input$conflevel_non, input$althyp_non, input$hypval_non)), collapse = "\n"))
+    output$nonparametrictest <- renderPrint({
+        return(nonparametrictest(intro.data(), input$group1_non, input$group2_non, input$conflevel_non, input$althyp_non, input$hypval_non))
     })
 

### 2016-06-15 modules/inference/nonparametric/ui.R

@@ -18,6 +18,6 @@ nonparametric_ui <- tabPanel("Nonparametric",
                                                   
                                                   column(8,
                                                          tags$b("Nonparametric Results"),
-                                                         verbatimTextOutput("nonparametrictable")
+                                                         verbatimTextOutput("nonparametrictest")
                                                   )
                                          )

### 2016-06-15 jcgs-paper-2016/paper.Rmd
Changes from mtg with andee

@@
 However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student.
-This additional time commitment may explain why introductory courses often utilize software applications such as JMP, even though instructors tend to use a tool like R in their own work.
-Still, some compromises must be made when using software like JMP, including dealing with software licenses and unsupported desktop platforms.
-**AK: This is heavy on the JMP**
+This additional time commitment may explain why introductory courses often utilize point-and-click applications, even though instructors tend to use a tool like R in their own work.
+Still, some compromises must be made when using many graphical applications, including dealing with software licenses and unsupported desktop platforms.
 
@@
 
-Multiple software packages have recently been written in an attempt to spur interest in introductory statistics and R programming.
+Multiple software packages have recently been written in an attempt to spur interest in  R programming and statistics.
 DataCamp's [@datacamp] courses are user-friendly ways to learning basic R programming and data analysis techniques.
@@
 Project MOSAIC's goal is to faciliate this learning, but using the package requires a knowledge of R programming that the introductory student may not have.
-iNZight Lite [@inzight] also attempts to expose students to data analysis without requiring programming knowledge.
-While this tool has a lot of functionality and a very broad scope, it does not include reproducible R code, and therefore has less of a focus on spurring interest in coding for students **AK: this seems a little out of place, did we even mention the interest thing yet?**.
 R Commander [@fox2005] and Deducer [@fellows2012] provide a graphical front-end to many statistical functions in R, but are not web-based and thus require local installation and configuration.
+iNZight Lite [@inzight] also attempts to expose students to data analysis without requiring programming knowledge, but does not include reproducible R code, and therefore has less of a focus on spurring interest in coding for students.
 
@@
 By default, the output will include the R code, but if the student elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
-This reproducibility framework has a number of key advantages:
 
-1.
-This eases a student who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
+This reproducibility framework has three key advantages.
+First, this eases a student who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
 Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many students may have.
-
-2. An analysis created by `intRo` can be reproduced in an R session to easily assess and extend the results.
-
-3. "Printing" the results of an `intRo` analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
+Second, an analysis created by `intRo` can be reproduced in an R session to easily assess and extend the results.
+Finally, "printing" the results of an `intRo` analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
 
@@
 A high-level diagram of this process is given in figure \ref{fig:app_creation_modules}.
-Based on the modules contained within the `modules` folder, the interface and functionality of the app is dynamically generated at runtime.
-**AK: We need a more detailed description of this image.**
+`intRo` modules are split up into multiple R scripts which are included either in Shiny's user interface or server definitions.
+At runtime, the `intRo` sources in the specified modules (contained in the modules folder) to dynamically generate the functionality available in the application.
+This allows for the specific functionality needed to be determined and adjusted by the individual course instructor.
+In this example, the instructor is electing to include a nonparametric module, which is not enabled by default, to allow the students to perform a wilcoxon rank sum test.
 
@@
 \includegraphics[width=\linewidth]{images/app_creation_modules.pdf}
-\caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.}
+\caption{This figure depicts how the Shiny \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.
+\texttt{intRo} modules are split up into multiple R scripts which are included either in Shiny's user interface or server definitions.
+At runtime, the \texttt{intRo} sources in the specified modules (contained in the modules folder) to dynamically generate the functionality available in the application.
+This allows for the specific functionality needed to be determined and adjusted by the individual course instructor.
+In this example, the instructor is electing to include a nonparametric module, which is not enabled by default, to allow the students to perform a wilcoxon rank sum test.}
 \label{fig:app_creation_modules}
@@
 Modularity was a design decision we focused on from the start of `intRo's` development.
-There are some practical benefits to thinking of related statistics and data science functionality in terms of modules: **AK: We have too many lists in this paper.**
-
-1.
+There are some practical benefits to thinking of related statistics and data science functionality in terms of modules.
 Course instructors can easily enable or disable functionality depending on the exact needs of the particular class being taught.
 Because this is done at run-time, including new modules is as simple as downloading and placing it within `intRo's` modules folder, or removing existing modules from that folder.
-2. Errors can be more easily isolated to specific components. For instance, if an error is encountered, simply disabling the module can provide a temporary workaround while the issue is identified.
-3. Modularity helps to organize the different pieces of code into functionality chunks that make it easier for developers to maintain.
+Furthermore, errors can be more easily isolated to specific components.
+For instance, if an error is encountered, simply disabling the module can provide a temporary workaround while the issue is identified.
+Finally, modularity helps to organize the different pieces of code into functionality chunks that make it easier for developers to maintain.
 
@@
 This folder houses the previously defined scripts.
-As an example, we will walk through the contents of the `contingency` module.
+As an example, we will walk through the contents of the `nonparametric` module as previously mentioned in this section.
 
-**AK: Should we walk through a difference models? Since this is not inference, it is confusing**
-Since the `contingency` module performs a statistical analysis, it is part of the inference category, and hence can be found in the `intRo` repository at **modules/inference/contingency**.
+Since the `nonparametric` module performs a statistical test, it is part of the inference category, and hence can be found in the `intRo` repository at **modules/inference/nonparametric**.
 Let's first examine *helper.R*:
@@
 ```{r, eval=FALSE}
-cont.table <- function(intro.data, x, y, type, digits) {  
-    interpolate(
-        ~ (my.tbl <- cbind(rbind(table(df$y, df$x), 
-                      Total = colSums(table(df$y, df$x))), 
-                Total = c(rowSums(table(df$y, df$x))[-(length(df$x))], 
-                          sum(table(df$y, df$x))))),
+nonparametrictest <- function(intro.data, x, y, 
+                                conflevel, althyp, hypval) {
+    interpolate(~(wilcox.test(x = df$x, y = df$y, 
+                              conf.level = conf, 
+                              alternative = althyp, 
+                              mu = hypval)),
                   df = quote(intro.data),
@@
                   y = y,
+                  conf = conflevel,
+                  althyp = althyp,
+                  hypval = hypval,
                   mydir = userdir, 
                   `_env` = environment(), 
-        file = "code_contingency.R"
-    )
-
-    # Additional formatting code suppressed for readability
-    ...
-
-    interpolate(~(format(round(my.tbl, digits = new.digits))), 
-                mydir = userdir, 
-                `_env` = environment(), 
-                file = "code_contingency.R",
-                append = TRUE)
+                  file = "code_nonparametric.R")
 }
@@
 This script is most immediately similar to standard R code.
-In this case, a function `cont.table` is created which, depending on the values of the parameters, ultimately returns a contingency table.
+In this case, a function `nonparametrictest` is created which, depending on the values of the parameters, ultimately returns the result of a wilcoxon rank sum test.
 One important difference from a typical R script is that each call in the script is wrapped in a function called `interpolate` [@interpolate].
@@
 
-Because all the code needed to implement a contingency table is found in the `base` and `stats` package, the `libraries.R` file is empty for the `contingency` module.
+Because all the code needed to implement a wilcoxon rank sum test is found in the `base` and `stats` package, the `libraries.R` file is empty for the `nonparametric` module.
+Additionally, there are no reactive objects defined, so `reactive.R` is also empty.
 `observe.R`, which defines the Shiny observers needed, is reproduced below:
@@
 observe({
-    xselect <- ifelse(checkVariable(intro.data(), input$xcont),
-                      input$xcont, 
-                      intro.categoricnames()[1])
-    yselect <- ifelse(checkVariable(intro.data(), input$ycont), 
-                      input$ycont, 
-                      intro.categoricnames()[2])
-
-    updateSelectizeInput(session, 
-                         "xcont", 
-                         choices = setdiff(intro.categoricnames(), 
-                                           yselect), 
-                         selected = xselect)
-    updateSelectizeInput(session, 
-                         "ycont", 
-                         choices = setdiff(intro.categoricnames(), 
-                                           xselect), 
-                         selected = yselect)
+    updateSelectizeInput(session, "group1_non", 
+                         choices = intro.numericnames(), 
+                         selected = ifelse(checkVariable(
+                             intro.data(), input$group1_non), 
+                             input$group1_non, 
+                             intro.numericnames()[1]))
+    updateSelectizeInput(session, "group2_non", 
+                         choices = intro.numericnames(), 
+                         selected = ifelse(checkVariable(
+                             intro.data(), input$group2_non), 
+                             input$group2_non, 
+                             intro.numericnames()[2]))
 })
 
-observeEvent(input$store_contingency, {
+observeEvent(input$store_nonparametric, {
     cat(paste0("\n\n", paste(readLines(
-        file.path(userdir, "code_contingency.R")), 
+        file.path(userdir, "code_nonparametric.R")), 
         collapse = "\n")), 
@@
 For further discussion of reactivity, see section \ref{reactive-design-choices}.
-In this example, observers are created to ensure that the choices of variable for the `contingency` module are only categorical variables.
-This is accomplished by utilizing the global reactive `intro.categoricnames()`, which returns a character vector containing the variables in the current dataset that are categorical.
+In this example, observers are created to ensure that the choices of variable for the `nonparametric` module are only numeric variables.
+This is accomplished by utilizing the global reactive `intro.numericnames()`, which returns a character vector containing the variables in the current dataset that are numeric.
 Finally, there is an event observer to store code generated from the module into the overall code script upon clicking the store button.
@@
 
-The `output.R` and `reactive.R` code is very simple:
+The `output.R` code is very simple:
 
 ```{r, eval=FALSE}
-output$conttable <- renderTable({
-    return(intro.contingency())
+output$nonparametrictest <- renderPrint({
+    return(nonparametrictable(intro.data(), input$group1_non, 
+                              input$group2_non, input$conflevel_non, 
+                              input$althyp_non, input$hypval_non))
 })
@@
 
-```{r, eval=FALSE}
-intro.contingency <- reactive({
-    if (is.null(intro.data())) return(NULL)
-    if (input$xcont == input$ycont || 
-        length(categoricNames(intro.data())) < 2) return(NULL)
-
-    my.conttable <- cont.table(intro.data(), 
-                               input$xcont, input$ycont, 
-                               input$conttype, input$contdigits)
-
-    return(my.conttable)
-})
-```
-
-`intro.contingency` is defined as a reactive object.
-Some small error checks are performed at the beginning to ensure that the data is present and available, and that two different categorical variables are selected.
-Assuming those conditions hold, a call to the previously defined `helper.R` function is made with the appropriate parameters, and the results are returned.
-The `output.R` script then simply uses Shiny's `renderTable` function to display the resulting table.
+The `output.R` script then simply uses Shiny's `renderPrint` function to display the resulting table.
 
@@
 ```{r, eval=FALSE}
-contingency_ui <- tabPanel("Contingency",
+nonparametric_ui <- tabPanel("Nonparametric",
     column(4,
          wellPanel(
-             selectizeInput("xcont", label = "X Variable (x)", 
-                            choices = categoricNames(mpg), 
-                            selected = categoricNames(mpg)[5]),
-
-             # Additional inputs suppressed for readability
-             ...
+             selectizeInput("group1_non", label = "Group 1 (x)", 
+                            choices = numericNames(mpg), 
+                            selected = numericNames(mpg)[1]),
+             selectizeInput("group2_non", "Group 2 (y)", 
+                            choices = numericNames(mpg), 
+                            selected = numericNames(mpg)[2]),
 
@@
 
-             tags$button("", id = "store_contingency", type = "button", 
-                         class = "btn action-button", 
-                         list(icon("save"), "Store Contingency Result"), 
+             selectizeInput("althyp_non", "Alternative Hypothesis", 
+                            c("Two-Sided" = "two.sided", 
+                              "Greater" = "greater", "Less" = "less")),
+             numericInput("hypval_non", "Hypothesized Value",
+                          value = 0),
+             sliderInput("conflevel_non", "Confidence Level",
+                         min=0.01, max=0.99, step=0.01, value=0.95),
+
+             hr(),
+
+             tags$button("", id = "store_nonparametric", type = "button", 
+                         class = "btn action-button", list(icon("save"), 
+                         "Store Nonparametric Result"), 
                          onclick = "$('#top-nav a:has(> .fa-print, 
@@
     column(8,
-         tableOutput("conttable")
+         tags$b("Nonparametric Results"),
+         verbatimTextOutput("nonparametrictest")
     )
@@
 This is an example of a decision made by the developers of Shiny to limit functionality in favor of usability of their package.
-
-**AK: I feel like this section could use a little more beef, but I'm not sure how**
 
@@
 
-We believe **AK: we believe is weak language, consider removing** `intRo` can be a powerful and effective tool for introductory statistics education.
+`intRo` can be a powerful and effective tool for introductory statistics education.
 Its modular structure allows it to be flexible enough for many different applications and curriculums.
@@
 If performance issues are encountered, the infrastructure used may need to be expanded to handle current and future load.
-An unknown quantity will be how feasible it is to increase adoption of `intRo` across the department **AK: not sure about "the department"**, as well as to other universities.
+An unknown quantity will be how feasible it is to increase adoption of `intRo` across Iowa State, as well as to other universities.
 One limitation of `intRo` is that uploading a dataset beyond about 30,000 rows tends to be slow.

### 2016-06-15 jcgs-paper-2016/paper.Rmd
Width cutoff for tidy opts

@@
 \footnotesize
-```{r ui, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
+```{r ui, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=50)}
 ## Source ui
@@
 \footnotesize
-```{r server, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
+```{r server, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=50)}
 shinyServer(function(input, output, session) {

### 2016-06-15 modules/data/sources/libraries.R
Load ggplot2 in the data module

@@ -1,3 +1,4 @@
 if (!("package:dplyr" %in% search())) interpolate(~(library(dplyr)), file = "code_All.R", mydir = userdir, append = TRUE, nodupes = TRUE)
 if (!("package:lubridate" %in% search())) interpolate(~(library(lubridate)), file = "code_All.R", mydir = userdir, append = TRUE, nodupes = TRUE)
 if (!("package:YaleToolkit" %in% search())) interpolate(~(library(YaleToolkit)), file = "code_All.R", mydir = userdir, append = TRUE, nodupes = TRUE)
+if (!("package:ggplot2" %in% search())) interpolate(~(library(ggplot2)), file = "code_All.R", mydir = userdir, append = TRUE, nodupes = TRUE)

### 2016-06-21 presentations/user-2016/index.Rmd
add useR! talk

--- 2015-08-11 jsm-2015/andee/index.Rmd
@@
 author: "Andee Kaplan & Eric Hare <br/> Iowa State University"
-date: "August 11, 2015 <br> http://j.mp/intRo-JSM2015"
+date: "June 28, 2016 <br> http://j.mp/intRo-JSM2015"
 output: ioslides_presentation
@@
  - `R` is great, but requires students to have some knowledge/interest in programming
- - `JMP` is powerful, but too big
+ - `JMP`, `Deducer`, `Rcmdr` are powerful, but too big
       - Licenses and inconsistent UI
@@
 ## What is `intRo`?
- - A simple **web-based** application for performing basic data analysis and statistical routines
+ - A simple **web-based** application for performing basic data analysis and statistical routines and accompanying utility package
  - Built using `R` and `Shiny`
@@
 
-# First look | http://intro-stats.com
+# Take a look | http://intro-stats.com
+
+# Deploying intRo
+
+## intRo package
+
+ - Installation: 
+    ```{r eval=FALSE}
+    devtools::install_github("rstudio/shinyapps") 
+    devtools::install_github("gammarama/intRo")`
+    ```
+
+ - Functions:
+    - `download_intRo` - Downloads a current revision of intRo to your machine
+    - `run_intRo` - Runs an intRo session locally with the specified options
+    - `deploy_intRo` - Deploys an instance of intRo to ShinyApps.io with the specified options
 
@@
 
-## Presenting choices (Cont'd)
-<img src="images/graphical_table.png">
-
 ## Producing clean code
@@
 ## Producing clean code (Cont'd)
-But how do we handle uploaded data?
+
+ - Each call is wrapped in a function called `interpolate` [@interpolate]
+    - Executes the given R code on the server
+    - Writes the code executed to the script window at the bottom of intRo
 
 ```{r, eval=FALSE}
-file.choose <- function () {
-    return(input$data_own[,"datapath"])
+nonparametric_test <- function(intro.data, x, y, ...) {
+  interpolate(~(wilcox.test(x = df$x, 
+                            y = df$y, ...),
+                  df = quote(intro.data),
+                  x = x,
+                  y = y,
+                  ...,  
+                  `_env` = environment(), 
+                  file = "code_nonparametric.R")
 }
-```
-
-Now we can provide to the user *and* the server:
-
-```{r, eval=FALSE}
-intro.data <- read.csv(file.choose())
 ```
@@
 
-## What's next?
- - Module creation package
- - Server load
- - Use in other classes
-
-## Module creation package
+## Module creation
 *Modularity* is a key feature of `intRo`, but module creation is currently:
@@
 
-**Idea**: R package (with associated Shiny app?) to automate creation of `intRo` modules
+**Idea**: Include funcitonality in current R package to automate creation of `intRo` modules
 

### 2016-06-21 presentations/user-2016/index.Rmd
fix bot.ly link

@@
 author: "Andee Kaplan & Eric Hare <br/> Iowa State University"
-date: "June 28, 2016 <br> http://j.mp/intRo-JSM2015"
+date: "June 28, 2016 <br> http://bit.ly/intRo-useR2016"
 output: ioslides_presentation

### 2016-06-23 user-2016/index.Rmd
cut some things per @erichare suggestions

--- 2016-06-21 presentations/user-2016/index.Rmd
@@
 
- - Storing the user's results
  - Presenting only relevant options to the user
  - Producing clean and executable R code
-
-## Storing results
- - Users must be able to print results, but only the results they want
- - Reactivity makes this somewhat challenging
-
-**Solution**: Store Buttons
-
-<img src="images/store.png" height="300"></img>
-
-## Storing results (Cont'd)
- - Each module contains, at the bottom, a Store button
- - When Store is pressed, code is stored in code panel
-     - Additional challenge: All options selected must also be stored!
- - When user goes to print results, only this particular code is executed
 
@@
 
- - Each call is wrapped in a function called `interpolate` [@interpolate]
+ - Each call is wrapped in a function called `interpolate`
     - Executes the given R code on the server
@@
 nonparametric_test <- function(intro.data, x, y, ...) {
+  ### <b>
   interpolate(~(wilcox.test(x = df$x, 
+  ### </b>
                             y = df$y, ...),
@@
 
-## Server load
-Limited testing has been done to assess how `intRo` handles heavy user load
-
- - Dozens of students accessing app at once
- - Simultaneous larger computations like printing results
-
-More server resources may need to be devoted to handle this, particularly if `intRo` is more widely adopted
-
 ## More classes

### 2016-06-24 jcgs-paper-2016/response.Rmd
Update piping structure comment

@@
 **We appreciate the comment.
-In teaching R Workshops at Iowa State, we observed that among the majority of the introductory R students we encountered, the pipe structure was often preferred to using traditional nested function calls.
-TODO: modify this to show that pipe structure is used more frequently.
+In teaching R Workshops at Iowa State, we observed that among many of the introductory R students we encountered, the pipe structure was often preferred to using traditional nested function calls.
 The pipe structure is also the standard by which `ggvis` plots are created, which we are using as the backend to our graphical module.
-It would also be a relatively simple task for an instructor to modify or remove the piping structure in a modified version of `intRo` for use in a particular class.**
+We do understand, however, that certain instructors may wish to include a modified version of the default modules removing the piping, which by following the procedure described in the paper would be a relatively simple task.**
 

### 2016-06-28 user-2016/index.Rmd
changed a word

@@
  - `JMP`, `Deducer`, `Rcmdr` are powerful, but too big
-      - Licenses and inconsistent UI
+      - Licenses and installation
  - New tools recently released to spark an interest in `R`

### 2016-07-19 jcgs-paper-2016/paper.Rmd
Finish up the revision

@@
 Leveraging the power of R and Shiny, `intRo` implements common statistical functions in an extensible modular structure, while including a point-and-click interface for the novice statistician.
-This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as Excel and JMP.
+This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages.
 `intRo` is currently deployed at the URL \url{http://www.intro-stats.com}.
@@
 
-The widespread adoption of R as a tool for statistical analysis has undoubtedly been an important development for the scientific community.
-However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student.
-This additional time commitment may explain why introductory courses often utilize point-and-click applications, even though instructors tend to use a tool like R in their own work.
+The widespread adoption of R [@r-stat] as a tool for statistical analysis has undoubtedly been an important development for the scientific community.
+However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student [@5359977].
+This additional time commitment may explain why introductory courses often utilize point-and-click applications, even if the instructor himself/herself uses R in their own work.
 Still, some compromises must be made when using many graphical applications, including dealing with software licenses and unsupported desktop platforms.
@@
 This in turn led us to create an introductory statistics application which we call `intRo`, available at http://www.intro-stats.com.
-`intRo` offers a number of key advantages over traditional statistics software, including ease of use and an aim to foster student interest in coding.
+`intRo` offers a number of key advantages over traditional statistics software, including ease of access and an aim to foster student interest in coding.
+Attempting to entirely hide the programming aspect from students, even in introductory classes, is a lost opportunity to get students interested in statistical computing.
+It is also a lost opportunity reaching students who learn differently or have a computational background.
+Another advantage is its modular structure, which allows course instructors to tailor the application towards the needs of a particular class, rather than accept a piece of software as is.
 Additionally, `intRo` stands apart from new tools in that it is a supplement to an existing class, fully usable by a beginning statistics student.
@@
 Finally, "printing" the results of an `intRo` analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
+These concepts are important because they encourage best practices with regards to disclosure of analysis methods in research [@baggerly2011reproducible; @xie2015].
 

### 2016-08-08 arxiv-2016/paper.tex
Arxiv submission

\documentclass[12pt,]{article}
%\usepackage{nips15submit_e,times}
\usepackage{url}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Introductory statistics with intRo},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
%\nipsfinalcopy

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.

\title{\bf Introductory statistics with \texttt{intRo}}
\author{Eric Hare \\ Iowa State University \\ \texttt{}  and \\ Andee Kaplan \\ Iowa State University \\ \texttt{} }
\date{}

\usepackage{graphicx}
\usepackage{multirow}
\usepackage{color}
\usepackage{float}

\pdfminorversion=4

\newcommand{\ak}[1]{{\color{magenta} #1}}
\bigskip

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\maketitle

\begin{abstract}
\texttt{intRo} is a web-based application for performing basic data
analysis and statistical routines. Leveraging the power of R and Shiny,
\texttt{intRo} implements common statistical functions in an extensible
modular structure, while including a point-and-click interface for the
novice statistician. This simplicity lends itself to a natural
presentation in an introductory statistics course as a substitute for
other commonly used statistical software packages. \texttt{intRo} is
currently deployed at the URL \url{http://www.intro-stats.com}. Within
this paper, we introduce the application and explore the design
decisions underlying \texttt{intRo}, as well as highlight some
challenges and advantages of reactive programming.
\end{abstract}

\noindent%
{\it Keywords:}  Interactivity, Programming Paradigms, Reactive Programming,
Reproducibility, Statistical Software
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\section{Introduction}\label{introduction}

The widespread adoption of R (R Core Team 2014) as a tool for
statistical analysis has undoubtedly been an important development for
the scientific community. However, using R in most cases still requires
a basic knowledge of programming concepts which may pose a steep
learning curve for the introductory statistics student (Tan, Ting, and
Ling 2009). This additional time commitment may explain why introductory
courses often utilize point-and-click applications, even if the
instructor himself/herself uses R in their own work. Still, some
compromises must be made when using many graphical applications,
including dealing with software licenses and unsupported desktop
platforms.

In teaching Introduction to Business Statistics at Iowa State
University, we witnessed profound struggles by students attempting to
practice introductory concepts discussed in class using current
software. Scrimshaw (2001) notes in his manuscript that ``open-ended
packages, like any others, may create obstacles to learning simply
through their lack of user-friendliness in the sheer mechanics of
operating them, rather than any intrinsic difficulty in the
content\ldots{}.'' In our own experience teaching, students' struggles
were often directly related to the use of the software and not any sort
of fundamental misunderstanding of the material, in agreement with
Scrimshaw's finding.

Multiple software packages have recently been written in an attempt to
spur interest in R programming and statistics. DataCamp's (DataCamp
2014) courses are user-friendly ways to learning basic R programming and
data analysis techniques. Swirl (Carchedi et al. 2014) is a similar
interactive tool to make learning R more fun by learning it within R
itself. Project MOSAIC (Pruim, Kaplan, and Horton 2014) has created a
suite of tools to simplify the teaching of statistics in the form of an
R package. The primary goal of DataCamp and Swirl is to teach R
programming, rather than facilitate the learning of introductory
statistics. Project MOSAIC's goal is to faciliate this learning, but
using the package requires a knowledge of R programming that the
introductory student may not have. R Commander (Fox 2005) and Deducer
(Fellows 2012) provide a graphical front-end to many statistical
functions in R, but are not web-based and thus require local
installation and configuration. iNZight Lite (Wild 2015) also attempts
to expose students to data analysis without requiring programming
knowledge, but does not include reproducible R code, and therefore has
less of a focus on spurring interest in coding for students.

Upon the release of RStudio's Shiny (RStudio and Inc. 2014) it became
easier for an R-based analysis to be converted to an interactive web
application. This in turn led us to create an introductory statistics
application which we call \texttt{intRo}, available at
\url{http://www.intro-stats.com}. \texttt{intRo} offers a number of key
advantages over traditional statistics software, including ease of
access and an aim to foster student interest in coding. Attempting to
entirely hide the programming aspect from students, even in introductory
classes, is a lost opportunity to get students interested in statistical
computing. It is also a lost opportunity reaching students who learn
differently or have a computational background. Another advantage is its
modular structure, which allows course instructors to tailor the
application towards the needs of a particular class, rather than accept
a piece of software as is. Additionally, \texttt{intRo} stands apart
from new tools in that it is a supplement to an existing class, fully
usable by a beginning statistics student. These advantages will be
discussed at length in this text. The paper is structured as follows:
section \ref{what-is-intro} introduces the application, its features and
its usability, and provides motivations for the design decisions
underlying it. Section \ref{understanding-intro} provides technical
details on the design decisions of \texttt{intRo}, in particular, its
modular and extensible structure. Section
\ref{deploying-intro-instances} describes the process of deploying a
classroom instance of \texttt{intRo} tailored to the individual
instructor. Finally, section \ref{conclusions-and-future-work} discusses
some future possibilities and limitations of the software.

\section{\texorpdfstring{What is
\texttt{intRo}?}{What is intRo?}}\label{what-is-intro}

\texttt{intRo} is a web-based tool to accompany an introductory
statistics class. It is meant to assist in the learning of statistics
rather than as a stand-alone delivery of statistics education, with the
intention of being used in conjunction with a guided class. An
accompanying R package, titled \texttt{intRo} and available on GitHub,
assists in the downloading, running, and deploying of \texttt{intRo}
instances.

Three fundamental philosophies that guided the creation of
\texttt{intRo}. In particular, \texttt{intRo} is \emph{easy} to use and
can be an \emph{exciting} part of learning statistics. Additionally,
\texttt{intRo} is an \emph{extensible} tool, allowing for a course
instructor using \texttt{intRo} to tailor the tool for his or her own
classroom needs.

In the development of \texttt{intRo}, we focused on aspects of the user
interface (UI) and output that make it easy to pick up without extensive
training. We used large, easy to click icons in the page header to help
students find what they need more easily. We also made the functionality
available the minimal necessary for an introductory statistics course.
Figure \ref{fig:user_experience} illustrates the simple steps a student
takes to generate a result in \texttt{intRo}. In this instance, a
student clicks on the Graphical tab to create a mosaic plot. The student
sees the plot, and elects to click the save button to store the plot and
its corresponding code to the final compendium.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{user_experience.pdf}
\caption{A typical student experience of generating a result in \texttt{intRo}. In this instance, a student clicks on the Graphical tab to create a mosaic plot. The student sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.}
\label{fig:user_experience}
\end{figure}

Beyond being simple, \texttt{intRo} is also consistent. The tool is
organized around specific tasks a student may perform in the process of
a data analysis, called modules. To the student, a module is simply a
page of statistics functionality that maintains a consistent layout,
helping the student to become familiar with the location of the options,
the results, and the code. Figure \ref{fig:ui} highlights the five
elements that comprise the \texttt{intRo} interface.

\begin{figure*}[ht!]
\centering
\includegraphics[width=\linewidth]{ui_annotate.pdf}
\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
\label{fig:ui}
\end{figure*}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Top Navigation} - The top navigation bar includes two sets of
  clickable icons. The left-aligned buttons are informational buttons.
  The first is a link to \texttt{intRo}. The second is a link to the
  documentation page. The third is a link to the GitHub repository where
  the code for \texttt{intRo} is housed. The final button is a link to
  our websites, which contain contact information if there are any
  questions or comments. The right-aligned buttons are \texttt{intRo}
  utilities. The first is a link to toggle the visibility of the code
  panel (5). The middle icon downloads an rmarkdown (Allaire et al.
  2014) document of the analysis performed. The last is a link to print
  the stored module results, and the associated code (if visible).
\item
  \textbf{Side Navigation} - The side navigation panel includes a list
  of data analysis tasks.
\item
  \textbf{Options Panel} - The options panel includes task-specific
  options which the student can use to customize their results.
\item
  \textbf{Results Panel} - The results pane displays the result of the
  selected module and options.
\item
  \textbf{Code Panel} - The code panel displays the R code used to
  generate the results from the student's \texttt{intRo} session. The
  code panel is shown by default to facilitate a transition to coding,
  but can be hidden by clicking the code toggle button in the Top
  Navigation bar.
\end{enumerate}

The modules included in \texttt{intRo} are split into three higher level
categories - data, summaries, and inference. Under each of these
categories, there are seven default modules, which perform specific data
analysis tasks that employ an easy to use point-and-click interface.
More modules can easily be added by an instructor, as detailed in
section \ref{modules}. The default modules support uploading and
downloading a dataset, transforming variables, graphical and numerical
summaries, simple linear regression, contingency tables, and T-tests.

We've also created a documentation website that is consistent with the
interface of the application. The documentation is available at
\url{http://gammarama.github.io/intRo} and hyperlinked within the
application itself. The documentation describes the specific
functionality provided by the default \texttt{intRo} modules.

Statistics courses with a lab component can most benefit from the use of
\texttt{intRo}. As students learn the aforementioned concepts, they can
practice using the tools to perform the statistical tests and summaries.
They simply upload a dataset and click the module appropriate to the
section of the class. Once they choose the options desired by the
instructor (variables, confidence levels, etc.), they can view the
results and ultimately print the output for turning in. Early labs can
walk the student through the limited number of button presses and clicks
necessary, while later labs can more vaguely describe the intended
output in hopes that the student can easily navigate the \texttt{intRo}
interface to obtain the results.

\texttt{intRo} has an ulterior motive as well: to get students excited
about programming. By navigating about the user interface of
\texttt{intRo}, students are actually creating a fully-executable R
script that they can download and run locally as well as viewing the
script change real-time within the application. This code creation
element of \texttt{intRo} is meant to generate excitement about
programming in R and empower students to feel that they can generate
code as well. \texttt{intRo} uses rmarkdown's render function in order
to print the results, by dymanically executing the student's R script.
By default, the output will include the R code, but if the student
elects to hide the source code by clicking the code toggle button at the
top, the code will not appear in the printed results.

This reproducibility framework has three key advantages. First, this
eases a student who may be intimidated by programming into the idea that
interacting with a user interface is really just a frontend for code.
Seeing the correspondence between graphical clicks and printed code will
hopefully also lessen the fear of coding that many students may have.
Second, an analysis created by \texttt{intRo} can be reproduced in an R
session to easily assess and extend the results. Finally, ``printing''
the results of an \texttt{intRo} analysis amounts to nothing more than
executing the R code on the server, adding another layer of
reproducibility. These concepts are important because they encourage
best practices with regards to disclosure of analysis methods in
research (Baggerly and Berry 2011; Xie 2015).

On the front end, user interaction with \texttt{intRo} is split into
bitesize chunks that we call modules. Modules are self-contained pieces
of functionality which implement common statistical procedures that are
often used in introductory statistics classes. These modules form the
core of \texttt{intRo} and are discussed at length in the next section.

\section{\texorpdfstring{Understanding
\texttt{intRo}}{Understanding intRo}}\label{understanding-intro}

To fully understand \texttt{intRo}, it is important to expand upon two
underlying concepts. The first is the idea of modularity in the context
of a Shiny application. The second is the idea of reactive programming
in the context of a web application.

\subsection{Modules}\label{modules}

An \texttt{intRo} module is a set of self-contained executible R scripts
that together produce a set of introductory statistics functionality.
\texttt{intRo} modules were designed in this way to allow for simple
dynamic creation of the user interface at run-time, as well as ease the
process of converting existing analysis code to the \texttt{intRo}
framework. A high-level diagram of this process is given in figure
\ref{fig:app_creation_modules}. \texttt{intRo} modules are split up into
multiple R scripts which are included either in Shiny's user interface
or server definitions. At runtime, the \texttt{intRo} sources in the
specified modules (contained in the modules folder) to dynamically
generate the functionality available in the application. This allows for
the specific functionality needed to be determined and adjusted by the
individual course instructor. In this example, the instructor is
electing to include a nonparametric module, which is not enabled by
default, to allow the students to perform a wilcoxon rank sum test.

\begin{figure*}[ht!]
\centering
\includegraphics[width=\linewidth]{app_creation_modules.pdf}
\caption{This figure depicts how the Shiny \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}. \texttt{intRo} modules are split up into multiple R scripts which are included either in Shiny's user interface or server definitions. At runtime, the \texttt{intRo} sources in the specified modules (contained in the modules folder) to dynamically generate the functionality available in the application. This allows for the specific functionality needed to be determined and adjusted by the individual course instructor. In this example, the instructor is electing to include a nonparametric module, which is not enabled by default, to allow the students to perform a wilcoxon rank sum test.}
\label{fig:app_creation_modules}
\end{figure*}

Section \ref{dynamic-ui-generation} provides some technical details on
how we implemented this. For the rest of this section, we focus on the
structure and development of the modules themselves, to aid in the
process of creating and deploying new modules.

Modularity was a design decision we focused on from the start of
\texttt{intRo\textquotesingle{}s} development. There are some practical
benefits to thinking of related statistics and data science
functionality in terms of modules. Course instructors can easily enable
or disable functionality depending on the exact needs of the particular
class being taught. Because this is done at run-time, including new
modules is as simple as downloading and placing it within
\texttt{intRo\textquotesingle{}s} modules folder, or removing existing
modules from that folder. Furthermore, errors can be more easily
isolated to specific components. For instance, if an error is
encountered, simply disabling the module can provide a temporary
workaround while the issue is identified. Finally, modularity helps to
organize the different pieces of code into functionality chunks that
make it easier for developers to maintain.

\texttt{intRo} modules are not to be confused with Shiny modules (Cheng
2015). Shiny modules are a recent feature added to Shiny which allows
the bundling of inputs and outputs into a single set of functionality.
They are more general and suitable for any application. \texttt{intRo}
modules are specifically for statistics functionality and work within
the \texttt{intRo} application only.

An \texttt{intRo} module consists of the following scripts:

\begin{itemize}
\tightlist
\item
  \emph{helper.R} - R code that performs some statistical analysis or
  transformation. This would typically be in the form of a function, and
  similar to any standard R script.
\item
  \emph{libraries.R} - Code to load any libraries which are not part of
  core R.
\item
  \emph{observe.R} - Shiny observer code typically used to update
  choices of an input box.
\item
  \emph{output.R} - Shiny output code defining the results of the
  analysis that should be displayed to the student.
\item
  \emph{reactive.R} - Shiny reactives, typically containing data that
  depend on inputs.
\item
  \emph{ui.R} - Shiny user interface definition, including the placement
  of the inputs and outputs.
\end{itemize}

The modules provided with \texttt{intRo} are contained in the modules
folder. The top level directory in the modules folder defines the
category of the module (currently \texttt{data}, \texttt{summaries}, or
\texttt{inference}). Within each of these categories is a folder named
according to the name of the module. This folder houses the previously
defined scripts. As an example, we will walk through the contents of the
\texttt{nonparametric} module as previously mentioned in this section.

Since the \texttt{nonparametric} module performs a statistical test, it
is part of the inference category, and hence can be found in the
\texttt{intRo} repository at \textbf{modules/inference/nonparametric}.
Let's first examine \emph{helper.R}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nonparametrictest <-}\StringTok{ }\NormalTok{function(intro.data, x, y, }
                                \NormalTok{conflevel, althyp, hypval) \{}
    \KeywordTok{interpolate}\NormalTok{(~(}\KeywordTok{wilcox.test}\NormalTok{(}\DataTypeTok{x =} \NormalTok{df$x, }\DataTypeTok{y =} \NormalTok{df$y, }
                              \DataTypeTok{conf.level =} \NormalTok{conf, }
                              \DataTypeTok{alternative =} \NormalTok{althyp, }
                              \DataTypeTok{mu =} \NormalTok{hypval)),}
                  \DataTypeTok{df =} \KeywordTok{quote}\NormalTok{(intro.data),}
                  \DataTypeTok{x =} \NormalTok{x,}
                  \DataTypeTok{y =} \NormalTok{y,}
                  \DataTypeTok{conf =} \NormalTok{conflevel,}
                  \DataTypeTok{althyp =} \NormalTok{althyp,}
                  \DataTypeTok{hypval =} \NormalTok{hypval,}
                  \DataTypeTok{mydir =} \NormalTok{userdir, }
                  \StringTok{`}\DataTypeTok{_env}\StringTok{`} \NormalTok{=}\StringTok{ }\KeywordTok{environment}\NormalTok{(), }
                  \DataTypeTok{file =} \StringTok{"code_nonparametric.R"}\NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

This script is most immediately similar to standard R code. In this
case, a function \texttt{nonparametrictest} is created which, depending
on the values of the parameters, ultimately returns the result of a
wilcoxon rank sum test. One important difference from a typical R script
is that each call in the script is wrapped in a function called
\texttt{interpolate} (Wickham 2015). \texttt{interpolate} both executes
the given R code on the server, and also writes the code executed to the
script window at the bottom of \texttt{intRo}.

Because all the code needed to implement a wilcoxon rank sum test is
found in the \texttt{base} and \texttt{stats} package, the
\texttt{libraries.R} file is empty for the \texttt{nonparametric}
module. Additionally, there are no reactive objects defined, so
\texttt{reactive.R} is also empty. \texttt{observe.R}, which defines the
Shiny observers needed, is reproduced below:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{observe}\NormalTok{(\{}
    \KeywordTok{updateSelectizeInput}\NormalTok{(session, }\StringTok{"group1_non"}\NormalTok{, }
                         \DataTypeTok{choices =} \KeywordTok{intro.numericnames}\NormalTok{(), }
                         \DataTypeTok{selected =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{checkVariable}\NormalTok{(}
                             \KeywordTok{intro.data}\NormalTok{(), input$group1_non), }
                             \NormalTok{input$group1_non, }
                             \KeywordTok{intro.numericnames}\NormalTok{()[}\DecValTok{1}\NormalTok{]))}
    \KeywordTok{updateSelectizeInput}\NormalTok{(session, }\StringTok{"group2_non"}\NormalTok{, }
                         \DataTypeTok{choices =} \KeywordTok{intro.numericnames}\NormalTok{(), }
                         \DataTypeTok{selected =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{checkVariable}\NormalTok{(}
                             \KeywordTok{intro.data}\NormalTok{(), input$group2_non), }
                             \NormalTok{input$group2_non, }
                             \KeywordTok{intro.numericnames}\NormalTok{()[}\DecValTok{2}\NormalTok{]))}
\NormalTok{\})}

\KeywordTok{observeEvent}\NormalTok{(input$store_nonparametric, \{}
    \KeywordTok{cat}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n\textbackslash{}n}\StringTok{"}\NormalTok{, }\KeywordTok{paste}\NormalTok{(}\KeywordTok{readLines}\NormalTok{(}
        \KeywordTok{file.path}\NormalTok{(userdir, }\StringTok{"code_nonparametric.R"}\NormalTok{)), }
        \DataTypeTok{collapse =} \StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{)), }
        \DataTypeTok{file =} \KeywordTok{file.path}\NormalTok{(userdir, }\StringTok{"code_All.R"}\NormalTok{), }
        \DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{\})}
\end{Highlighting}
\end{Shaded}

Shiny observers are a class of reactive objects within the Shiny
paradigm which do not return a value (RStudio and Inc. 2014). For
further discussion of reactivity, see section
\ref{reactive-design-choices}. In this example, observers are created to
ensure that the choices of variable for the \texttt{nonparametric}
module are only numeric variables. This is accomplished by utilizing the
global reactive \texttt{intro.numericnames()}, which returns a character
vector containing the variables in the current dataset that are numeric.
Finally, there is an event observer to store code generated from the
module into the overall code script upon clicking the store button. The
presence of this observer code and the definition of the button in the
user interface are enforced, and must be present in any \texttt{intRo}
module.

The \texttt{output.R} code is very simple:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{output$nonparametrictest <-}\StringTok{ }\KeywordTok{renderPrint}\NormalTok{(\{}
    \KeywordTok{return}\NormalTok{(}\KeywordTok{nonparametrictable}\NormalTok{(}\KeywordTok{intro.data}\NormalTok{(), input$group1_non, }
                              \NormalTok{input$group2_non, input$conflevel_non, }
                              \NormalTok{input$althyp_non, input$hypval_non))}
\NormalTok{\})}
\end{Highlighting}
\end{Shaded}

The \texttt{output.R} script then simply uses Shiny's
\texttt{renderPrint} function to display the resulting table.

Finally, the \texttt{ui.R} code is shown below:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nonparametric_ui <-}\StringTok{ }\KeywordTok{tabPanel}\NormalTok{(}\StringTok{"Nonparametric"}\NormalTok{,}
    \KeywordTok{column}\NormalTok{(}\DecValTok{4}\NormalTok{,}
         \KeywordTok{wellPanel}\NormalTok{(}
             \KeywordTok{selectizeInput}\NormalTok{(}\StringTok{"group1_non"}\NormalTok{, }\DataTypeTok{label =} \StringTok{"Group 1 (x)"}\NormalTok{, }
                            \DataTypeTok{choices =} \KeywordTok{numericNames}\NormalTok{(mpg), }
                            \DataTypeTok{selected =} \KeywordTok{numericNames}\NormalTok{(mpg)[}\DecValTok{1}\NormalTok{]),}
             \KeywordTok{selectizeInput}\NormalTok{(}\StringTok{"group2_non"}\NormalTok{, }\StringTok{"Group 2 (y)"}\NormalTok{, }
                            \DataTypeTok{choices =} \KeywordTok{numericNames}\NormalTok{(mpg), }
                            \DataTypeTok{selected =} \KeywordTok{numericNames}\NormalTok{(mpg)[}\DecValTok{2}\NormalTok{]),}

             \KeywordTok{hr}\NormalTok{(),}

             \KeywordTok{selectizeInput}\NormalTok{(}\StringTok{"althyp_non"}\NormalTok{, }\StringTok{"Alternative Hypothesis"}\NormalTok{, }
                            \KeywordTok{c}\NormalTok{(}\StringTok{"Two-Sided"} \NormalTok{=}\StringTok{ "two.sided"}\NormalTok{, }
                              \StringTok{"Greater"} \NormalTok{=}\StringTok{ "greater"}\NormalTok{, }\StringTok{"Less"} \NormalTok{=}\StringTok{ "less"}\NormalTok{)),}
             \KeywordTok{numericInput}\NormalTok{(}\StringTok{"hypval_non"}\NormalTok{, }\StringTok{"Hypothesized Value"}\NormalTok{,}
                          \DataTypeTok{value =} \DecValTok{0}\NormalTok{),}
             \KeywordTok{sliderInput}\NormalTok{(}\StringTok{"conflevel_non"}\NormalTok{, }\StringTok{"Confidence Level"}\NormalTok{,}
                         \DataTypeTok{min=}\FloatTok{0.01}\NormalTok{, }\DataTypeTok{max=}\FloatTok{0.99}\NormalTok{, }\DataTypeTok{step=}\FloatTok{0.01}\NormalTok{, }\DataTypeTok{value=}\FloatTok{0.95}\NormalTok{),}

             \KeywordTok{hr}\NormalTok{(),}

             \NormalTok{tags$}\KeywordTok{button}\NormalTok{(}\StringTok{""}\NormalTok{, }\DataTypeTok{id =} \StringTok{"store_nonparametric"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"button"}\NormalTok{, }
                         \DataTypeTok{class =} \StringTok{"btn action-button"}\NormalTok{, }\KeywordTok{list}\NormalTok{(}\KeywordTok{icon}\NormalTok{(}\StringTok{"save"}\NormalTok{), }
                         \StringTok{"Store Nonparametric Result"}\NormalTok{), }
                         \DataTypeTok{onclick =} \StringTok{"$('#top-nav a:has(> .fa-print, }
\StringTok{                         .fa-code, .fa-download)').highlight();"}\NormalTok{)}
         \NormalTok{)}
    \NormalTok{),}

    \KeywordTok{column}\NormalTok{(}\DecValTok{8}\NormalTok{,}
         \NormalTok{tags$}\KeywordTok{b}\NormalTok{(}\StringTok{"Nonparametric Results"}\NormalTok{),}
         \KeywordTok{verbatimTextOutput}\NormalTok{(}\StringTok{"nonparametrictest"}\NormalTok{)}
    \NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

This script defines all the inputs and outputs that the student will
see. The only requirements from \texttt{intRo\textquotesingle{}s}
perspective are (1) that there exist a store button at the bottom of the
middle panel for storing the results of the analysis in the code script,
and (2) that configuration options appear in the width 4 column in the
middle, and output appears in the width 8 column on the right. The
remaining input and output definitions depend on the statistical
analysis or transformation being performed.

Although the structure of an \texttt{intRo} module is relatively
straightforward, producing the code needed in a more seamless fashion
would certainly help open up the creation of such modules to a wider
audience. As we discuss in the conclusions and future work section,
providing an \texttt{intRo} module creation tool to abstract away some
of the less common coding paradigms, like the use of
\texttt{interpolate}, is an important effort that will continue to be
pursued.

\subsection{Reactive design choices}\label{reactive-design-choices}

Reactive programming is a programming paradigm that ``tackles issues
posed by event-driven applications by providing abstractions to express
programs as reactions to external events and having the language
automatically manage the flow of time (by conceptually supporting
simultaneity), and data and computation dependencies'' (Bainomugisha et
al. 2012). As implemented by Shiny, results automatically update when
users interact with the interface.

intRo leverages the reactive programming nature of Shiny, and as such is
designed around the idea of user input cascading through the entire
application. In a typical Shiny application, users interact with inputs
that act as parameters to function, which in turn yield different
results. Within \texttt{intRo}, the students are able to interact with
and manipulate the data underlying the entire application. This posed
many challenges in the creation of \texttt{intRo} and drove design
decisions, namely timely save points according to the student's
workflow, and reactive updating of variable lists tied to inputs across
the entire application. Because the student may experiment with
different configurations or select different variables, we did not want
to store all actions taken in the intRo session. Rather, each module
includes a button allowing the student to explicitly store the output
visible in the results panel into the R script. This way, output is only
stored when the student is satisfied, and the resulting output is not
cluttered with unnecessary information.

In the creation of \texttt{intRo} we walked a fine line between giving
the student flexibility and having realistic usability. At the same
time, \texttt{intRo} was created as a consumer of another package,
Shiny, in which we as developers were the beneficiaries of another team
of developers' decision to balance flexibility and usability. For a
tangible example, consider the graphical summaries module. We only allow
variables of a type consistent with the selected plot to be displayed.
This is a conscious decision that limits an \texttt{intRo} user's
flexibility, while maximizing the usability (by minimizing crashes) of
the application. On the flip side of this, Shiny allows much higher
flexibility. For instance, the entire application (including user
interface) is created dynamically upon load, based on the modules
currently housed within \texttt{intRo}. However, Shiny does have limits
on its flexibility based on the designers decisions for usability. One
current example is the slider element. This element allows for fixed
width steps from its minimum to its maximum. The JavaScript library
being utilized in Shiny allows for arbitrary function calls to to
generate these steps, however they must be written in plain JavaScript.
This is an example of a decision made by the developers of Shiny to
limit functionality in favor of usability of their package.

\section{\texorpdfstring{Deploying \texttt{intRo}
Instances}{Deploying intRo Instances}}\label{deploying-intro-instances}

Although students can access intRo from
\url{http://www.intro-stats.com}, course instructors may wish to
download, customize, and deploy their own instance, perhaps with new
modules or modified theming or functionality. \texttt{intRo} can be
downloaded, ran, and deployed on ShinyApps.io through the use of the R
package \texttt{intRo}. Currently, the package is only available on
GitHub, and can be installed using the devtools package as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{devtools::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"gammarama/intRo"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

After installing the \texttt{intRo} package, the first function one
should call is \texttt{download\_intRo}. \texttt{download\_intRo} takes
as an argument a directory in which to store the application. By
default, it selects the working directory of the R session. This
function clones the application branch of the \texttt{intRo} repository
on GitHub, and hence will pull the latest version of the code whenever
it is ran.

Running \texttt{download\_intRo} will produce an \texttt{intRo} folder
in the specified folder. It can then be ran as any Shiny application,
using Shiny's \texttt{runApp} command. However, we have provided a
wrapper function \texttt{run\_intRo} which adds some additional
customization options to the execution process. \texttt{run\_intRo}
takes as argument the path to the folder containing the \texttt{intRo}
application. It also takes several more optional arguments:

\begin{itemize}
\tightlist
\item
  \texttt{enabled\_modules}: A character vector containing the modules
  to enable
\item
  \texttt{theme}: A string representing a shinythemes theme to use
\item
  \texttt{...}: Additional arguments passed to Shiny's \texttt{runApp}
  function
\end{itemize}

The package provides help documentation which explains in further detail
the format that these arguments would take, but as an example, suppose I
wanted to download \texttt{intRo} to my working directory, execute an
\texttt{intRo} session with only the data sources, data transform, and
numerical summaries modules enabled, and apply the cerulean theme. The
series of calls to do so would be as follows:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{download_intRo}\NormalTok{()}
\KeywordTok{run_intRo}\NormalTok{(}\DataTypeTok{enabled_modules =} \KeywordTok{c}\NormalTok{(}\StringTok{"data/transform"}\NormalTok{, }\StringTok{"summaries/numerical"}\NormalTok{), }
          \DataTypeTok{theme =} \StringTok{"cerulean"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Note that the data sources module is required, and hence must be
included in all intRo sessions and need not be specified in the
enabled\_modules argument.

If the intent is to use a specific instance of \texttt{intRo} where many
students will access it at the same time, such as in an introductory
statistics class, it may be preferable to deploy a custom instance of
\texttt{intRo} to a publicly accessible URL. The package provides a
function \texttt{deploy\_intRo} which is a wrapper for the
\texttt{deployApp} function contained in the shinyapps package. Once the
shinyapps package is installed and configured, \texttt{deploy\_intRo}
will upload \texttt{intRo} as an application on the instructor's
ShinyApps.io account. The function takes the same arguments as
\texttt{run\_intRo}, so it can be deployed with a custom selection of
modules, and a customized theme. It also takes an additional argument
\texttt{google\_analytics}, which allows the specification of a Google
Analytics tracking ID. It also takes \texttt{...} as additional
arguments to be passed into the \texttt{deployApp} routine. For example,
if we wished to deploy the instance of \texttt{intRo} we ran previously,
we would call it like so:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{deploy_intRo}\NormalTok{(}\DataTypeTok{enabled_modules =} \KeywordTok{c}\NormalTok{(}\StringTok{"data/transform"}\NormalTok{, }\StringTok{"summaries/numerical"}\NormalTok{), }
             \DataTypeTok{theme =} \StringTok{"cerulean"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Once the process finished, the app will become available at
\url{http://<user>.shinyapps.io/intRo}, where
\texttt{\textless{}user\textgreater{}} is the username of the
ShinyApps.io account configured.

\section{Conclusions and future work}\label{conclusions-and-future-work}

\texttt{intRo} can be a powerful and effective tool for introductory
statistics education. Its modular structure allows it to be flexible
enough for many different applications and curriculums. Its ease-of-use
allows the student to focus her attention on the statistics task at
hand, rather than struggling with software licenses and confusing
interface navigation. Reproducible code generated from each analysis can
be used to spark an interest in R programming in those who might
otherwise not be exposed to it.

In addition to the current functionality, there are some practical
improvements in the works that will make \texttt{intRo} more useful to
both students and instructors. In particular, we have begun development
on an R package which will allow \texttt{intRo} modules to be created
automatically from user written R code. This package will generate the
necessary file structure to allow the module's incorporation into
\texttt{intRo} as well as translate user code to \texttt{intRo}
compatible code and populate the necessary files. This will vastly
improve \texttt{intRo\textquotesingle{}s} flexibility and allow it to be
used in a wider range of curricula, including more advanced statistics
courses. Additionally, we would like to utilize the expanded interactive
capabilities of ggvis in order to make \texttt{intRo's} plots more
engaging to students. One way to do this would be implementing linked
plots, in which interactions with one plot are reflected in other plots
that illustrate the same data. This would be particularly useful in the
regression module so that students could explore observations with high
influence and leverage.

We hope to use \texttt{intRo} in courses to collect feedback regarding
the ease of use and functionality. This will allow us to assess its
usefulness relative to software used in the past, as well as gauge areas
for improvement. Furthermore, we can determine the effectiveness of code
printing on generating excitement from the students about programming in
R.

Challenges do exist with regards to the wider adoption of
\texttt{intRo}. For instance, we will need to monitor how well the
server hosting \texttt{intRo} handles the load of dozens of students
performing data analyses at once. If performance issues are encountered,
the infrastructure used may need to be expanded to handle current and
future load. An unknown quantity will be how feasible it is to increase
adoption of \texttt{intRo} across Iowa State, as well as to other
universities. One limitation of \texttt{intRo} is that uploading a
dataset beyond about 30,000 rows tends to be slow. Even once the data is
successfully uploaded, the default modules produce results more slowly
than with smaller datasets. This is a limitation that should be further
investigated if and when \texttt{intRo} sees wider adoption.

Regardless, tools that focus on usability and extensibility in
statistics education, such as \texttt{intRo}, are sure to encourage the
next round of innovators to be interested and excited about statistical
computing.

\section{Supplementary material}\label{supplementary-material}

All code and documents related to this manuscript are available at
\url{https://github.com/gammarama/intRo}.

\section{Appendix}\label{appendix}

\subsection{Dynamic UI Generation}\label{dynamic-ui-generation}

\texttt{intRo\textquotesingle{}s} user interface and functionality is
dynamically generated depending on the set of modules enabled. The key
driver to populating \texttt{server.R} and \texttt{ui.R} is the modules
folder, the directory structure of which defines the placement of each
module. The interface is then created with the following statement.

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Source ui}
\NormalTok{mylist <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
\NormalTok{old_heading <-}\StringTok{ ""}
\NormalTok{for (i in }\KeywordTok{seq_along}\NormalTok{(modules)) \{}
    \NormalTok{my.module <-}\StringTok{ }\KeywordTok{strsplit}\NormalTok{(modules[i], }\StringTok{"/"}\NormalTok{)[[}\DecValTok{1}\NormalTok{]]}
    \NormalTok{if (my.module[}\DecValTok{1}\NormalTok{] !=}\StringTok{ }\NormalTok{old_heading) \{}
        \NormalTok{mylist[[}\KeywordTok{length}\NormalTok{(mylist) +}\StringTok{ }\DecValTok{1}\NormalTok{]] <-}\StringTok{ }\NormalTok{Hmisc::}\KeywordTok{capitalize}\NormalTok{(my.module[}\DecValTok{1}\NormalTok{])}
        \NormalTok{old_heading <-}\StringTok{ }\NormalTok{my.module[}\DecValTok{1}\NormalTok{]}
    \NormalTok{\}}
    \NormalTok{mylist[[}\KeywordTok{length}\NormalTok{(mylist) +}\StringTok{ }\DecValTok{1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{get}\NormalTok{(}\KeywordTok{paste}\NormalTok{(my.module[}\DecValTok{2}\NormalTok{], }
        \StringTok{"ui"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{"_"}\NormalTok{))}
\NormalTok{\}}

\NormalTok{## mylist is a list containing the different ui}
\NormalTok{## module code Create the UI}
\KeywordTok{shinyUI}\NormalTok{(}\KeywordTok{navbarPage}\NormalTok{(}\StringTok{"intRo"}\NormalTok{, }\DataTypeTok{id =} \StringTok{"top-nav"}\NormalTok{, }\DataTypeTok{theme =} \StringTok{"bootstrap.min.css"}\NormalTok{, }
    \KeywordTok{tabPanel}\NormalTok{(}\DataTypeTok{title =} \StringTok{""}\NormalTok{, }\DataTypeTok{icon =} \KeywordTok{icon}\NormalTok{(}\StringTok{"home"}\NormalTok{), }\KeywordTok{fluidRow}\NormalTok{(}\KeywordTok{do.call}\NormalTok{(navlistPanel, }
        \KeywordTok{c}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{id =} \StringTok{"side-nav"}\NormalTok{, }\DataTypeTok{widths =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{10}\NormalTok{)), }
            \NormalTok{mylist)))), ...))}
\end{Highlighting}
\end{Shaded}

\normalsize

The key piece of code being the \texttt{do.call} statement loading the
list of ui elements from the module's \texttt{ui.R} file. The server
functions are then dynamically generated using a similar method.

\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{shinyServer}\NormalTok{(function(input, output, session) \{}
    \NormalTok{types <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"helper.R"}\NormalTok{, }\StringTok{"observe.R"}\NormalTok{, }\StringTok{"reactive.R"}\NormalTok{, }
        \StringTok{"output.R"}\NormalTok{)}

    \NormalTok{modules_tosource <-}\StringTok{ }\KeywordTok{file.path}\NormalTok{(}\StringTok{"modules"}\NormalTok{, }\KeywordTok{apply}\NormalTok{(}\KeywordTok{expand.grid}\NormalTok{(modules, }
        \NormalTok{types), }\DecValTok{1}\NormalTok{, paste, }\DataTypeTok{collapse =} \StringTok{"/"}\NormalTok{))}

    \NormalTok{for (mod in modules_tosource) \{}
        \KeywordTok{source}\NormalTok{(mod, }\DataTypeTok{local =} \OtherTok{TRUE}\NormalTok{)}
    \NormalTok{\}}
\NormalTok{\})}
\end{Highlighting}
\end{Shaded}

\normalsize
In this way, we were able to have \texttt{intRo} be fully extensible,
its structure and functionality dependent entirely on the modules
present within the application.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-rmarkdown}{}
Allaire, JJ, Jonathan McPherson, Yihui Xie, Hadley Wickham, Joe Cheng,
and Jeff Allen. 2014. \emph{Rmarkdown: Dynamic Documents for R}.
\url{http://CRAN.R-project.org/package=rmarkdown}.

\hypertarget{ref-baggerly2011reproducible}{}
Baggerly, Keith A, and Donald A Berry. 2011. ``Reproducible Research.''
\emph{AMSTAT News: The Membership Magazine of the American Statistical
Association}, no. 403. American Statistical Association: 16--17.

\hypertarget{ref-bainomugisha2012survey}{}
Bainomugisha, Engineer, Andoni Lombide Carreton, Tom Van Cutsem, Stijn
Mostinckx, and Wolfgang De Meuter. 2012. ``A Survey on Reactive
Programming.'' In \emph{ACM Computing Surveys}. Citeseer.

\hypertarget{ref-swirl}{}
Carchedi, Nick, Bill Bauer, Gina Grdina, and Sean Kross. 2014.
\emph{Swirl: Learn R, in R.}
\url{http://CRAN.R-project.org/package=swirl}.

\hypertarget{ref-shinymodules}{}
Cheng, Joe. 2015. ``Shiny - Modularizing Shiny App Code.''
\url{http://shiny.rstudio.com/articles/modules.html}.

\hypertarget{ref-datacamp}{}
DataCamp. 2014. ``Online R Tutorials and Data Science Courses -
Datacamp.'' \url{https://www.datacamp.com/}.

\hypertarget{ref-fellows2012}{}
Fellows, Ian. 2012. ``Deducer: A Data Analysis Gui for R.''
\emph{Journal of Statistical Software} 49 (8).

\hypertarget{ref-fox2005}{}
Fox, John. 2005. ``The R Commander: A Basic-Statistics Graphical User
Interface to R.'' \emph{Journal of Statistical Software} 14 (9).

\hypertarget{ref-mosaic}{}
Pruim, Randall, Daniel Kaplan, and Nicholas Horton. 2014. \emph{Mosaic:
Project Mosaic (Mosaic-Web.org) Statistics and Mathematics Teaching
Utilities}. \url{http://CRAN.R-project.org/package=mosaic}.

\hypertarget{ref-r-stat}{}
R Core Team. 2014. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{http://www.R-project.org/}.

\hypertarget{ref-shiny}{}
RStudio, and Inc. 2014. \emph{Shiny: Web Application Framework for R}.
\url{http://CRAN.R-project.org/package=shiny}.

\hypertarget{ref-scrimshaw2001computers}{}
Scrimshaw, Peter. 2001. ``Computers and the Teacher's Role.''
\emph{Knowledge, Power and Learning}. London, Paul Chapman Publishing
Ltd.

\hypertarget{ref-5359977}{}
Tan, P. H., C. Y. Ting, and S. W. Ling. 2009. ``Learning Difficulties in
Programming Courses: Undergraduates' Perspective and Perception.'' In
\emph{Computer Technology and Development, 2009. Icctd '09.
International Conference on}, 1:42--46.
doi:\href{https://doi.org/10.1109/ICCTD.2009.188}{10.1109/ICCTD.2009.188}.

\hypertarget{ref-interpolate}{}
Wickham, Hadley. 2015. ``Graphics \& Computing Student Paper Winners @
Jsm 2015.'' \url{https://github.com/hadley/15-student-papers}.

\hypertarget{ref-inzight}{}
Wild, Chris. 2015. ``INZight Lite.''
\url{http://lite.docker.stat.auckland.ac.nz}.

\hypertarget{ref-xie2015}{}
Xie, Yihui. 2015. \emph{Dynamic Documents with R and Knitr}. Vol. 29.
CRC Press.

\end{document}

### 2016-09-07 jcgs-paper-2016/paper.Rmd
Comments from meeting
Notes from meeting with Dr. Hofmann

@@
 ---
-title: "Introductory statistics with `intRo`"
+title: "Designing modular software for reproducible analyses: a case study in introductory statistics"
 author:
@@
 abstract: |
-  `intRo` is a web-based application for performing basic data analysis and statistical routines.
-Leveraging the power of R and Shiny, `intRo` implements common statistical functions in an extensible modular structure, while including a point-and-click interface for the novice statistician.
-This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages.
-`intRo` is currently deployed at the URL \url{http://www.intro-stats.com}.
-Within this paper, we introduce the application and explore the design decisions underlying `intRo`, as well as highlight some challenges and advantages of reactive programming.
+   Modular programming is a development paradigm that emphasizes self-contained, flexible, and independent pieces of functionality.
+However, the rise of web-based software applications has presented new challenges for designing an extensible, modular software system.
+In this paper, we outline a framework for designing such a system, with a focus on reproducibility of the results.
+We present as a case study a Shiny-based web application called `intRo`, that allows the user to perform basic data analyses and statistical routines.
+Finally, we highlight some challenges we encountered when combining modular programming concepts with reactive programming as used by Shiny.
 keywords: Interactivity, Programming Paradigms, Reactive Programming, Reproducibility, Statistical Software
@@
 
-# Introduction
+# Background
 
-The widespread adoption of R [@r-stat] as a tool for statistical analysis has undoubtedly been an important development for the scientific community.
+Modularity is a pervasive concept in computer science, extending from the design of systems
+[@parnas1972criteria], to the design of software [@szyperski1996independently].
+Modularity offers several advantages to both a developer and a user.
+In particular, functionality can be dynamically loaded and unloaded depending on the particular use case.
+Open source modular software opens up the possibility of extensions contributed by a wide array of programmers, which can allow the software to morph into areas that weren't anticipated early in development.
+In the statistics realm, R [@r-stat] is a prime example of the virtues of modular programming.
+As of this writing, The Comprehensive R Archive Network (CRAN) contains over 9000 source packages which can be installed and dynamically loaded in a particular session as needed.
+
+Other statistics software also makes use of a number of these ideas.
+Microsoft Excel and JMP both include support for extensions, called macros and add-ins respectively, which allow programming routines to be written extending the base functionality of these programs.
+Compared with R, however, these programs don't maintain a large central repository of publicly available extensions on the level of CRAN.
+There are also software packages building upon R, and thus gaining the advantages of CRAN natively, such as R Commander [@fox2005] and Deducer [@fellows2012], which each provide a graphical front-end to many statistical functions in R.
+One thing these software packages all have in common is the requirement of local installation and configuration, which means certain operating systems and platforms will not support their use.
+
+Modern web technologies have enabled a new generation of software packages that reside solely on the web, which eliminates the issue of local installation.
+Upon the release of RStudio's Shiny [@shiny] it became easier for an R-based analysis to be converted to an interactive web application.
+Several recent software packages have built up Shiny to provide a web-based system based on R.
+One such package is iNZight Lite [@inzight] which attempts to expose students to data analysis without requiring programming knowledge.
+Like most web-based systems, this does not include reproducible R code, which limits its useful in a scientific or academic setting.
+Another package is called Radiant [@radiant]
+
+XXX Say something about reproducibility here - it has to be a goal to make an analysis reproducibility (baggerly cancer trials, forensic bioinformatics)
+
+In our view, a modern software system should be **modular**, **extensible**, **web-based**, and foster **reproducibility**.
+We have developed a case-study application called intRo which we will use to illustrate our method of developing a system meeting these criteria.
+The paper is structured as follows: Section \ref{case-study-intro} describes the motivation behind intRo as an extensible system.
+section \ref{what-is-intro} introduces the application, its features and its usability, and provides motivations for the design decisions underlying it.
+Section \ref{understanding-intro} provides technical details on how we build intRo, and how it can be used to develop other modular systems.
+Section \ref{deploying-intro-instances} describes the process of deploying a classroom instance of `intRo` tailored to the individual instructor.
+Finally, section \ref{conclusions-and-future-work} discusses some future possibilities and limitations of both intRo, and modular systems in general.
+
+# Case Study: intRo
+
+The widespread adoption of R as a tool for statistical analysis has undoubtedly been an important development for the scientific community.
 However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student [@5359977].
@@
 The primary goal of DataCamp and Swirl is to teach R programming, rather than facilitate the learning of introductory statistics.
-Project MOSAIC's goal is to faciliate this learning, but using the package requires a knowledge of R programming that the introductory student may not have.
-R Commander [@fox2005] and Deducer [@fellows2012] provide a graphical front-end to many statistical functions in R, but are not web-based and thus require local installation and configuration.
-iNZight Lite [@inzight] also attempts to expose students to data analysis without requiring programming knowledge, but does not include reproducible R code, and therefore has less of a focus on spurring interest in coding for students.
 
-Upon the release of RStudio's Shiny [@shiny] it became easier for an R-based analysis to be converted to an interactive web application.
-This in turn led us to create an introductory statistics application which we call `intRo`, available at http://www.intro-stats.com.
+These limitations led us to create an introductory statistics application which we call `intRo`, available at http://www.intro-stats.com.
 `intRo` offers a number of key advantages over traditional statistics software, including ease of access and an aim to foster student interest in coding.
@@
 These advantages will be discussed at length in this text.
-The paper is structured as follows: section \ref{what-is-intro} introduces the application, its features and its usability, and provides motivations for the design decisions underlying it.
-Section \ref{understanding-intro} provides technical details on the design decisions of `intRo`, in particular, its modular and extensible structure.
-Section \ref{deploying-intro-instances} describes the process of deploying a classroom instance of `intRo` tailored to the individual instructor.
-Finally, section \ref{conclusions-and-future-work} discusses some future possibilities and limitations of the software.
 
@@
 The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.
-
-We've also created a documentation website that is consistent with the interface of the application.
-The documentation is available at http://gammarama.github.io/intRo and hyperlinked within the application itself.
-The documentation describes the specific functionality provided by the default `intRo` modules.
-
-Statistics courses with a lab component can most benefit from the use of `intRo`.
-As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries.
-They simply upload a dataset and click the module appropriate to the section of the class.
-Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in.
-Early labs can walk the student through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the student can easily navigate the `intRo` interface to obtain the results.
 

### 2016-09-08 jcgs-paper-2016/paper.Rmd
Change title

@@
 ---
-title: "Designing modular software for reproducible analyses: a case study in introductory statistics"
+title: "Designing modular software: a case study in introductory statistics"
 author:

### 2016-09-09 jcgs-paper-2016/paper.Rmd
Restructure the section headers

@@
 
-Modern web technologies have enabled a new generation of software packages that reside solely on the web, which eliminates the issue of local installation.
+With the advantages of R clear, an approach to building statistical software and statistical learning tools would be to attempt to generate interest in programming, which could help naturally ease the transition into the use of R.
+Multiple software packages have recently been written in an attempt to spur this interest in R programming and statistics.
+DataCamp's [@datacamp] courses are user-friendly ways to learning basic R programming and data analysis techniques.
+Swirl [@swirl] is a similar interactive tool to make learning R more fun by learning it within R itself.
+Project MOSAIC [@mosaic] has created a suite of tools to simplify the teaching of statistics in the form of an R package.
+The primary goal of DataCamp and Swirl is to teach R programming, rather than facilitate the learning of introductory statistics.
+
+Modern web technologies have enabled a new generation of software packages that reside solely on the web, which eliminates the issue of local installation and helps abstract away some of the more challenging programming aspects of working directly with R.
 Upon the release of RStudio's Shiny [@shiny] it became easier for an R-based analysis to be converted to an interactive web application.
@@
 
-XXX Say something about reproducibility here - it has to be a goal to make an analysis reproducibility (baggerly cancer trials, forensic bioinformatics)
+Though challenging in a Graphical User Interface, a reproducibility framework has three key advantages.
+First, it eases a student who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
+Seeing the correspondence between graphical clicks and printed code should help lessen the fear of coding that many students may have.
+Second, an analysis created by a reproducible software system can be brought in an R session to easily assess and extend the results.
+Finally, with the help of knitr and RMarkdown, "printing" the results of a reproducible software system analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
+These concepts are important because they encourage best practices with regards to disclosure of analysis methods in research [@baggerly2011reproducible; @xie2015].
 
-In our view, a modern software system should be **modular**, **extensible**, **web-based**, and foster **reproducibility**.
+Based on this, we believe a modern software system should be **modular**, **extensible**, **web-based**, and foster **reproducibility**.
 We have developed a case-study application called intRo which we will use to illustrate our method of developing a system meeting these criteria.
-The paper is structured as follows: Section \ref{case-study-intro} describes the motivation behind intRo as an extensible system.
-section \ref{what-is-intro} introduces the application, its features and its usability, and provides motivations for the design decisions underlying it.
-Section \ref{understanding-intro} provides technical details on how we build intRo, and how it can be used to develop other modular systems.
-Section \ref{deploying-intro-instances} describes the process of deploying a classroom instance of `intRo` tailored to the individual instructor.
+The paper is structured as follows:.
+Section \ref{case-study-intro} introduces the application, its features and its usability, and provides motivations for why it was built.
+Section \ref{intro-design-decisions} provides technical details on how we built intRo, by walking through the underlying modularity, reproducibility, and reactive framework, as well as how it can be used to develop other software systems with these properties.
 Finally, section \ref{conclusions-and-future-work} discusses some future possibilities and limitations of both intRo, and modular systems in general.
@@
 
-Multiple software packages have recently been written in an attempt to spur interest in  R programming and statistics.
-DataCamp's [@datacamp] courses are user-friendly ways to learning basic R programming and data analysis techniques.
-Swirl [@swirl] is a similar interactive tool to make learning R more fun by learning it within R itself.
-Project MOSAIC [@mosaic] has created a suite of tools to simplify the teaching of statistics in the form of an R package.
-The primary goal of DataCamp and Swirl is to teach R programming, rather than facilitate the learning of introductory statistics.
-
-These limitations led us to create an introductory statistics application which we call `intRo`, available at http://www.intro-stats.com.
+These challenges led us to create an introductory statistics application which we call `intRo`, available at http://www.intro-stats.com.
 `intRo` offers a number of key advantages over traditional statistics software, including ease of access and an aim to foster student interest in coding.
@@
 Additionally, `intRo` stands apart from new tools in that it is a supplement to an existing class, fully usable by a beginning statistics student.
-These advantages will be discussed at length in this text.
-
-# What is `intRo`?
-
-`intRo` is a web-based tool to accompany an introductory statistics class.
-It is meant to assist in the learning of statistics rather than as a stand-alone delivery of statistics education, with the intention of being used in conjunction with a guided class.
-An accompanying R package, titled `intRo` and available on GitHub, assists in the downloading, running, and deploying of `intRo` instances.
+An accompanying R package, titled `intRo` and available on GitHub, assists in the downloading, running, and deploying of `intRo` instances (See section \ref{deploying-intro-instances}).
 
@@
 `intRo` has an ulterior motive as well: to get students excited about programming.
-By navigating about the user interface of `intRo`, students are actually creating a fully-executable R script that they can download and run locally as well as viewing the script change real-time within the application.
+By navigating about the user interface of `intRo`, students are actually creating a fully-executable, reproducible R script that they can download and run locally as well as viewing the script change real-time within the application.
 This code creation element of `intRo` is meant to generate excitement about programming in R and empower students to feel that they can generate code as well.
@@
 By default, the output will include the R code, but if the student elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results.
-
-This reproducibility framework has three key advantages.
-First, this eases a student who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
-Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many students may have.
-Second, an analysis created by `intRo` can be reproduced in an R session to easily assess and extend the results.
-Finally, "printing" the results of an `intRo` analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
-These concepts are important because they encourage best practices with regards to disclosure of analysis methods in research [@baggerly2011reproducible; @xie2015].
 
@@
 
-# Understanding `intRo`
+# intRo Design Decisions
 
-To fully understand `intRo`, it is important to expand upon two underlying concepts.
-The first is the idea of modularity in the context of a Shiny application.
-The second is the idea of reactive programming in the context of a web application.
+In this section, we detail the design choices surrounding `intRo`'s extensibility. We have designed it in such a way that these ideas can be used in other Shiny-based software projects.
 
-## Modules
+## Designing for Modularity
 
@@
 
-## Reactive design choices
+## Designing for Reproducibility
+
+## Designing for Reactivity
 
@@
 This is an example of a decision made by the developers of Shiny to limit functionality in favor of usability of their package.
-
-# Deploying `intRo` Instances
-
-Although students can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and deploy their own instance, perhaps with new modules or modified theming or functionality.
-`intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.
-Currently, the package is only available on GitHub, and can be installed using the devtools package as follows:
-
-```{r, eval=FALSE}
-devtools::install_github("gammarama/intRo")
-```
-
-After installing the `intRo` package, the first function one should call is `download_intRo`.
-`download_intRo` takes as an argument a directory in which to store the application.
-By default, it selects the working directory of the R session.
-This function clones the application branch of the `intRo` repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
-
-Running `download_intRo` will produce an `intRo` folder in the specified folder.
-It can then be ran as any Shiny application, using Shiny's `runApp` command.
-However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
-`run_intRo` takes as argument the path to the folder containing the `intRo` application.
-It also takes several more optional arguments:
-
-* `enabled_modules`: A character vector containing the modules to enable
-* `theme`: A string representing a shinythemes theme to use
-* `...`: Additional arguments passed to Shiny's `runApp` function
-
-The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download `intRo` to my working directory, execute an `intRo` session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
-The series of calls to do so would be as follows:
-
-```{r, eval=FALSE}
-download_intRo()
-run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
-          theme = "cerulean")
-```
-
-Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
-
-If the intent is to use a specific instance of `intRo` where many students will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
-The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package.
-Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the instructor's ShinyApps.io account.
-The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
-It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
-It also takes `...` as additional arguments to be passed into the `deployApp` routine.
-For example, if we wished to deploy the instance of `intRo` we ran previously, we would call it like so:
-
-```{r, eval=FALSE}
-deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
-             theme = "cerulean")
-```
-
-Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
 
@@
 
+## Deploying `intRo` Instances
+
+Although students can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and deploy their own instance, perhaps with new modules or modified theming or functionality.
+`intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`.
+Currently, the package is only available on GitHub, and can be installed using the devtools package as follows:
+
+```{r, eval=FALSE}
+devtools::install_github("gammarama/intRo")
+```
+
+After installing the `intRo` package, the first function one should call is `download_intRo`.
+`download_intRo` takes as an argument a directory in which to store the application.
+By default, it selects the working directory of the R session.
+This function clones the application branch of the `intRo` repository on GitHub, and hence will pull the latest version of the code whenever it is ran.
+
+Running `download_intRo` will produce an `intRo` folder in the specified folder.
+It can then be ran as any Shiny application, using Shiny's `runApp` command.
+However, we have provided a wrapper function `run_intRo` which adds some additional customization options to the execution process.
+`run_intRo` takes as argument the path to the folder containing the `intRo` application.
+It also takes several more optional arguments:
+
+* `enabled_modules`: A character vector containing the modules to enable
+* `theme`: A string representing a shinythemes theme to use
+* `...`: Additional arguments passed to Shiny's `runApp` function
+
+The package provides help documentation which explains in further detail the format that these arguments would take, but as an example, suppose I wanted to download `intRo` to my working directory, execute an `intRo` session with only the data sources, data transform, and numerical summaries modules enabled, and apply the cerulean theme.
+The series of calls to do so would be as follows:
+
+```{r, eval=FALSE}
+download_intRo()
+run_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+          theme = "cerulean")
+```
+
+Note that the data sources module is required, and hence must be included in all intRo sessions and need not be specified in the enabled_modules argument.
+
+If the intent is to use a specific instance of `intRo` where many students will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
+The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package.
+Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the instructor's ShinyApps.io account.
+The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.
+It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID.
+It also takes `...` as additional arguments to be passed into the `deployApp` routine.
+For example, if we wished to deploy the instance of `intRo` we ran previously, we would call it like so:
+
+```{r, eval=FALSE}
+deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"), 
+             theme = "cerulean")
+```
+
+Once the process finished, the app will become available at \url{http://<user>.shinyapps.io/intRo}, where `<user>` is the username of the ShinyApps.io account configured.
+
 # References

### 2016-09-09 jcgs-paper-2016/paper.Rmd
Small title case tweaks

@@
 ---
-title: "Designing modular software: a case study in introductory statistics"
+title: "Designing Modular Software: A Case Study in Introductory Statistics"
 author:
@@
 Finally, we highlight some challenges we encountered when combining modular programming concepts with reactive programming as used by Shiny.
-keywords: Interactivity, Programming Paradigms, Reactive Programming, Reproducibility, Statistical Software
+keywords: Interactivity, Modularity, Programming Paradigms, Reactive Programming, Reproducibility, Statistical Software
 bibliography: refs.bib
@@
 
-# Case Study: intRo
+# Case Study: `intRo`
 
@@
 
-# intRo Design Decisions
+# `intRo` Design Decisions
 
@@
 
-# Conclusions and future work
+# Conclusions and Future Work
 
@@
 
-# Supplementary material
+# Supplementary Material
 

### 2016-09-29 jcgs-paper-2016/paper.Rmd
Reproducibility section

@@
 
+While web-based tools written using Shiny (including `intRo`) have appealing characteristics such as being multi-platform, requiring no installation, and requiring no software licenses, one limitation immediately presents itself.
+The actions taken in the application are typically not reproducible as in a standard R script.
+We have designed `intRo` to overcome this limitation by capturing the unevaluated expression of all actions taken by the user in the interface.
+This expression is then parsed and printed in a code window at the bottom, while simultaneously being executed by the R process running on the server.
+
+In essence, this procedure transcribes user actions in a GUI to R code.
+When run in a standard R session, the results produced will be identical to the results shown in `intRo`.
+The full series of actions taken by the user are transcribed and can then be exchanged by researchers, students, and developers in a manner similar to normal scripting.
+Even "printing" the results of an `intRo` session amounts to nothing more than executing the given code on the server, and then storing the results in an rmarkdown document, weaving the code with the results to produce a full compendium.
+While not strictly necessary, this lends credibility to the results produced by `intRo` in the sense that they are directly reproduced by the server every time the user clicks the print button.
+
+Reproducibility in `intRo` is accomplished with the previously mentioned `interpolate` function.
+`interpolate` accepts an expression and an arbitrary number of arguments as an argument, substitutes the arguments into the expression, prints the results to the console, and evaluates the parsed expression.
+This allows for all modules to be shoe-horned into the framework by wrapping the resulting R code in calls to `interpolate`.
+One potential drawback of this solution is that it requires module developers to manually wrap their functions in this call, but this could be mitigated by a package that creates modules automatically (See Section \ref{conclusions-and-future-work}).
+
+One potential enhancement to this framework would be the inclusion of state-saving and state-resuming.
+Because an `intRo` session is uniquely represented by the series of commands stored as code, the code itself could represent a checkpoint for resuming a new `intRo` session.
+Currently, beginning a new session will start the application with no memory of previous sessions.
+In real-world usage, state saving could allow a user to continue work later.
+At this time, this can only be done by taking the code and pasting it into a standard R session, although such an enhancement would likely involve minimal changes to `intRo`'s underlying structure.s
+
 ## Designing for Reactivity

### 2016-09-30 jcgs-paper-2016/paper.Rmd
add table

--- 2016-09-09 jcgs-paper-2016/paper.Rmd
@@
 Like most web-based systems, this does not include reproducible R code, which limits its useful in a scientific or academic setting.
-Another package is called Radiant [@radiant]
+Another package is called Radiant [@radiant], which is a web-based application with the aim of furthering business education and financial analysis.
+While the application is modular and extensible, it does require installation and hosting and is inundated with more features than necessary for an introductory student.
+An overview of the comparison between the features of these statistical software packages is presented in table \ref{tab:compare}.
+Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default.
+For example, `R` does have an associated GUI, however this interface is very limited, thus only partially fulfilling the behavior of a graphical user interface.
+
+```{r, echo=FALSE}
+software_compare <- data.frame("Software package" = "intRo",
+                               "GUI-based" = "Yes",
+                               "Installation required" = "No",
+                               "Modular" = "Yes",
+                               "Web-based" = "Yes",
+                               "Extensible" = "Yes",
+                               "Fosters Reproducibility" = "Yes",
+                               "Open Source" = "Yes",
+                               "Functionality" = "Limited",
+                               stringsAsFactors = FALSE)
+
+software_compare <- rbind(software_compare,
+                          c("JMP", "Yes", "Yes", "Partial", "No", "Partial", "No", "No", "Full"),
+                          c("R", "Partial", "Yes", "Yes", "No", "Yes", "Yes", "Yes", "Full"),
+                          c("R Commander", "Yes", "Yes", "No", "No", "Yes", "Yes", "Yes", "Moderate"),
+                          c("Deducer", "Yes", "Yes", "No", "No", "Yes", "Yes", "Yes", "Moderate"),
+                          c("MOSAIC", "No", "Yes", "No", "No", "No", "Yes", "Yes", "Limited"),
+                          c("iNZight Lite", "Yes", "No", "Yes", "Yes", "Partial", "No", "Yes", "Limited"),
+                          c("Radiant", "Yes", "Yes", "Yes", "Yes", "Partial", "Yes", "Yes", "Moderate"))
+
+software_compare <- software_compare[, -8]
+
+knitr::kable(software_compare, 
+             caption="\\label{tab:compare}A comparison of statistical software packages across the metrics of usability, modularity, extensibility, and reproducibility.
+Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default.",
+             col.names = c("Software", "GUI", "Install", "Modular", "Web-based", "Extensible", "Reproducible", "Functionality"))
+```
 

### 2016-09-30 jcgs-paper-2016/paper.Rmd
format table

--- 2016-09-29 jcgs-paper-2016/paper.Rmd
@@
 Like most web-based systems, this does not include reproducible R code, which limits its useful in a scientific or academic setting.
-Another package is called Radiant [@radiant]
+Another package is called Radiant [@radiant], which is a web-based application with the aim of furthering business education and financial analysis.
+While the application is modular and extensible, it does require installation and hosting and is inundated with more features than necessary for an introductory student.
+An overview of the comparison between the features of these statistical software packages is presented in table \ref{tab:compare}.
+Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default.
+For example, `R` does have an associated GUI, however this interface is very limited, thus only partially fulfilling the behavior of a graphical user interface.
+
+```{r, echo=FALSE, results='asis'}
+software_compare <- data.frame("Software" = "intRo",
+                               "GUI" = "Yes",
+                               "Install" = "No",
+                               "Modular" = "Yes",
+                               "Web" = "Yes",
+                               "Extensible" = "Yes",
+                               "Reproducible" = "Yes",
+                               "Open Source" = "Yes",
+                               "Features" = "Limited",
+                               stringsAsFactors = FALSE)
+
+software_compare <- rbind(software_compare,
+                          c("JMP", "Yes", "Yes", "Partial", "No", "Partial", "No", "No", "Full"),
+                          c("R", "Partial", "Yes", "Yes", "No", "Yes", "Yes", "Yes", "Full"),
+                          c("Rcmdr", "Yes", "Yes", "No", "No", "Yes", "Yes", "Yes", "Moderate"),
+                          c("Deducer", "Yes", "Yes", "No", "No", "Yes", "Yes", "Yes", "Moderate"),
+                          c("MOSAIC", "No", "Yes", "No", "No", "No", "Yes", "Yes", "Limited"),
+                          c("iNZight Lite", "Yes", "No", "Yes", "Yes", "Partial", "No", "Yes", "Limited"),
+                          c("Radiant", "Yes", "Yes", "Yes", "Yes", "Partial", "Yes", "Yes", "Moderate"))
+
+software_compare <- software_compare[, -8]
+
+print(xtable::xtable(software_compare, 
+               label="tab:compare",
+               caption="A comparison of statistical software packages across the metrics of usability, modularity, extensibility, and reproducibility.
+Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default."),
+      comment = FALSE,
+      include.rownames = FALSE)
+```
 

### 2016-10-10 jcgs-paper-2016/paper.Rmd
Add to conclusion

@@
 
-`intRo` can be a powerful and effective tool for introductory statistics education.
+In this paper, we have outlined a framework for designing a web-based, modular, extensible system which reproduces user actions into R code.
+We believe that the development strategies we've outlined can and should be applied to other software systems, as each of these characteristics aids in the ease-of-use and functionality of the overall product.
+Although we present them in the context of an introductory statistics application, these ideas are generalizable and we hope that they will gain traction in many other modern software systems.
+
+With regards to `intRo` itself, we believe it can be a powerful and effective tool for introductory statistics education.
 Its modular structure allows it to be flexible enough for many different applications and curriculums.
@@
 
-Regardless, tools that focus on usability and extensibility in statistics education, such as `intRo`, are sure to encourage the next round of innovators to be interested and excited about statistical computing.
+Regardless, tools that focus on usability and extensibility, such as `intRo`, are sure to encourage the next round of innovators to be interested and excited about statistical computing.
 

### 2016-10-10 jcgs-paper-2016/paper.Rmd
Typo fixes

@@
 Upon the release of RStudio's Shiny [@shiny] it became easier for an R-based analysis to be converted to an interactive web application.
-Several recent software packages have built up Shiny to provide a web-based system based on R.
+Several recent software packages have built upon Shiny to provide a web-based system based on R.
 One such package is iNZight Lite [@inzight] which attempts to expose students to data analysis without requiring programming knowledge.
@@
 Under each of these categories, there are seven default modules, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
-More modules can easily be added by an instructor, as detailed in section \ref{modules}.
+More modules can easily be added by an instructor, as detailed in section \ref{designing-for-modularity}.
 The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.
@@
 Shiny observers are a class of reactive objects within the Shiny paradigm which do not return a value [@shiny].
-For further discussion of reactivity, see section \ref{reactive-design-choices}.
+For further discussion of reactivity, see section \ref{designing-for-reactivity}.
 In this example, observers are created to ensure that the choices of variable for the `nonparametric` module are only numeric variables.
@@
 This allows for all modules to be shoe-horned into the framework by wrapping the resulting R code in calls to `interpolate`.
-One potential drawback of this solution is that it requires module developers to manually wrap their functions in this call, but this could be mitigated by a package that creates modules automatically (See Section \ref{conclusions-and-future-work}).
+A possible drawback of this solution is that it requires module developers to manually wrap their functions in this call, but this could be mitigated by a package that creates modules automatically (See Section \ref{conclusions-and-future-work}).
 

### 2016-10-10 jcgs-paper-2016/paper.Rmd
Small fixes

@@
 Still, some compromises must be made when using many graphical applications, including dealing with software licenses and unsupported desktop platforms.
+From the instructor's perspective in particular, managing a large group of software licenses for students with various computing environments and versions could wind up being extremely cumbersome.
 
@@
 We also made the functionality available the minimal necessary for an introductory statistics course.
-Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in `intRo`.
+Figure \ref{fig:user_experience} presents a schematic of the simple steps a student takes to generate a result in `intRo`.
 In this instance, a student clicks on the Graphical tab to create a mosaic plot.
@@
 \includegraphics[width=\linewidth]{images/user_experience.pdf}
-\caption{A typical student experience of generating a result in \texttt{intRo}.
+\caption{A schematic of the typical student experience of generating a result in \texttt{intRo}.
 In this instance, a student clicks on the Graphical tab to create a mosaic plot.
@@
 On the front end, user interaction with `intRo` is split into bitesize chunks that we call modules.
-Modules are self-contained pieces of functionality which implement common statistical procedures that are often used in introductory statistics classes.
-These modules form the core of `intRo` and are discussed at length in the next section.
+In `intRo`'s context, modules are self-contained pieces of functionality which implement common statistical procedures.
+These modules form the core functionality of `intRo` and are discussed at length in the next section.
 
@@
 They are more general and suitable for any application.
-`intRo` modules are specifically for statistics functionality and work within the `intRo` application only.
+`intRo` modules are for statistics functionality and work within the `intRo` application only.
 
@@
 This folder houses the previously defined scripts.
-As an example, we will walk through the contents of the `nonparametric` module as previously mentioned in this section.
+As an example, we will walk through the process of creating a new module called `nonparametric`, as previously mentioned in this section, which will perform a wilcoxon rank sum test.
 
-Since the `nonparametric` module performs a statistical test, it is part of the inference category, and hence can be found in the `intRo` repository at **modules/inference/nonparametric**.
+Since the `nonparametric` module performs a statistical test, it is part of the inference category, and hence would be placed in the `intRo` repository at **modules/inference/nonparametric**.
 Let's first examine *helper.R*:

### 2016-10-10 jcgs-paper-2016/response-1strev.Rmd
add reviewer responses

--- 2016-06-24 jcgs-paper-2016/response.Rmd
@@
 title: "Response to reviewers"
-author: "Eric Hare, Andee Kaplan"
-date: "May 26, 2016"
-output: html_document
+date: "July 19, 2016"
+output: pdf_document
 ---
@@
 
-The “Article Assessment” questions on the report form don’t work for a paper like this.
+(1) The “Article Assessment” questions on the report form don’t work for a paper like this.
 The authors report on some very impressive work.
@@
 
-I think the authors need to decide, “What sort of paper am I writing?” Included in the manuscript are elements of “a paper about teaching introductory statistics using a software system”, elements of “a paper about designing and building a simple extensible system”, and “a paper about how to use and extend a particular system”.
+(2) I think the authors need to decide, “What sort of paper am I writing?” Included in the manuscript are elements of “a paper about teaching introductory statistics using a software system”, elements of “a paper about designing and building a simple extensible system”, and “a paper about how to use and extend a particular system”.
 I think the most compelling story for the Journal of Computational and Graphical Statistics is the designing and building of a simple online system for intro stat that is easy to learn, use and extend – and teaches R as a side effect.
@@
 
-Everywhere – distinguish between user=student (who uses what is served up) and user=teacher (who customises what is served up)
+(3) Everywhere – distinguish between user=student (who uses what is served up) and user=teacher (who customises what is served up)
 
@@
 
-p.
+(4) p.
 2: Another online analysis system implemented in shiny is iNZight Lite at http://lite.docker.stat.auckland.ac.nz/.
@@
 
-p 3 bottom: Remove parentheses around “(and its corresponding code)” because this is actually one of the key features.
+(5) p 3 bottom: Remove parentheses around “(and its corresponding code)” because this is actually one of the key features.
 
@@
 
-R Commander should enter the discussion around here (even though it isn’t online)
+(6) R Commander should enter the discussion around here (even though it isn’t online)
 
@@
 
-P 5: Fig 2 shows a simple design that should be easy for casual users (as intro students really are) to get their heads around.
+(7) P 5: Fig 2 shows a simple design that should be easy for casual users (as intro students really are) to get their heads around.
 Not all that keen on the use of “statistics” as a category heading in Panel 2 because it reinforces the ghettoizing of the word conflicting with a message that says everything you are doing here is statistics.
@@
 
-P 6 top: When you look at the actual R code generated it can be quite complex with its pipe structures etc which I would have thought does rather limit its introductory value.
+(8) P 6 top: When you look at the actual R code generated it can be quite complex with its pipe structures etc which I would have thought does rather limit its introductory value.
 
@@
 
-P 6-7: discussion of particular modules should be in a “for example” vein. There is no need to justify what particular features are and are not included. (That would be in a teaching paper.)
+(9) P 6-7: discussion of particular modules should be in a “for example” vein. There is no need to justify what particular features are and are not included. (That would be in a teaching paper.)
 
@@
 
-P 7 top: Being able to choose only plot types that are “compatible” with the variables used does, however, preclude learning experiences of the form “this thing you want to do is not appropriate for this type of data.”
+(10) P 7 top: Being able to choose only plot types that are “compatible” with the variables used does, however, preclude learning experiences of the form “this thing you want to do is not appropriate for this type of data.”
 
@@
 
-P 8: I think this descends to details too quickly.
+(11) P 8: I think this descends to details too quickly.
 The big features are designing for extensibility, with conceptions about switching modules on and off and adding modules, and the interface as presented being assembled on the fly when the system starts up.
@@
 
-P 9-10 looks like appendix stuff to me.
+(12) P 9-10 looks like appendix stuff to me.
 
@@
 
-Section 3 discusses downloading onto a local machine before customising the hosted online system which should be the focus (again conception and issues involved before any details).
+(13) Section 3 discusses downloading onto a local machine before customising the hosted online system which should be the focus (again conception and issues involved before any details).
 
@@
 
-P 11 near bottom: “as an application on the user’s account” puzzled me -- needs to say shinyapps.io account
+(14) P 11 near bottom: “as an application on the user’s account” puzzled me -- needs to say shinyapps.io account
 
@@
 
-Section 3.2 mainly belongs in up-front motivation. Table 1 is an implementation mechanism.
+(15) Section 3.2 mainly belongs in up-front motivation. Table 1 is an implementation mechanism.
 
@@
 
-P 14 top: “currently unsupported” justifications are more relevant to a teaching paper than a software design paper.
+(16) P 14 top: “currently unsupported” justifications are more relevant to a teaching paper than a software design paper.
 
@@
 
-P 14 Last para Section 3: could go into motivational discussion.
+(17) P 14 Last para Section 3: could go into motivational discussion.
 
@@
 
-Section 4.1 is a big part of what I see as the real story but descends to details too quickly.
+(18) Section 4.1 is a big part of what I see as the real story but descends to details too quickly.
 
@@
 
-Most of Section 4.2 looks to me like more motivation, except the interpolate stuff which is an implementation mechanism
+(19) Most of Section 4.2 looks to me like more motivation, except the interpolate stuff which is an implementation mechanism
 
@@
 
-Section 4.3: more “main story”.
+(20) Section 4.3: more “main story”.
 

### 2016-10-10 response.Rmd
add reviewer responses

---
title: "Response to second review"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

# Reviewer 1

(1) The paper seems to me inappropriate for the JCGS. Because it focuses purely on software, I think that (once the other issues are addressed) it would be more appropriate for the Journal of Statistical Software or the R Journal.

    **We have refocused the paper on more general computational concepts of reproducibility, reactivity, and modularity, in the context of designing modern statistical software systems. This fits nicely within the aim of JCGS to "extend the use of computational and graphical methods in statistics and data analysis".**

(2) Although the software looks promising, unless I've missed something, it currently has very limited functionality -- indeed, not enough to sustain most introductory statistics courses. This isn't a permanent limitation, however, since the authors have created interesting infrastructure for extending the software. In its current state, the software doesn't quite look ready for prime time.

    **We further highlight in the introduction as well as the beginning of section two that this was our end goal, simplifying software for the user when necessary.**

(3) The authors ignore other (and at present, much more capable) GUIs for R -- the R Commander and Deducer. Since these are the most direct competitors to intRo -- more so than, e.g., Excel and JMP -- it seems obvious that the authors should address similarities, differences, advantages, and disadvantages. One obvious difference is that these other GUIs run on individual computers rather than being web-based, and it would also make sense to address more generally the advantages and disadvantages of the two approaches. (For example, in my tests of intRo, I noticed slow response, even with very small data sets. Of course, it's hard to know to what to attribute that -- it could as easily have been a network issue as slowness at the server end.)

    **We have added comparisons to more software, including Radiant, to the introduction. Within these comparisons, we focused on extensibility, modularity, reproducibility, licensing, and the functionality of each piece of software. We have also included a table that summarises these comparisons.**

# Reviewer 2

(1) While the authors have made a pretty good fist of addressing everything I asked for and it is pretty
well good to go from my point of view. But I think they can still strengthen their case and their appeal somewhat. I think a title like the following would suggest better the new focus of the paper:
"Customisable introductory statistics software via intRo"

    **We agree, thanks for the suggestion. We have changed the title to "Designing Modular Software: A Case Study in Introductory Statistics", which we feel highlights the customizable aspects of intRo.**

(3) Also there are 2 big ideas in here about software that go beyond the particular application:
extensibility (the ability to “easily” add functionality) and customisation/simplification (the ability to
easily hide functionality) and a way of implementing these. I think these 2 elements and their
advantages should be highlighted from the get go, starting with the abstract. The advantages of
extensibility are obvious but the desirability of being able to simplify the view of software a
particular audience gets needs highlighting. (Also the idea that they enable this by the way in which
the instance of the software that will be served up to the user is assembled on the fly at run time.)
Detailed comments follow. (Perhaps unfortunately, many draw on what I have said above.):

    **We have updated the abstract to reflect the new focus of the paper, which is writing extensible software through the use of modular programming within the framework of shiny. Additionally, we've highlighted the ability to remove functionality and present a simplified view to the user in the abstract. The paper itself now addressed the "big ideas" of extensibility and customization much more thoroughly.**

(4) Page 2 bottom: could note that having a large group of students install software on their own
machines can be a nightmare with differing hardware configurations, versioning etc.

    **We added a sentence in section 2, page 5 top to address this.**

(5) Page 3, last para pf Section 1 needs breaking up.

    **The content of this paragraph is no longer in the paper as a single paragraph, but rather redispersed into multiple sections.**

(6) Page 4, Note that Fig 1 is a schematic, not a view of the software.

    **We have added the verbiage "schematic" in the text and figure caption to note this.**

(7) Page 7 last line: no need for “introductory” here. These ideas apply generally.

    **Good point. We removed the word "introductory" from the sentence.**

(8) Page 8 ff (also page 10-11): I think this story would be better spun in terms of someone writing an
incorporating a new module (a few tiny wording changes).

    **We changed the wording to indicate the example walks through the process of creating a module, as well as including it.**

# Associate Editor

(1) The authors have responded to some of the reviewers' comments, but there is still more work to do. The motivation for intRo is as a simpler interface to R (for teaching statistics AND generating interest in R), but this should just be background information for this article.  There are many other R GUIs that are MUCH more developed for teaching statistics (that also generate R code as well).  The features of intRo that are for the student/user are less interesting and do not add anything beyond existing software. The main focus of the article should be on the novel aspects of this work, which is the modularity/extensibility - the features of intRo that are for the teacher/developer.

    **We have refocused the article onto the novel aspects of modularity and extensibility by reqriting the introduction and reframing the article as a case study in modularity. Accordingly, we have moved details about intRo as an interface to R to section 2.**

(2) The comparison with existing software (Rcmdr, Deducer, iNZight, Shiny) should be more in-depth, but focused on the modularity/extensibility issue.

    **We have added comparisons to more software, including Radiant, to the introduction. Within these comparisons, we focused on extensibility, modularity, reproducibility, licensing, and the functionality of each piece of software. We have also included a table that summarises these comparisons.**

### 2016-10-10 jcgs-paper-2016/paper.Rmd

@@
    Modular programming is a development paradigm that emphasizes self-contained, flexible, and independent pieces of functionality.
+This practice allows new functionality to be seamlessly added when new functionality is desired, and unwanted functionality to be removed, thus simplify the user-facing view of the software.
 However, the rise of web-based software applications has presented new challenges for designing an extensible, modular software system.
@@
 We present as a case study a Shiny-based web application called `intRo`, that allows the user to perform basic data analyses and statistical routines.
-Finally, we highlight some challenges we encountered when combining modular programming concepts with reactive programming as used by Shiny.
+Finally, we highlight some challenges we encountered, and how to address them, when combining modular programming concepts with reactive programming as used by Shiny.
 keywords: Interactivity, Modularity, Programming Paradigms, Reactive Programming, Reproducibility, Statistical Software

### 2016-10-20 jcgs-paper-2016/cover_letter-1strev.Rmd
New cover letter

---
author: 
 - Eric Hare
 - PhD Student 
 - Department of Statistics 
 - Iowa State University
date: "`r format(Sys.time(), '%B %d, %Y')`"
opening: Dear Dr. Cook,
closing: Best regards,
address: 
 - Dianne Cook
 - Editor
 - Journal of Computational and Graphical Statistics
return-address:
 - Eric Hare
 - Department of Statistics
 - Iowa State University
 - Ames, IA 50011
output:
  pdf_document:
    template: template-letter.tex
geometry: margin=1in
---

Re: Manuscript reference No. JCGS-16-045

Thank you for the reviewers' comments and the opportunity to resubmit our manuscript, "Introductory statistics with intRo". We feel the paper is stronger as a result of incorporating the feedback.

We look forward to hearing your thoughts on our changes, and we appreciate your consideration of our resubmission.

### 2016-10-20 jcgs-paper-2016/cover_letter.Rmd
New cover letter

@@
 date: "`r format(Sys.time(), '%B %d, %Y')`"
-opening: To whom it may concern,
+opening: Dear Dr. Cook,
 closing: Best regards,
 address: 
+ - Dianne Cook
  - Editor
@@
 
-Thank you in advance for reviewing our manuscript "Introductory statistics with `intRo`", which I have submitted for consideration of publication in the Journal of Computational and Graphical Statistics.
+Re: Manuscript reference No. JCGS-16-045
 
-This paper details an interactive web-based statistical software application intended for use in an introductory statistics classroom.
-We introduce the tool as well as an accompanying R package to ease the process of deploying intRo within a curriculum.
-Additionally, the paper describes the process for extending the functionality in intRo in order to personalize the software.
+Attached you will find the newest revision of our paper, written in response to the feedback we received from the reviewers, the associate editor, and yourself.
 
-Thank you for your consideration of our submission.
+We have attempted to restructure the paper to focus on general techniques for designing a modular software system, using intRo itself as a case study.
+In particular, we have re-written the title, abstract, and a substantial portion of the background to highlight this new direction.
+
+In addition to the refocusing, we have also extended our review of other software systems, and emphasized the need that intRo addresses.
+Table 1 provides a somewhat subjective but thorough comparison of many "competing" software systems in terms of a number of qualities, including whether or not they are web-based, the level of extensibility, and the amount of functionality.
+We have also done our best to address the other minor comments that were raised in the second round of reviews.
+
+We hope that this latest revision addresses the majority of the issues with the previous, and we welcome any further critiques or comments.

### 2016-10-20 jcgs-paper-2016/paper.tex
We have submitted our revision!

--- 2016-08-08 arxiv-2016/paper.tex
@@
             pdfauthor={},
-            pdftitle={Introductory statistics with intRo},
+            pdftitle={Designing Modular Software: A Case Study in Introductory Statistics},
             colorlinks=true,
@@
 
-\title{\bf Introductory statistics with \texttt{intRo}}
+\title{\bf Designing Modular Software: A Case Study in Introductory Statistics}
 \author{Eric Hare \\ Iowa State University \\ \texttt{}  and \\ Andee Kaplan \\ Iowa State University \\ \texttt{} }
@@
 \begin{abstract}
-\texttt{intRo} is a web-based application for performing basic data
-analysis and statistical routines. Leveraging the power of R and Shiny,
-\texttt{intRo} implements common statistical functions in an extensible
-modular structure, while including a point-and-click interface for the
-novice statistician. This simplicity lends itself to a natural
-presentation in an introductory statistics course as a substitute for
-other commonly used statistical software packages. \texttt{intRo} is
-currently deployed at the URL \url{http://www.intro-stats.com}. Within
-this paper, we introduce the application and explore the design
-decisions underlying \texttt{intRo}, as well as highlight some
-challenges and advantages of reactive programming.
+Modular programming is a development paradigm that emphasizes
+self-contained, flexible, and independent pieces of functionality. This
+practice allows new features to be seamlessly added when desired, and
+unwanted features to be removed, thus simplifying the user-facing view
+of the software. The recent rise of web-based software applications has
+presented new challenges for designing an extensible, modular software
+system. In this paper, we outline a framework for designing such a
+system, with a focus on reproducibility of the results. We present as a
+case study a Shiny-based web application called \texttt{intRo}, that
+allows the user to perform basic data analyses and statistical routines.
+Finally, we highlight some challenges we encountered, and how to address
+them, when combining modular programming concepts with reactive
+programming as used by Shiny.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:}  Interactivity, Programming Paradigms, Reactive Programming,
+{\it Keywords:}  Interactivity, Modularity, Programming Paradigms, Reactive Programming,
 Reproducibility, Statistical Software
@@
 
-\section{Introduction}\label{introduction}
-
-The widespread adoption of R (R Core Team 2014) as a tool for
-statistical analysis has undoubtedly been an important development for
-the scientific community. However, using R in most cases still requires
-a basic knowledge of programming concepts which may pose a steep
-learning curve for the introductory statistics student (Tan, Ting, and
-Ling 2009). This additional time commitment may explain why introductory
-courses often utilize point-and-click applications, even if the
-instructor himself/herself uses R in their own work. Still, some
-compromises must be made when using many graphical applications,
-including dealing with software licenses and unsupported desktop
-platforms.
+\section{Background}\label{background}
+
+Modularity is a pervasive concept in computer science, extending from
+the design of systems (Parnas 1972), to the design of software
+(Szyperski 1996). Modularity offers several advantages to both a
+developer and a user. In particular, functionality can be dynamically
+loaded and unloaded depending on the particular use case. Open source
+modular software precipitates the possibility of extensions contributed
+by a wide array of programmers, which can allow the software to morph
+into areas that weren't anticipated early in development. In the
+statistics realm, R (R Core Team 2014) is a prime example of the virtues
+of modular programming. As of this writing, The Comprehensive R Archive
+Network (CRAN) contains over 9000 source packages which can be installed
+and dynamically loaded in a particular session as needed.
+
+Other statistics software also makes use of a number of these ideas.
+Microsoft Excel and JMP both include support for extensions, called
+macros and add-ins respectively, which allow programming routines to be
+written extending the base functionality of these programs. Compared
+with R, however, these programs don't maintain a large central
+repository of publicly available extensions on the level of CRAN. There
+are also software packages building upon R, and thus gaining the
+advantages of CRAN natively, such as R Commander (Fox 2005) and Deducer
+(Fellows 2012), which each provide a graphical front-end to many
+statistical functions in R. One thing these software packages all have
+in common is the requirement of local installation and configuration,
+which means certain operating systems and platforms will not support
+their use.
+
+With the advantages of R clear, an approach to building statistical
+software and statistical learning tools would be to attempt to generate
+interest in programming, which could help naturally ease the transition
+into the use of R. Multiple software packages have recently been written
+in an attempt to spur this interest in R programming and statistics.
+DataCamp's (DataCamp 2014) courses are a user-friendly way to learn
+basic R programming and data analysis techniques. Swirl (Carchedi et al.
+2014) is a similar interactive tool to make learning R more fun by
+learning it within R itself. Project MOSAIC (Pruim, Kaplan, and Horton
+2014) has created a suite of tools to simplify the teaching of
+statistics in the form of an R package. The primary goal of DataCamp and
+Swirl is to teach R programming, rather than facilitate the learning of
+introductory statistics.
+
+Modern web technologies have enabled a new generation of software
+packages that reside solely on the web, which eliminates the issue of
+local installation and helps abstract away some of the more challenging
+programming aspects of working directly with R. Upon the release of
+RStudio's Shiny (RStudio and Inc. 2014) it became easier for an R-based
+analysis to be converted to an interactive web application. Several
+recent software packages have built upon Shiny to provide a web-based
+system based on R. One such package is iNZight Lite (Wild 2015) which
+attempts to expose students to data analysis without requiring
+programming knowledge. Like most web-based systems, this does not
+include reproducible R code, which limits its usefulness in a scientific
+or academic setting. Another package is called Radiant (Nijs 2016),
+which is a web-based application with the aim of furthering business
+education and financial analysis. While the application is modular and
+extensible, it does require installation and hosting and is inundated
+with more features than necessary for an introductory student. An
+overview of the comparison between the features of these statistical
+software packages is presented in Table \ref{tab:compare}. Partial
+fulfillment of requirements is noted in the table, as well as a measure
+of the complexity of functionality offered by default. For example,
+\texttt{R} does have an associated Graphical User Interface (GUI),
+however this interface is very limited, thus only partially fulfilling
+the behavior of a GUI.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{llllllll}
+  \hline
+Software & GUI & Install & Modular & Web & Extensible & Reproducible & Features \\ 
+  \hline
+intRo & Yes & No & Yes & Yes & Yes & Yes & Limited \\ 
+  JMP & Yes & Yes & Partial & No & Partial & No & Full \\ 
+  R & Partial & Yes & Yes & No & Yes & Yes & Full \\ 
+  Rcmdr & Yes & Yes & No & No & Yes & Yes & Moderate \\ 
+  Deducer & Yes & Yes & No & No & Yes & Yes & Moderate \\ 
+  MOSAIC & No & Yes & No & No & No & Yes & Limited \\ 
+  iNZight Lite & Yes & No & Yes & Yes & Partial & No & Limited \\ 
+  Radiant & Yes & Yes & Yes & Yes & Partial & Yes & Moderate \\ 
+   \hline
+\end{tabular}
+\caption{A comparison of statistical software packages across the metrics of usability, modularity, extensibility, and reproducibility.
+Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default.}
+\label{tab:compare}
+\end{table}
+
+Though challenging in a GUI, a reproducibility framework has three key
+advantages. First, it eases a student who may be intimidated by
+programming into the idea that interacting with a user interface is
+really just a frontend for code. Seeing the correspondence between
+graphical clicks and printed code should help lessen the fear of coding
+that many students may have. Second, an analysis created by a
+reproducible software system can be brought in an R session to easily
+assess and extend the results. Finally, with the help of knitr (Xie
+2015) and rmarkdown (Allaire et al. 2014), ``printing'' the results of a
+reproducible software system analysis amounts to nothing more than
+executing the R code on the server, adding another layer of
+reproducibility. These concepts are important because they encourage
+best practices with regards to disclosure of analysis methods in
+research (Baggerly and Berry 2011; Xie 2015).
+
+Based on the above, we believe a modern software system should be
+\textbf{modular}, \textbf{extensible}, \textbf{web-based}, and foster
+\textbf{reproducibility}. We have developed a case-study application
+called \texttt{intRo} which we will use to illustrate our method of
+developing a system meeting these criteria. The paper is structured as
+follows: Section \ref{case-study-intro} introduces the application, its
+features and its usability, and provides motivations for why it was
+built. Section \ref{intro-design-decisions} provides technical details
+on how we built \texttt{intRo}, by walking through the underlying
+modularity, reproducibility, and reactive framework, as well as how it
+can be used to develop other software systems with these properties.
+Finally, Section \ref{conclusions-and-future-work} discusses some future
+possibilities and limitations of both \texttt{intRo}, and modular
+systems in general.
+
+\section{\texorpdfstring{Case Study:
+\texttt{intRo}}{Case Study: intRo}}\label{case-study-intro}
+
+The widespread adoption of R as a tool for statistical analysis has
+undoubtedly been an important development for the scientific community.
+However, using R in most cases still requires a basic knowledge of
+programming concepts which may pose a steep learning curve for the
+introductory statistics student (Tan, Ting, and Ling 2009). This
+additional time commitment may explain why introductory courses often
+utilize point-and-click applications, even if the instructor
+himself/herself uses R in their own work. Still, some compromises must
+be made when using many graphical applications, including dealing with
+software licenses and unsupported desktop platforms. From the
+instructor's perspective in particular, managing a large group of
+software licenses for students with various computing environments and
+versions could wind up being extremely cumbersome.
 
@@
 
-Multiple software packages have recently been written in an attempt to
-spur interest in R programming and statistics. DataCamp's (DataCamp
-2014) courses are user-friendly ways to learning basic R programming and
-data analysis techniques. Swirl (Carchedi et al. 2014) is a similar
-interactive tool to make learning R more fun by learning it within R
-itself. Project MOSAIC (Pruim, Kaplan, and Horton 2014) has created a
-suite of tools to simplify the teaching of statistics in the form of an
-R package. The primary goal of DataCamp and Swirl is to teach R
-programming, rather than facilitate the learning of introductory
-statistics. Project MOSAIC's goal is to faciliate this learning, but
-using the package requires a knowledge of R programming that the
-introductory student may not have. R Commander (Fox 2005) and Deducer
-(Fellows 2012) provide a graphical front-end to many statistical
-functions in R, but are not web-based and thus require local
-installation and configuration. iNZight Lite (Wild 2015) also attempts
-to expose students to data analysis without requiring programming
-knowledge, but does not include reproducible R code, and therefore has
-less of a focus on spurring interest in coding for students.
-
-Upon the release of RStudio's Shiny (RStudio and Inc. 2014) it became
-easier for an R-based analysis to be converted to an interactive web
-application. This in turn led us to create an introductory statistics
-application which we call \texttt{intRo}, available at
+These challenges led us to create an introductory statistics application
+which we call \texttt{intRo}, available at
 \url{http://www.intro-stats.com}. \texttt{intRo} offers a number of key
@@
 from new tools in that it is a supplement to an existing class, fully
-usable by a beginning statistics student. These advantages will be
-discussed at length in this text. The paper is structured as follows:
-section \ref{what-is-intro} introduces the application, its features and
-its usability, and provides motivations for the design decisions
-underlying it. Section \ref{understanding-intro} provides technical
-details on the design decisions of \texttt{intRo}, in particular, its
-modular and extensible structure. Section
-\ref{deploying-intro-instances} describes the process of deploying a
-classroom instance of \texttt{intRo} tailored to the individual
-instructor. Finally, section \ref{conclusions-and-future-work} discusses
-some future possibilities and limitations of the software.
-
-\section{\texorpdfstring{What is
-\texttt{intRo}?}{What is intRo?}}\label{what-is-intro}
-
-\texttt{intRo} is a web-based tool to accompany an introductory
-statistics class. It is meant to assist in the learning of statistics
-rather than as a stand-alone delivery of statistics education, with the
-intention of being used in conjunction with a guided class. An
-accompanying R package, titled \texttt{intRo} and available on GitHub,
-assists in the downloading, running, and deploying of \texttt{intRo}
-instances.
+usable by a beginning statistics student. An accompanying R package,
+titled \texttt{intRo} and available on GitHub, assists in the
+downloading, running, and deploying of \texttt{intRo} instances (See
+Section \ref{deploying-intro-instances}).
 
@@
 available the minimal necessary for an introductory statistics course.
-Figure \ref{fig:user_experience} illustrates the simple steps a student
-takes to generate a result in \texttt{intRo}. In this instance, a
-student clicks on the Graphical tab to create a mosaic plot. The student
-sees the plot, and elects to click the save button to store the plot and
-its corresponding code to the final compendium.
+Figure \ref{fig:user_experience} presents a schematic of the simple
+steps a student takes to generate a result in \texttt{intRo}. In this
+instance, a student clicks on the Graphical tab to create a mosaic plot.
+The student sees the plot, and elects to click the save button to store
+the plot and its corresponding code to the final compendium.
 
@@
 \centering
-\includegraphics[width=\linewidth]{user_experience.pdf}
-\caption{A typical student experience of generating a result in \texttt{intRo}.
+\includegraphics[width=\linewidth]{images/user_experience.pdf}
+\caption{A schematic of the typical student experience of generating a result in \texttt{intRo}.
 In this instance, a student clicks on the Graphical tab to create a mosaic plot.
@@
 \centering
-\includegraphics[width=\linewidth]{ui_annotate.pdf}
+\includegraphics[width=\linewidth]{images/ui_annotate.pdf}
 \caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
@@
   utilities. The first is a link to toggle the visibility of the code
-  panel (5). The middle icon downloads an rmarkdown (Allaire et al.
-  2014) document of the analysis performed. The last is a link to print
-  the stored module results, and the associated code (if visible).
+  panel (5). The middle icon downloads an rmarkdown document of the
+  analysis performed. The last is a link to print the stored module
+  results, and the associated code (if visible).
 \item
@@
 More modules can easily be added by an instructor, as detailed in
-section \ref{modules}. The default modules support uploading and
-downloading a dataset, transforming variables, graphical and numerical
-summaries, simple linear regression, contingency tables, and T-tests.
-
-We've also created a documentation website that is consistent with the
-interface of the application. The documentation is available at
-\url{http://gammarama.github.io/intRo} and hyperlinked within the
-application itself. The documentation describes the specific
-functionality provided by the default \texttt{intRo} modules.
-
-Statistics courses with a lab component can most benefit from the use of
-\texttt{intRo}. As students learn the aforementioned concepts, they can
-practice using the tools to perform the statistical tests and summaries.
-They simply upload a dataset and click the module appropriate to the
-section of the class. Once they choose the options desired by the
-instructor (variables, confidence levels, etc.), they can view the
-results and ultimately print the output for turning in. Early labs can
-walk the student through the limited number of button presses and clicks
-necessary, while later labs can more vaguely describe the intended
-output in hopes that the student can easily navigate the \texttt{intRo}
-interface to obtain the results.
+Section \ref{designing-for-modularity}. The default modules support
+uploading and downloading a dataset, transforming variables, graphical
+and numerical summaries, simple linear regression, contingency tables,
+and T-tests.
 
@@
 about programming. By navigating about the user interface of
-\texttt{intRo}, students are actually creating a fully-executable R
-script that they can download and run locally as well as viewing the
-script change real-time within the application. This code creation
-element of \texttt{intRo} is meant to generate excitement about
+\texttt{intRo}, students are actually creating a fully-executable,
+reproducible R script that they can download and run locally as well as
+viewing the script change real-time within the application. This code
+creation element of \texttt{intRo} is meant to generate excitement about
 programming in R and empower students to feel that they can generate
@@
 
-This reproducibility framework has three key advantages. First, this
-eases a student who may be intimidated by programming into the idea that
-interacting with a user interface is really just a frontend for code.
-Seeing the correspondence between graphical clicks and printed code will
-hopefully also lessen the fear of coding that many students may have.
-Second, an analysis created by \texttt{intRo} can be reproduced in an R
-session to easily assess and extend the results. Finally, ``printing''
-the results of an \texttt{intRo} analysis amounts to nothing more than
-executing the R code on the server, adding another layer of
-reproducibility. These concepts are important because they encourage
-best practices with regards to disclosure of analysis methods in
-research (Baggerly and Berry 2011; Xie 2015).
-
 On the front end, user interaction with \texttt{intRo} is split into
-bitesize chunks that we call modules. Modules are self-contained pieces
-of functionality which implement common statistical procedures that are
-often used in introductory statistics classes. These modules form the
-core of \texttt{intRo} and are discussed at length in the next section.
-
-\section{\texorpdfstring{Understanding
-\texttt{intRo}}{Understanding intRo}}\label{understanding-intro}
-
-To fully understand \texttt{intRo}, it is important to expand upon two
-underlying concepts. The first is the idea of modularity in the context
-of a Shiny application. The second is the idea of reactive programming
-in the context of a web application.
-
-\subsection{Modules}\label{modules}
+bitesize chunks that we call modules. In \texttt{intRo}'s context,
+modules are self-contained pieces of functionality which implement
+common statistical procedures. These modules form the core functionality
+of \texttt{intRo} and are discussed at length in the next section.
+
+\section{\texorpdfstring{\texttt{intRo} Design
+Decisions}{intRo Design Decisions}}\label{intro-design-decisions}
+
+In this section, we detail the design choices surrounding
+\texttt{intRo}'s extensibility. We have designed it in such a way that
+these ideas can be used in other Shiny-based software projects.
+
+\subsection{Designing for Modularity}\label{designing-for-modularity}
 
@@
 process of converting existing analysis code to the \texttt{intRo}
-framework. A high-level diagram of this process is given in figure
+framework. A high-level diagram of this process is given in Figure
 \ref{fig:app_creation_modules}. \texttt{intRo} modules are split up into
@@
 \centering
-\includegraphics[width=\linewidth]{app_creation_modules.pdf}
+\includegraphics[width=\linewidth]{images/app_creation_modules.pdf}
 \caption{This figure depicts how the Shiny \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.
@@
 Modularity was a design decision we focused on from the start of
-\texttt{intRo\textquotesingle{}s} development. There are some practical
-benefits to thinking of related statistics and data science
-functionality in terms of modules. Course instructors can easily enable
-or disable functionality depending on the exact needs of the particular
-class being taught. Because this is done at run-time, including new
-modules is as simple as downloading and placing it within
-\texttt{intRo\textquotesingle{}s} modules folder, or removing existing
-modules from that folder. Furthermore, errors can be more easily
-isolated to specific components. For instance, if an error is
-encountered, simply disabling the module can provide a temporary
-workaround while the issue is identified. Finally, modularity helps to
-organize the different pieces of code into functionality chunks that
-make it easier for developers to maintain.
+\texttt{intRo}'s development. There are some practical benefits to
+thinking of related statistics and data science functionality in terms
+of modules. Because modules are enabled at run-time, including new
+functionality is as simple as downloading and placing a module within
+\texttt{intRo}'s modules folder, or removing existing modules from that
+folder. Furthermore, errors can be more easily isolated to specific
+components. For instance, if an error is encountered, simply disabling
+the module can provide a temporary workaround while the issue is
+identified. Finally, modularity helps to organize the different pieces
+of code into functionality chunks that make it easier for developers to
+maintain.
 
@@
 They are more general and suitable for any application. \texttt{intRo}
-modules are specifically for statistics functionality and work within
-the \texttt{intRo} application only.
+modules are for statistics functionality and work within the
+\texttt{intRo} application only.
 
@@
 according to the name of the module. This folder houses the previously
-defined scripts. As an example, we will walk through the contents of the
-\texttt{nonparametric} module as previously mentioned in this section.
+defined scripts. As an example, we will walk through the process of
+creating a new module called \texttt{nonparametric}, as previously
+mentioned in this section, which will perform a wilcoxon rank sum test.
 
 Since the \texttt{nonparametric} module performs a statistical test, it
-is part of the inference category, and hence can be found in the
+fits within the inference category, and hence should be placed in the
 \texttt{intRo} repository at \textbf{modules/inference/nonparametric}.
-Let's first examine \emph{helper.R}:
+Let's first create \emph{helper.R}:
 
@@
 found in the \texttt{base} and \texttt{stats} package, the
-\texttt{libraries.R} file is empty for the \texttt{nonparametric}
-module. Additionally, there are no reactive objects defined, so
-\texttt{reactive.R} is also empty. \texttt{observe.R}, which defines the
-Shiny observers needed, is reproduced below:
+\texttt{libraries.R} file will be empty for the \texttt{nonparametric}
+module. Additionally, no reactive objects need be defined, so
+\texttt{reactive.R} will also be empty. \texttt{observe.R}, which
+defines the Shiny observers needed, can be written as follows:
 
@@
 paradigm which do not return a value (RStudio and Inc. 2014). For
-further discussion of reactivity, see section
-\ref{reactive-design-choices}. In this example, observers are created to
-ensure that the choices of variable for the \texttt{nonparametric}
+further discussion of reactivity, see Section
+\ref{designing-for-reactivity}. In this example, observers are created
+to ensure that the choices of variable for the \texttt{nonparametric}
 module are only numeric variables. This is accomplished by utilizing the
@@
 
-The \texttt{output.R} code is very simple:
+The \texttt{output.R} code can be very simple:
 
@@
 
-Finally, the \texttt{ui.R} code is shown below:
+Finally, a possible \texttt{ui.R} file is shown below:
 
@@
 This script defines all the inputs and outputs that the student will
-see. The only requirements from \texttt{intRo\textquotesingle{}s}
-perspective are (1) that there exist a store button at the bottom of the
-middle panel for storing the results of the analysis in the code script,
-and (2) that configuration options appear in the width 4 column in the
-middle, and output appears in the width 8 column on the right. The
-remaining input and output definitions depend on the statistical
-analysis or transformation being performed.
+see. The only requirements from \texttt{intRo}'s perspective are (1)
+that there exist a store button at the bottom of the middle panel for
+storing the results of the analysis in the code script, and (2) that
+configuration options appear in the width 4 column in the middle, and
+output appears in the width 8 column on the right. The remaining input
+and output definitions depend on the statistical analysis or
+transformation being performed.
 
@@
 
-\subsection{Reactive design choices}\label{reactive-design-choices}
+\subsection{Designing for
+Reproducibility}\label{designing-for-reproducibility}
+
+While web-based tools written using Shiny (including \texttt{intRo})
+have appealing characteristics such as being multi-platform, requiring
+no installation, and requiring no software licenses, one limitation
+immediately presents itself. The actions taken in the application are
+typically not reproducible as in a standard R script. We have designed
+\texttt{intRo} to overcome this limitation by capturing the unevaluated
+expression of all actions taken by the user in the interface. This
+expression is then parsed and printed in a code window at the bottom,
+while simultaneously being executed by the R process running on the
+server.
+
+In essence, this procedure transcribes user actions in a GUI to R code.
+When run in a standard R session, the results produced will be identical
+to the results shown in \texttt{intRo}. The full series of actions taken
+by the user are transcribed and can then be exchanged by researchers,
+students, and developers in a manner similar to normal scripting. Even
+``printing'' the results of an \texttt{intRo} session amounts to nothing
+more than executing the given code on the server, and then storing the
+results in an rmarkdown document, weaving the code with the results to
+produce a full compendium. While not strictly necessary, this lends
+credibility to the results produced by \texttt{intRo} in the sense that
+they are directly reproduced by the server every time the user clicks
+the print button.
+
+Reproducibility in \texttt{intRo} is accomplished with the previously
+mentioned \texttt{interpolate} function. \texttt{interpolate} accepts an
+expression and an arbitrary number of arguments as an argument,
+substitutes the arguments into the expression, prints the results to the
+console, and evaluates the parsed expression. This allows for all
+modules to be shoe-horned into the framework by wrapping the resulting R
+code in calls to \texttt{interpolate}. A possible drawback of this
+solution is that it requires module developers to manually wrap their
+functions in this call, but this could be mitigated by a package that
+creates modules automatically (See Section
+\ref{conclusions-and-future-work}).
+
+One potential enhancement to this framework would be the inclusion of
+state-saving and state-resuming. Because an \texttt{intRo} session is
+uniquely represented by the series of commands stored as code, the code
+itself could represent a checkpoint for resuming a new \texttt{intRo}
+session. Currently, beginning a new session will start the application
+with no memory of previous sessions. In real-world usage, state saving
+could allow a user to continue work later. At this time, this can only
+be done by taking the code and pasting it into a standard R session,
+although such an enhancement would likely involve minimal changes to
+\texttt{intRo}'s underlying structure.s
+
+\subsection{Designing for Reactivity}\label{designing-for-reactivity}
 
@@
 
-\section{\texorpdfstring{Deploying \texttt{intRo}
+\section{Conclusions and Future Work}\label{conclusions-and-future-work}
+
+In this paper, we have outlined a framework for designing a web-based,
+modular, extensible system which reproduces user actions into R code. We
+believe that the development strategies we've outlined can and should be
+applied to other software systems, as each of these characteristics aids
+in the ease-of-use and functionality of the overall product. Although we
+present them in the context of an introductory statistics application,
+these ideas are generalizable and we hope that they will gain traction
+in many other modern software systems.
+
+With regards to \texttt{intRo} itself, we believe it can be a powerful
+and effective tool for introductory statistics education. Its modular
+structure allows it to be flexible enough for many different
+applications and curriculums. Its ease-of-use allows the student to
+focus her attention on the statistics task at hand, rather than
+struggling with software licenses and confusing interface navigation.
+Reproducible code generated from each analysis can be used to spark an
+interest in R programming in those who might otherwise not be exposed to
+it.
+
+In addition to the current functionality, there are some practical
+improvements in the works that will make \texttt{intRo} more useful to
+both students and instructors. In particular, we have begun development
+on an R package which will allow \texttt{intRo} modules to be created
+automatically from user written R code. This package will generate the
+necessary file structure to allow the module's incorporation into
+\texttt{intRo} as well as translate user code to \texttt{intRo}
+compatible code and populate the necessary files. This will vastly
+improve \texttt{intRo}'s flexibility and allow it to be used in a wider
+range of curricula, including more advanced statistics courses.
+Additionally, we would like to expand the interactive capabilities of
+our graphics in order to make \texttt{intRo}'s plots more engaging to
+students. One way to do this would be implementing linked plots, in
+which interactions with one plot are reflected in other plots that
+illustrate the same data. This would be particularly useful in the
+regression module so that students could explore observations with high
+influence and leverage.
+
+We hope to use \texttt{intRo} in courses to collect feedback regarding
+the ease of use and functionality. This will allow us to assess its
+usefulness relative to software used in the past, as well as gauge areas
+for improvement. Furthermore, we can determine the effectiveness of code
+printing on generating excitement from the students about programming in
+R.
+
+Challenges do exist with regards to the wider adoption of
+\texttt{intRo}. For instance, we will need to monitor how well the
+server hosting \texttt{intRo} handles the load of dozens of students
+performing data analyses at once. If performance issues are encountered,
+the infrastructure used may need to be expanded to handle current and
+future load. An unknown quantity will be how feasible it is to increase
+adoption of \texttt{intRo} across Iowa State, as well as to other
+universities. One limitation of \texttt{intRo} is that uploading a
+dataset beyond about 30,000 rows tends to be slow. Even once the data is
+successfully uploaded, the default modules produce results more slowly
+than with smaller datasets. This is a limitation that should be further
+investigated if and when \texttt{intRo} sees wider adoption.
+
+Regardless, tools that focus on usability and extensibility, such as
+\texttt{intRo}, are sure to encourage the next round of innovators to be
+interested and excited about statistical computing.
+
+\section{Supplementary Material}\label{supplementary-material}
+
+All code and documents related to this manuscript are available at
+\url{https://github.com/gammarama/intRo}.
+
+\section{Appendix}\label{appendix}
+
+\subsection{Dynamic UI Generation}\label{dynamic-ui-generation}
+
+\texttt{intRo}'s user interface and functionality is dynamically
+generated depending on the set of modules enabled. The key driver to
+populating \texttt{server.R} and \texttt{ui.R} is the modules folder,
+the directory structure of which defines the placement of each module.
+The interface is then created with the following statement.
+
+\footnotesize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{## Source ui}
+\NormalTok{mylist <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
+\NormalTok{old_heading <-}\StringTok{ ""}
+\NormalTok{for (i in }\KeywordTok{seq_along}\NormalTok{(modules)) \{}
+    \NormalTok{my.module <-}\StringTok{ }\KeywordTok{strsplit}\NormalTok{(modules[i], }\StringTok{"/"}\NormalTok{)[[}\DecValTok{1}\NormalTok{]]}
+    \NormalTok{if (my.module[}\DecValTok{1}\NormalTok{] !=}\StringTok{ }\NormalTok{old_heading) \{}
+        \NormalTok{mylist[[}\KeywordTok{length}\NormalTok{(mylist) +}\StringTok{ }\DecValTok{1}\NormalTok{]] <-}\StringTok{ }\NormalTok{Hmisc::}\KeywordTok{capitalize}\NormalTok{(my.module[}\DecValTok{1}\NormalTok{])}
+        \NormalTok{old_heading <-}\StringTok{ }\NormalTok{my.module[}\DecValTok{1}\NormalTok{]}
+    \NormalTok{\}}
+    \NormalTok{mylist[[}\KeywordTok{length}\NormalTok{(mylist) +}\StringTok{ }\DecValTok{1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{get}\NormalTok{(}\KeywordTok{paste}\NormalTok{(my.module[}\DecValTok{2}\NormalTok{], }
+        \StringTok{"ui"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{"_"}\NormalTok{))}
+\NormalTok{\}}
+
+\NormalTok{## mylist is a list containing the different ui}
+\NormalTok{## module code Create the UI}
+\KeywordTok{shinyUI}\NormalTok{(}\KeywordTok{navbarPage}\NormalTok{(}\StringTok{"intRo"}\NormalTok{, }\DataTypeTok{id =} \StringTok{"top-nav"}\NormalTok{, }\DataTypeTok{theme =} \StringTok{"bootstrap.min.css"}\NormalTok{, }
+    \KeywordTok{tabPanel}\NormalTok{(}\DataTypeTok{title =} \StringTok{""}\NormalTok{, }\DataTypeTok{icon =} \KeywordTok{icon}\NormalTok{(}\StringTok{"home"}\NormalTok{), }\KeywordTok{fluidRow}\NormalTok{(}\KeywordTok{do.call}\NormalTok{(navlistPanel, }
+        \KeywordTok{c}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{id =} \StringTok{"side-nav"}\NormalTok{, }\DataTypeTok{widths =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{10}\NormalTok{)), }
+            \NormalTok{mylist)))), ...))}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+The key piece of code being the \texttt{do.call} statement loading the
+list of ui elements from the module's \texttt{ui.R} file. The server
+functions are then dynamically generated using a similar method.
+
+\footnotesize
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{shinyServer}\NormalTok{(function(input, output, session) \{}
+    \NormalTok{types <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"helper.R"}\NormalTok{, }\StringTok{"observe.R"}\NormalTok{, }\StringTok{"reactive.R"}\NormalTok{, }
+        \StringTok{"output.R"}\NormalTok{)}
+
+    \NormalTok{modules_tosource <-}\StringTok{ }\KeywordTok{file.path}\NormalTok{(}\StringTok{"modules"}\NormalTok{, }\KeywordTok{apply}\NormalTok{(}\KeywordTok{expand.grid}\NormalTok{(modules, }
+        \NormalTok{types), }\DecValTok{1}\NormalTok{, paste, }\DataTypeTok{collapse =} \StringTok{"/"}\NormalTok{))}
+
+    \NormalTok{for (mod in modules_tosource) \{}
+        \KeywordTok{source}\NormalTok{(mod, }\DataTypeTok{local =} \OtherTok{TRUE}\NormalTok{)}
+    \NormalTok{\}}
+\NormalTok{\})}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+In this way, we were able to have \texttt{intRo} be fully extensible,
+its structure and functionality dependent entirely on the modules
+present within the application.
+
+\subsection{\texorpdfstring{Deploying \texttt{intRo}
 Instances}{Deploying intRo Instances}}\label{deploying-intro-instances}
@@
 
-\section{Conclusions and future work}\label{conclusions-and-future-work}
-
-\texttt{intRo} can be a powerful and effective tool for introductory
-statistics education. Its modular structure allows it to be flexible
-enough for many different applications and curriculums. Its ease-of-use
-allows the student to focus her attention on the statistics task at
-hand, rather than struggling with software licenses and confusing
-interface navigation. Reproducible code generated from each analysis can
-be used to spark an interest in R programming in those who might
-otherwise not be exposed to it.
-
-In addition to the current functionality, there are some practical
-improvements in the works that will make \texttt{intRo} more useful to
-both students and instructors. In particular, we have begun development
-on an R package which will allow \texttt{intRo} modules to be created
-automatically from user written R code. This package will generate the
-necessary file structure to allow the module's incorporation into
-\texttt{intRo} as well as translate user code to \texttt{intRo}
-compatible code and populate the necessary files. This will vastly
-improve \texttt{intRo\textquotesingle{}s} flexibility and allow it to be
-used in a wider range of curricula, including more advanced statistics
-courses. Additionally, we would like to utilize the expanded interactive
-capabilities of ggvis in order to make \texttt{intRo's} plots more
-engaging to students. One way to do this would be implementing linked
-plots, in which interactions with one plot are reflected in other plots
-that illustrate the same data. This would be particularly useful in the
-regression module so that students could explore observations with high
-influence and leverage.
-
-We hope to use \texttt{intRo} in courses to collect feedback regarding
-the ease of use and functionality. This will allow us to assess its
-usefulness relative to software used in the past, as well as gauge areas
-for improvement. Furthermore, we can determine the effectiveness of code
-printing on generating excitement from the students about programming in
-R.
-
-Challenges do exist with regards to the wider adoption of
-\texttt{intRo}. For instance, we will need to monitor how well the
-server hosting \texttt{intRo} handles the load of dozens of students
-performing data analyses at once. If performance issues are encountered,
-the infrastructure used may need to be expanded to handle current and
-future load. An unknown quantity will be how feasible it is to increase
-adoption of \texttt{intRo} across Iowa State, as well as to other
-universities. One limitation of \texttt{intRo} is that uploading a
-dataset beyond about 30,000 rows tends to be slow. Even once the data is
-successfully uploaded, the default modules produce results more slowly
-than with smaller datasets. This is a limitation that should be further
-investigated if and when \texttt{intRo} sees wider adoption.
-
-Regardless, tools that focus on usability and extensibility in
-statistics education, such as \texttt{intRo}, are sure to encourage the
-next round of innovators to be interested and excited about statistical
-computing.
-
-\section{Supplementary material}\label{supplementary-material}
-
-All code and documents related to this manuscript are available at
-\url{https://github.com/gammarama/intRo}.
-
-\section{Appendix}\label{appendix}
-
-\subsection{Dynamic UI Generation}\label{dynamic-ui-generation}
-
-\texttt{intRo\textquotesingle{}s} user interface and functionality is
-dynamically generated depending on the set of modules enabled. The key
-driver to populating \texttt{server.R} and \texttt{ui.R} is the modules
-folder, the directory structure of which defines the placement of each
-module. The interface is then created with the following statement.
-
-\footnotesize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## Source ui}
-\NormalTok{mylist <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
-\NormalTok{old_heading <-}\StringTok{ ""}
-\NormalTok{for (i in }\KeywordTok{seq_along}\NormalTok{(modules)) \{}
-    \NormalTok{my.module <-}\StringTok{ }\KeywordTok{strsplit}\NormalTok{(modules[i], }\StringTok{"/"}\NormalTok{)[[}\DecValTok{1}\NormalTok{]]}
-    \NormalTok{if (my.module[}\DecValTok{1}\NormalTok{] !=}\StringTok{ }\NormalTok{old_heading) \{}
-        \NormalTok{mylist[[}\KeywordTok{length}\NormalTok{(mylist) +}\StringTok{ }\DecValTok{1}\NormalTok{]] <-}\StringTok{ }\NormalTok{Hmisc::}\KeywordTok{capitalize}\NormalTok{(my.module[}\DecValTok{1}\NormalTok{])}
-        \NormalTok{old_heading <-}\StringTok{ }\NormalTok{my.module[}\DecValTok{1}\NormalTok{]}
-    \NormalTok{\}}
-    \NormalTok{mylist[[}\KeywordTok{length}\NormalTok{(mylist) +}\StringTok{ }\DecValTok{1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{get}\NormalTok{(}\KeywordTok{paste}\NormalTok{(my.module[}\DecValTok{2}\NormalTok{], }
-        \StringTok{"ui"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{"_"}\NormalTok{))}
-\NormalTok{\}}
-
-\NormalTok{## mylist is a list containing the different ui}
-\NormalTok{## module code Create the UI}
-\KeywordTok{shinyUI}\NormalTok{(}\KeywordTok{navbarPage}\NormalTok{(}\StringTok{"intRo"}\NormalTok{, }\DataTypeTok{id =} \StringTok{"top-nav"}\NormalTok{, }\DataTypeTok{theme =} \StringTok{"bootstrap.min.css"}\NormalTok{, }
-    \KeywordTok{tabPanel}\NormalTok{(}\DataTypeTok{title =} \StringTok{""}\NormalTok{, }\DataTypeTok{icon =} \KeywordTok{icon}\NormalTok{(}\StringTok{"home"}\NormalTok{), }\KeywordTok{fluidRow}\NormalTok{(}\KeywordTok{do.call}\NormalTok{(navlistPanel, }
-        \KeywordTok{c}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{id =} \StringTok{"side-nav"}\NormalTok{, }\DataTypeTok{widths =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{10}\NormalTok{)), }
-            \NormalTok{mylist)))), ...))}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-
-The key piece of code being the \texttt{do.call} statement loading the
-list of ui elements from the module's \texttt{ui.R} file. The server
-functions are then dynamically generated using a similar method.
-
-\footnotesize
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{shinyServer}\NormalTok{(function(input, output, session) \{}
-    \NormalTok{types <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"helper.R"}\NormalTok{, }\StringTok{"observe.R"}\NormalTok{, }\StringTok{"reactive.R"}\NormalTok{, }
-        \StringTok{"output.R"}\NormalTok{)}
-
-    \NormalTok{modules_tosource <-}\StringTok{ }\KeywordTok{file.path}\NormalTok{(}\StringTok{"modules"}\NormalTok{, }\KeywordTok{apply}\NormalTok{(}\KeywordTok{expand.grid}\NormalTok{(modules, }
-        \NormalTok{types), }\DecValTok{1}\NormalTok{, paste, }\DataTypeTok{collapse =} \StringTok{"/"}\NormalTok{))}
-
-    \NormalTok{for (mod in modules_tosource) \{}
-        \KeywordTok{source}\NormalTok{(mod, }\DataTypeTok{local =} \OtherTok{TRUE}\NormalTok{)}
-    \NormalTok{\}}
-\NormalTok{\})}
-\end{Highlighting}
-\end{Shaded}
-
-\normalsize
-In this way, we were able to have \texttt{intRo} be fully extensible,
-its structure and functionality dependent entirely on the modules
-present within the application.
-
 \section*{References}\label{references}
@@
 
+\hypertarget{ref-radiant}{}
+Nijs, Vincent. 2016. ``Radiant - Business Analytics Using R and Shiny.''
+\url{https://radiant-rstats.github.io/docs}.
+
+\hypertarget{ref-parnas1972criteria}{}
+Parnas, David Lorge. 1972. ``On the Criteria to Be Used in Decomposing
+Systems into Modules.'' \emph{Communications of the ACM} 15 (12). ACm:
+1053--8.
+
 \hypertarget{ref-mosaic}{}
@@
 Ltd.
+
+\hypertarget{ref-szyperski1996independently}{}
+Szyperski, Clemens. 1996. ``Independently Extensible Systems-Software
+Engineering Potential and Challenges.'' \emph{Australian Computer
+Science Communications} 18. UNIVERSITY OF CANTERBURY: 203--12.
 

### 2016-12-02 jcgs-paper-2016/paper-2ndrev.Rmd
Start on the edits

--- 2016-10-10 jcgs-paper-2016/paper.Rmd
@@
     template: jcgs_article.tex
+    keep_tex: true
 fontsize: 12pt
@@
    Modular programming is a development paradigm that emphasizes self-contained, flexible, and independent pieces of functionality.
-This practice allows new functionality to be seamlessly added when new functionality is desired, and unwanted functionality to be removed, thus simplify the user-facing view of the software.
-However, the rise of web-based software applications has presented new challenges for designing an extensible, modular software system.
+This practice allows new features to be seamlessly added when desired, and unwanted features to be removed, thus simplifying the user-facing view of the software.
+The recent rise of web-based software applications has presented new challenges for designing an extensible, modular software system.
 In this paper, we outline a framework for designing such a system, with a focus on reproducibility of the results.
@@
 In particular, functionality can be dynamically loaded and unloaded depending on the particular use case.
-Open source modular software opens up the possibility of extensions contributed by a wide array of programmers, which can allow the software to morph into areas that weren't anticipated early in development.
+Open source modular software precipitates the possibility of extensions contributed by a wide array of programmers, which can allow the software to morph into areas that weren't anticipated early in development.
 In the statistics realm, R [@r-stat] is a prime example of the virtues of modular programming.
@@
 Multiple software packages have recently been written in an attempt to spur this interest in R programming and statistics.
-DataCamp's [@datacamp] courses are user-friendly ways to learning basic R programming and data analysis techniques.
+DataCamp's [@datacamp] courses are a user-friendly way to learn basic R programming and data analysis techniques.
 Swirl [@swirl] is a similar interactive tool to make learning R more fun by learning it within R itself.
@@
 One such package is iNZight Lite [@inzight] which attempts to expose students to data analysis without requiring programming knowledge.
-Like most web-based systems, this does not include reproducible R code, which limits its useful in a scientific or academic setting.
+Like most web-based systems, this does not include reproducible R code, which limits its usefulness in a scientific or academic setting.
 Another package is called Radiant [@radiant], which is a web-based application with the aim of furthering business education and financial analysis.
 While the application is modular and extensible, it does require installation and hosting and is inundated with more features than necessary for an introductory student.
-An overview of the comparison between the features of these statistical software packages is presented in table \ref{tab:compare}.
+An overview of the comparison between the features of these statistical software packages is presented in Table \ref{tab:compare}.
 Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default.
-For example, `R` does have an associated GUI, however this interface is very limited, thus only partially fulfilling the behavior of a graphical user interface.
+For example, `R` does have an associated Graphical User Interface (GUI), however this interface is very limited, thus only partially fulfilling the behavior of a GUI.
 
@@
 
-Though challenging in a Graphical User Interface, a reproducibility framework has three key advantages.
+Though challenging in a GUI, a reproducibility framework has three key advantages.
 First, it eases a student who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
@@
 Second, an analysis created by a reproducible software system can be brought in an R session to easily assess and extend the results.
-Finally, with the help of knitr and RMarkdown, "printing" the results of a reproducible software system analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
+Finally, with the help of knitr [@xie2015] and rmarkdown [@rmarkdown], "printing" the results of a reproducible software system analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
 These concepts are important because they encourage best practices with regards to disclosure of analysis methods in research [@baggerly2011reproducible; @xie2015].
 
-Based on this, we believe a modern software system should be **modular**, **extensible**, **web-based**, and foster **reproducibility**.
-We have developed a case-study application called intRo which we will use to illustrate our method of developing a system meeting these criteria.
-The paper is structured as follows:.
-Section \ref{case-study-intro} introduces the application, its features and its usability, and provides motivations for why it was built.
-Section \ref{intro-design-decisions} provides technical details on how we built intRo, by walking through the underlying modularity, reproducibility, and reactive framework, as well as how it can be used to develop other software systems with these properties.
-Finally, section \ref{conclusions-and-future-work} discusses some future possibilities and limitations of both intRo, and modular systems in general.
+Based on the above, we believe a modern software system should be **modular**, **extensible**, **web-based**, and foster **reproducibility**.
+We have developed a case-study application called `intRo` which we will use to illustrate our method of developing a system meeting these criteria.
+The paper is structured as follows: Section \ref{case-study-intro} introduces the application, its features and its usability, and provides motivations for why it was built.
+Section \ref{intro-design-decisions} provides technical details on how we built `intRo`, by walking through the underlying modularity, reproducibility, and reactive framework, as well as how it can be used to develop other software systems with these properties.
+Finally, Section \ref{conclusions-and-future-work} discusses some future possibilities and limitations of both `intRo`, and modular systems in general.
 
@@
 Additionally, `intRo` stands apart from new tools in that it is a supplement to an existing class, fully usable by a beginning statistics student.
-An accompanying R package, titled `intRo` and available on GitHub, assists in the downloading, running, and deploying of `intRo` instances (See section \ref{deploying-intro-instances}).
+An accompanying R package, titled `intRo` and available on GitHub, assists in the downloading, running, and deploying of `intRo` instances (See Section \ref{deploying-intro-instances}).
 
@@
 The first is a link to toggle the visibility of the code panel (5).
-The middle icon downloads an rmarkdown [@rmarkdown] document of the analysis performed.
+The middle icon downloads an rmarkdown document of the analysis performed.
 The last is a link to print the stored module results, and the associated code (if visible).
@@
 Under each of these categories, there are seven default modules, which perform specific data analysis tasks that employ an easy to use point-and-click interface.
-More modules can easily be added by an instructor, as detailed in section \ref{designing-for-modularity}.
+More modules can easily be added by an instructor, as detailed in Section \ref{designing-for-modularity}.
 The default modules support uploading and downloading a dataset, transforming variables, graphical and numerical summaries, simple linear regression, contingency tables, and T-tests.
@@
 `intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework.
-A high-level diagram of this process is given in figure \ref{fig:app_creation_modules}.
+A high-level diagram of this process is given in Figure \ref{fig:app_creation_modules}.
 `intRo` modules are split up into multiple R scripts which are included either in Shiny's user interface or server definitions.
@@
 
-Modularity was a design decision we focused on from the start of `intRo's` development.
+Modularity was a design decision we focused on from the start of `intRo`'s development.
 There are some practical benefits to thinking of related statistics and data science functionality in terms of modules.
-Course instructors can easily enable or disable functionality depending on the exact needs of the particular class being taught.
-Because this is done at run-time, including new modules is as simple as downloading and placing it within `intRo's` modules folder, or removing existing modules from that folder.
+Because modules are enabled at run-time, including new functionality is as simple as downloading and placing a module within `intRo`'s modules folder, or removing existing modules from that folder.
 Furthermore, errors can be more easily isolated to specific components.
@@
 
-Since the `nonparametric` module performs a statistical test, it is part of the inference category, and hence would be placed in the `intRo` repository at **modules/inference/nonparametric**.
-Let's first examine *helper.R*:
+Since the `nonparametric` module performs a statistical test, it fits within the inference category, and hence should be placed in the `intRo` repository at **modules/inference/nonparametric**.
+Let's first create *helper.R*:
 
@@
 
-Because all the code needed to implement a wilcoxon rank sum test is found in the `base` and `stats` package, the `libraries.R` file is empty for the `nonparametric` module.
-Additionally, there are no reactive objects defined, so `reactive.R` is also empty.
-`observe.R`, which defines the Shiny observers needed, is reproduced below:
+Because all the code needed to implement a wilcoxon rank sum test is found in the `base` and `stats` package, the `libraries.R` file will be empty for the `nonparametric` module.
+Additionally, no reactive objects need be defined, so `reactive.R` will also be empty.
+`observe.R`, which defines the Shiny observers needed, can be written as follows:
 
@@
 Shiny observers are a class of reactive objects within the Shiny paradigm which do not return a value [@shiny].
-For further discussion of reactivity, see section \ref{designing-for-reactivity}.
+For further discussion of reactivity, see Section \ref{designing-for-reactivity}.
 In this example, observers are created to ensure that the choices of variable for the `nonparametric` module are only numeric variables.
@@
 
-The `output.R` code is very simple:
+The `output.R` code can be very simple:
 
@@
 
-Finally, the `ui.R` code is shown below:
+Finally, a possible `ui.R` file is shown below:
 
@@
 This script defines all the inputs and outputs that the student will see.
-The only requirements from `intRo's` perspective are (1) that there exist a store button at the bottom of the middle panel for storing the results of the analysis in the code script, and (2) that configuration options appear in the width 4 column in the middle, and output appears in the width 8 column on the right.
+The only requirements from `intRo`'s perspective are (1) that there exist a store button at the bottom of the middle panel for storing the results of the analysis in the code script, and (2) that configuration options appear in the width 4 column in the middle, and output appears in the width 8 column on the right.
 The remaining input and output definitions depend on the statistical analysis or transformation being performed.
@@
 This package will generate the necessary file structure to allow the module's incorporation into `intRo` as well as translate user code to `intRo` compatible code and populate the necessary files.
-This will vastly improve `intRo's` flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
-Additionally, we would like to utilize the expanded interactive capabilities of ggvis in order to make \texttt{intRo's} plots more engaging to students.
+This will vastly improve `intRo`'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses.
+Additionally, we would like to expand the interactive capabilities of our graphics in order to make `intRo`'s plots more engaging to students.
 One way to do this would be implementing linked plots, in which interactions with one plot are reflected in other plots that illustrate the same data.
@@
 
-`intRo's` user interface and functionality is dynamically generated depending on the set of modules enabled.
+`intRo`'s user interface and functionality is dynamically generated depending on the set of modules enabled.
 The key driver to populating \texttt{server.R} and \texttt{ui.R} is the modules folder, the directory structure of which defines the placement of each module.

### 2016-12-02 jcgs-paper-2016/paper.tex
Start on the edits

@@
 attempts to expose students to data analysis without requiring
-programming knowledge. Like most web-based systems, this does not
-include reproducible R code, which limits its usefulness in a scientific
-or academic setting. Another package is called Radiant (Nijs 2016),
-which is a web-based application with the aim of furthering business
-education and financial analysis. While the application is modular and
-extensible, it does require installation and hosting and is inundated
-with more features than necessary for an introductory student. An
-overview of the comparison between the features of these statistical
-software packages is presented in Table \ref{tab:compare}. Partial
-fulfillment of requirements is noted in the table, as well as a measure
-of the complexity of functionality offered by default. For example,
-\texttt{R} does have an associated Graphical User Interface (GUI),
-however this interface is very limited, thus only partially fulfilling
-the behavior of a GUI.
+programming knowledge. inZight Lite is currently partially but not fully
+reproducible. Another package is called Radiant (Nijs 2016), which is a
+web-based application with the aim of furthering business education and
+financial analysis. While the application is modular and extensible, it
+does require installation and hosting and is inundated with more
+features than necessary for an introductory student. An overview of the
+comparison between the features of these statistical software packages
+is presented in Table \ref{tab:compare}. Partial fulfillment of
+requirements is noted in the table, as well as a measure of the
+complexity of functionality offered by default. For example, \texttt{R}
+does have an associated Graphical User Interface (GUI), however this
+interface is very limited, thus only partially fulfilling the behavior
+of a GUI.
 
@@
   MOSAIC & No & Yes & No & No & No & Yes & Limited \\ 
-  iNZight Lite & Yes & No & Yes & Yes & Partial & No & Limited \\ 
+  iNZight Lite & Yes & No & Yes & Yes & Partial & Partial & Limited \\ 
   Radiant & Yes & Yes & Yes & Yes & Partial & Yes & Moderate \\ 
@@
 \caption{This figure depicts how the Shiny \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.
-\texttt{intRo} modules are split up into multiple R scripts which are included either in Shiny's user interface or server definitions.
-At runtime, the \texttt{intRo} sources in the specified modules (contained in the modules folder) to dynamically generate the functionality available in the application.
+\texttt{intRo} modules are split up into multiple R scripts which are included either in Shiny's user interface (\texttt{ui.R}) or server (\texttt{helper.R}, \texttt{libraries.R}, \texttt{observe.R}, \texttt{reactive.R}, \texttt{output.R}) definitions.
+At runtime, \texttt{intRo} sources in the specified modules (contained in the modules folder) to dynamically generate the functionality available in the application.
 This allows for the specific functionality needed to be determined and adjusted by the individual course instructor.

### 2016-12-02 jcgs-paper-2016/paper.Rmd
Some more small minor edits

@@
 One thing these software packages all have in common is the requirement of local installation and configuration, which means certain operating systems and platforms will not support their use.
+For example, Excel and JMP are not supported on Linux.
 
@@
 One such package is iNZight Lite [@inzight] which attempts to expose students to data analysis without requiring programming knowledge.
-Like most web-based systems, this does not include reproducible R code, which limits its useful in a scientific or academic setting.
+inZight Lite is currently partially but not fully reproducible.
 Another package is called Radiant [@radiant], which is a web-based application with the aim of furthering business education and financial analysis.
-While the application is modular and extensible, it does require installation and hosting and is inundated with more features than necessary for an introductory student.
-An overview of the comparison between the features of these statistical software packages is presented in table \ref{tab:compare}.
+While the application is modular and extensible, it does require local installation and hosting.
+An overview of the comparison between the features of these statistical software packages is presented in Table \ref{tab:compare}.
 Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default.
-For example, `R` does have an associated GUI, however this interface is very limited, thus only partially fulfilling the behavior of a graphical user interface.
+For example, `R` does have an associated Graphical User Interface (GUI), however this interface is very limited, thus only partially fulfilling the behavior of a GUI.
 
@@
                           c("MOSAIC", "No", "Yes", "No", "No", "No", "Yes", "Yes", "Limited"),
-                          c("iNZight Lite", "Yes", "No", "Yes", "Yes", "Partial", "No", "Yes", "Limited"),
+                          c("iNZight Lite", "Yes", "No", "Yes", "Yes", "Partial", "Partial", "Yes", "Limited"),
                           c("Radiant", "Yes", "Yes", "Yes", "Yes", "Partial", "Yes", "Yes", "Moderate"))
@@
 However, using R in most cases still requires a basic knowledge of programming concepts which may pose a steep learning curve for the introductory statistics student [@5359977].
-This additional time commitment may explain why introductory courses often utilize point-and-click applications, even if the instructor himself/herself uses R in their own work.
+This additional time commitment may explain why introductory courses often utilize point-and-click applications, even if the instructors use R in their own work.
 Still, some compromises must be made when using many graphical applications, including dealing with software licenses and unsupported desktop platforms.
@@
 Additionally, `intRo` stands apart from new tools in that it is a supplement to an existing class, fully usable by a beginning statistics student.
-An accompanying R package, titled `intRo` and available on GitHub, assists in the downloading, running, and deploying of `intRo` instances (See section \ref{deploying-intro-instances}).
+An accompanying R package, titled `intRo` and available on GitHub, assists in the downloading, running, and deploying of `intRo` instances (see Section \ref{deploying-intro-instances}).
 
@@
 This allows for the specific functionality needed to be determined and adjusted by the individual course instructor.
-In this example, the instructor is electing to include a nonparametric module, which is not enabled by default, to allow the students to perform a wilcoxon rank sum test.
+In this example, the instructor is electing to include a nonparametric module, which is not enabled by default, to allow the students to perform a Wilcoxon rank sum test.
 
@@
 \caption{This figure depicts how the Shiny \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.
-\texttt{intRo} modules are split up into multiple R scripts which are included either in Shiny's user interface or server definitions.
-At runtime, the \texttt{intRo} sources in the specified modules (contained in the modules folder) to dynamically generate the functionality available in the application.
+\texttt{intRo} modules are split up into multiple R scripts which are included either in Shiny's user interface (\texttt{ui.R}) or server (\texttt{helper.R}, \texttt{libraries.R}, \texttt{observe.R}, \texttt{reactive.R}, \texttt{output.R}) definitions.
+At runtime, \texttt{intRo} sources in the specified modules (contained in the modules folder) to dynamically generate the functionality available in the application.
 This allows for the specific functionality needed to be determined and adjusted by the individual course instructor.
-In this example, the instructor is electing to include a nonparametric module, which is not enabled by default, to allow the students to perform a wilcoxon rank sum test.}
+In this example, the instructor is electing to include a nonparametric module, which is not enabled by default, to allow the students to perform a Wilcoxon rank sum test.}
 \label{fig:app_creation_modules}
@@
 This folder houses the previously defined scripts.
-As an example, we will walk through the process of creating a new module called `nonparametric`, as previously mentioned in this section, which will perform a wilcoxon rank sum test.
+As an example, we will walk through the process of creating a new module called `nonparametric`, as previously mentioned in this section, which will perform a Wilcoxon rank sum test.
 
-Since the `nonparametric` module performs a statistical test, it is part of the inference category, and hence would be placed in the `intRo` repository at **modules/inference/nonparametric**.
-Let's first examine *helper.R*:
+Since the `nonparametric` module performs a statistical test, it fits within the inference category, and hence should be placed in the `intRo` repository at **modules/inference/nonparametric**.
+Let's first create *helper.R*:
 
@@
 This script is most immediately similar to standard R code.
-In this case, a function `nonparametrictest` is created which, depending on the values of the parameters, ultimately returns the result of a wilcoxon rank sum test.
+In this case, a function `nonparametrictest` is created which, depending on the values of the parameters, ultimately returns the result of a Wilcoxon rank sum test.
 One important difference from a typical R script is that each call in the script is wrapped in a function called `interpolate` [@interpolate].
-`interpolate` both executes the given R code on the server, and also writes the code executed to the script window at the bottom of `intRo`.
+`interpolate` both executes the given R code on the server, and also writes the code executed to the script window at the bottom of `intRo` (see Section \ref{designing-for-reproducibility}).
 
-Because all the code needed to implement a wilcoxon rank sum test is found in the `base` and `stats` package, the `libraries.R` file is empty for the `nonparametric` module.
-Additionally, there are no reactive objects defined, so `reactive.R` is also empty.
-`observe.R`, which defines the Shiny observers needed, is reproduced below:
+Because all the code needed to implement a Wilcoxon rank sum test is found in the `base` and `stats` package, the `libraries.R` file will be empty for the `nonparametric` module.
+Additionally, no reactive objects need be defined, so `reactive.R` will also be empty.
+`observe.R`, which defines the Shiny observers needed, can be written as follows:
 
@@
 This allows for all modules to be shoe-horned into the framework by wrapping the resulting R code in calls to `interpolate`.
-A possible drawback of this solution is that it requires module developers to manually wrap their functions in this call, but this could be mitigated by a package that creates modules automatically (See Section \ref{conclusions-and-future-work}).
+A possible drawback of this solution is that it requires module developers to manually wrap their functions in this call, but this could be mitigated by a package that creates modules automatically (see Section \ref{conclusions-and-future-work}).
 

### 2016-12-02 jcgs-paper-2016/paper.tex
Some more small minor edits

@@
 into areas that weren't anticipated early in development. In the
-statistics realm, R (R Core Team 2014) is a prime example of the virtues
+statistics realm, R (R Core Team 2016) is a prime example of the virtues
 of modular programming. As of this writing, The Comprehensive R Archive
@@
 which means certain operating systems and platforms will not support
-their use.
+their use. For example, Excel and JMP are not supported on Linux.
 
@@
 financial analysis. While the application is modular and extensible, it
-does require installation and hosting and is inundated with more
-features than necessary for an introductory student. An overview of the
+does require local installation and hosting. An overview of the
 comparison between the features of these statistical software packages
@@
 additional time commitment may explain why introductory courses often
-utilize point-and-click applications, even if the instructor
-himself/herself uses R in their own work. Still, some compromises must
-be made when using many graphical applications, including dealing with
-software licenses and unsupported desktop platforms. From the
-instructor's perspective in particular, managing a large group of
-software licenses for students with various computing environments and
-versions could wind up being extremely cumbersome.
+utilize point-and-click applications, even if the instructors use R in
+their own work. Still, some compromises must be made when using many
+graphical applications, including dealing with software licenses and
+unsupported desktop platforms. From the instructor's perspective in
+particular, managing a large group of software licenses for students
+with various computing environments and versions could wind up being
+extremely cumbersome.
 
@@
 titled \texttt{intRo} and available on GitHub, assists in the
-downloading, running, and deploying of \texttt{intRo} instances (See
+downloading, running, and deploying of \texttt{intRo} instances (see
 Section \ref{deploying-intro-instances}).
@@
 electing to include a nonparametric module, which is not enabled by
-default, to allow the students to perform a wilcoxon rank sum test.
+default, to allow the students to perform a Wilcoxon rank sum test.
 
@@
 This allows for the specific functionality needed to be determined and adjusted by the individual course instructor.
-In this example, the instructor is electing to include a nonparametric module, which is not enabled by default, to allow the students to perform a wilcoxon rank sum test.}
+In this example, the instructor is electing to include a nonparametric module, which is not enabled by default, to allow the students to perform a Wilcoxon rank sum test.}
 \label{fig:app_creation_modules}
@@
 creating a new module called \texttt{nonparametric}, as previously
-mentioned in this section, which will perform a wilcoxon rank sum test.
+mentioned in this section, which will perform a Wilcoxon rank sum test.
 
@@
 on the values of the parameters, ultimately returns the result of a
-wilcoxon rank sum test. One important difference from a typical R script
+Wilcoxon rank sum test. One important difference from a typical R script
 is that each call in the script is wrapped in a function called
@@
 the given R code on the server, and also writes the code executed to the
-script window at the bottom of \texttt{intRo}.
-
-Because all the code needed to implement a wilcoxon rank sum test is
+script window at the bottom of \texttt{intRo} (see Section
+\ref{designing-for-reproducibility}).
+
+Because all the code needed to implement a Wilcoxon rank sum test is
 found in the \texttt{base} and \texttt{stats} package, the
@@
 functions in this call, but this could be mitigated by a package that
-creates modules automatically (See Section
+creates modules automatically (see Section
 \ref{conclusions-and-future-work}).
@@
 \hypertarget{ref-r-stat}{}
-R Core Team. 2014. \emph{R: A Language and Environment for Statistical
+R Core Team. 2016. \emph{R: A Language and Environment for Statistical
 Computing}. Vienna, Austria: R Foundation for Statistical Computing.
-\url{http://www.R-project.org/}.
+\url{https://www.R-project.org/}.
 

### 2016-12-09 jcgs-paper-2016/response.Rmd
New response and small edits

--- 2016-05-26 jcgs-paper-2016/response.Rmd
@@
 ---
-title: "Response to reviewers"
-author: "Eric Hare, Andee Kaplan"
-date: "May 26, 2016"
-output: html_document
+title: "Response to third review"
+date: "`r format(Sys.Date(), '%B %d, %Y')`"
+output: pdf_document
 ---
 
-# Reviewer 1
+# Associate Editor
 
-This is an intelligent paper describing interesting software, but it (and the associated software) suffer from three problems:
+1. two brief comments from one reviewer should both be addressed
 
-(1) The paper seems to me inappropriate for the JCGS.
-Because it focuses purely on software, I think that (once the other issues are addressed) it would be more appropriate for the Journal of Statistical Software or the R Journal.
+    **We have addressed these two comments**
 
-(2) Although the software looks promising, unless I've missed something, it currently has very limited functionality -- indeed, not enough to sustain most introductory statistics courses.
-This isn't a permanent limitation, however, since the authors have created interesting infrastructure for extending the software.
-In its current state, the software doesn't quite look ready for prime time.
+2. from other reviewer, address the minor fixes
 
-**Our intent was not for `intRo` to sustain an introductory statistics course, but rather to supplement it.
-In other words, limited functionality was actually a design goal from the beginning so as not to overwhelm a student seeing statistics for the first time.
-While we agree that it is likely there will be course topics that `intRo` doesn't support, if additional functionality is desired by the instructor, more modules can be added to support this.**
+    **As noted in the comments in this document, we have addressed what we felt were the more minor fixes**
 
-(3) The authors ignore other (and at present, much more capable) GUIs for R -- the R Commander and Deducer.
-Since these are the most direct competitors to intRo -- more so than, e.g., Excel and JMP -- it seems obvious that the authors should address similarities, differences, advantages, and disadvantages.
-One obvious difference is that these other GUIs run on individual computers rather than being web-based, and it would also make sense to address more generally the advantages and disadvantages of the two approaches.
-(For example, in my tests of intRo, I noticed slow response, even with very small data sets.
-Of course, it's hard to know to what to attribute that -- it could as easily have been a network issue as slowness at the server end.)
+3. remove the comments about Radiant being "inundated with more features than necessary"
+
+    **This comment has been removed**
+
+4. a sentence or two more on comparison with Radiant, e.g., mention Radiant's less focus on *extensibility* (if that is true)
+
+    **We have extended the comparison with Radiant by noting Radiant's more extensive functionality and the differences in the goals of each**
 
@@
 
-The “Article Assessment” questions on the report form don’t work for a paper like this.
-The authors report on some very impressive work.
-The simplicity of conception and execution is one of its biggest strengths.
-Their motivations are compelling, as is their system design and execution but I think the story-telling can be improved considerably.
+5.
+Page 3 re iNZight Lite.
+The model fitting module in iNZight writes R reproducible code (and has since its inception) and it all will eventually so tone back this criticism a little.
+Given the sentence that comes before, I think this sentence could be just deleted without weakening the motivation for this paper and project.
 
-**We appreciate the very kind comments and hope that this revision has improved the story being told.**
+    **We have clarified the differences in reproducibility as requested**
 
-I think the authors need to decide, “What sort of paper am I writing?” Included in the manuscript are elements of “a paper about teaching introductory statistics using a software system”, elements of “a paper about designing and building a simple extensible system”, and “a paper about how to use and extend a particular system”.
-I think the most compelling story for the Journal of Computational and Graphical Statistics is the designing and building of a simple online system for intro stat that is easy to learn, use and extend – and teaches R as a side effect.
-This would leave room for a sister publication in something like the Journal of Statistics Education which actually focusses on teaching aspects.
+6. Figure 3: I take it that the contents of helper.R, libraries.R, ... get sucked in to populate/form server.R. This could be made clearer (or this misconception countered).
 
-**We have taken care to significantly shift the focus away from the teaching aspects, and towards the extensible software system aspects.
-In particular, we have removed the lengthy teaching discussion and moved elements of it to the motivational section.
-We have also trimmed much of the content involving supported and unsupported functionality.
-Specific changes are noted in the responses to the further points raised below.**
+    **We have updated the figure caption to better explain how these files populate both the server and the ui definitions.**
 
-If you buy this conception, the teaching-related aspects should be shortened, shaved of much detail and form the basis of an introductory motivational section that leads into a design brief and a plan of attack (using shiny etc).
-There needs to be high-level discussion of how the system is conceived and put together accompanied by high-level diagrams before descending into code fragments (much detailed code-fragment discussion could go to appendices).
-The paper just needs to convey the sense of how things are done.
-Actual details of how to customise and extend the system are provided in the system’s documentation.
+# Reviewer 3
 
-Everywhere – distinguish between user=student (who uses what is served up) and user=teacher (who customises what is served up)
+7. p1. Not sure when this package was originally written, but R Core Team 2014 probably needs to be updated to 2016.
 
-p.
-2: Another online analysis system implemented in shiny is iNZight Lite at http://lite.docker.stat.auckland.ac.nz/.
-It is more ambitious and therefore more complex.
-Also it does not have the “exposing the R code” capability.
+    **We have updated the citation for R**
 
-p 3 bottom: Remove parentheses around “(and its corresponding code)” because this is actually one of the key features.
+8. I don't follow the logic here. Perhaps give an example of which platform does not support them.
 
-**Done**
+    **We have provided an example of this in the paper (Microsoft Excel and JMP are not supported on Linux)**
 
-R Commander should enter the discussion around here (even though it isn’t online)
+9. p2. "RStudio and Inc. 2014": the .bib entry should be author = {{RStudio, Inc.}}
 
-P 5: Fig 2 shows a simple design that should be easy for casual users (as intro students really are) to get their heads around.
-Not all that keen on the use of “statistics” as a category heading in Panel 2 because it reinforces the ghettoizing of the word conflicting with a message that says everything you are doing here is statistics.
+    **Fixed**
 
-P 6 top: When you look at the actual R code generated it can be quite complex with its pipe structures etc which I would have thought does rather limit its introductory value.
+10. perhaps change to "even if the instructors use R in their own work"?
 
-**We appreciate the comment.
-In teaching R Workshops at Iowa State, we observed that among the majority of the introductory R students we encountered, the pipe structure was often preferred to using traditional nested function calls.
-TODO: modify this to show that pipe structure is used more frequently.
-The pipe structure is also the standard by which `ggvis` plots are created, which we are using as the backend to our graphical module.
-It would also be a relatively simple task for an instructor to modify or remove the piping structure in a modified version of `intRo` for use in a particular class.**
+    **Fixed**
 
-P 6-7: discussion of particular modules should be in a “for example” vein. There is no need to justify what particular features are and are not included. (That would be in a teaching paper.)
+11. p9. wilcoxon -> Wilcoxon (also on following pages)
 
-**We have removed the in-depth discussion about the particular modules, and instead more vaguely described the functionality.
-We have also made sure to indicate that the full documentation is available online, which covers the supported functionality of every provided module.**
+    **We have corrected this throughout the paper**
 
-P 7 top: Being able to choose only plot types that are “compatible” with the variables used does, however, preclude learning experiences of the form “this thing you want to do is not appropriate for this type of data.”
+12.
+p10.
+I think pseudocode suffices in Figure 3; no need to put actual code in it.
+In server.R, typically the arguments are input, output, session, and I'm not sure if session, input, output is the same.
 
-**We have also gone back and forth on this issue.
-The reasoning behind choosing only plot types compatible is that the user will see the different plot types available as they select different variables.
-So while the teaching moment could take place post-selection (by, for instance, producing an error if a plot type is used with incompatible variable types), we chose to have the teaching moment be a step earlier, where the student may see the list and wonder why the plot type they thought they wanted is not actually an option.
-We felt that for a lot of introductory students, seeing a red error message is a frustrating moment and can discourage them from having any sort of autonomy within the software.
-Further, by doing a transformation of a variable, the plot types will automatically appear if they are compatible.**
+    **We have updated the order of the arguments in figure 3.**
 
-P 8: I think this descends to details too quickly.
-The big features are designing for extensibility, with conceptions about switching modules on and off and adding modules, and the interface as presented being assembled on the fly when the system starts up.
+13. p15. Okay, interpolate() was explained in detail here. Then p12 should have referenced this section.
 
-**We agree.
-We have restructured this portion of the paper to merely discuss the motivation behind an extensible system, with mentions of disabling and enabling modules.
-We have moved the details to the Understanding intRo section with the details on modules.**
-
-P 9-10 looks like appendix stuff to me.
-
-Section 3 discusses downloading onto a local machine before customising the hosted online system which should be the focus (again conception and issues involved before any details).
-
-**We have moved this section to after the discussion on customizing intRo. It now comes just before the conclusion.**
-
-P 11 near bottom: “as an application on the user’s account” puzzled me -- needs to say shinyapps.io account
-
-**Done**
-
-Section 3.2 mainly belongs in up-front motivation. Table 1 is an implementation mechanism.
-
-**We have merged remnants of Section 3.2 into Section 2 / motivation.
-We have also drastically cut down the content to focus this paper more on the extensible systems side, rather than the teaching side.
-As such, we have also removed Table 1.**
-
-P 14 top: “currently unsupported” justifications are more relevant to a teaching paper than a software design paper.
-
-**We have removed this comment.**
-
-P 14 Last para Section 3: could go into motivational discussion.
-
-**We have moved this paragraph to the motivational discussion.**
-
-Section 4.1 is a big part of what I see as the real story but descends to details too quickly.
-
-Most of Section 4.2 looks to me like more motivation, except the interpolate stuff which is an implementation mechanism
-
-**We have merged most of the content from the former Section 4.2 (on reproducibility) with Section 2, and removed a discussion on interpolation.**
-
-Section 4.3: more “main story”.
-
-**We agree and have emphasized this section as one of the two primary pieces to the main story, the other being modularity.**
+    **We have referenced this section on page 12**

### 2016-12-09 jcgs-paper-2016/paper.Rmd
New response and small edits

@@
 inZight Lite is currently partially but not fully reproducible.
-Another package is called Radiant [@radiant], which is a web-based application with the aim of furthering business education and financial analysis.
+Another package is called Radiant [@radiant], which is a very promising web-based application with the aim of furthering business education and financial analysis.
 While the application is modular and extensible, it does require local installation and hosting.
+Radiant is a business analytics platform with a host of functionality, while **intRo** focuses on being a simple and extensible web tool for learning.
 An overview of the comparison between the features of these statistical software packages is presented in Table \ref{tab:compare}.

### 2016-12-09 jcgs-paper-2016/paper.Rmd
Rebuild with fixed cite

@@
 First, it eases a student who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code.
-Seeing the correspondence between graphical clicks and printed code should help lessen the fear of coding that many students may have.
+Seeing the correspondence between graphical clicks and printed code lessens the fear of coding that many students may have.
 Second, an analysis created by a reproducible software system can be brought in an R session to easily assess and extend the results.
-Finally, with the help of knitr [@xie2015] and rmarkdown [@rmarkdown], "printing" the results of a reproducible software system analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
+Finally, with the help of knitr [@xie2015] and rmarkdown [@rmarkdown], "printing" the results of the analysis amounts to nothing more than executing the R code on the server, adding another layer of reproducibility.
 These concepts are important because they encourage best practices with regards to disclosure of analysis methods in research [@baggerly2011reproducible; @xie2015].

### 2016-12-09 jcgs-paper-2016/paper.tex
Rebuild with fixed cite

@@
 programming aspects of working directly with R. Upon the release of
-RStudio's Shiny (RStudio and Inc. 2014) it became easier for an R-based
+RStudio's Shiny (RStudio, Inc. 2014) it became easier for an R-based
 analysis to be converted to an interactive web application. Several
@@
 reproducible. Another package is called Radiant (Nijs 2016), which is a
-web-based application with the aim of furthering business education and
-financial analysis. While the application is modular and extensible, it
-does require local installation and hosting. An overview of the
-comparison between the features of these statistical software packages
-is presented in Table \ref{tab:compare}. Partial fulfillment of
-requirements is noted in the table, as well as a measure of the
-complexity of functionality offered by default. For example, \texttt{R}
-does have an associated Graphical User Interface (GUI), however this
-interface is very limited, thus only partially fulfilling the behavior
-of a GUI.
+very promising web-based application with the aim of furthering business
+education and financial analysis. While the application is modular and
+extensible, it does require local installation and hosting. Radiant is a
+business analytics platform with a host of functionality, while
+\textbf{intRo} focuses on being a simple and extensible web tool for
+learning. An overview of the comparison between the features of these
+statistical software packages is presented in Table \ref{tab:compare}.
+Partial fulfillment of requirements is noted in the table, as well as a
+measure of the complexity of functionality offered by default. For
+example, \texttt{R} does have an associated Graphical User Interface
+(GUI), however this interface is very limited, thus only partially
+fulfilling the behavior of a GUI.
 
@@
 really just a frontend for code. Seeing the correspondence between
-graphical clicks and printed code should help lessen the fear of coding
-that many students may have. Second, an analysis created by a
-reproducible software system can be brought in an R session to easily
-assess and extend the results. Finally, with the help of knitr (Xie
-2015) and rmarkdown (Allaire et al. 2014), ``printing'' the results of a
-reproducible software system analysis amounts to nothing more than
-executing the R code on the server, adding another layer of
-reproducibility. These concepts are important because they encourage
-best practices with regards to disclosure of analysis methods in
-research (Baggerly and Berry 2011; Xie 2015).
+graphical clicks and printed code lessens the fear of coding that many
+students may have. Second, an analysis created by a reproducible
+software system can be brought in an R session to easily assess and
+extend the results. Finally, with the help of knitr (Xie 2015) and
+rmarkdown (Allaire et al. 2014), ``printing'' the results of the
+analysis amounts to nothing more than executing the R code on the
+server, adding another layer of reproducibility. These concepts are
+important because they encourage best practices with regards to
+disclosure of analysis methods in research (Baggerly and Berry 2011; Xie
+2015).
 
@@
 Shiny observers are a class of reactive objects within the Shiny
-paradigm which do not return a value (RStudio and Inc. 2014). For
-further discussion of reactivity, see Section
-\ref{designing-for-reactivity}. In this example, observers are created
-to ensure that the choices of variable for the \texttt{nonparametric}
-module are only numeric variables. This is accomplished by utilizing the
-global reactive \texttt{intro.numericnames()}, which returns a character
-vector containing the variables in the current dataset that are numeric.
+paradigm which do not return a value (RStudio, Inc. 2014). For further
+discussion of reactivity, see Section \ref{designing-for-reactivity}. In
+this example, observers are created to ensure that the choices of
+variable for the \texttt{nonparametric} module are only numeric
+variables. This is accomplished by utilizing the global reactive
+\texttt{intro.numericnames()}, which returns a character vector
+containing the variables in the current dataset that are numeric.
 Finally, there is an event observer to store code generated from the
@@
 simultaneity), and data and computation dependencies'' (Bainomugisha et
-al. 2012). As implemented by Shiny, results automatically update when
+al. 2013). As implemented by Shiny, results automatically update when
 users interact with the interface.
@@
 \hypertarget{ref-bainomugisha2012survey}{}
-Bainomugisha, Engineer, Andoni Lombide Carreton, Tom Van Cutsem, Stijn
-Mostinckx, and Wolfgang De Meuter. 2012. ``A Survey on Reactive
-Programming.'' In \emph{ACM Computing Surveys}. Citeseer.
+Bainomugisha, Engineer, Andoni Lombide Carreton, Tom van Cutsem, Stijn
+Mostinckx, and Wolfgang de Meuter. 2013. ``A Survey on Reactive
+Programming.'' In \emph{ACM Comput. Surv.}, 45:52:1--52:34. 4. New York,
+NY, USA: ACM.
 
@@
 \hypertarget{ref-parnas1972criteria}{}
-Parnas, David Lorge. 1972. ``On the Criteria to Be Used in Decomposing
-Systems into Modules.'' \emph{Communications of the ACM} 15 (12). ACm:
+Parnas, D. L. 1972. ``On the Criteria to Be Used in Decomposing Systems
+into Modules.'' \emph{Commun. ACM} 15 (12). New York, NY, USA: ACM:
 1053--8.
@@
 \hypertarget{ref-shiny}{}
-RStudio, and Inc. 2014. \emph{Shiny: Web Application Framework for R}.
+RStudio, Inc. 2014. \emph{Shiny: Web Application Framework for R}.
 \url{http://CRAN.R-project.org/package=shiny}.
@@
 Engineering Potential and Challenges.'' \emph{Australian Computer
-Science Communications} 18. UNIVERSITY OF CANTERBURY: 203--12.
+Science Communications} 18. University of Cambridge: 203--12.
 

### 2016-12-09 jcgs-paper-2016/cover_letter.Rmd
Final submission

@@
 
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
 Re: Manuscript reference No. JCGS-16-045
 
-Attached you will find the newest revision of our paper, written in response to the feedback we received from the reviewers, the associate editor, and yourself.
+Attached you will find the newest revision of our paper. Contained is:
 
-We have attempted to restructure the paper to focus on general techniques for designing a modular software system, using intRo itself as a case study.
-In particular, we have re-written the title, abstract, and a substantial portion of the background to highlight this new direction.
+- **paper.pdf**, the non-blinded version of our paper
+- **paper_source.zip**, the source files for our paper, which include:
+    - **paper.tex**, the tex source
+    - **images**, a directory of figures
+    - **refs.bib**, the references
+    - **paper.pdf**, a copy of the non-blinded version of our paper
 
-In addition to the refocusing, we have also extended our review of other software systems, and emphasized the need that intRo addresses.
-Table 1 provides a somewhat subjective but thorough comparison of many "competing" software systems in terms of a number of qualities, including whether or not they are web-based, the level of extensibility, and the amount of functionality.
-We have also done our best to address the other minor comments that were raised in the second round of reviews.
+We also have included **forms.zip**, which has the two requested forms signed, as well as **response.pdf**, which has our final reviewer response.
 
-We hope that this latest revision addresses the majority of the issues with the previous, and we welcome any further critiques or comments.
+Let us know if there are any further concerns or requests.

### 2016-12-09 jcgs-paper-2016/final_submission/paper.tex
Final final submission

--- 2016-12-09 jcgs-paper-2016/paper.tex
@@
 business analytics platform with a host of functionality, while
-\textbf{intRo} focuses on being a simple and extensible web tool for
+\texttt{intRo} focuses on being a simple and extensible web tool for
 learning. An overview of the comparison between the features of these
@@
 measure of the complexity of functionality offered by default. For
-example, \texttt{R} does have an associated Graphical User Interface
-(GUI), however this interface is very limited, thus only partially
-fulfilling the behavior of a GUI.
+example, R does have an associated Graphical User Interface (GUI),
+however this interface is very limited, thus only partially fulfilling
+the behavior of a GUI.
 

### 2016-12-09 jcgs-paper-2016/paper.Rmd
Final final submission

@@
 While the application is modular and extensible, it does require local installation and hosting.
-Radiant is a business analytics platform with a host of functionality, while **intRo** focuses on being a simple and extensible web tool for learning.
+Radiant is a business analytics platform with a host of functionality, while `intRo` focuses on being a simple and extensible web tool for learning.
 An overview of the comparison between the features of these statistical software packages is presented in Table \ref{tab:compare}.
 Partial fulfillment of requirements is noted in the table, as well as a measure of the complexity of functionality offered by default.
-For example, `R` does have an associated Graphical User Interface (GUI), however this interface is very limited, thus only partially fulfilling the behavior of a GUI.
+For example, R does have an associated Graphical User Interface (GUI), however this interface is very limited, thus only partially fulfilling the behavior of a GUI.
 

### 2017-01-04 R/deploy_intRo.R
RSConnect instead of shinyapps

@@ -4,9 +4,9 @@
 #' @param enabled_modules The modules to enable
 #' @param theme The shinythemes theme to use
 #' @param google_analytics The google analytics tracking ID to use
-#' @param ... Additional arguments passed to the shinyapps deployApp function
+#' @param ... Additional arguments passed to the rsconnect deployApp function
 #' @export
-#' @importFrom shinyapps deployApp
+#' @importFrom rsconnect deployApp
 #' @examples
 #' \dontrun{
 #'     deploy_intRo()

### 2017-01-04 jcgs-paper-2016/final_submission/paper.tex
Use rsconnect instead of shinyapps

@@
 function \texttt{deploy\_intRo} which is a wrapper for the
-\texttt{deployApp} function contained in the shinyapps package. Once the
-shinyapps package is installed and configured, \texttt{deploy\_intRo}
+\texttt{deployApp} function contained in the rsconnect package. Once the
+rsconnect package is installed and configured, \texttt{deploy\_intRo}
 will upload \texttt{intRo} as an application on the instructor's

### 2017-01-04 jcgs-paper-2016/paper.Rmd
Use rsconnect instead of shinyapps

@@
 If the intent is to use a specific instance of `intRo` where many students will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL.
-The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package.
-Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the instructor's ShinyApps.io account.
+The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the rsconnect package.
+Once the rsconnect package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the instructor's ShinyApps.io account.
 The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme.

### 2019-07-28 jsm-2019/index.Rmd
started slides for jsm

---
title: "Designing Modular Software:"
subtitle: "A Case Study in Introductory Statistics"
author: "Eric Hare and **Andee Kaplan**"
institute: "Omni Analytics and Colorado State University"
date: "2019/07/29"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: center, middle

# Background

### Some things to think about when writing software

---

# What is modularity?

### The degree to which a system's components may be separated and recombined.

#### Advantages

- Functionality can be dynamically loaded and unloaded depending on the **particular use case**

- Open source modularity allows **extensions** contributed by many people

- A **flexible system** that can be tailored to keep software easy to use

---

# What else is important?

### Reproducibility

Encourage **best practices** with regards to disclosure of analysis methods in research

### Extensibility

**Expands** into areas that weren't anticipated early in development

- e.g. packages in `R`, macros in Excel, add-ins in JMP

### Ease of use

Help users **avoid frustration**

- e.g., web technologies eliminate local installation, easy to learn GUI

---
class: center, middle

# Case study: `intRo`

### Statistical Analysis Software for Teaching

---

# Full disclosure

.pull-left[
![](images/attn.png)
<p style="font-size: 0.9rem;font-style: italic;"><a href="https://svgsilh.com/e91e63/image/98739.html">"warning"</a> is licensed under <a href="https://creativecommons.org/licenses/cc0/1.0/?ref=ccsearch&atype=html" style="margin-right: 5px;">CC0 1.0</a><a href="https://creativecommons.org/licenses/cc0/1.0/?ref=ccsearch&atype=html" target="_blank" rel="noopener noreferrer" style="display: inline-block;white-space: none;opacity: .7;margin-top: 2px;margin-left: 3px;height: 22px !important;"><img style="height: inherit;margin-right: 3px;display: inline-block;" src="https://search.creativecommons.org/static/img/cc_icon.svg" /><img style="height: inherit;margin-right: 3px;display: inline-block;" src="https://search.creativecommons.org/static/img/cc-cc0_icon.svg" /></a></p>
]

.pull-right[
We have an **ulterior motive** to get students excited about programming!

- By navigating about the user interface of `intRo`, students are creating a fully-executable `R` script that they can download and run locally

- Viewing their script change real-time within the application

- This is a sneaky way to **spark curiosity** of students

]

---
class: center, inverse, middle

# Live demo

### http://intro-stats.com

---

# Design decisions

---

# Challenges (that have now been solved)

---

# Get started

---

class: center, middle

# Thanks!

http://www.intro-stats.com

.footnote[Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).]

### 2019-07-28 modules/data/sources/helper.R
things broke...

@@ -4,7 +4,8 @@ super_bowl_salaries <- read.csv("data/superbowl-salaries.csv")
 process_logical <- function(data, x) {
     if (is.null(x) || all(lapply(x, nchar) == 0)) return(data)
     
-    relevant_cols <- names(data)[nchar(x) > 0]
+    #relevant_cols <- names(data)[nchar(x) > 0]
+    relevant_cols <- names(data)[which(nchar(x) > 0) - 1]
     if (length(relevant_cols) == 0) return(data)
     
     my_strs <- strsplit(gsub(",", " , ", x), split = ",")

### 2019-07-28 scripts/top-nav-links.js

@@ -27,6 +27,7 @@ $('.fa-code').parent().parent().toggleClass('active');
 
 $('a:contains("hide_me")').parent().parent().parent().css("width", "100%");
 
+/*
 var space_width = (
 $('.container').innerWidth() - 
 $('.brand').outerWidth() -
@@ -34,6 +35,10 @@ eval($('.navbar a').map(function(){
     return $(this).parent().width();
 }).get().join("+")) + 
 $('a:contains("hide_me")').parent().width() - 94);
+*/
+
+var space_width = $('a:contains("hide_me")').parent().parent().parent().width() - 600;
+
 
 //spacing of navbar
 $('a:contains("hide_me")').css("visibility", "hidden");

### 2019-07-28 presentations/jsm-2019/index.Rmd
add presentation to web

--- 2019-07-28 jsm-2019/index.Rmd
@@
 ---
+background-image: url(images/app_creation_modules.png)
+background-position: 50% 100%
+background-size: contain
 
-# Design decisions
+# Modularity
+
+---
+# Modularity (Cont'd)
+The interface is created with the following statement
+
+```{r, tidy=FALSE, eval=FALSE}
+## Source ui
+module_info <- read.table("modules/modules.txt", header = TRUE, 
+                          sep=",")
+sapply(file.path("modules", dir("modules")[dir("modules") 
+                      != "modules.txt"], "ui.R"), source)
+## Create the UI
+shinyUI(navbarPage("intRo", theme = "bootstrap.min.css",
+               tabPanel(title="", icon=icon("home"),
+                        fluidRow(
+                          do.call(navlistPanel, #<<
+                              c(list(id = "side-nav", #<<
+                              widths = c(2, 10)), mylist)) #<<
+                        )
+               ), ...
+))
+```
+
+---
+# Modularity (Cont'd)
+The server functions are dynamically generated using a similar method
+
+```{r, eval=FALSE, tidy=FALSE}
+shinyServer(function(input, output, session) {
+  ## Module info
+  module_info <- read.table("modules/modules.txt", 
+                            header = TRUE, sep=",")
+  ## Modules
+  types <- c("helper.R", "static.R", "observe.R", 
+             "reactive.R", "output.R")
+  modules_tosource <- file.path("modules", 
+                                apply(expand.grid(
+                                module_info$module, types)
+                                , 1, paste, collapse = "/"))
+  ## Source the modules
+  for (mod in modules_tosource) {
+    source(mod, local = TRUE) #<<
+  }
+})
+```
 
@@
 
-# Challenges (that have now been solved)
+# Other design decisions
+
+**Reproducibility** and **Ease of Use**
+
+![](images/ui_annotate.png)
+
+---
+
+# But what about Extensibility?
+
+![](https://media.giphy.com/media/3og0INAY5MLmEBubyU/giphy.gif)
+
+Module creation is currently:
+
+- Undocumented
+- Entirely manual
+- Unnecessarily lengthy
+
+---
+
+# Challenges
+
+### Storing the user's results
+
+Users must be able to print results, but only the results they want.
+
+### Presenting only relevant options to the user
+
+![](images/graphical_table.png)
+
+---
+
+# Challenges (with even better solutions)
+
+### Producing clean and executable `R` code
+
+Code generated must be executable on the server, executable on the user's machine, and clean for a human to read.
+
+```{r, eval=FALSE}
+nonparametrictest <- function(intro.data, x, y, conflevel, althyp, hypval) {
+  interpolate(~(wilcox.test(x = df$x, y = df$y, #<<
+                            conf.level = conf,
+                            alternative = althyp,
+                            mu = hypval)),
+              df = quote(intro.data), x = x, y = y,
+              conf = conflevel, althyp = althyp, hypval = hypval,
+              mydir = userdir, `_env` = environment(),
+              file = "code_nonparametric.R")
+}
+```
+
+See https://rstudio.github.io/shinymeta for a new (**better**) implementation.
 
@@
 # Get started
+
+Access at http://intro-stats.com or **host your own**!
+
+```{r eval=FALSE}
+devtools::install_github("gammarama/intRo")
+```
+
+- `download_intRo()` - pull the latest version of the application.
+- `run_intRo()` - run the application with specified options, like `enabled_modules`.
+- `deploy_intRo()`- deploy to ShinyApps.io using your account and options.
+
+```{r eval=FALSE}
+download_intRo()
+run_intRo(enabled_modules = c("data/transform", "summaries/numerical"),
+          theme = "cerulean")
+deploy_intRo(enabled_modules = c("data/transform", "summaries/numerical"),
+          theme = "cerulean")
+```
 
